### 2021-04-06 chapters/introduction.tex
Initial Overleaf Import

\chapter{Introduction}
\documentclass[../main]{subfiles}
\begin{document}

\label{ch:introduction}
Context: what is the bigger scope of the problem you are trying to solve? Try to connect to societal/economical challenges.
Problem Analysis: Here you present your analysis of the problem situation that your research will address.
How does this problem manifest itself at your host organisation?
Also summarises existing scientific insight into the problem.
\glo

\section{Problem statement}

\subsection{Research questions}
To tackle these issues, we investigate ...

\subsection{Research method}

\section{Contributions}
Our research makes the following contributions:
\begin{enumerate}
	\item 1
	\item 2
	\item 3
\end{enumerate}

\section{Outline}
In Chapter~\ref{ch:background} we describe the background of this thesis. 
Chapter~\ref{ch:research} describes ... 
Results are shown in Chapter~\ref{ch:results} and discussed in Chapter~\ref{ch:discussion}. Chapter~\ref{ch:related_work}, contains the work related to this thesis.
Finally, we present our concluding remarks in Chapter~\ref{ch:conclusion} together with future work.

\end{document}

### 2021-04-13 acronym.tex
Added introduction + first draft of rq1.1

%helpful acronym examples
\newacronym{poc}{POC}{proof-of-concept}

\newacronym{sut}{SUT}{System Under Test}

\newacronym{strew}{STREW}{Software Testing and Reliability Early Warning}

\newacronym{sat}{SAT}{Software Analysis Toolkit}

\newacronym{tloc}{TLOC}{Lines Of Test Code}

\newacronym{loc}{LOC}{Lines of Code}

\newacronym{taime}{TAIME}{Test Suite Assessment and Improvement Method}

\newacronym{tqm}{TQM}{Test Quality Model}

\newacronym{tdd}{TDD}{Test-Driven Development}

\newacronym{cor}{COR}{Conditional Operator Replacement}

\newacronym{ror}{ROR}{Relational Operator Replacement}

\newacronym{ast}{AST}{Abstract Syntax Tree}

### 2021-04-13 chapters/2_introduction.tex
Added introduction + first draft of rq1.1

\documentclass[../main]{subfiles}
\begin{document}

\chapter{Introduction}
\label{ch:introduction}
Software testing is a critical part in the software engineering process. It is widely used and accepted in the industry for quality assurance. 
Software testing can detect software bugs during the development process and can also serve for regression purposes\cite{Bertolino2007SoftwareDreams}. 
During the software testing process developers write automated tests called unit tests. 
These tests can be run by testing frameworks such as JUnit and TestNG\cite{JUnit, TestNG}.
Monitoring the quality of test code has shown that tests with lower quality leads to more defect-prone production code\cite{Spadini2018OnQuality}. 
Research has also shown that as the size of production code grows, the size of the test code grows along\cite{Zaidman2008MiningCode}. 
Quality control on test suites is therefore important as the maintenance on tests can be difficult and generate risks if done incorrectly\cite{Horvath2015TestMetrics}.
These risks are related to the growing complexity and size which consequently lead to incomprehensible tests. 
One of these risks is the occurrence of test bugs i.e., test that fail even tough the program is correct(\textit{false positive}). 
A different one, which is even worse, are tests that do not fail when the program is not working as desired(\textit{false negative}).
The \textit{false negative} is a problem when breaking changes or not detected by the test suite.
To solve this issue we can measure the fault detecting capability of a test suite i.e., test suite effectiveness.
\newline
Test suite effectiveness is measured by the number of faulty versions of a \acrfull{sut} that are detected by a test suite. 
Real faults are unknown in advance, we can use mutation testing as a proxy measurement. 
There is statistically significant evidence that mutant detection correlates with real fault detection\cite{Just2014AreTesting}.
\newline
Mutation testing tools generate faulty versions of the program and then run the tests to determine
if the fault was detected. These faults, called mutants, are created by so-called mutators which mutate
specific statements in the source code. Each mutant represents a very small change to prevent changing
the overall functionality of the program\cite{Lipton1978HintsProgrammer}.
\newline
Mutation testing tools generate faulty versions of the program and then run the tests to determine if the fault was detected.
After all the mutants are executed the mutation score can be calculated. 
The mutation score is the percentage of mutants killed divided by the total amount of mutants executed.
If a large amount of mutants survives, it might be an indication that the test quality is insufficient as programming errors remain undetected.
Mutation testing has considerable drawbacks, such as the equivalent mutant problem and being resource expensive.

\section{Problem statement}
Mutation testing requires a significant amount of computational resources\cite{Kintis2016EffectiveMutation}. 
The high cost requirement is often a barrier for adopting mutation testing \cite{Pizzoleto2019}.
A lot of techniques and methods have been developed to improve the performance, however, most of these approaches are not as effective as mutation testing a full set of mutants\cite{Pizzoleto2019,Yao2014}. 
\newline
A technique to reduce the number of mutants or the number of test case executions is mutation clustering.
Mutation clustering aims to reduce the amount of mutants to be executed by clustering the mutants\cite{Ma2016,Yu2019PossibilityScope}.
Clustering is a technique used for the grouping of data, for pattern recognition and classification, based on some similar characteristics\cite{Alsabti1997}.
The clustering of mutants has been researched with promising results\cite{Ji2009,Wilinski2015,Ma2016}. For example Ma et al., \cite{Ma2016} clustered the mutants for expressions. 
However in their paper they indicate that the choosing expressions as the domain of their cluster is a limitation.
A recurring problem in the research of mutation clustering is to determine the means of domains as the centre of a cluster\cite{Ji2009,Wilinski2015,Ma2016,Wei2021SpectralTesting}.

\subsection{Research questions}
To reduce the cost of mutation testing, we try to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants. 
Our goal is to cluster any mutant that is generated.
We will do this with two different approaches.
Our first approach will be a qualitative approach and second a quantitative approach. 
We structure our research on the following research questions:
\newline
\textbf{Research Question 1}: What set of characteristics can we identify for clustering generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
\textbf{Research Question 1.1}: How do the existing mutation testing tools for Java compare to each other?
\newline
\textbf{Research Question 2}: How can we train a machine learning model to recognize and cluster generated mutants to reduce the amount executed, while maintaining effectiveness?

\subsection{Research method}
We start with comparing existing mutation testing tools. 
We will use the existing literature as a starting point and research what would be the best tool to use for this thesis.
\newline
After we have selected a tool we will research existing clustering techniques.
Based on the results of this research we select and implement one or more clustering algorithm that will cluster mutants. 
One or more mutants of each cluster will be executed by the mutation tooling.
We will then compare the mutation score of the clustered mutants with the mutation score of a full set of mutants.
\newline
To answer our second research question we will train a machine learning model to cluster mutants.
We can then repeat the comparison we did for research question one. 
\newline
The projects we choose for generating mutants will be open source and selected by \acrfull{loc}. 
We want to see how the algorithm/model performs on big and small projects.
These project must also require a test suite.
\newline
We focus on Java based projects because Java is one of the most popular programming languages\cite{Tiobe}.

\section{Contributions}
Our research makes the following contributions:
\begin{enumerate}
 \item A methodology to cluster generated mutants based on chosen characteristics to reduce the cost of mutation testing.
 \item A \acrfull{poc} which implements the qualitative methodology chosen and elaborated within the thesis.
 \item A \acrfull{poc} which implements the quantitative approach to cluster mutants. 
\end{enumerate}

\section{Outline}
In Chapter~\ref{ch:background} we describe the background of this thesis. 
\todo{finish this at the end}
Chapter~\ref{ch:research} describes ... 
Results are shown in Chapter~\ref{ch:results} and discussed in Chapter~\ref{ch:discussion}. Chapter~\ref{ch:related_work}, contains the work related to this thesis.
Finally, we present our concluding remarks in Chapter~\ref{ch:conclusion} together with future work.

\end{document}

### 2021-04-13 chapters/3_background.tex
Added introduction + first draft of rq1.1

\documentclass[../main]{subfiles}
\begin{document}

\chapter{Background}
\label{ch:background}
This chapter will present the necessary background information for this thesis. First, we define some basic terminology that will be used throughout this thesis.
Next, we will compare existing mutation testing tools. We start of by reviewing existing literature and then extend the comparison. \todo{extend this later on?}
\section{Terminology}
\todo{do this when we finished experiments and all}
\todo{subsuming mutants}

\end{document}

### 2021-04-13 chapters/4_mutation_tool.tex
Added introduction + first draft of rq1.1

\documentclass[../main]{subfiles}
\begin{document}

\chapter{Mutation testing tools}
\label{ch:mutation-tool}

Before we can start clustering mutants we need to select a tool that helps us generate mutants. 
Our goal is to find the tool that generates the greatest amount of mutants. 
\newline
First, we review existing literature. 
Secondly we extend the comparison by adding information about the current state of the tools.
With the combination of the information gathered we make a decision on what tool to use for this thesis.

\section{State of the art documented}
\label{ch:tooling-literature}
The literature used several criteria for comparing mutation testing tools for Java:
\begin{itemize}
  \item Execution time of each tool.
  \item Tool's application cost in terms of the number of test cases needs to be generated and the the number of equivalent mutants that would have to be inspected.
  \item Effectiveness of the mutation adequate test suite of each tool. Each test of the test suite contributes to the
effectiveness score, i.e., if one test is removed, less than 100\% effectiveness score is achieved. A
cross-testing technique is applied to evaluate the effectiveness each tool’s mutation adequate test
suite. The adequate test suite of each tool is run on the set of mutants generated by each other tool.
\end{itemize}
Kintis et al. analysed and compared the effectiveness of PIT,  MuJava and Major\cite{Kintis2016AnalysingStudy}. 
They started out with performing a mini literature survey on mutation testing tools used in research on test effectiveness in Java. 
With a data set of twelve methods and six Java projects they evaluated each tool using a cross-testing technique.
They found that the mutation adequate test suite of  MuJava was the most effective, followed by Major and PIT.
The score of the application cost was the inverse of the previous ranking: PIT required the least test cases and generated the smallest set of equivalent mutants.
During their research they also compared the different mutant operators the tools supported. 
An overview of mutant operators per tool was created. 
They compared the mutant operators and reported overlapping operators.
\newline
Marki and Lindstrom performed the research on the same set of mutation tools\cite{Marki2017MutationJava}. 
Similar to Kintis et al., they applied the same cross-testing technique for three small Java programs. 
These programs are often used in testing literature.
They found that none of the mutation tools subsume \todo{add subsumtion to word list} each other. 
If the tools were ranked,  MuJava would generate the strongest mutants followed by Major and PIT.
Furthermore they found that  MuJava generated significantly more equivalent mutants.
MuJava also had an execution time twice the amount of that of Major and PIT combined.
\newline
Laurent et al., introduced PIT+, an version of PIT with extended set of mutantors\cite{Laurent2017AssessingPIT}.
They used the tests suites generated by kintis et al., and combined these into an adequate test suite that would detect the combined set of mutants generated by PIT, muJava and Major. 
They found that the set of mutants generated by PIT+ was equally strong as the combined sets of mutants of all three tools. 
The extended set of mutants used in PIT+ is now integrated in the PIT project\cite{pit+}.

\section{State of the art tools}
The literature is clear about how the mutation testing tools compare to each other.
However the most recent study is from 2017.
At the time of writing this thesis the studies are at least four years old. 
Some of the tools have been developed further or are still actively developed\cite{pit-releases,Major}.
\newline
There are three candidate tools; Major,  MuJava and PIT.
We extended the existing mutant operator comparison for these tools with the operators that have been added since the publishing date of the literature.
We also researched the overlap for the new mutant operators.
\newpage
Table \ref{tab:mutators-current} shows the mutant operators that were developed after the publication of the literature. The amount of mutant operators for  MuJava and PIT have increased significantly while there were no new mutant operators for Major.
\begin{table}[htb]
\centering
\begin{tabular}{|l|l|l|l|}
\hline
\multicolumn{2}{|c|}{\textbf{ MuJava}}                           & \multicolumn{2}{|c|}{\textbf{PIT}}                      \\ \hline
IHD & Hiding variable deletion                                  & EM                    & Empty Returns                   \\ \hline
IHI & Hiding variable   insertion                               & FR                    & False Returns                   \\ \hline
IOD & Overriding method   deletion                              & NR                    & Null Returns                    \\ \hline
IOP & Overriding method   calling position change               & TR                    & True Returns                    \\ \hline
IOR & Overriding method   rename                                & PR                    & Primitive Returns               \\ \hline
ISI & super keyword insertion                                   & ER                    & Experimental Switch             \\ \hline
ISD & super keyword deletion                                    & BI                    & Big Integer                     \\ \hline
IPC & Explicit call to a   parent’s constructor deletion        & NRC                   & Naked Receiver                  \\ \hline
PNC & new method call with   child class type                   & N                     & Negation                        \\ \hline
PMD & Member variable   declaration with parent class type      & AOR                   & Arithmic Operator Replacement   \\ \hline
PPD & Parameter variable   declaration with child class type    & AOD                   & Arithmic Operator Deletion      \\ \hline
PCI & Type cast operator   insertion                            & CR                    & Constant replacement            \\ \hline
PCC & Cast type change                                          & BO                    & Bitwise Operator                \\ \hline
PCD & Type cast operator   deletion                             & ROR                   & Relational Operator Replacement \\ \hline
PRV & Reference assignment   with other comparable variable     & UOI                   & Unary Operation Insertion       \\ \hline
OMR & Overloading method   contents replace                     & \multicolumn{1}{l|}{} &                                 \\ \hline
OMD & Overloading method   deletion                             & \multicolumn{1}{l|}{} &                                 \\ \hline
OAC & Arguments of   overloading method call change             & \multicolumn{1}{l|}{} &                                 \\ \hline
JTI & this keyword insertion                                    & \multicolumn{1}{l|}{} &                                 \\ \hline
JTD & this keyword deletion                                     & \multicolumn{1}{l|}{} &                                 \\ \hline
JSI & static modifier   insertion                               & \multicolumn{1}{l|}{} &                                 \\ \hline
JSD & static modifier   deletion                                & \multicolumn{1}{l|}{} &                                 \\ \hline
JID & Member variable   initialization deletion                 & \multicolumn{1}{l|}{} &                                 \\ \hline
JDC & Java-supported default   constructor deletion             & \multicolumn{1}{l|}{} &                                 \\ \hline
EOA & Reference assignment   and content assignment replacement & \multicolumn{1}{l|}{} &                                 \\ \hline
EOC & Reference comparison   and content comparison replacement & \multicolumn{1}{l|}{} &                                 \\ \hline
EAM & Acessor method change                                     & \multicolumn{1}{l|}{} &                                 \\ \hline
EMM & Modifier method change                                    & \multicolumn{1}{l|}{} &                                 \\ \hline
\end{tabular}
\caption{\label{tab:mutators-current}Current mutators supported.}
\end{table}
\newpage
Tools sometimes implement the same mutator differently, resulting in different mutant sets. Table \ref{tab:mutators-overlap} shows an overview over the relations between the mutators of MuJava and PIT. This overview takes into account the mutators that are not already in the studies of Kintis et al., and Marki and Lindstrom. However some of the new mutators map to the existing mutators.

\begin{table}[htb]
\centering
\begin{tabular}{|l|l|l|l|}
\hline
\multicolumn{2}{|c|}{\textbf{ MuJava mutator}}      & \multicolumn{2}{c|}{\textbf{PIT mutator}}                         \\ \hline
AORB & Arithmetic Operator Replacement Binary      & \multirow{4}{*}{M}   & \multirow{4}{*}{Math}                      \\ \cline{1-2}
ASRS & Short-Cut Assignment   Operator Replacement &                      &                                            \\ \cline{1-2}
SOR  & Shift Operator   Replacement                &                      &                                            \\ \cline{1-2}
COR  & Conditional Operator   Replacement          &                      &                                            \\ \hline
AOIU & Arithmetic Operator   Insertion Unary       & \multirow{2}{*}{UOI} & \multirow{2}{*}{Unary Operation Insertion} \\ \cline{1-2}
AOIS & Arithmetic Operator   Insertion Short-cut   &                      &                                            \\ \hline
ODL  & Operator Deletion                           & AOD                  & Arithmic Operator   Deletion               \\ \hline
AODS & Arithmetic Operator   Deletion Short-cut    & RI                   & Remove Increments                          \\ \hline
ROR  & Relational Operator   Replacement           & CB                   & Conditionals Boundary                      \\ \hline
AORS & Arithmetic Operator   Replacement Short-Cut & I                    & Increments                                 \\ \hline
COD  & Conditional Operator   Deletion             & RC                   & Remove Conditionials                       \\ \hline
\end{tabular}
\caption{\label{tab:mutators-overlap}Overview of mutator overlaps between  MuJavav and PIT}
\end{table}
Next we tried to generate mutants with all three tools. 
All of the sources that we used for generating mutants are build with the build tools Maven or Gradle. 
Marki and Lindstrom\cite{Marki2017MutationJava} mention that Major supports Maven, but it is not documented anywhere.
For this reason we could not generate mutants with Major.
Table \ref{tab:mutants-generated} shows the amount of mutants generated per project and per tool. The results show that in all of the cases PIT generates more mutants than MuJava. In some cases the difference is significant. This difference can partially be explained by the fact that MuJava does not support source projects with Java version 1.7 or higher\cite{mujava}. All the the source code that uses features from this version of Java and higher resulted in an error during mutant generation.

\begin{table}[htb]
\centering
\begin{tabular}{|l|l|c|c|c|}
\hline
                                & \textbf{version} & \textbf{\begin{tabular}[c]{@{}c@{}}LOC\\ (Lines of code)\end{tabular}} & \multicolumn{1}{l|}{\textbf{MuJava}} & \multicolumn{1}{l|}{\textbf{PIT}} \\ \hline
org.apache.commons.numbers.core & 1.0-beta1        & 450                                                                    & 988                                  & 4384                              \\ \hline
Jodatime                        & 2.10.10          & 28811                                                                  & 52925                                & 112772                            \\ \hline
Zxing 3.4.1                     & 3.4.1            & 24792                                                                  & 65983                                & 161409                            \\ \hline
Google Auto Common              & 0.11             & 2338                                                                   & 20                                   & 5219                              \\ \hline
Google Auto Factory             & 1.0-beta9        & 1507                                                                   & 69                                   & 5832                              \\ \hline
Google Auto Value               & 1.7.5            & 7466                                                                   & 745                                  & 16746                             \\ \hline
Google Auto Service             & 1.0-rc6          & 750                                                                    & 0                                    & 475                               \\ \hline
ScribeJava-Core                 & 8.1.0            & 5709                                                                   & 1358                                 & 5746                              \\ \hline
Checkstyle                      & 8.41.1           & 38491                                                                  & 4529                                 & 100952                            \\ \hline
fastjson                        & 1.2.75           & 43405                                                                  & 70577                                & 116188                            \\ \hline
\end{tabular}
\caption{\label{tab:mutants-generated}Amount of mutants generated by MuJava and PIT}
\end{table}
\newpage
\section{Mutation tool}
We have studied literature on three different tools and extended the research with more recent data.
MuJava is not actively maintained and has not been updated in the last few years.
It does not support JUnit 4 and all versions of TestNG\cite{mujava}.
These are test engines\todo{add to vocabulary} and are crucial for executing tests.
MuJava also does not support source projects with java version 1.7 or higher\cite{mujava}.
Conforming to these requirements would decrease the set of projects we could use in or experiments.
\newline
While Major does support JUnit 4,
we could not generate mutants with this tool.
It would be too time consuming to customize all source projects to work with Major.
Furthermore there is not much documentation available for this tool.
\newline
PIT targets industry, is open source and actively maintained\cite{Kintis2016AnalysingStudy}.
It supports a wide scale of build tooling and is significantly faster than the other tools.
For example PIT provides a plugin system in which you can inject your own code in various stages of mutation testing\cite{pit}.
PIT also generated significantly more mutants in every project we have tested.
Based on these observations we decided that PIT would be the best choice for generating and executing mutants.
\end{document}

### 2021-04-17 chapters/3_background.tex
Wip until this date

@@
 \todo{subsuming mutants}
+\todo{levensteihn distance}
+\section{Project selection}
+We chose two main requirements for selecting software projects; the projects should have a test suite and the test suite should not contain failing tests.
+We started out by selecting projects that were also used in other research within the mutation testing and testing domain\cite{Pizzoleto2019,Yu2019PossibilityScope,Wei2021SpectralTesting, Zhang2019PredictiveTesting, Chen2018SpeedingStudy, Laurent2017AssessingPIT}.
+This resulted in six projects that satisfied the requirements.
+To extend our sample of projects we selected projects from the first six pages of the most popular Java projects on GitHub\footnote{\url{https://github.com/search?p=7&q=language\%3Ajava+stars\%3A\%3E10000&type=Repositories}}.
+From this selection we filtered out all the projects that contained failing tests that weren't libraries or applications.
+We ended up with a total sample of 21 projects. 
+These projects consists of Java libraries, Android libraries and Java applications.
+
+\section{Clustering algorithms}
 

### 2021-04-19 chapters/2_introduction.tex
Wip until this date
Added writeup for clustering algorithms + created chapter for rq1

@@
 Monitoring the quality of test code has shown that tests with lower quality leads to more defect-prone production code\cite{Spadini2018OnQuality}. 
-Research has also shown that as the size of production code grows, the size of the test code grows along\cite{Zaidman2008MiningCode}. 
+Research has also found that if the production code grows, the test code grows along with it\cite{Zaidman2008MiningCode}. 
 Quality control on test suites is therefore important as the maintenance on tests can be difficult and generate risks if done incorrectly\cite{Horvath2015TestMetrics}.
-These risks are related to the growing complexity and size which consequently lead to incomprehensible tests. 
-One of these risks is the occurrence of test bugs i.e., test that fail even tough the program is correct(\textit{false positive}). 
-A different one, which is even worse, are tests that do not fail when the program is not working as desired(\textit{false negative}).
+\newline
+The growing complexity and size of code bases consequently lead to incomprehensible tests.
+As a result test bugs can occur. 
+A test bug is a test that fail even tough the program is correct(\textit{false positive}). 
+Another bug, which is even worse, are tests that do not fail when the program is not working as desired(\textit{false negative}).
 The \textit{false negative} is a problem when breaking changes or not detected by the test suite.
-To solve this issue we can measure the fault detecting capability of a test suite i.e., test suite effectiveness.
 \newline
+To solve this issue we can measure the fault detecting capability of a test suite, this is called the test suite effectiveness.
 Test suite effectiveness is measured by the number of faulty versions of a \acrfull{sut} that are detected by a test suite.
 Real faults are unknown in advance, we can use mutation testing as a proxy measurement. 
-There is statistically significant evidence that mutant detection correlates with real fault detection\cite{Just2014AreTesting}.
+Research has found statistically significant evidence which correlates mutant detection with real fault detection\cite{Just2014AreTesting}.
 \newline
-Mutation testing tools generate faulty versions of the program and then run the tests to determine
-if the fault was detected. These faults, called mutants, are created by so-called mutators which mutate
-specific statements in the source code. Each mutant represents a very small change to prevent changing
-the overall functionality of the program\cite{Lipton1978HintsProgrammer}.
+Mutation testing tools create small faulty versions of the program.
+The test suite is then run to verify if these faults are detected.
+These faults are called mutants and are created by mutators.
+Mutators mutate specific expressions or statements in the source code. 
+To prevent changing the overall functionality of the program, mutants are represented as very small changes\cite{Lipton1978HintsProgrammer}.
 \newline
-Mutation testing tools generate faulty versions of the program and then run the tests to determine if the fault was detected.
 After all the mutants are executed the mutation score can be calculated. 
 The mutation score is the percentage of mutants killed divided by the total amount of mutants executed.
-If a large amount of mutants survives, it might be an indication that the test quality is insufficient as programming errors remain undetected.
+If a large amount of mutants survives, it might be an indication that the quality of the test suite is poor as programming errors remain undetected.
 Mutation testing has considerable drawbacks, such as the equivalent mutant problem and being resource expensive.
@@
 \newline
-The projects we choose for generating mutants will be open source and selected by \acrfull{loc}. 
 We want to see how the algorithm/model performs on big and small projects.
@@
 \newline
-We focus on Java based projects because Java is one of the most popular programming languages\cite{Tiobe}.
+We chose to focus on projects that use Java as main programming language because it is one of the most popular languages\cite{Tiobe}.
 

### 2021-04-19 chapters/3_background.tex
Added writeup for clustering algorithms + created chapter for rq1

@@
 \todo{levensteihn distance}
+\todo{maven and gradle}
 \section{Project selection}
@@
 \section{Clustering algorithms}
-
+This research focuses on clustering mutants and use of existing clustering algorithms. 
+Many different types of clustering algorithms have been proposed\cite{Rodriguez2019}. While there is a lot of diversity, some methods are more frequently used than others\cite{Wu2008TopMining}. 
+Rodriguez et al., compared the performance of nine different clustering algorithms:
+\begin{itemize}
+    \item k-means
+    \item CLARA
+    \item hierarchical
+    \item EM
+    \item hcmodel
+    \item spectral
+    \item subscpace
+    \item optics
+    \item dbscan
+\end{itemize}
+The goal of their study was to guide researchers, who have little experience in data mining techniques, to the application of clustering methods.
+They evaluated the algorithms in three distinct situations: default parameters, single parameter variation and random variation of parameters.
+They used 400 generated artificial data sets which were normally distributed.
+\newline
+The results reported in their research are respective to specific configurations of normally distributed data and algorithmic implementations.
+Nonetheless they do give a good overview on how the algorithms compare to each other.
+\newline
+For the default parameter situation the spectral clustering algorithm had the best performance and the hierarchical algorithm had the worst performance.
+\newline
+Regarding single parameter variations, for data sets containing 2 columns, the hierarchical, optics and EM methods showed significant performance variation.
+\newline
+With respect to the multidimensional analysis for data sets, the performance of the algorithms for the multidimensional selection of parameters was similar to that using the default parameters.
+They conclude their research with observing that, for data sets with 10 or more columns the spectral algorithm consistently provided the best performance.
+However the EM, hierarchical, k- means and subspace algorithms can also achieve similar performance with some parameter tuning.
+The optics and dbscan algorithms aim at different data distributions than performed in this study. There different results could be obtained for non-normally distributed data.
 \end{document}

### 2021-04-19 chapters/4_mutation_tool.tex
Added writeup for clustering algorithms + created chapter for rq1

@@
   \item Execution time of each tool.
-  \item Tool's application cost in terms of the number of test cases needs to be generated and the the number of equivalent mutants that would have to be inspected.
-  \item Effectiveness of the mutation adequate test suite of each tool. Each test of the test suite contributes to the
-effectiveness score, i.e., if one test is removed, less than 100\% effectiveness score is achieved. A
-cross-testing technique is applied to evaluate the effectiveness each tool’s mutation adequate test
-suite. The adequate test suite of each tool is run on the set of mutants generated by each other tool.
+  \item The cost as the number of test cases that needs to be generated and the the number of equivalent mutants that would have to be inspected.
+  \item Effectiveness of the mutation adequate test suite of each tool.
+Every test of the suite adds to the effectiveness score.
+If a test is removed less than 100\% effectiveness can be achieved.
+To evaluate the effectiveness of each tool's mutation adequate test suite, a cross-testing technique is applied.
+The adequate test suite per tool is run on the set of mutants generated by the other tools.
 \end{itemize}
-Kintis et al. analysed and compared the effectiveness of PIT,  MuJava and Major\cite{Kintis2016AnalysingStudy}. 
+Kintis et al. analysed and compared the performance of PIT,  MuJava and Major\cite{Kintis2016AnalysingStudy}. 
 They started out with performing a mini literature survey on mutation testing tools used in research on test effectiveness in Java. 
 With a data set of twelve methods and six Java projects they evaluated each tool using a cross-testing technique.
-They found that the mutation adequate test suite of  MuJava was the most effective, followed by Major and PIT.
-The score of the application cost was the inverse of the previous ranking: PIT required the least test cases and generated the smallest set of equivalent mutants.
+As a result they found that the most effective mutation adequate test suite was from MuJava, followed by Major and PIT.
+The score of the application cost was the inverse of the previous ranking: PIT generated the smallest set of equivalent mutants and required the least test cases.
 During their research they also compared the different mutant operators the tools supported. 
@@
 \newline
-Marki and Lindstrom performed the research on the same set of mutation tools\cite{Marki2017MutationJava}. 
-Similar to Kintis et al., they applied the same cross-testing technique for three small Java programs. 
-These programs are often used in testing literature.
-They found that none of the mutation tools subsume \todo{add subsumtion to word list} each other. 
-If the tools were ranked,  MuJava would generate the strongest mutants followed by Major and PIT.
+Marki and Lindstrom performed their research on the same three mutation testing tools.\cite{Marki2017MutationJava}. 
+The same cross-testing technique used by Kintis et al., was applied for three small Java programs. 
+These programs are popular programs used in testing literature.
+They found that the mutation tools do not subsume \todo{add subsumption to word list} each other. 
+According to the research a ranking of the tools would be as follows; MuJava generated the strongest mutants followed by Major and PIT.
 Furthermore they found that MuJava generated significantly more equivalent mutants.
@@
 \newline
-Laurent et al., introduced PIT+, an version of PIT with extended set of mutantors\cite{Laurent2017AssessingPIT}.
-They used the tests suites generated by kintis et al., and combined these into an adequate test suite that would detect the combined set of mutants generated by PIT, muJava and Major. 
-They found that the set of mutants generated by PIT+ was equally strong as the combined sets of mutants of all three tools. 
-The extended set of mutants used in PIT+ is now integrated in the PIT project\cite{pit+}.
+Laurent et al., introduced PIT+, a version of PIT with extended set of mutators\cite{Laurent2017AssessingPIT}.
+They used the same tests suites that were generated by kintis et al., and combined these into an adequate test suite that would detect the combined set of mutants generated by PIT, MuJava and Major. 
+They discovered that the set of mutants generated by PIT+ was stronger than the combined sets generated by the other three tools.
+Fortunately the extended set of mutants used in PIT+ is now integrated in the PIT project\cite{pit+}.
 
@@
 We also researched the overlap for the new mutant operators.
-\newpage
 Table \ref{tab:mutators-current} shows the mutant operators that were developed after the publication of the literature.
@@
 All of the sources that we used for generating mutants are build with the build tools Maven or Gradle. 
-Marki and Lindstrom\cite{Marki2017MutationJava} mention that Major supports Maven, but it is not documented anywhere.
+Marki and Lindstrom\cite{Marki2017MutationJava} mention that Major includes support for Maven, but it is not documented anywhere.
 For this reason we could not generate mutants with Major.
@@
 MuJava also does not support source projects with java version 1.7 or higher\cite{mujava}.
-Conforming to these requirements would decrease the set of projects we could use in or experiments.
+Conforming to these requirements would recude the set of projects we could use in or experiments.
 \newline
-While Major does support JUnit 4,
-we could not generate mutants with this tool.
+While Major supports JUnit 4,
+we did not succeed in generating mutants with this tool.
 It would be too time consuming to customize all source projects to work with Major.
@@
 \newline
-PIT targets industry, is open source and actively maintained\cite{Kintis2016AnalysingStudy}.
-It supports a wide scale of build tooling and is significantly faster than the other tools.
+PIT targets the industry, is actively maintained and is open source\cite{Kintis2016AnalysingStudy}.
+It supports Maven, Gradle, has a \acrfull{cli} and has a faster execution time than the other tools.
 For example PIT provides a plugin system in which you can inject your own code in various stages of mutation testing\cite{pit}.
 PIT also generated significantly more mutants in every project we have tested.
-Based on these observations we decided that PIT would be the best choice for generating and executing mutants.
+Based on the information presented in this chapter we dcecided that PIT would be the best choice for generating and executing mutants.
 \end{document}

### 2021-04-19 chapters/5_clustering_similarity.tex
Added writeup for clustering algorithms + created chapter for rq1

\documentclass[../main]{subfiles}
\begin{document}

\chapter{Clustering on similarity}\todo{do we add dissimilarity?}
\label{ch:clustering_similarity}
The similarity between a mutant and the original code it is mutated from can be defined as a characteristic. 
We can then cluster on similarity. 
In this chapter we implement a hierarchical clustering algorithm to cluster on the similarity of a mutant.

\section{Similarity measure}

\section{Hierarchical clustering}
\end{document}

### 2021-05-28 chapters/2_introduction.tex
v1.0 RQ1

@@
 
-\subsection{Research questions}
 To reduce the cost of mutation testing, we try to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants. 
@@
 \textbf{Research Question 2}: How can we train a machine learning model to recognize and cluster generated mutants to reduce the amount executed, while maintaining effectiveness?
-
-\subsection{Research method}
+\newline
 We start with comparing existing mutation testing tools. 
@@
 After we have selected a tool we will research existing clustering techniques.
-Based on the results of this research we select and implement one or more clustering algorithm that will cluster mutants. 
+Based on the results of this research we select and implement a clustering algorithm that will cluster a characteristic of the mutants. 
 One or more mutants of each cluster will be executed by the mutation tooling.

### 2021-05-28 chapters/3_background.tex
Starting with rq1 writeup
v1.0 RQ1

@@
 \section{Project selection}
-We chose two main requirements for selecting software projects; the projects should have a test suite and the test suite should not contain failing tests.
-We started out by selecting projects that were also used in other research within the mutation testing and testing domain\cite{Pizzoleto2019,Yu2019PossibilityScope,Wei2021SpectralTesting, Zhang2019PredictiveTesting, Chen2018SpeedingStudy, Laurent2017AssessingPIT}.
-This resulted in six projects that satisfied the requirements.
-To extend our sample of projects we selected projects from the first six pages of the most popular Java projects on GitHub\footnote{\url{https://github.com/search?p=7&q=language\%3Ajava+stars\%3A\%3E10000&type=Repositories}}.
-From this selection we filtered out all the projects that contained failing tests that weren't libraries or applications.
-We ended up with a total sample of 21 projects. 
-These projects consists of Java libraries, Android libraries and Java applications.
+We choose three main requirements for selecting software projects; the projects should have a test suite, the test suite should not contain failing tests and the mutation testing tool should be able to execute mutants for the sample project.
+We selected projects that were also used in other research within the mutation testing and testing domain\cite{Pizzoleto2019,Yu2019PossibilityScope,Wei2021SpectralTesting, Zhang2019PredictiveTesting, Chen2018SpeedingStudy, Laurent2017AssessingPIT}.
+We extend our sample by selecting projects from the first six pages of the most popular Java projects on GitHub\footnote{\url{https://github.com/search?p=7&q=language\%3Ajava+stars\%3A\%3E10000&type=Repositories}}.
+The unfiltered sample contains 50+ projects.
+From there we filter out all the projects that contain failing tests and that aren't libraries or applications.
+The result consists of a sample with sixteen projects in total.
 

### 2021-05-28 chapters/5_identifying_characteristics.tex
v1.0 RQ1

\documentclass[../main]{subfiles}
\begin{document}

\chapter{Identifying characteristics}
\label{ch:clustering_characteristics}
Our goal is to cluster mutants based on characteristics. 
To achieve this we need to identify characteristics that the mutants contain. 
In this chapter we identify characteristics which we can use for clustering.

\section{Overview of characteristics}
Zhang et al., identifies several mutant characteristics\cite{Zhang2019PredictiveTesting}. 
Their research focuses on detecting if a mutant survives or not before executing the mutant.
While the goal of their research is different than ours, we can still use the characteristics they identified.
We select characteristics relevant to our research from the research of Zhang et al.
We extend the list of characteristics with data we can extract from our selected mutation testing tool. 
\newline
PIT generates and executes mutants on byte code level\cite{pitestBytecode}.
By mutating in Java byte code we can identify characteristics specific to Java byte code.
PIT also gathers mutant metadata.
We can extract this mutant metadata and identify characteristics. 
With the combination of information extracted from PIT and characteristics selected from existing research we identify the following characteristics:
\begin{enumerate}
    \item Mutant similarity.
    \item Amount of tests the mutant is challenged by.
    \item Mutator identifier.
    \item Mutant opcode.
    \item Mutant return type.
    \item Mutant contains local variables.
    \item Mutant is in try catch block.
    \item Mutant is in finally block.
    \item Number of mutated byte code instructions.
    \item Name of class that contains the mutant.
    \item Name of function that contains the mutant.
    \item Line number of block that contains the mutant.
    \item Line number on which the mutation occur.
    \item Number of lines of byte code of mutated class.
\end{enumerate}
In the following subsections we will elaborate more on these characteristics.

\section{Mutant similarity}
A piece of code is a textual representation of instructions for a computer.
A mutant is a transformation of a piece of code.
This transformation is predefined.
The difference between a mutant and its parent can be defined as similarity.
We identify the similarity between a mutant and parent as a characteristic.
\newline
To measure this characteristic we select a similarity measure.
There are existing similarities measures that base their similarity in different ways\cite{Gomma2013AApproaches}. 
We select the Levenshtein distance as similarity measure for our research\cite{Gomma2013AApproaches}.
The Levenshtein distance defines the distance between two strings by counting the minimum number of operations(create/edit/delete) needed to transform one string into the other\cite{Gomma2013AApproaches}.
The more the textual representation of a mutant differs from its parent the bigger the Levenshtein distance will be.

\section{Amount of tests challenged by mutant}
A mutant may be challenged by tests.
If the test fails, the mutant is considered killed.
If the test passes, the mutant is considered survived.
If the mutant is not challenged by any tests it is also considered survived.
The amount of tests a mutant is challenged by is dependant on the coverage and test effectiveness of the test suite.
We identify this number as a characteristic.

\section{Mutator identifier}
At the time of writing the selected mutation testing tool PIT supports more than 100 mutators\cite{pitestMutators}.
A mutator is the definition of how the code is mutated from its original\cite{Offutt2001}.
Each mutator mutates the source code in a dfferent way.
PIT assigns identifiers to their mutators for execution and reporting purposes. 
Each mutant has a mutator identifier that matches with how the mutant is mutated from the source code.
We can extract this data from PIT and use it as a characteristic.

\section{Mutant opcode}
\say{A Java Virtual Machine instruction consists of a one-byte opcode specifying the operation to be performed, followed by zero or more operands supplying arguments or data that are used by the operation.} (Java 12 SE docs ch.12).
When PIT mutates the source code it uses Java opcodes to alter the original byte code. 
A record of which opcode is being used is stored in memory during the mutation testing process. 
This opcode can differ per mutator and is dependant on how the code is written.
During the mutation testing process of PIT we can extract the opcode.
We identify the opcode as a characteristic.

\section{Mutant return type}
A mutant may contain a return statement. 
This is dependant on the source code that is mutated.
A return statement can be identified as a characteristic.
If a mutant contains a return statement, we can extract its type from PIT.
There is the possibility that a mutant does not contain a return statement, in this case we will set the mutant return type to a default value.

\section{Mutant location}
The location of a mutant in PIT is defined by multiple properties;
the name of the class, the name of the function, the line number of the start of the block the mutant is located in and the line number the mutation occurs on.
These properties refer to the location in Java byte code.
The location can be identified as a characteristic of a mutant.
PIT uses this location in the mutation testing and reporting processes.
We can extract the location as defined by the properties in PIT.
Each property is identified as a characters

\section{Extracting characteristics}
PIT exposes some of the characteristics with its plugin system.
Unfortunately not all characteristics are readily available for extraction.
To extract all the characteristics from PIT we create a modified version of the application.
We adjust the existing code to export the characteristics per mutant in various stages of the mutation testing process.
We create a new model that stores the listed characteristics trough out the mutation testing process. 
The data can be written to a file during the reporting stage of the process.

\end{document}

### 2021-05-28 chapters/6_experiment_design.tex
v1.0 RQ1

\documentclass[../main]{subfiles}
\begin{document}
\chapter{Experiment design}
\todo{expand this}
We hypothesise that there is a we can cluster on the characteristics of a mutant such that each mutant should be a representative for all of the mutants in its cluster.
Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.

We design an experiment in which we implement the following steps:
\begin{enumerate}
    \item Extract all the characteristics.
    \item For each mutant calculate the Levenshtein distance.
    \item Cluster the mutants based on the characteristics.
    \item Randomly select one mutant from each cluster.
    \item Execute the selected mutants in PIT and calculate the mutation score.
\end{enumerate}
We elaborate on these steps in the following subsections.

\subsection{Levenshtein distance on Java byte code}
A mutant is a piece of code that differs in a predefined way from its parent.
Java source code compiles into Java byte code.
During the compilation of Java code certain context is abstracted away\cite{byteCodeEngineering}.
Optimisations are applied which also changes the Java byte code\cite{byteCodeEngineering}.
As a result the textual similarity between a Java code mutant and its parent is different from the similarity between a Java byte code mutant and its parent.
While the textual similarity may be different between the Java code and Java byte code the functionality remains the same.
Java byte code reflects more of the semantic nature of the source code than the source code itself does. 
\newline
PIT generates and executes mutants on byte code level\cite{pitestBytecode}. 
In other words PIT executes the unit tests of a source against the byte code of a mutant.
\newline
For this experiment we will use the Levenshtein distance between a Java byte code mutant and its parent.
By calculating the distance on byte code we filter out the context that is present in Java code.
This gives us a Levenshtein distance that represents more of the semantic difference, between a mutant and parent, than calculating the distance for Java code.

\subsection{Levenshtein distance implementation}
write here that we use a memory efficient algorithm

\subsection{Hierarchical clustering}
Clustering Levenshtein distances has been done before with hierarchical clustering\cite{Rajalingam2011HierarchicalStudy, Gothai2010PerformanceAlgorithms}. 
Hierarchical clustering is subdivided into agglomerative and divisive. 
The agglomerative hierarchical technique follows bottom up approach whereas divisive follows top-down approaches.
Hierarchical clustering uses different metrics which measures the the distance between two clusters and the linkage criteria\cite{Rajalingam2011HierarchicalStudy}. 
The linkage criteria specifies the dissimilarity in the sets as a function of the pair-wise distances of observations in those sets\cite{Rajalingam2011HierarchicalStudy}.
\newline
Research shows that the complete linkage outperforms the single linkage method\cite{Vijaya2019ComparativeClustering}.
The ward linkage and complete linkage methods perform the same when clusters are well separated\cite{Vijaya2019ComparativeClustering}.
However if the clusters overlap the ward linkage outperforms the complete linkage\cite{Vijaya2019ComparativeClustering}.
\newline
We identify each Levenshtein distance between a mutant and its parent as a separate cluster.
Starting out with each mutant as a separate cluster we can use the agglomerative form of hierarchical clustering.
Since we cannot assume that our clusters are well separated we chose to use the ward linkage method for our clustering algorithm.
The agglormorative hierarchical clustering algorithm with ward linkage will cluster our mutants represented as Levenshtein distances.
\todo{write about implementation}

\subsection{Weighed mutation score}
Our hypothesis states that each mutants should be a representative for all the mutants in its cluster. 
This translates to if a mutant survives or is killed all mutants in that cluster should survive or be killed respectively.
To reflect this in our experiment a weighed mutation score is calculated for the clustered mutants.
This weighed mutation score is the product of the result of a mutant(1 for killed and 0 for survived) and the amount of mutants in the cluster it represents.
The weighed mutation score is then comparable to the score of a full set as the total number of mutants will be the same.
\newline
For example, take a full set with a score of 75/100 killed mutants. \todo{example too much?}
This gives us a mutation score of 75\%. 
We then cluster the mutants in four clusters consisting of 12, 30, 38 and 20 mutants, respectively.
We randomly select four mutants of each cluster and execute them.
The mutants representing cluster one and four survive.
If we calculate the weighed score we get 68/100 which is 68\%.
We can then compare this to the score of a full set because the amount of mutants executed is the same: 75/100(75\%) and 68/100(68\%).

\subsection{PIT plugin system}
PIT offers a plugin system in which developers can inject their own functionality in PIT\cite{pitestPlugin}. 
There are two main types of plugins; a Mutation Result Listener and a Mutation interceptor\cite{pitestPlugin}.
A mutation result listener  receives the details of analysed mutations as they arrive\cite{pitestPlugin}.
A Mutation interceptor passed a complete list of all mutation that will be generated to each class before the mutation are challenged by tests\cite{pitestPlugin}.
The implementation of our experiment makes use of this plugins system.
\newline
We developed three separate plugins for our experiment.
The first plugin is a Mutation interceptor.
It analyses each mutant, calculates the Levenshtein distance and writes a mutant identifier and distance to a \acrshort{csv} file. 
It also clears the list of mutants provided by PIT to prevent execution. 
The second plugin is also a Mutation interceptor.
It reads the clustered mutants from a \acrshort{csv} file and compares them to the list of mutants provided by PIT.
Only the clustered mutants are passed back to PIT for execution.
The last plugin is a mutation result listener. 
This plugin calculates the weighed mutation score and outputs it to a text file.
\newline
The plugins in PIT are executed based on type and per source class.
This means that a plugin will never receive all the mutants generated for a whole project during execution.
Because of how PIT executes its plugins we have to execute PIT twice.
In the first run we calculate the Levenshtein distance for all mutants and in the second run we execute the selected mutants and calculate the weighed mutation score.
\end{document}

### 2021-06-04 chapters/4_mutation_tool.tex
Starting with rq1 writeup
WIP exp1 design + finished characteristics

@@
 \chapter{Mutation testing tools}
-\label{ch:mutation-tool}
+\label{ch:mutation∂-tool}
 
@@
 However the most recent study is from 2017.
-At the time of writing this thesis the studies are at least four years old. 
+At the time of writing this thesis these studies are at least four years old. 
 Some of the tools have been developed further or are still actively developed\cite{pit-releases,Major}.
@@
 Table \ref{tab:mutators-overlap} shows an overview over the relations between the mutators of MuJava and PIT.
-This overview takes into account the mutators that are not already in the studies of Kintis et al., and Marki and Lindstrom.
+This overview does include the mutators that have been reviewed in the studies of Kintis et al., and Marki and Lindstrom.
 However some of the new mutators map to the existing mutators.
@@
 COD  & Conditional Operator   Deletion             & RC                   & Remove Conditionials                       \\ \hline
+
 \end{tabular}
@@
                                 & \textbf{version} & \textbf{\begin{tabular}[c]{@{}c@{}}LOC\\ (Lines of code)\end{tabular}} & \multicolumn{1}{l|}{\textbf{MuJava}} & \multicolumn{1}{l|}{\textbf{PIT}} \\ \hline
-org.apache.commons.numbers.core & 1.0-beta1        & 450                                                                    & 988                                  & 4384                              \\ \hline
+Commons-numbers.core            & 1.0-beta1        & 450                                                                    & 988                                  & 4384                              \\ \hline
 Jodatime                        & 2.10.10          & 28811                                                                  & 52925                                & 112772                            \\ \hline
 Zxing 3.4.1                     & 3.4.1            & 24792                                                                  & 65983                                & 161409                            \\ \hline
-Google Auto Common              & 0.11             & 2338                                                                   & 20                                   & 5219                              \\ \hline
+Google Auto Common              & 0.11.0           & 2338                                                                   & 20                                   & 5219                              \\ \hline
 Google Auto Factory             & 1.0-beta9        & 1507                                                                   & 69                                   & 5832                              \\ \hline
@@
 Checkstyle                      & 8.41.1           & 38491                                                                  & 4529                                 & 100952                            \\ \hline
-fastjson                        & 1.2.75           & 43405                                                                  & 70577                                & 116188                            \\ \hline
+Fastjson                        & 1.2.75           & 43405                                                                  & 70577                                & 116188                            \\ \hline
+Jfreechart                      & 1.5.3            & 91876                                                                  & 0                                    & 350741                            \\ \hline
+Commons-lang                    & 3.12.0           & 29836                                                                  & 31145                                & 134764                            \\ \hline
+Commons-codec                   & 1.15.0           & 9656                                                                   & 21719                                & 54804                            \\ \hline
+Commons-text                   & 1.9.0             & 9781                                                                   & 10403                                & 48490                            \\ \hline
+Commons-io                     & 2.8.0             & 13947                                                                  & 13245                                & 44631                            \\ \hline
+Gson                           & 2.8.6             & 8078                                                                   & 9198                                 & 28485                            \\ \hline
+Commons-cli                    & 1.4.0             & 2782                                                                   & 831                                  & 7193                             \\ \hline
+Commons-csv                    & 1.8.0             & 1855                                                                   & 2081                                 & 6906                             \\ \hline
 \end{tabular}

### 2021-06-04 chapters/5_identifying_characteristics.tex
WIP exp1 design + finished characteristics

@@
 To achieve this we need to identify characteristics that the mutants contain. 
-In this chapter we identify characteristics which we can use for clustering.
+In this chapter we identify characteristics that are relevant for defining clusters.
 
@@
 Their research focuses on detecting if a mutant survives or not before executing the mutant.
+The mutant characteristics identified by Zhang et al., have been used in other research with promising results\cite{Oonk2021}.
 While the goal of their research is different than ours, we can still use the characteristics they identified.
-We select characteristics relevant to our research from the research of Zhang et al.
+We select characteristics relevant to our research from the research of Zhang et al and Oonk.
 We extend the list of characteristics with data we can extract from our selected mutation testing tool. 
@@
     \item Mutant is in finally block.
-    \item Number of mutated byte code instructions.
     \item Name of class that contains the mutant.
     \item Name of function that contains the mutant.
-    \item Line number of block that contains the mutant.
+    \item Line number of the start of the block that contains the mutant.
     \item Line number on which the mutation occur.
-    \item Number of lines of byte code of mutated class.
 \end{enumerate}
@@
 The more the textual representation of a mutant differs from its parent the bigger the Levenshtein distance will be.
+A unit test tests a small piece of code.
+If this piece of codes differs too much from what is being tested the test is more likely fail. 
+The Levenshtein distance measures this difference.
+This could be a useful characteristic because mutants that have the same similarity could be grouped together.
 
@@
 We identify this number as a characteristic.
+Every test the mutant is challenged by is an occasion for the mutant to be killed or not.
+This could be a useful characteristic because mutants that have more chances to be killed could be grouped together.
 
@@
 We can extract this data from PIT and use it as a characteristic.
+It could be a useful characteristic because mutants that are changed in the same predefined way could be grouped together.
 
@@
 We identify the opcode as a characteristic.
+The opcode could be a useful characteristic because mutants with the same opcode may be grouped together.
 
@@
 There is the possibility that a mutant does not contain a return statement, in this case we will set the mutant return type to a default value.
+The return type is a distinctive property of a piece of code, this distinction a useful characteristic because mutants with the same return type may be grouped together.
 
-\section{Mutant location}
+\section{Mutant location}\label{ch:characteristics_location}
 The location of a mutant in PIT is defined by multiple properties;
@@
 Each property is identified as a characters
+The location of a piece of code is a distinctive property. 
+The same piece of code in a different location may have a different effect on the source code or unit tests.
+A mutant may not be unique if the location is not taken into account.
+It could be a useful characteristic because mutants that are in the same location may be grouped together.
 
-\section{Extracting characteristics}
-PIT exposes some of the characteristics with its plugin system.
-Unfortunately not all characteristics are readily available for extraction.
-To extract all the characteristics from PIT we create a modified version of the application.
-We adjust the existing code to export the characteristics per mutant in various stages of the mutation testing process.
-We create a new model that stores the listed characteristics trough out the mutation testing process. 
-The data can be written to a file during the reporting stage of the process.
+\section{Error handling capability}
+A technique to handle errors in Java is to use try-catch blocks\cite{javaDocs}. 
+The presence of try-catch blocks are an indication of the error handling capabilities of certain pieces of code.
+An addition to the try-catch blocks are finally blocks.
+A finally block is executed after a try and/or catch block\cite{javaDocs}.
+There is a possibility that a mutant may be generated inside these blocks.
+If a mutant is located in a try, catch or finally block can be defined as a characteristic.
+It could be a useful characteristic because the mutants that are located in the same type of block may be grouped together.
+
+\section{Local variables}
+A mutant may be located in a location(see Chapter \ref{ch:characteristics_location}) that may contain local variables.
+There is a chance a mutant modifies a local variable.
+A variable modified by a mutant increases the chance for a mutant to be killed.
+This is because a unit test may check for a specific value for a specific local variable.
+The amount of local variables a location contains can be identified as a characteristic.
+We can extract the amount of local variables a location contains from PIT.
+This could be a useful characteristic as mutants with the same amount of local variables may be grouped together.
 

### 2021-06-04 chapters/6_experiment_design.tex
WIP exp1 design + finished characteristics

@@
 \begin{document}
-\chapter{Experiment design}
-\todo{expand this}
-We hypothesise that there is a we can cluster on the characteristics of a mutant such that each mutant should be a representative for all of the mutants in its cluster.
+\chapter{Experiment design rq1}\todo{discuss how to split this with exp design for rq2}
+In this chapter we will design an experiment to answer our first research question.
+We hypothesise that we can cluster mutants with the identified set of characteristics 
+while maintaining effectiveness when executing one mutant from each cluster that is in the majority of that cluster.
+If a cluster contains ten mutants with eight mutants killed and two mutants survived, all the mutants that are killed are in the majority of that cluster.
 Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.
+\newline
+We will start with extracting the characteristics from the selected mutation testing tool. 
+Next we select and implement a clustering algorithm for which we can cluster mutants.
+We can then cluster the mutants and compare the results of the clusters with the results of the mutants.
 
-We design an experiment in which we implement the following steps:
-\begin{enumerate}
-    \item Extract all the characteristics.
-    \item For each mutant calculate the Levenshtein distance.
-    \item Cluster the mutants based on the characteristics.
-    \item Randomly select one mutant from each cluster.
-    \item Execute the selected mutants in PIT and calculate the mutation score.
-\end{enumerate}
-We elaborate on these steps in the following subsections.
+\section{Extracting characteristics}
+To extract all the characteristics we identified we need to run PIT and configure it to generate all mutants that are possible within the tool. 
+PIT works in phases, the first phase is the generation phase where mutants are generated.
+The second phase is the execution phase where mutants are challenged by the test suite.
+The last phase is the reporting phase where a report is generated based on the users preferences and data generated by PIT.
+During the various stages we can extract characteristics by using PIT's plugin system.
+\newline
+PIT offers a plugin system in which developers can inject their own code in PIT\cite{pitestPlugin}.
+There are two main types of plugins; a Mutation Result Listener and a Mutation interceptor\cite{pitestPlugin}.
+A mutation result listener receives the details of analysed mutations as they arrive\cite{pitestPlugin}.
+A Mutation interceptor is passed a complete list of all mutation that will be generated to each class before the mutation are challenged by tests\cite{pitestPlugin}.
+The implementation of our experiment makes use of this plugins system.
+\newline
+We developed two plugins to extract characteristics.
+The first plugin is of type Mutation interceptor.
+It is provided a list of details per class. 
+We can extract all characteristics but the Levensthein distance and the amount of tests a mutant is challenged by.
+To calculate the Levensthein distance we need the byte code of the original code and the mutated code.
+We generate the byte code of the original code based on the location provided in the list of details.
+We make use of existing functionality in PIT for generating byte code.
+The mutator is also provided in the list of details.
+With the combination of location and mutator we can generate the byte code of the mutant.
+With both pieces of code available we can calculate the Levenshtein distance between them.
+\newline
+The characteristics gathered until this phase of the mutation testing process are written, together with a unique identifier, to a \acrshort{csv} file.
+\newline
+The second plugin is a mutation result listener. 
+This plugin gets passed a list of the results per mutant.
+This list contains data about the mutant survival status.
+It also contains data about the number of tests the mutant is challenged by.
+We store this data together with the identifier to a different \acrshort{csv} file.
+\newline
+When the mutation testing process has finished we need to merge the characteristic gathered in the second plugin.
+We add the characteristic; the number of tests the mutant is challenged by to the data file containing all the other characteristics.
+We can merge the data because both files contain the same unique identifier per mutant.
 
-\subsection{Levenshtein distance on Java byte code}
+\section{Levenshtein distance on Java byte code}
 A mutant is a piece of code that differs in a predefined way from its parent.
@@
 Java byte code reflects more of the semantic nature of the source code than the source code itself does. 
-\newline
 PIT generates and executes mutants on byte code level\cite{pitestBytecode}. 
@@
 \newline
-For this experiment we will use the Levenshtein distance between a Java byte code mutant and its parent.
+For our characteristic we will use the Levenshtein distance between a Java byte code mutant and its parent.
 By calculating the distance on byte code we filter out the context that is present in Java code.
@@
 
-\subsection{Hierarchical clustering}
+\section{Hierarchical clustering}\todo{elaborate on why hierarchical clustering}
 Clustering Levenshtein distances has been done before with hierarchical clustering\cite{Rajalingam2011HierarchicalStudy, Gothai2010PerformanceAlgorithms}. 
@@
 \newline
-We identify each Levenshtein distance between a mutant and its parent as a separate cluster.
+We identify all characteristics per mutant as a separate cluster.
 Starting out with each mutant as a separate cluster we can use the agglomerative form of hierarchical clustering.
@@
 
-\subsection{Weighed mutation score}
-Our hypothesis states that each mutants should be a representative for all the mutants in its cluster. 
-This translates to if a mutant survives or is killed all mutants in that cluster should survive or be killed respectively.
-To reflect this in our experiment a weighed mutation score is calculated for the clustered mutants.
+\section{Weighed mutation score}
+Our hypothesis states that each mutants executed should be in the majority of mutants in its cluster. 
+With a mutant executed from each cluster we can calculate a mutation score.
+This mutation score is a weighed mutation score.
 This weighed mutation score is the product of the result of a mutant(1 for killed and 0 for survived) and the amount of mutants in the cluster it represents.
@@
 
-\subsection{PIT plugin system}
-PIT offers a plugin system in which developers can inject their own functionality in PIT\cite{pitestPlugin}. 
-There are two main types of plugins; a Mutation Result Listener and a Mutation interceptor\cite{pitestPlugin}.
-A mutation result listener  receives the details of analysed mutations as they arrive\cite{pitestPlugin}.
-A Mutation interceptor passed a complete list of all mutation that will be generated to each class before the mutation are challenged by tests\cite{pitestPlugin}.
-The implementation of our experiment makes use of this plugins system.
-\newline
-We developed three separate plugins for our experiment.
-The first plugin is a Mutation interceptor.
-It analyses each mutant, calculates the Levenshtein distance and writes a mutant identifier and distance to a \acrshort{csv} file. 
-It also clears the list of mutants provided by PIT to prevent execution. 
-The second plugin is also a Mutation interceptor.
-It reads the clustered mutants from a \acrshort{csv} file and compares them to the list of mutants provided by PIT.
-Only the clustered mutants are passed back to PIT for execution.
-The last plugin is a mutation result listener. 
-This plugin calculates the weighed mutation score and outputs it to a text file.
-\newline
-The plugins in PIT are executed based on type and per source class.
-This means that a plugin will never receive all the mutants generated for a whole project during execution.
-Because of how PIT executes its plugins we have to execute PIT twice.
-In the first run we calculate the Levenshtein distance for all mutants and in the second run we execute the selected mutants and calculate the weighed mutation score.
 \end{document}

### 2021-06-07 chapters/5_identifying_characteristics.tex
Finished experiment design

@@
 
-\chapter{Identifying characteristics}
+\section{Identifying characteristics}
 \label{ch:clustering_characteristics}
@@
 
-\section{Overview of characteristics}
+\subsection{Overview of characteristics}
 Zhang et al., identifies several mutant characteristics\cite{Zhang2019PredictiveTesting}. 
@@
 \end{enumerate}
-In the following subsections we will elaborate more on these characteristics.
+In the following sections we will elaborate more on these characteristics.
 
-\section{Mutant similarity}
+\subsection{Mutant similarity}
 A piece of code is a textual representation of instructions for a computer.
@@
 
-\section{Amount of tests challenged by mutant}
+\subsection{Amount of tests challenged by mutant}
 A mutant may be challenged by tests.
@@
 
-\section{Mutator identifier}
+\subsection{Mutator identifier}
 At the time of writing the selected mutation testing tool PIT supports more than 100 mutators\cite{pitestMutators}.
@@
 
-\section{Mutant opcode}
+\subsection{Mutant opcode}
 \say{A Java Virtual Machine instruction consists of a one-byte opcode specifying the operation to be performed, followed by zero or more operands supplying arguments or data that are used by the operation.} (Java 12 SE docs ch.12).
@@
 
-\section{Mutant return type}
+\subsection{Mutant return type}
 A mutant may contain a return statement. 
@@
 
-\section{Mutant location}\label{ch:characteristics_location}
+\subsection{Mutant location}\label{ch:characteristics_location}
 The location of a mutant in PIT is defined by multiple properties;
@@
 
-\section{Error handling capability}
+\subsection{Error handling capability}
 A technique to handle errors in Java is to use try-catch blocks\cite{javaDocs}. 
@@
 
-\section{Local variables}
+\subsection{Local variables}
 A mutant may be located in a location(see Chapter \ref{ch:characteristics_location}) that may contain local variables.

### 2021-06-07 chapters/6_experiment_design.tex
Finished experiment design

@@
 \begin{document}
-\chapter{Experiment design rq1}\todo{discuss how to split this with exp design for rq2}
+\section{Experiment design}\todo{discuss how to split this with exp design for rq2}
 In this chapter we will design an experiment to answer our first research question.
-We hypothesise that we can cluster mutants with the identified set of characteristics 
-while maintaining effectiveness when executing one mutant from each cluster that is in the majority of that cluster.
+We hypothesise that we can cluster mutants with the set of characteristics we identified
+while maintaining effectiveness and reducing the amount executed when executing one mutant from each cluster that is in the majority of that cluster.
 If a cluster contains ten mutants with eight mutants killed and two mutants survived, all the mutants that are killed are in the majority of that cluster.
@@
 
-\section{Extracting characteristics}
+\subsection{Extracting characteristics}
 To extract all the characteristics we identified we need to run PIT and configure it to generate all mutants that are possible within the tool. 
@@
 We can merge the data because both files contain the same unique identifier per mutant.
+We do not add the information about the survival of a mutant.
+This is data we need to validate our results.
 
-\section{Levenshtein distance on Java byte code}
+\subsection{Levenshtein distance on Java byte code}
 A mutant is a piece of code that differs in a predefined way from its parent.
@@
 
-\subsection{Levenshtein distance implementation}
+\subsubsection{Levenshtein distance implementation}
 write here that we use a memory efficient algorithm
 
-\section{Hierarchical clustering}\todo{elaborate on why hierarchical clustering}
+\subsection{Hierarchical clustering}\todo{elaborate on why hierarchical clustering}
 Clustering Levenshtein distances has been done before with hierarchical clustering\cite{Rajalingam2011HierarchicalStudy, Gothai2010PerformanceAlgorithms}. 
+Research shows that hierarchical clustering performs better when clustering with at least ten features\cite{Rodriguez2019}.
+It also states that varying the parameters of hierarchical clustering improves the performance compared to that of the default settings of the algorithm\cite{Rodriguez2019}.
+We selected hierarchical clustering to cluster the mutants.
+We have more than ten features and adjust the parameters of the algorithm based on the characteristics in our data set.
+Next we will explain the configuration we use for clustering the mutants. 
+\newline
 Hierarchical clustering is subdivided into agglomerative and divisive. 
@@
 Since we cannot assume that our clusters are well separated we chose to use the ward linkage method for our clustering algorithm.
-The agglormorative hierarchical clustering algorithm with ward linkage will cluster our mutants represented as Levenshtein distances.
+The agglormorative hierarchical clustering algorithm with ward linkage will cluster our mutants represented by the characteristics we gathered.
 \todo{write about implementation}
 
-\section{Weighed mutation score}
+\subsection{Weighed mutation score}
+\label{ch:weighed_score}
 Our hypothesis states that each mutants executed should be in the majority of mutants in its cluster. 
@@
 We then cluster the mutants in four clusters consisting of 12, 30, 38 and 20 mutants, respectively.
-We randomly select four mutants of each cluster and execute them.
-The mutants representing cluster one and four survive.
+We select four mutants of each cluster that are in the majority and execute them.
+The mutants representing cluster one and four survive and two and three are killed.
 If we calculate the weighed score we get 68/100 which is 68\%.
@@
 
+\subsection{Categorical data}
+The hierarchical clustering algorithm needs all characteristics in a numerical form\cite{Vijaya2019ComparativeClustering}.
+The characteristics mutator identifier, class name, method name and return type are non numerical.
+To deal with this problem we need to apply categorical variable encoding to these specific features.
+There are different categorical variable encoding techniques available\cite{Potdar2017AClassifiers}.
+The categorical characteristics we use have no particular ranking compared to each other.
+There is also no specific order to the characteristics.
+I.e. a return type void is not better or worse than a return type string.
+The same goes for the location characteristics, there is no location that should weigh more during the clustering than the other locations.
+The individual characteristics do contain a finite set of values.
+For example multiple mutants may contain the same class name.
+Taking into account the properties of our categorical characteristics we chose to use a nominal variable encoding.
+Nominal encoding comprises a finite set of discrete values with no relationship between values\cite{Potdar2017AClassifiers}.
+
+\subsection{Cluster size}
+We decide our cluster size based on the amount of mutants generated in a full set.
+As a baseline we will reduce the amount of mutants by 75\%, 50\% and 25\%.
+This results in a cluster size that will be equal to 75\%, 50\% and 25\% of the total amount of mutants generated in a full set.
+The amount of mutants inside a cluster will be decided by the clustering algorithm.
+
+\subsection{Validation}
+The most efficient way to measure test effectiveness with mutation testing is by running all mutants that a tool can possibly generate.
+The goal of this experiment is to reduce the amount of mutants executed while maintaining effectiveness.
+To reduce the amounts executed we cluster the mutants. 
+We can measure this by counting the amount of clusters we generate and compare it to the number of total mutants generated by the selected tool.
+\newline
+The baseline for our experiment is the mutation score of the set of all mutants generated by our selected mutation testing tool.
+\newline
+To validate how effective our clustering is we can compare the weighed mutation score(see Chapter \ref{ch:weighed_score}) of our clustered set to the mutation score of the full set.
+The closer the weighed score is to that of a full set the more effective our set of characteristics and clustering algorithm proves to be.
+In other words we want achieve a mutation score that is as close as possible as to that of a full set.
+We select a significance level of $\leq 0.05$.
+This is the conventional threshold for declaring statistical significance\cite{Kirk1996PracticalCome}.
+\newline
+Depending on the effectiveness of our clustering algorithm we may loose accuracy.
+This can happen if a cluster contains mutants both types. 
+We can measure this accuracy by calculating a percentage of all mutants that have survived against the ones that have been killed in a cluster.
+If the majority of the mutants in a cluster is killed then we consider that cluster to represent a killed cluster and the other way around for survived mutants.
+We consider the mutants that are not in the majority of the cluster as inaccuracy.
+\newline
+To decide if we accept or reject our hypothesis we use statistical significance\cite{Kirk1996PracticalCome}.
+If the weighed mutation score of our clustered set deviates more than 5\% from the score of a full set we will reject our hypothesis.
+We also compare the inaccuracy to the significance level to decide if our inaccuracy is statistical significant.
 \end{document}

### 2021-06-20 chapters/6_experiment_design_rq1.tex
Final results exp1 added

--- 2021-06-07 chapters/6_experiment_design.tex
@@
 \begin{document}
-\section{Experiment design}\todo{discuss how to split this with exp design for rq2}
+% Because we use sections but these sections should be interpreted as chapters for rq1 and rq2 sectioning.
+\clearpage
+\section{Experiment design}
 In this chapter we will design an experiment to answer our first research question.
 We hypothesise that we can cluster mutants with the set of characteristics we identified
-while maintaining effectiveness and reducing the amount executed when executing one mutant from each cluster that is in the majority of that cluster.
-If a cluster contains ten mutants with eight mutants killed and two mutants survived, all the mutants that are killed are in the majority of that cluster.
+while maintaining effectiveness and reducing the amount executed when executing one mutant from each cluster that is randomly selected.
 Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.
@@
 
-\subsubsection{Levenshtein distance implementation}
-write here that we use a memory efficient algorithm
+\subsection{Levenshtein distance implementation}
+There are different implementations of the Levensthein distance.
+For our experiment we need to calculate the Levenshtein distance on the byte code generated in PIT.
+Since we make use of the plugin system it was natural to use an implementation that was written in Java.
+The Levenshtein distance can be calculated in different ways for example with recursion.
+The implementation with recursion is has a complexity of O(N\^{}2) and a memory need of O(N\^{}2).
+There is also an improved implementation that does not use recursion.
+While the complexity is the same the memory need is reduced to O(N).
+This last implementation is the implementation we selected for our experiment.
 
-\subsection{Hierarchical clustering}\todo{elaborate on why hierarchical clustering}
+\subsection{Hierarchical clustering}
 Clustering Levenshtein distances has been done before with hierarchical clustering\cite{Rajalingam2011HierarchicalStudy, Gothai2010PerformanceAlgorithms}. 
@@
 We selected hierarchical clustering to cluster the mutants.
-We have more than ten features and adjust the parameters of the algorithm based on the characteristics in our data set.
+We have more than ten characteristics which we can use as features and adjust the parameters of the algorithm based on the characteristics in our data set.
 Next we will explain the configuration we use for clustering the mutants. 
@@
 The agglormorative hierarchical clustering algorithm with ward linkage will cluster our mutants represented by the characteristics we gathered.
-\todo{write about implementation}
 
@@
 \label{ch:weighed_score}
-Our hypothesis states that each mutants executed should be in the majority of mutants in its cluster. 
+Our hypothesis states that each mutants executed should represent that whole cluster.
 With a mutant executed from each cluster we can calculate a mutation score.
@@
 We then cluster the mutants in four clusters consisting of 12, 30, 38 and 20 mutants, respectively.
-We select four mutants of each cluster that are in the majority and execute them.
+We randomly select four mutants of each cluster and execute them.
 The mutants representing cluster one and four survive and two and three are killed.

### 2021-06-23 chapters/3_background.tex
redo structure

@@
 The optics and dbscan algorithms aim at different data distributions than performed in this study. There different results could be obtained for non-normally distributed data.
+
+\section{Neural networks for clustering}
+There are many different types of neural networks.
+These neural networks have different applications based on their properties.
+For this thesis we will focus on neural networks designed for clustering.
+\newline
+Neural networks consists of machine learning algorithms\cite{supervisedUnsupervised}. 
+These machine learning algorithms are divided in supervised and unsupervised learning algorithms\cite{supervisedUnsupervised}.
+Supervised machine learning algorithms are commonly used for classification and categorization\cite{supervisedUnsupervised}.
+Unsupervised machine learning algorithms are commonly used for clustering and dimensionality reduction\cite{supervisedUnsupervised}.
+\newline
+K.L. Du surveyed different neural networks and algorithms used for clustering:
+\begin{itemize}
+    \item Self organizing map.
+    \item Learning vector quantization.
+    \item C-means clustering.
+    \item Mountain and Substractive clustering.
+    \item Neural gas.
+    \item ART networks.
+    \item Fuzzy c-means.
+\end{itemize}
+In his paper he also touches on subjects relevant to clustering and neural networks such as the under-utilization problem, fuzzy clustering, robust clustering, clustering based on non- Euclidean distance measures, supervised clustering, and hierarchical clustering.
+Neural network variants and their references are also discussed.
+\newline
+K.L. Du describes the mathematical principles on which the neural networks and algorithms are based.
+He closes his paper with computer simulations of some of the neural networks he surveyed.
+
 \end{document}

### 2021-06-23 chapters/5_research_question_1.tex
redo structure

\documentclass[../main]{subfiles}
\begin{document}
\chapter{Research question 1}
\label{ch:reserach_question1}
\subfile{rq1/1_identifying_characteristics}
\subfile{rq1/2_experiment_design}
\subfile{rq1/3_results}
\end{document}

### 2021-06-23 chapters/rq1/2_experiment_design.tex
redo structure

--- 2021-06-20 chapters/6_experiment_design_rq1.tex
@@
 We also compare the inaccuracy to the significance level to decide if our inaccuracy is statistical significant.
+
+\subsection{Mutant selection}
+As stated in our hypothesis we will randomly select a mutant from each cluster to be executed.
+To make our results reproducible we will select a random mutant based on a generated seed.
+The seed generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.
+Achieving consistent results while applying random selection will contribute to the validity of the experiment.
+We will use the same statistical significance threshold to decide if the variation by the random sampling is acceptable or not.
+
 \end{document}

### 2021-06-23 chapters/rq2/1_experiment_design.tex
redo structure

\documentclass[../main]{subfiles}
\begin{document}
% Because we use sections but these sections should be interpreted as chapters for rq1 and rq2 sectioning.
\clearpage
\section{Experiment design}
In this chapter we will design an experiment to answer our second research question.
HYOPTHESIS HERE\todo{do hypothesis}
Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.
\newline
\toto{finish this}

\subsection{Selecting a neural network}

\end{document}

### 2021-06-25 chapters/rq1/2_experiment_design.tex
latest changes
Fixed reference error

@@
 \subsection{Hierarchical clustering}
-Clustering Levenshtein distances has been done before with hierarchical clustering\cite{Rajalingam2011HierarchicalStudy, Gothai2010PerformanceAlgorithms}. 
+Clustering Levenshtein distances has been done before with hierarchical clustering\cite{Rajalingam2011, Gothai2010PerformanceAlgorithms}. 
 Research shows that hierarchical clustering performs better when clustering with at least ten features\cite{Rodriguez2019}.
@@
 The agglomerative hierarchical technique follows bottom up approach whereas divisive follows top-down approaches.
-Hierarchical clustering uses different metrics which measures the the distance between two clusters and the linkage criteria\cite{Rajalingam2011HierarchicalStudy}. 
-The linkage criteria specifies the dissimilarity in the sets as a function of the pair-wise distances of observations in those sets\cite{Rajalingam2011HierarchicalStudy}.
+Hierarchical clustering uses different metrics which measures the the distance between two clusters and the linkage criteria\cite{Rajalingam2011}. 
+The linkage criteria specifies the dissimilarity in the sets as a function of the pair-wise distances of observations in those sets\cite{Rajalingam2011}.
 \newline

### 2021-06-25 chapters/3_background.tex

@@
 \section{Neural networks for clustering}
+\label{ch:neuralNetworkSurvey}
 There are many different types of neural networks.
@@
 \newline
-K.L. Du surveyed different neural networks and algorithms used for clustering:
+K.L. Du surveyed different neural networks and algorithms used for clustering\cite{Du2010Clustering:Approach}:
 \begin{itemize}
@@
     \item C-means clustering.
-    \item Mountain and Substractive clustering.
+    \item Mountain clustering.
     \item Neural gas.

### 2021-06-25 chapters/rq2/1_experiment_design.tex

@@
-\documentclass[../main]{subfiles}
+\documentclass[../../main]{subfiles}
 \begin{document}
@@
 \newline
-\toto{finish this}
+\todo{finish this}
 
 \subsection{Selecting a neural network}
+Our starting point for selecting a neural network is the survey done by K.L. Du\cite{Du2010Clustering:Approach}.
+We start with a process of elimination to filter the neural networks named in chapter \ref{ch:background}.
+\newline
+supervised networkd\todo{elaborate why no supervised}
+\newline
+The moutain clustering networks is effective method for estimating the number of clusters\cite{Du2010Clustering:Approach}. 
+The goal of this experiment is to cluster mutants, while finding out the number of clusters we need might be valuable it is not what we want to achieve.
+Therefore we will not select mountain clustering.
+Subtractive clustering is a modified form of mountain c\cite{Du2010Clustering:Approach}. 
+\newline
+
+\subsection{Training method}
+
+\subsection{Validation}
 

### 2021-06-25 chapters/3_background.tex
changed neural net to machine learning model

--- 2021-05-28 chapters/3_background.tex
@@
 The optics and dbscan algorithms aim at different data distributions than performed in this study. There different results could be obtained for non-normally distributed data.
+
+\section{Machine learning models for clustering}
+\label{ch:neuralNetworkSurvey}
+There are many different types of machine learning models.
+These models have different applications based on their properties.
+For this thesis we will focus on machine learning models designed for clustering.
+\newline
+Machine learning algorithms are divided in supervised and unsupervised learning algorithms\cite{supervisedUnsupervised}.
+Supervised machine learning algorithms are commonly used for classification and categorization\cite{supervisedUnsupervised}.
+Unsupervised machine learning algorithms are commonly used for clustering and dimensionality reduction\cite{supervisedUnsupervised}.
+\newline
+K.L. Du surveyed different machine learning algorithms used for clustering\cite{Du2010Clustering:Approach}:
+\begin{itemize}
+    \item Self organizing map.
+    \item Learning vector quantization.
+    \item C-means clustering.
+    \item Mountain clustering.
+    \item Neural gas.
+    \item ART networks.
+    \item Fuzzy c-means.
+\end{itemize}
+In his paper he also touches on subjects relevant to clustering and machine learning models such as the under-utilization problem, fuzzy clustering, robust clustering, clustering based on non- Euclidean distance measures, supervised clustering, and hierarchical clustering.
+Machine learning model variants and their references are also discussed.
+\newline
+K.L. Du describes the mathematical principles on which the machine learning algorithms are based.
+He closes his paper with computer simulations of some of the machine learning models he surveyed.
+
 \end{document}

### 2021-06-25 chapters/rq2/1_experiment_design.tex
changed neural net to machine learning model

@@
 In this chapter we will design an experiment to answer our second research question.
-HYOPTHESIS HERE\todo{do hypothesis}
+We hypothesise that we train a machine learning model with the set of characteristics we identified
+while maintaining effectiveness and reducing the amount executed when executing one mutant from each cluster that is randomly selected.
 Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.
@@
 
-\subsection{Selecting a neural network}
-Our starting point for selecting a neural network is the survey done by K.L. Du\cite{Du2010Clustering:Approach}.
-We start with a process of elimination to filter the neural networks named in chapter \ref{ch:background}.
+\subsection{Selecting a machine learning model}
+Our starting point for selecting a machine learning model is the survey done by K.L. Du\cite{Du2010Clustering:Approach}.
+We start with a process of elimination to filter the machine learning models named in chapter \ref{ch:background}.
 \newline
-supervised networkd\todo{elaborate why no supervised}
+supervised learning\todo{elaborate why no supervised}
 \newline
-The moutain clustering networks is effective method for estimating the number of clusters\cite{Du2010Clustering:Approach}. 
+The moutain clustering model is effective method for estimating the number of clusters\cite{Du2010Clustering:Approach}. 
 The goal of this experiment is to cluster mutants, while finding out the number of clusters we need might be valuable it is not what we want to achieve.

### 2021-06-25 chapters/3_background.tex
Added machine learning model selection

@@
 \label{ch:neuralNetworkSurvey}
-There are many different types of machine learning models.
-These models have different applications based on their properties.
+There are many different types of machine learning models with their respective applications.
 For this thesis we will focus on machine learning models designed for clustering.
@@
 Machine learning algorithms are divided in supervised and unsupervised learning algorithms\cite{supervisedUnsupervised}.
-Supervised machine learning algorithms are commonly used for classification and categorization\cite{supervisedUnsupervised}.
-Unsupervised machine learning algorithms are commonly used for clustering and dimensionality reduction\cite{supervisedUnsupervised}.
+Supervised learning is the machine learning task of learning a function that maps an input to an output based on example input-output pairs\cite{supervisedUnsupervised}.
+In contrast to supervised learning, unsupervised learning shows self-organization that captures patterns as neuronal predilections or probability densities\cite{supervisedUnsupervised}.
+Unsupervised learning algorithms are commonly used for classification and categorization\cite{supervisedUnsupervised}.
+Unsupervised learning is a type of algorithm that learns patterns from untagged data\cite{supervisedUnsupervised}.
+These machine learning algorithms are commonly used for clustering and dimensionality reduction\cite{supervisedUnsupervised}.
 \newline
-K.L. Du surveyed different machine learning algorithms used for clustering\cite{Du2010Clustering:Approach}:
+K.L. Du surveyed different machine learning models used for clustering\cite{Du2010Clustering:Approach}.
+The models he discusses are:
 \begin{itemize}
@@
     \item Learning vector quantization.
-    \item C-means clustering.
+    \item C-means(K-means) clustering.
     \item Mountain clustering.
@@
 \end{itemize}
-In his paper he also touches on subjects relevant to clustering and machine learning models such as the under-utilization problem, fuzzy clustering, robust clustering, clustering based on non- Euclidean distance measures, supervised clustering, and hierarchical clustering.
+K.L. Du describes the mathematical principles on which the machine learning algorithms are based as well as the origin, pros and cons of each model.
+\newline
+He also touches on subjects relevant to clustering and machine learning models such as the under-utilization problem, fuzzy clustering, robust clustering, clustering based on non- Euclidean distance measures, supervised clustering, and hierarchical clustering.
 Machine learning model variants and their references are also discussed.
 \newline
-K.L. Du describes the mathematical principles on which the machine learning algorithms are based.
 He closes his paper with computer simulations of some of the machine learning models he surveyed.

### 2021-06-25 chapters/rq2/1_experiment_design.tex
Added machine learning model selection

--- 2021-06-23 chapters/rq2/1_experiment_design.tex
@@
 In this chapter we will design an experiment to answer our second research question.
-HYOPTHESIS HERE\todo{do hypothesis}
+We hypothesise that we can train a machine learning model with the set of characteristics we identified
+while maintaining effectiveness and reducing the amount executed when executing one mutant from each cluster that is randomly selected.
 Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.
 \newline
-\toto{finish this}
+\todo{finish this}
 
-\subsection{Selecting a neural network}
+\subsection{Selecting a machine learning model}
+The first choice we have to make on selecting a machine learning model is to decide on supervised or unsupervised learning.
+The application and description of these two types of learning are described in Chapter \ref{ch:background}.
+Supervised machine learning models require a labeled data set.
+Our data set consists of mutants represented by the characteristics identified in Chapter \ref{ch:clustering_characteristics}.
+This data set is unlabeled as it does not map the mutants to categories.
+We also do now know what the categories would be or how to map a mutant to a specific category.
+Instead of manually evaluating each mutant and identifying categories to map them to we can train a machine learning model to do this for us.
+To do this we need an unsupervised machine learning model.
+\newline
+Our starting point for selecting an unsupervised machine learning model are the models discussed in the survey done by K.L. Du\cite{Du2010Clustering:Approach}.
+We will select a model by a process of elimination.
+
+\subsubsection{Mountain model}
+The mountain clustering model is an effective method for estimating the number of clusters\cite{Du2010Clustering:Approach}. 
+While finding out the number of clusters we need might be valuable it is not what we want to achieve. 
+Therefore we will not select mountain clustering.
+
+\subsubsection{ART models}
+ART models are characterized by systems of differential equations that formulate stable self-organizing learning methods\cite{Du2010Clustering:Approach}.
+The stability and plasticity properties as well as the ability to efficiently process dynamic data make the ART attractive for clustering large, rapidly changing sequences of input patterns\cite{Du2010Clustering:Approach}.
+While the ART models are very good at what they do, they do not fit our purpose of clustering mutants.
+As described above the ART models expect rapidly changing sequences of input patterns our data set is not that.
+Therefore we will not select an ART model.
+
+\subsubsection{Neural gas}
+The neural gas is a topology-preserving network, and can be treated as an extension to the C-means\cite{Du2010Clustering:Approach}.
+A data optimal topological ordering is achieved by using neighborhood ranking within the input space at each training step\cite{Du2010Clustering:Approach}.
+To find its neighborhood rank, each neuron compares its distance to the input vector with those of all the other neurons to the input vector\cite{Du2010Clustering:Approach}. 
+Unlike the self organizing map, which uses predefined static neighborhood relations, the neural gas network determines a dynamical neighborhood relation as learning proceeds\cite{Du2010Clustering:Approach}.
+A big pro of the neural gas network is that it is able to spawn neurons.
+This means that we do not need to know the amount of clusters in advance.
+Unfortunately the neural gas network is a learning network.
+One of the properties of a learning network is that it is not designed to fit new data points\cite{supervisedUnsupervised}.
+For this reason we cannot use the neural gas network to predict in what cluster a new mutant should be placed.
+Therefore we will not select the neural gas network.
+
+\subsubsection{Learning vector quantization}
+The learning vector quantization model is a supervised learning model.
+We established that we need an unsupervised learning model to cluster the mutants.
+Therefore we will not select the learning vector quantization model.
+
+\subsubsection{Self organizing map, C-means and fuzzy C-means}
+The self organizing map, C-means and fuzzy C-means models are all models that would fit our purpose. 
+To choose between these models we will look at the performance of each model.
+Mingoti et al., compared the performance of these three models \cite{Mingoti2006ComparingAlgorithms}.
+They did a Monte Carlo simulation where the cluster sizes and amount of random numbers in the data set varied each simulation\cite{Mingoti2006ComparingAlgorithms}.
+Other variables such as cluster boundaries and overlap were also controlled in the experiment.
+They observe that all the C-means and fuzzy C-means had good performance for non overlapping situations\cite{Mingoti2006ComparingAlgorithms}.
+The best results for average recovery and internal dispersion rates were found for Fuzzy C-means which was very stable in all situations achieving recovery averages over 90\%\cite{Mingoti2006ComparingAlgorithms}
+The C-means method was very affected by the presence of a large amount of outliers.
+They conclude that the self organizing map did not perform well in many cases being very affected by the amount of variables and clusters even for the non overlapping cases\cite{Mingoti2006ComparingAlgorithms}.
+They also conclude that their results partially overlap with other research in the same domain.
+\newline
+Based on these results we can conclude that the fuzzy c-means model has the best performance overall.
+For this experiment we will select the fuzzy c-means model.
+\subsection{Training strategy}
+
+\subsection{Validation}
 

### 2021-06-28 chapters/rq2/1_experiment_design.tex
WIP exp2 design

@@
 While the ART models are very good at what they do, they do not fit our purpose of clustering mutants.
-As described above the ART models expect rapidly changing sequences of input patterns our data set is not that.
+As described above the ART models expect rapidly changing sequences of input patterns, our data set does not constists of this kind of data.
 Therefore we will not select an ART model.
@@
 Based on these results we can conclude that the fuzzy c-means model has the best performance overall.
-For this experiment we will select the fuzzy c-means model.
+For this experiment we select the fuzzy c-means model.
+
 \subsection{Training strategy}
+Our data set contains of mutants represented by their characteristics. 
+The characteristics of the mutants are generated per project and have a project specific context.
+The relations between the different mutants are also specific per project.
+For example mutant with a mutator in a certain project can have a different effect on the source code and can give other results than a mutator in a different project.
+To keep the relations and context we decided to train our model per project.
+\newline
+To validate our model we need to split our data set into a training set and a validation set.
+Our total data set consists of approximately 1 million mutants generated over fourteen projects.
+Research has shown that if the data set is big enough a ratio of 80/20 is sufficient\cite{Guyon1997ARatio}.
+We will divide our projects into training and validation groups to reflect the 80/20 ratio.
+
+\subsection{Implementation of algorithm}
+The origin of the fuzzy c-means algorithm is referenced in the survey paper of Du et al\cite{Du2010Clustering:Approach}.
+The fuzzy c-means algorithm is developed by Bezdek et al.\cite{Bezdek1984FCM:Algorithm}.
+Based on their research an implementation can be developed.
+There are existing libraries that contain the implementation of the fuzzy c-means algorithm.
+Developing our own version of the fuzzy c-means algorithm is out of scope for this research.
+For this reason we choose to use a library that implements the fuzzy c-means algorithm.
+The library we will use for this experiment is the fuzzy c-means implementation build by Dias et al.\cite{dias2019fuzzy}.
+The theory developed by Bezdek et al., is implemented in this library.
+The library has been used in other research\cite{DeAlmeidaNeto2020, Kopf2019, Nwadiugwu2020} and is actively maintained.
+
+\subsection{Algorithm parameters}
 

### 2021-07-03 chapters/rq1/2_experiment_design.tex
minor tweaks exp2 design + results

@@
 \subsection{Validation}
+\label{ch:exp1_validation}
 The most efficient way to measure test effectiveness with mutation testing is by running all mutants that a tool can possibly generate.
@@
 \newline
-To decide if we accept or reject our hypothesis we use statistical significance\cite{Kirk1996PracticalCome}.
 If the weighed mutation score of our clustered set deviates more than 5\% from the score of a full set we will reject our hypothesis.
@@
 The seed generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.
+We will repeat our experiment 30 times with 30 different seeds.
 Achieving consistent results while applying random selection will contribute to the validity of the experiment.

### 2021-07-03 chapters/rq2/1_experiment_design.tex
minor tweaks exp2 design + results

@@
 Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.
-\newline
-\todo{finish this}
+We will select a machine learning model, decide on a training strategy, train the model and validate the results.
 
 \subsection{Selecting a machine learning model}
-The first choice we have to make on selecting a machine learning model is to decide on supervised or unsupervised learning.
+We will first choose between supervised or unsupervised learning.
+Our starting point for selecting a machine learning model are the models discussed in the survey done by K.L. Du.
+Based on the choice of supervised or unsupervised learning we will review the corresponding models and decide if they fit our use case by a process of elimination.
+
+\subsubsection{Supervised or unsupervised learning}
+In this subsection we will decide on supervised or unsupervised learning.
 The application and description of these two types of learning are described in Chapter \ref{ch:background}.
+\newline
 Supervised machine learning models require a labeled data set.
-Our data set consists of mutants represented by the characteristics identified in Chapter \ref{ch:clustering_characteristics}.
-This data set is unlabeled as it does not map the mutants to categories.
-We also do now know what the categories would be or how to map a mutant to a specific category.
-Instead of manually evaluating each mutant and identifying categories to map them to we can train a machine learning model to do this for us.
-To do this we need an unsupervised machine learning model.
+We can argue that we can label our data set with the results of our first experiments.
+Each mutant would then be labeled with the identity of the cluster it is clustered in by the hierarchical clustering model.
+If we would train a supervised model with the data labeled as described, it may result in a model that can do the same as hierarchical clustering.
+Our research is not to prove if we can train a model to cluster mutants the same way as hierarchical clustering.
+We want to train a model to find the relations between the characteristics without having to define these relations manually.
 \newline
-Our starting point for selecting an unsupervised machine learning model are the models discussed in the survey done by K.L. Du\cite{Du2010Clustering:Approach}.
-We will select a model by a process of elimination.
+Unsupervised learning models do not need a labeled data set and can recognize relations or discover hidden patterns in the characteristics we identified.
+This corresponds to what we want to achieve with this experiment.
+Therefore we choose unsupervised learning to continue with in our experiment.
 
-\subsubsection{Mountain model}
-The mountain clustering model is an effective method for estimating the number of clusters\cite{Du2010Clustering:Approach}. 
-While finding out the number of clusters we need might be valuable it is not what we want to achieve. 
-Therefore we will not select mountain clustering.
+\subsubsection{Mountain and subtractive clustering}
+The mountain and subtractive clustering models are an effective method for estimating the number of clusters\cite{Du2010Clustering:Approach}. 
+The subtractive clustering model is an extension of the mountain model\cite{Du2010Clustering:Approach}. 
+\newline
+The method grids the data space and computes a potential value for each grid point based on its distance to the actual data points.
+Each grid point is a potential cluster center.
+The potential for each grid is calculated based on the density of the surrounding data points.
+The grid with the highest potential is selected as the first cluster center and then the potential values of all the other grids are reduced according to their distances to the first cluster center.
+\newline
+Subtractive clustering is designed for learning fuzzy systems models from data\cite{Chiu1994FuzzyEstimation}. 
+To find out the number of clusters we need for our model we can make use of these models.
 
@@
 Other variables such as cluster boundaries and overlap were also controlled in the experiment.
-They observe that all the C-means and fuzzy C-means had good performance for non overlapping situations\cite{Mingoti2006ComparingAlgorithms}.
+They observe that the C-means and fuzzy C-means had good performance for non overlapping situations\cite{Mingoti2006ComparingAlgorithms}.
 The best results for average recovery and internal dispersion rates were found for Fuzzy C-means which was very stable in all situations achieving recovery averages over 90\%\cite{Mingoti2006ComparingAlgorithms}
@@
 \subsection{Training strategy}
+\label{ch:training_strategy}
 Our data set contains of mutants represented by their characteristics. 
@@
 For example mutant with a mutator in a certain project can have a different effect on the source code and can give other results than a mutator in a different project.
-To keep the relations and context we decided to train our model per project.
+To keep the relations and context we decided to train a model per project.
 \newline
-To validate our model we need to split our data set into a training set and a validation set.
+To validate the model we need to split our data set into a training set and a validation set.
 Our total data set consists of approximately 1 million mutants generated over fourteen projects.
 Research has shown that if the data set is big enough a ratio of 80/20 is sufficient\cite{Guyon1997ARatio}.
-We will divide our projects into training and validation groups to reflect the 80/20 ratio.
+We will divide each source projects into training and validation groups to reflect the 80/20 ratio as best as possible.
+Depending on the amount of mutants per project we might not achieve a perfect 80/20 ratio.
 
 \subsection{Implementation of algorithm}
-The origin of the fuzzy c-means algorithm is referenced in the survey paper of Du et al\cite{Du2010Clustering:Approach}.
+The research paper of the fuzzy c-means algorithm is referenced in the survey paper of Du et al\cite{Du2010Clustering:Approach}.
 The fuzzy c-means algorithm is developed by Bezdek et al.\cite{Bezdek1984FCM:Algorithm}.
@@
 \subsection{Algorithm parameters}
+The fuzzy c-means algorithm has to initialize the centroids of the clusters.
+Our selected library does this based on a random number.
+It also allows us to control this random number.
+To make the initialization of the centroid reproducible and deterministic we will use a generated seed.
+The seed generated is included in the Github repository\cite{rbasarat-repo}.
+\newline
+Bezdek et al., also researched cluster validity for the fuzzy c-means model\cite{Bezdek1995}.
+They researched what effect which parameters have on the validity of the clusters.
+They specifically analyzed the role of weighting exponent \textit{m}(fuzziness parameter).
+They concluded that the best choice for \textit{m} is probably in the interval [1.5, 2.5], whose mean and midpoint is \textit{m}=2.
+Wu proposed a new guideline for selecting parameter \textit{m}\cite{Wu2012}.
+His point of view was that a large \textit{m} value will make the fuzzy c-means more robust to noise and outliers.
+He suggests implementing fuzzy c-means with \textit{m}=[1.5, 4].
+When the data set contains noise and outliers, the fuzzifier\textit{m}=4 is recommended in a theoretical upper bound case.
+Since our data set does not contain noise and outliers we choose to select the midpoint concluded in the research of Bezdek et al.\cite{Bezdek1995OnModel}.
+The value of \textit{m} in this experiment is 2.
+\newline
+Finding the correct amount of clusters has been an ongoing problem for clustering algorithms\cite{Du2010Clustering:Approach}.
+Fortunately there are some techniques available to estimate the cluster amount.
+During model selection we discussed the mountain and subtractive models. 
+These models are designed to estimate the amount of clusters in a data set.
+The subtractive clustering model is specifically designed to approximate the number of clusters for fuzzy system models\cite{Chiu1994FuzzyEstimation}.
+We will select and implement the subtractive clustering model to estimate the amount of clusters.
+We can use this amount for the cluster amount parameter in the fuzzy c-means algorithm.
+
+\subsection{Implementation of subtractive clustering}
+The research paper of the mountain and subtractive algorithms are referenced in the survey paper of Du et al\cite{Du2010Clustering:Approach}.
+Based on the paper of the theory of the subtractive algorithm\cite{Chiu1994FuzzyEstimation} we found an existing implementation\cite{matlabSubtractive}.
+The implementation is professionally build and maintained by Matlab\cite{matlabSubtractive}.
+Matlab is a tool commonly used in the academic world and is professionally maintained for this reason we decided to use their implementation of subtractive clustering.
 
 \subsection{Validation}
-
+To validate our machine learning model we will calculate the weighed mutation score and cluster accuracy based on the clusters calculated by the model.
+This is the same validation method as described in Chapter \ref{ch:exp1_validation}.
+We will use the mutants of projects that are not selected for training as input for validating the model.
+As described in Chapter \ref{ch:training_strategy} the projects should represent approximately 20\% of the total amount of mutants in our data set.
+It may be possible the model has to cluster mutants it has never seen before.
+We will use the same statistical significance threshold as in our firsts experiment to decide whether we will accept or reject or hypothesis.
+To make our results reproducible we will select a random mutant based on a generated seed.
+The seed generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.
 \end{document}

### 2021-07-03 chapters/1_abstract.tex
Added abstract

\documentclass[../main]{subfiles}
\begin{document}
\chapter*{Abstract}
Software testing is a crucial part of the software engineering process. 
It is widely used in the industry.
A part of software testing is building test suites which contain unit tests.
These unit tests are written by developers. 
As projects grow the test suite grows along.
Maintaining and monitoring these test suites is important because they influence the cost of maintenance.
Monitoring the quality of test code has shown that tests with lower quality leads to more defect-prone production code.
Measuring the effectiveness of test suites detecting faults is part of the monitoring process.
A technique to measure the test effectiveness is mutation testing.
Mutation testing is a computationally expensive technique that requires the ability to run tests.
Different techniques have been proposed to reduce the cost of mutation testing.
One of these techniques is the clustering of mutants.
By clustering the mutants we can execute less mutants to reduces the cost.
We conducted two experiment in which we cluster the mutants with two different techniques.
Before conducting the experiment we did a state of the art comparison between three mutation testing tools for Java.
Based on this research we selected a mutation testing tool to use in our experiments.
\newline
We started our experiment by identifying characteristics to represent the mutants.
For the first experiment we used hierarchical clustering and in the second experiment we trained the fuzzy c-means model.
We calculated a weighed mutation score and compared this with the mutation score of a full set of mutants executed.
Results show that with hierarchical clustering we can reduce the amount of mutants executed while maintaining the effectiveness.
The clusters generated by the machine learning model were less accurate and showed a significant effectiveness decrease.
The machine learning model didn't perform as well as the hierarchical clustering.
However, more work is needed in finding the right parameters for the fuzzy c-means algorithm. 
\end{document}

### 2021-07-03 chapters/7_discussion.tex
Added abstract

\documentclass[../main]{subfiles}
\begin{document}

\chapter{Discussion}
\label{ch:discussion}
In this chapter, we discuss the results of our experiment(s) on ...

\begin{finding}
	Highlight like this an important finding of your analysis of the results.
	\label{find:important1}
\end{finding}

Refer to Finding~\ref{find:important1}.

\section{Threats to validity}
What could affect the validity of your research? Think of pitfalls of your research method, experimental setup, interpreting the results.

\end{document}

### 2021-07-03 chapters/8_related_work.tex
Added abstract

\documentclass[../main]{subfiles}
\begin{document}

\chapter{Related work}
\label{ch:related_work}
We divide the related work into the following categories:  ...

\end{document}

### 2021-07-05 chapters/7_discussion.tex
added results exp1

@@
 \label{ch:discussion}
-In this chapter, we discuss the results of our experiment(s) on ...
+In this chapter, we discuss the results of our experiments on clustering the mutants.
 
@@
 What could affect the validity of your research? Think of pitfalls of your research method, experimental setup, interpreting the results.
+- pit generated mutants -> it's hard to know if we can cluster ALL posible mutants
+- Selected sources
+- equivalent mutants
 

### 2021-07-05 chapters/7_discussion.tex
rq1 discussion finished

@@
 In this chapter, we discuss the results of our experiments on clustering the mutants.
+We will first discuss the results of the first experiment. 
+For the first experiment will discuss the all the results together and compare the results between the variants.
+Secondly we will discuss the results of the second experiment.
+Thirdly we will discuss how the results both experiments compare to each other.
+And last we will discuss the practicality of our research.s
+
+\section{RQ 1: Identifying and clustering mutant characteristics}
+The maximum deviation from the original mutation score, for all variants(with and without Levenshtein distance) of the experiment, are below the statistical 
+As hypothesised we observe that it is possible to cluster mutants with the set of characteristics we identified while maintaining effectiveness and reducing the amount executed.
+The hypothesis holds for all the projects and variants of the experiment.
+Most of the projects had a deviation of less than 1\% for all variations.
 
 \begin{finding}
-	Highlight like this an important finding of your analysis of the results.
-	\label{find:important1}
+    With a reduction up to 75\% of the total amount of mutants we can still maintain effectiveness.
+\end{finding}
+In both variants we observe a reduction in average accuracy when reducing the amount of clusters.
+This reductions is 0.1\% for clustering with all characteristics and 0.04\% for clustering without Levenshtein distance.
+We can also see that the average cluster accuracy is lower when reducing the amount executed.
+\begin{finding}
+    The more you reduce the amount of mutants executed the more accuracy is lost.
+\end{finding}
+The results show that the average accuracy for reductions of 25\% and 50\% is the same.
+This observation shows true for both variants of the experiment which is remarkable.
+\begin{finding}
+    In our data set it is safe to reduce the mutants by half without reducing the accuracy.
+\end{finding}
+It is also remarkable to see the min and max accuracy of the all clusters are 50\% to 100\% respectively.
+Based on these results we can state the following:
+\begin{itemize}
+    \item For every project there was at least one cluster that consisted only of killed or survived mutants.
+    \item For every project there was at least one cluster that evenly divided between killed and survived mutants.
+\end{itemize}
+With the exception of all Google Auto Service and Commons Text with a reduction of 25\% and all characteristics all projects showed an average accuracy above 90\%.
+The combination of hierarchical clustering and the identified characteristics results in high accuracy clusters.
+As logic dictates we expect that a higher accuracy means a score closer to that of the original.
+However we do not observe this behaviour in our results.
+This is due to our random sampling method.
+If we would select a mutant in the majority of each cluster the accuracy per cluster would correlate to the deviation in mutation score.
+Unfortunately we do not know the majority of a cluster as we do now know if a mutant survives or is killed before execution.
+\begin{finding}
+    Random sampling a mutant per cluster may result in a lower mutation score.
+\end{finding}
+As stated in Chapter \ref{ch:results_rq1} the calculation of the Levenshtein distance took a long time.
+For this reason we repeated the experiment without the Levenshtein distance.
+Comparing the results between these variants we can observe a max deviation of 0.05\% between average difference of the mutation score of.
+In 66\% of the results displayed in Chapter \ref{ch:results_rq1} clustering without the Levenshtein distance is more accurate.
+The biggest difference measured between all the projects and their reductions is 2.13\% where of clustering without Levenshtein distance was more accurate.
+\begin{finding}
+    While clustering without the Levenshtein distance was more accurate in our results, the Levenshtein distance characteristic does not significantly increase or decrease the cluster accuracy.
 \end{finding}
 
-Refer to Finding~\ref{find:important1}.
+The application of our research is limited to all software projects written in the language Java and the limitations of our mutation testing tool PIT\cite{pit}.
+This scope is significantly larger than most of the research done in the domain of mutation clustering(Chapter \ref{ch:related_work}).
 
@@
 What could affect the validity of your research? Think of pitfalls of your research method, experimental setup, interpreting the results.
-- pit generated mutants -> it's hard to know if we can cluster ALL posible mutants
+- pit generated mutants -> it's hard to know if we can cluster ALL possible mutants
 - Selected sources

### 2021-07-05 chapters/rq2/1_experiment_design.tex
Added abstract
rq1 discussion finished

@@
 For example mutant with a mutator in a certain project can have a different effect on the source code and can give other results than a mutator in a different project.
-To keep the relations and context we decided to train a model per project.
+To keep the relations and context we decided to train a separate model per project.
 \newline
@@
 Depending on the amount of mutants per project we might not achieve a perfect 80/20 ratio.
+The division of the mutants will be done randomly. 
+To make our results reproducible we will select a random division based on a generated seed.
+The seed generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.
 
@@
 To make our results reproducible we will select a random mutant based on a generated seed.
+We will repeat our experiment 30 times with 30 different seeds.
+Achieving consistent results while applying random selection and cluster initialization will contribute to the validity of the experiment.
 The seed generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.
+
 \end{document}

### 2021-07-06 chapters/1_abstract.tex
added correct results of exp2

@@
 One of these techniques is the clustering of mutants.
-By clustering the mutants we can execute less mutants to reduces the cost.
+By clustering the mutants we can execute less mutants to reduce the cost.
 We conducted two experiment in which we cluster the mutants with two different techniques.
@@
 For the first experiment we used hierarchical clustering and in the second experiment we trained the fuzzy c-means model.
-We calculated a weighed mutation score and compared this with the mutation score of a full set of mutants executed.
+We calculated a weighted mutation score and compared this with the mutation score of a full set of mutants executed.
 Results show that with hierarchical clustering we can reduce the amount of mutants executed while maintaining the effectiveness.

### 2021-07-06 chapters/7_discussion.tex
added correct results of exp2
discussion without threats to validity

@@
 In this chapter, we discuss the results of our experiments on clustering the mutants.
-We will first discuss the results of the first experiment. 
-For the first experiment will discuss the all the results together and compare the results between the variants.
-Secondly we will discuss the results of the second experiment.
-Thirdly we will discuss how the results both experiments compare to each other.
-And last we will discuss the practicality of our research.s
+First we discuss the results of the first experiment. 
+We will discuss the all the variant results together and then compare the results between the variants.
+Second we will discuss the results of the second experiment and compare them to those of the first.
+
+And last we will discuss the practicality of our research.
 
 \section{RQ 1: Identifying and clustering mutant characteristics}
-The maximum deviation from the original mutation score, for all variants(with and without Levenshtein distance) of the experiment, are below the statistical 
+The maximum deviation from the original mutation score, for all variants(with and without Levenshtein distance) of the experiment, is below the statistical threshold.
 As hypothesised we observe that it is possible to cluster mutants with the set of characteristics we identified while maintaining effectiveness and reducing the amount executed.
 The hypothesis holds for all the projects and variants of the experiment.
-Most of the projects had a deviation of less than 1\% for all variations.
+66\% of the displayed results had a deviation of less than 1\%. 
+This includes both variations of the experiment.
 
@@
 In both variants we observe a reduction in average accuracy when reducing the amount of clusters.
-This reductions is 0.1\% for clustering with all characteristics and 0.04\% for clustering without Levenshtein distance.
+These reductions are 0.1\% and 0.04\% for clustering with and without Levenshtein distance respectively.
 We can also see that the average cluster accuracy is lower when reducing the amount executed.
@@
 \begin{finding}
-    In our data set it is safe to reduce the mutants by half without reducing the accuracy.
+    We can reduce the amount of mutants executed in our data up to half of the original amount before reducing the accuracy even more.
 \end{finding}
@@
 \end{itemize}
-With the exception of all Google Auto Service and Commons Text with a reduction of 25\% and all characteristics all projects showed an average accuracy above 90\%.
+With the exception of Google Auto Service and Commons Text, with a reduction of 25\% and all characteristics, all projects showed an average accuracy above 90\%.
 The combination of hierarchical clustering and the identified characteristics results in high accuracy clusters.
@@
 \begin{finding}
-    Random sampling a mutant per cluster may result in a lower mutation score.
+    Random sampling a mutant per cluster may result in a bigger deviation between original score and weighted score.
 \end{finding}
@@
 In 66\% of the results displayed in Chapter \ref{ch:results_rq1} clustering without the Levenshtein distance is more accurate.
-The biggest difference measured between all the projects and their reductions is 2.13\% where of clustering without Levenshtein distance was more accurate.
+The biggest difference in mutation score measured between all the projects and their reductions is 3.63\%.
+Clustering without Levenshtein distance is on average more accurate.
 \begin{finding}
@@
 
+\section{RQ 2: Training a machine learning model and clustering mutant characteristics}
+The maximum deviation from the original mutation score, for all variants(with and without Levenshtein distance) of the experiment, is above the statistical threshold.
+The biggest deviation measured is 22.39\%.
+This is significantly higher than the biggest deviation measured in the first experiment.
+56\% of the results of both variants had a deviation in mutation score bigger than 5\%.
+\begin{finding}
+Our hypothesis does not hold for the clustering the mutants with the machine learning model trained in our experiment.
+\end{finding}
+We observe that the number of clusters does not have an obvious relation to the total amount of mutants. 
+This is expected as the subtractive clustering does not look at the total amount of data points(mutants) but to the contents of the data points.
+The number of clusters compared to the cluster amounts of the hierarchical clustering is significantly lower.
+While we do not know the exact clusters centroids the subtractive model calculated, a low cluster amount indicates close relationships between certain mutants.
+\begin{finding}
+The mutants represented as characteristics in our data set per project have, according to the subtractive clustering model, close relationships to certain other mutants in the same data set.
+\end{finding}
+We observe that the minimum lowest minimum cluster accuracy measured over all variants and projects is 50.54\% with and average of 62\%. 
+This is higher than the minimum of the hierarchical clustering.
+This means that in all project there was not one cluster that was evenly divided.
+While the minimum accuracy was higher, the maximum accuracy was lower compared to the maximum accuracy of the hierarchical clustering results.
+The highest measured accuracy  measured over all variants and projects is 97.80\% with an average of 83.79\%
+This means that in all project there was not one cluster that consisted of only survived or killed mutants.
+\begin{finding}
+The overall accuracy of the clusters generated by the machine learning model in both variants if lower than that of the hierarchical clustering.
+\end{finding}
+When we compare the accuracy per project between the machine learning model and hierarchical clustering we can observe that the lower accuracy of the machine learning model clusters have a higher impact on the deviation in mutation score.
+This is caused by the relatively low amount of clusters.
+One cluster contains more mutants which results in a bigger weight per cluster.
+\begin{finding}
+The inaccuracy of clusters has a bigger impact when the cluster amount is lower.
+\end{finding}
+
+\section{Application}
 The application of our research is limited to all software projects written in the language Java and the limitations of our mutation testing tool PIT\cite{pit}.
+We can repeat our experiment for every java project that can be run by PIT.
 This scope is significantly larger than most of the research done in the domain of mutation clustering(Chapter \ref{ch:related_work}).
+The extraction of mutants is build within the PIT plugin system which is a plug and play system.
+Applying our research can be done in three steps.
+The first step is to run PIT with the plugin to gather the characteristics.
+This plugin makes sure PIT exits without executing mutants after gathering the characteristics.
+The second step is to run the script for either the machine learning model or the hierarchical clustering.
+The last step is to run PIT again with the plugin to filter the mutants based on the results of the clustering.
+It is possible to build all these steps into a DevOps pipeline to automate the process. 
+The plugins and scripts needed for these steps are available on Github\cite{rbasarat-repo}.
 

### 2021-07-06 chapters/rq2/1_experiment_design.tex
discussion without threats to validity
fixed wrong cite

@@
 \newline
-Bezdek et al., also researched cluster validity for the fuzzy c-means model\cite{Bezdek1995}.
+Bezdek et al., also researched cluster validity for the fuzzy c-means model\cite{Bezdek1995OnModel}.
 They researched what effect which parameters have on the validity of the clusters.
@@
 The subtractive clustering model is specifically designed to approximate the number of clusters for fuzzy system models\cite{Chiu1994FuzzyEstimation}.
+It approximates the centroids and amount of the clusters based on the data points in the data set\cite{Chiu1994FuzzyEstimation}.
 We will select and implement the subtractive clustering model to estimate the amount of clusters.

### 2021-07-12 chapters/2_introduction.tex
changes

@@
 A recurring problem in the research of mutation clustering is to determine the means of domains as the centre of a cluster\cite{Ji2009,Wilinski2015,Ma2016,Wei2021SpectralTesting}.
-
+\newline
 To reduce the cost of mutation testing, we try to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants. 
@@
 Based on the results of this research we select and implement a clustering algorithm that will cluster a characteristic of the mutants. 
-One or more mutants of each cluster will be executed by the mutation tooling.
+We will randomly select on mutant per of each cluster to be executed by the mutation tooling.
 We will then compare the mutation score of the clustered mutants with the mutation score of a full set of mutants.
@@
 In Chapter~\ref{ch:background} we describe the background of this thesis. 
-\todo{finish this at the end}
-Chapter~\ref{ch:research} describes ... 
-Results are shown in Chapter~\ref{ch:results} and discussed in Chapter~\ref{ch:discussion}. Chapter~\ref{ch:related_work}, contains the work related to this thesis.
+Chapter \ref{ch:mutation-tool} compares the different mutation testing tools to decide on a tool to use for this research.
+In chapter \ref{ch:reserach_question1} we identify characteristics to represent mutants and design an experiment to cluster them.
+In chapter \ref{ch:reserach_question2} we select a machine learning model and design an experiment to cluster the mutants with this model.
+Results are shown in Chapters \ref{ch:results_rq1} and \ref{ch:results_rq2} for the experiments in described Chapters \ref{ch:reserach_question1} and \ref{ch:reserach_question2} respectively. 
+The results are then discussed in Chapter~\ref{ch:discussion}. Chapter~\ref{ch:related_work}, contains the work related to this thesis.
 Finally, we present our concluding remarks in Chapter~\ref{ch:conclusion} together with future work.

### 2021-07-12 chapters/3_background.tex
minor tweaks exp2 design + results
changes

@@
 \section{Project selection}
+\label{ch:project_selection}
 We choose three main requirements for selecting software projects; the projects should have a test suite, the test suite should not contain failing tests and the mutation testing tool should be able to execute mutants for the sample project.
@@
 From there we filter out all the projects that contain failing tests and that aren't libraries or applications.
-The result consists of a sample with sixteen projects in total.
+The result consists of a sample with fifteen projects in total.
 
@@
     \item C-means(K-means) clustering.
-    \item Mountain clustering.
+    \item Mountain and subtractive clustering.
     \item Neural gas.

### 2021-07-12 chapters/7_discussion.tex
changes

@@
 Second we will discuss the results of the second experiment and compare them to those of the first.
-
 And last we will discuss the practicality of our research.
@@
 This includes both variations of the experiment.
-
+\newline
+The results prove that the relations between the characteristics we identified can be used to cluster mutants efficiently. 
+This can be explained by the correlation between the result of a mutant and the euclidean distance used to decide whether a mutants should be clustered together.
+If two mutants are grouped it is \textit{n} percent likely that the mutants have the same result where \textit{n} is the cluster accuracy.
 \begin{finding}
@@
 While the minimum accuracy was higher, the maximum accuracy was lower compared to the maximum accuracy of the hierarchical clustering results.
-The highest measured accuracy  measured over all variants and projects is 97.80\% with an average of 83.79\%
+The highest accuracy measured over all variants and projects is 97.80\% with an average of 83.79\%
 This means that in all project there was not one cluster that consisted of only survived or killed mutants.
 \begin{finding}
-The overall accuracy of the clusters generated by the machine learning model in both variants if lower than that of the hierarchical clustering.
+The overall accuracy of the clusters generated by the machine learning model in both variants is lower compared to that of the hierarchical clustering.
 \end{finding}
@@
 \begin{finding}
-The inaccuracy of clusters has a bigger impact when the cluster amount is lower.
+The inaccuracy of clusters has a bigger impact on mutation score when the cluster amount is lower.
 \end{finding}
@@
 \section{Threats to validity}
-What could affect the validity of your research? Think of pitfalls of your research method, experimental setup, interpreting the results.
-- pit generated mutants -> it's hard to know if we can cluster ALL possible mutants
-- Selected sources
-- equivalent mutants
+In this section, we discuss the threats to internal validity, external validity and reliability.
+\subsection{Internal validity}
+\subsubsection{Equivalent mutants}
+Equivalent mutants are mutants that are syntactically different but functionally the same.
+These mutants influence the mutation score while giving the same result as their equivalent counterpart.
+In our research equivalent mutants could give a false image of the accuracy of the clusters and the overall accuracy if such a mutant was randomly selected.
+PIT contains some functionality for reducing equivalent mutants\cite{pit}.
+We also reduced the chance of inaccuracy by repeating our experiment.
+
+\subsubsection{Test set representativity}
+The software projects we used in our research were systematically selected.
+Chapter \ref{ch:project_selection} provides an indication but not a wide representation of industrial software projects.
+The conclusions drawn regarding the increase performance by reducing the mutants are based on the comparison between our own baseline measurements.
+Another threat which is the result of the representation of our test set is the representation of the mutants.
+While our data set contains around 1 million mutants we do not know if we have seen all mutants that exist.
+This will always be a problem as it would be nearly impossible to cluster and mutation test every piece of Java code, that conforms to the requirements of PIT, ever written.
+
+\subsubsection{Project configuration}
+Some tests may not be recognized by PIT due to some external data required for the test.
+Another reason may be that the test uses global variables that are not available to PIT.
+We encountered some of these tests in our data set.
+While these tests did pass when run with only the unit test framework they would not when run by PIT.
+These projects still met the requirements stated in Chapter \ref{ch:project_selection}.
+To solve this issue we removed these tests from the test suite.
+By removing these specific tests we may have slightly, tough negligible, deviated from the individual real-world settings for respective projects.
+
+\subsubsection{Expertise}
+While this research is in the domain of software engineering it borrows techniques from data science.
+The researcher's field of expertise is software engineering.
+Since the researcher had to use techniques from different domains we can only guarantee that the research executed was done to the best of their ability.
+Thus the research may contain sub optimal design choices which may have been made from a software engineering perspective.
+
+\subsection{External validity}
+We clustered mutants with two different techniques.
+Our results are not generalisable to projects written in other programming languages as well as projects that cannot be run by PIT.
+As a result of using PIT we limit ourselves to two testing frameworks; JUnit and TestNG.
+PIT is still under active development and may expand its applicability, which in turn increases the applicability of this research.
 

### 2021-07-12 chapters/8_related_work.tex
changes

@@
 \label{ch:related_work}
-We divide the related work into the following categories:  ...
+Clustering mutants has been done before.
+The research that exists tries to cluster mutants by defining the centroid of a cluster and cluster the mutants on that definition.
+In our research extends mutation clustering by defining a centroid that is dynamic.
+In other words the centroids of our clusters are defined by the mutants itself.
+
+\subsection{Clustering overlapped mutants}
+\label{ch:overlapping_mutants}
+Ma et al., clustered mutants by overlap\cite{Ma2016}.
+They defined the term overlap as as mutants that are functionally equivalent.
+This is close to an equivalent mutant but not the same.
+They explain that an equivalent mutants is functionally identical to the original source code while an overlapped mutant is functionally identical to at least one other mutant.
+If a mutant does not have an overlapped mutant that mutant would be a cluster on its own.
+The achieved a reduction, in mutants executed, of 10\%.
+They write in their research that it is limited to the mutator operators on expression level. 
+If a mutator did not generate two or more mutants they could not detect overlapping mutants.
+Their future work includes adding more mutators and widening the scope level to statements or blocks.
+\newline
+The research does not include efficiency.
+While they show that they can cluster mutants based on overlap they do not show how efficient the mutation testing process is compared to executing a full set of mutants.
+A threat to their validity, which is not addressed, is the fact that two overlapping mutants may not be challenged by the same tests and thus can result in different results.
+Our research solves this problem by taking into account the location of the mutants.
+Chapter \ref{ch:characteristics_location} describes what characteristics we use.
+
+\subsection{Clustering by scope}
+Yu et al., extended the research about overlapping mutants\cite{Yu2019PossibilityScope}.
+They added more mutators to the scope.
+The also extended the research to include statements and blocks.
+Their results shows an increase in clustered mutants.
+The problems mentioned in Chapter \ref{ch:overlapping_mutants} are not addressed in this research.
+
+\subsection{Clustering Hamming distance}
+Ji et al., did a qualitative domain analysis on mutants\cite{Ji2009}.
+As a result they identified the Hamming distance for to cluster with. 
+The Hamming distance is, like the Levenshtein distance, a similarity measure.
+They use the k-means to cluster the mutants with the Hamming distance.
+They write that the reduced test set in their experiment is still as strong as the original test set\cite{Ji2009}.
+Ji et al., were successful in clustering mutants with Hamming distance and the k-means algorithm.
+But they do remark that their research still has problems with rationally determining the domains of their clusters.
+Our research does not have this problem.
+In our reserach the centre of of the domains are decided by the characteristics of the mutants.
+
+\subsection{Spectral clustering}
+Wei et al., makes use of an intelligent technique, namely spectral clustering, to improve the efficacy of mutant reduction\cite{Wei2021SpectralTesting}.
+They defined multiple definitions for the mutants; similarity distance, distance between the mutants and a killing matrix of the mutants.
+With these definitions they reduce the mutants based on their proposed reduction method.
+This method is based on spectral clustering (SCMT), including the determination method of the number of clusters, spectral clustering of mutants, and selection of representative mutants.
+The reduced were then clustered with a classical clustering algorithm.
+Their results are promising and show good cluster accuracy.
+They write that there is still work left to do in optimizing the matrices and clustering algorithms.
+Our goal overlaps with that of Wei, the difference is that we used a different methodology to achieve that goal.
+
+\subsection{Clustering similarity}
+Hussain et al., used the k-means and agglomerative clustering algorithm to cluster the mutants according to a similarity measure. 
+They used the Hamming distance a similarity measure.
+They calculated the distance and used this as data te feed into the clustering algorithms.
+The number of clusters and the initial position of the cluster center in the k-means algorithm are difficult to determine, and the process of the agglomerative clustering algorithm makes it difficult to correct the existing cluster formation\cite{Hussain2008}.
+They did take into account efficiency.
+Hussain et al, had the same goals as our research. 
+The difference is that our research used different and more elaborated methods to cluster the mutants.
+
+\subsection{Generalizing mutants}
+Wilinski et al., tried to generalize the mutants by defining metrics.
+Each metric is calculated for a specific mutator.
+The three metrics are usefulness, frequency and dependency.
+Combining the results of usefulness and frequency metrics, they observed that reducing the number of generated mutants gives noticeable mutant cost reduction without a loss of the mutation score accuracy.
+Wilinski et al., their research is very narrowly scoped to the specific mutators they decided to do research on.
+Our research does not contain this limitation.
+The mutator is used as a characteristic in our research instead of limiting our research it helps define the the mutant.
 

### 2021-07-12 chapters/9_conclusion.tex
changes

\documentclass[../main]{subfiles}
\begin{document}

\chapter{Conclusion}
\label{ch:conclusion}
Our research started with comparing three mutation testing tools for Java and found that the state of the art was outdated.
We replicated the state of the art research with current data available on the mutation testing tools.
The state of art showed that two out of three mutation testing tools were not actively maintained anymore.
The mutation testing tool PIT scored best on the metrics measured in our replication study, this is also the tool we decided to use for our research.
\newline
We created a list of characteristics based on existing research. 
We then extended this list by identifying characteristic with logic and data available.
We used the identified characteristics to represent the mutants.
\newline
We clustered the mutants with two different clustering techniques; hierarchical clustering and by training a machine learning clustering model.
The hierarchical clustering showed that we can cluster mutants while maintaining effectiveness.
We achieved reductions of up to 75\% of the total amount of mutants without significantly reducing the accuracy.
\newline
The machine learning model showed a significant reduction in effectiveness.
We trained the machine learning model with parameters that were optimal according to existing research. 
We also used a machine learning model, specialized in calculating the number of clusters for fuzzy clustering, to calculate the amount of clusters.
While the minimum clustering accuracy was higher than that of the hierarchical clustering, the maximum cluster accuracy was lower than that of the hierarchical clustering.
The cluster accuracy had a greater impact on the clusters of the machine learning model because there was a relatively small amount of clusters.
Our research proves that it is possible to determine the means of domains as the centre of a cluster within a programming language without narrowing down the scope to specific pieces of code.
It can be applied to all Java projects that can be run with PIT. 
Which is a larger scope than existing mutation clustering research used.

\section{Future work}
\label{sec:future_work}
\subsubsection{Optimizing parameters of machine learning model}
We trained the machine learning model with specific parameters guided by science. 
These proved to be non optimal for our purpose.
This research can be extended by finding the optimal parameters for the fuzzy c-means clustering algorithm.
The results showed a higher minimum accuracy which can indicate that there is potential for clustering the mutants with the machine learning model.
Further research and fine-tuning is required to find out if the model can be trained to achieve more accurate scores.

\subsubsection{Training strategy}
Our approach for training the model proved to be non optimal.
During the training of the model we experienced that the more clusters you select the more memory is needed for training.
This research did not have sufficient hardware available.
A different approach in training strategy may be explored to mitigate the hardware limitation.

\subsubsection{Training a single machine learning model}
For this researched we trained a model for each source project. 
A different way is to train one machine learning model with the total set of mutants.
The model could be continuously trained each time a new source is added.
Further research is recommended in this direction as it will improve the applicability of this research.

\subsubsection{Different sets of characteristics}
We have proven that this set of characteristics can achieve accurate results in combination with hierarchical clustering.
It would be valuable to find out if it possible to reduce or expand the set of characteristics and still maintain accurate results.
Further experimentation can be done with different sets of characteristics.
\end{document}

### 2021-07-12 chapters/rq1/2_experiment_design.tex
added results exp1
rq1 discussion finished
changes

@@
 The agglomerative hierarchical technique follows bottom up approach whereas divisive follows top-down approaches.
-Hierarchical clustering uses different metrics which measures the the distance between two clusters and the linkage criteria\cite{Rajalingam2011}. 
+Hierarchical clustering uses different metrics which calculates the euclidean distance between two clusters and the linkage criteria\cite{Rajalingam2011}. 
 The linkage criteria specifies the dissimilarity in the sets as a function of the pair-wise distances of observations in those sets\cite{Rajalingam2011}.
@@
 Since we cannot assume that our clusters are well separated we chose to use the ward linkage method for our clustering algorithm.
-The agglormorative hierarchical clustering algorithm with ward linkage will cluster our mutants represented by the characteristics we gathered.
+The agglomerative hierarchical clustering algorithm with ward linkage will cluster our mutants represented by the characteristics we gathered.
 
@@
 \subsection{Cluster size}
-We decide our cluster size based on the amount of mutants generated in a full set.
-As a baseline we will reduce the amount of mutants by 75\%, 50\% and 25\%.
-This results in a cluster size that will be equal to 75\%, 50\% and 25\% of the total amount of mutants generated in a full set.
+Agglomerative hierarchical clustering will continue to cluster until there is one cluster left. 
+Naturally this cluster will contain all the mutants.
+We can cut off the clustering algorithm at any point.
+We decide on a cluster amount based on the amount of mutants generated in a full set.
+As a baseline we will cut off clustering in steps of 25\%.
+The steps of 25\% will give a good global indication in reduction.
+If necessary we can decide to cut of the clustering at different points.
+The baseline will consist of three experiments where the cluster size is 25\%, 50\% and 75\% of the total amount of mutants respectively.
 The amount of mutants inside a cluster will be decided by the clustering algorithm.
@@
 In other words we want achieve a mutation score that is as close as possible as to that of a full set.
-We select a significance level of $\leq 0.05$.
+We select a statistical significance level of $\leq 0.05$.
 This is the conventional threshold for declaring statistical significance\cite{Kirk1996PracticalCome}.

### 2021-07-13 chapters/rq1/2_experiment_design.tex
fixex typo

@@
 
-\subsection{Weighed mutation score}
-\label{ch:weighed_score}
+\subsection{weighted mutation score}
+\label{ch:weighted_score}
 Our hypothesis states that each mutants executed should represent that whole cluster.
 With a mutant executed from each cluster we can calculate a mutation score.
-This mutation score is a weighed mutation score.
-This weighed mutation score is the product of the result of a mutant(1 for killed and 0 for survived) and the amount of mutants in the cluster it represents.
-The weighed mutation score is then comparable to the score of a full set as the total number of mutants will be the same.
+This mutation score is a weighted mutation score.
+This weighted mutation score is the product of the result of a mutant(1 for killed and 0 for survived) and the amount of mutants in the cluster it represents.
+The weighted mutation score is then comparable to the score of a full set as the total number of mutants will be the same.
 \newline
@@
 The mutants representing cluster one and four survive and two and three are killed.
-If we calculate the weighed score we get 68/100 which is 68\%.
+If we calculate the weighted score we get 68/100 which is 68\%.
 We can then compare this to the score of a full set because the amount of mutants executed is the same: 75/100(75\%) and 68/100(68\%).
@@
 \newline
-To validate how effective our clustering is we can compare the weighed mutation score(see Chapter \ref{ch:weighed_score}) of our clustered set to the mutation score of the full set.
-The closer the weighed score is to that of a full set the more effective our set of characteristics and clustering algorithm proves to be.
+To validate how effective our clustering is we can compare the weighted mutation score(see Chapter \ref{ch:weighted_score}) of our clustered set to the mutation score of the full set.
+The closer the weighted score is to that of a full set the more effective our set of characteristics and clustering algorithm proves to be.
 In other words we want achieve a mutation score that is as close as possible as to that of a full set.
@@
 \newline
-If the weighed mutation score of our clustered set deviates more than 5\% from the score of a full set we will reject our hypothesis.
+If the weighted mutation score of our clustered set deviates more than 5\% from the score of a full set we will reject our hypothesis.
 We also compare the inaccuracy to the significance level to decide if our inaccuracy is statistical significant.

### 2021-07-13 chapters/rq2/1_experiment_design.tex
fixex typo

@@
 \subsection{Validation}
-To validate our machine learning model we will calculate the weighed mutation score and cluster accuracy based on the clusters calculated by the model.
+To validate our machine learning model we will calculate the weighted mutation score and cluster accuracy based on the clusters calculated by the model.
 This is the same validation method as described in Chapter \ref{ch:exp1_validation}.

### 2021-07-13 chapters/7_discussion.tex
fixed our tool to the tool

@@
 \section{Application}
-The application of our research is limited to all software projects written in the language Java and the limitations of our mutation testing tool PIT\cite{pit}.
+The application of our research is limited to all software projects written in the language Java and the limitations of the selected mutation testing tool PIT\cite{pit}.
 We can repeat our experiment for every java project that can be run by PIT.

### 2021-07-13 chapters/7_discussion.tex
number of clusters

@@
 \end{finding}
-In both variants we observe a reduction in average accuracy when reducing the amount of clusters.
+In both variants we observe a reduction in average accuracy when reducing the number of clusters.
 These reductions are 0.1\% and 0.04\% for clustering with and without Levenshtein distance respectively.
@@
 This is expected as the subtractive clustering does not look at the total amount of data points(mutants) but to the contents of the data points.
-The number of clusters compared to the cluster amounts of the hierarchical clustering is significantly lower.
-While we do not know the exact clusters centroids the subtractive model calculated, a low cluster amount indicates close relationships between certain mutants.
+The number of clusters compared to the that of the hierarchical clustering is significantly lower.
+While we do not know the exact clusters centroids the subtractive model calculated, a low number of clusters indicates close relationships between certain mutants.
 \begin{finding}
@@
 When we compare the accuracy per project between the machine learning model and hierarchical clustering we can observe that the lower accuracy of the machine learning model clusters have a higher impact on the deviation in mutation score.
-This is caused by the relatively low amount of clusters.
+This is caused by the relatively low number of clusters.
 One cluster contains more mutants which results in a bigger weight per cluster.
 \begin{finding}
-The inaccuracy of clusters has a bigger impact on mutation score when the cluster amount is lower.
+The inaccuracy of clusters has a bigger impact on mutation score when the number of clusters is lower.
 \end{finding}

### 2021-07-13 chapters/9_conclusion.tex
number of clusters

@@
 We trained the machine learning model with parameters that were optimal according to existing research. 
-We also used a machine learning model, specialized in calculating the number of clusters for fuzzy clustering, to calculate the amount of clusters.
+We also used a machine learning model, specialized in calculating the number of clusters for fuzzy clustering, to calculate the number of clusters.
 While the minimum clustering accuracy was higher than that of the hierarchical clustering, the maximum cluster accuracy was lower than that of the hierarchical clustering.
-The cluster accuracy had a greater impact on the clusters of the machine learning model because there was a relatively small amount of clusters.
+The cluster accuracy had a greater impact on the clusters of the machine learning model because there was a relatively small number of clusters.
 Our research proves that it is possible to determine the means of domains as the centre of a cluster within a programming language without narrowing down the scope to specific pieces of code.
 It can be applied to all Java projects that can be run with PIT. 
-Which is a larger scope than existing mutation clustering research used.
+Which is a larger scope than in existing mutation clustering research.
 

### 2021-07-13 chapters/rq1/2_experiment_design.tex
fixed our tool to the tool
number of clusters

@@
 We can cut off the clustering algorithm at any point.
-We decide on a cluster amount based on the amount of mutants generated in a full set.
+We decide on the number of clusters based on the amount of mutants generated in a full set.
 As a baseline we will cut off clustering in steps of 25\%.
@@
 To reduce the amounts executed we cluster the mutants. 
-We can measure this by counting the amount of clusters we generate and compare it to the number of total mutants generated by the selected tool.
+We can measure this by counting the number of clusters we generate and compare it to the number of total mutants generated by the selected tool.
 \newline
-The baseline for our experiment is the mutation score of the set of all mutants generated by our selected mutation testing tool.
+The baseline for our experiment is the mutation score of the set of all mutants generated by the selected mutation testing tool.
 \newline

### 2021-07-13 chapters/rq2/1_experiment_design.tex
number of clusters

@@
 A big pro of the neural gas network is that it is able to spawn neurons.
-This means that we do not need to know the amount of clusters in advance.
+This means that we do not need to know the number of clusters in advance.
 Unfortunately the neural gas network is a learning network.
@@
 Finding the correct amount of clusters has been an ongoing problem for clustering algorithms\cite{Du2010Clustering:Approach}.
-Fortunately there are some techniques available to estimate the cluster amount.
+Fortunately there are some techniques available to estimate the number of clusters.
 During model selection we discussed the mountain and subtractive models. 
-These models are designed to estimate the amount of clusters in a data set.
+These models are designed to estimate the number of clusters in a data set.
 The subtractive clustering model is specifically designed to approximate the number of clusters for fuzzy system models\cite{Chiu1994FuzzyEstimation}.
 It approximates the centroids and amount of the clusters based on the data points in the data set\cite{Chiu1994FuzzyEstimation}.
-We will select and implement the subtractive clustering model to estimate the amount of clusters.
-We can use this amount for the cluster amount parameter in the fuzzy c-means algorithm.
+We will select and implement the subtractive clustering model to estimate the number of clusters.
+We can use this number for the number of clusters parameter in the fuzzy c-means algorithm.
 

### 2021-07-13 chapters/1_abstract.tex
Fixing typos styling etc..

@@
 Monitoring the quality of test code has shown that tests with lower quality leads to more defect-prone production code.
+\newline
 Measuring the effectiveness of test suites detecting faults is part of the monitoring process.
@@
 One of these techniques is the clustering of mutants.
-By clustering the mutants we can execute less mutants to reduce the cost.
+By clustering mutants we can execute less mutants to reduce the cost.
+\newline
 We conducted two experiment in which we cluster the mutants with two different techniques.
-Before conducting the experiment we did a state of the art comparison between three mutation testing tools for Java.
+Before conducting the experiment we conducted a state of the art comparison between three mutation testing tools for Java.
 Based on this research we selected a mutation testing tool to use in our experiments.
@@
 Results show that with hierarchical clustering we can reduce the amount of mutants executed while maintaining the effectiveness.
-The clusters generated by the machine learning model were less accurate and showed a significant effectiveness decrease.
+The clusters generated by the machine learning model were less accurate and showed a significant decrease in effectiveness.
 The machine learning model didn't perform as well as the hierarchical clustering.

### 2021-07-13 chapters/2_introduction.tex
Fixing typos styling etc..

@@
 Research has also found that if the production code grows, the test code grows along with it\cite{Zaidman2008MiningCode}. 
-Quality control on test suites is therefore important as the maintenance on tests can be difficult and generate risks if done incorrectly\cite{Horvath2015TestMetrics}.
+Quality control on test suites is therefore important as the maintenance on tests can be difficult and generates risks if done incorrectly\cite{Horvath2015TestMetrics}.
 \newline
@@
 Another bug, which is even worse, are tests that do not fail when the program is not working as desired(\textit{false negative}).
-The \textit{false negative} is a problem when breaking changes or not detected by the test suite.
+The \textit{false negative} is a problem when breaking changes are not detected by the test suite.
 \newline
@@
 After we have selected a tool we will research existing clustering techniques.
-Based on the results of this research we select and implement a clustering algorithm that will cluster a characteristic of the mutants. 
+Based on the results of this research we select and implement a clustering algorithm that will cluster one ore more characteristics of the mutants. 
 We will randomly select on mutant per of each cluster to be executed by the mutation tooling.
@@
  \item A \acrfull{poc} which implements the qualitative methodology chosen and elaborated within the thesis.
- \item A \acrfull{poc} which implements the quantitative approach to cluster mutants. 
+ \item A \acrfull{poc} which implements the quantitative approach to cluster mutants and is elaborated within the thesis. 
 \end{enumerate}
@@
 In chapter \ref{ch:reserach_question2} we select a machine learning model and design an experiment to cluster the mutants with this model.
-Results are shown in Chapters \ref{ch:results_rq1} and \ref{ch:results_rq2} for the experiments in described Chapters \ref{ch:reserach_question1} and \ref{ch:reserach_question2} respectively. 
-The results are then discussed in Chapter~\ref{ch:discussion}. Chapter~\ref{ch:related_work}, contains the work related to this thesis.
+Results are shown in Chapters \ref{ch:results_rq1} and \ref{ch:results_rq2} for the experiments described in Chapters \ref{ch:reserach_question1} and \ref{ch:reserach_question2} respectively. 
+The results are then discussed in Chapter~\ref{ch:discussion}.
+\newline
+Chapter~\ref{ch:related_work}, contains the work related to this thesis.
 Finally, we present our concluding remarks in Chapter~\ref{ch:conclusion} together with future work.

### 2021-07-13 chapters/3_background.tex
fixed our tool to the tool
Fixing typos styling etc..

@@
 This chapter will present the necessary background information for this thesis. First, we define some basic terminology that will be used throughout this thesis.
-Next, we will compare existing mutation testing tools. We start of by reviewing existing literature and then extend the comparison. \todo{extend this later on?}
+Next, we will compare existing mutation testing tools. We start of by reviewing existing literature and then extend the comparison. 
+This thesis relies on clustering techniques.
+We review the state of the art on clustering algorithms and secondly the state of the art on machine learning models specialized in clustering.
+
 \section{Terminology}
-\todo{do this when we finished experiments and all}
-\todo{subsuming mutants}
-\todo{levensteihn distance}
-\todo{maven and gradle}
+Subsuming mutants.
+Mutation testing.
+Levenshtein distance.
+Maven.
+Gradle.
+Mutator.
+test engine??
 \section{Project selection}
@@
 We selected projects that were also used in other research within the mutation testing and testing domain\cite{Pizzoleto2019,Yu2019PossibilityScope,Wei2021SpectralTesting, Zhang2019PredictiveTesting, Chen2018SpeedingStudy, Laurent2017AssessingPIT}.
+\newline
 We extend our sample by selecting projects from the first six pages of the most popular Java projects on GitHub\footnote{\url{https://github.com/search?p=7&q=language\%3Ajava+stars\%3A\%3E10000&type=Repositories}}.
@@
 \newline
-For the default parameter situation the spectral clustering algorithm had the best performance and the hierarchical algorithm had the worst performance.
+For the default parameter experiment, the spectral clustering algorithm had the best performance and the hierarchical algorithm had the worst performance.
 \newline
@@
 K.L. Du surveyed different machine learning models used for clustering\cite{Du2010Clustering:Approach}.
-The models he discusses are:
+Relevant models he discusses are:
 \begin{itemize}

### 2021-07-13 chapters/4_mutation_tool.tex
changes
Fixing typos styling etc..

@@
 \chapter{Mutation testing tools}
-\label{ch:mutation∂-tool}
+\label{ch:mutation-tool}
 
 Before we can start clustering mutants we need to select a tool that helps us generate mutants. 
-Our goal is to find the tool that generates the greatest amount of mutants. 
+Our goal is to find the tool that generates the most amount of mutants. 
 \newline
@@
 Kintis et al. analysed and compared the performance of PIT,  MuJava and Major\cite{Kintis2016AnalysingStudy}. 
-They started out with performing a mini literature survey on mutation testing tools used in research on test effectiveness in Java. 
+They started out with performing a mini literature survey on mutation testing tools used in existing research on test effectiveness in Java. 
 With a data set of twelve methods and six Java projects they evaluated each tool using a cross-testing technique.
@@
 During their research they also compared the different mutant operators the tools supported. 
-An overview of mutant operators per tool was created. 
+An overview of mutant operators(also known as mutators) per tool was created. 
 They compared the mutant operators and reported overlapping operators.
@@
 These programs are popular programs used in testing literature.
-They found that the mutation tools do not subsume \todo{add subsumption to word list} each other. 
+They found that the mutation tools do not subsume each other. 
 According to the research a ranking of the tools would be as follows; MuJava generated the strongest mutants followed by Major and PIT.
-Furthermore they found that MuJava generated significantly more equivalent mutants.
+Furthermore they found that MuJava generated significantly more equivalent mutants\cite{Marki2017MutationJava}.
 MuJava also had an execution time twice the amount of that of Major and PIT combined.
@@
 At the time of writing this thesis these studies are at least four years old. 
-Some of the tools have been developed further or are still actively developed\cite{pit-releases,Major}.
+Some of the tools were updated or are still under active development\cite{pit-releases,Major}.
 \newline
@@
 IOR & Overriding method   rename                                & PR                    & Primitive Returns               \\ \hline
-ISI & super keyword insertion                                   & ER                    & Experimental Switch             \\ \hline
-ISD & super keyword deletion                                    & BI                    & Big Integer                     \\ \hline
+ISI & Super keyword insertion                                   & ER                    & Experimental Switch             \\ \hline
+ISD & Super keyword deletion                                    & BI                    & Big Integer                     \\ \hline
 IPC & Explicit call to a   parent’s constructor deletion        & NRC                   & Naked Receiver                  \\ \hline
-PNC & new method call with   child class type                   & N                     & Negation                        \\ \hline
+PNC & New method call with   child class type                   & N                     & Negation                        \\ \hline
 PMD & Member variable   declaration with parent class type      & AOR                   & Arithmic Operator Replacement   \\ \hline
@@
 OAC & Arguments of   overloading method call change             & \multicolumn{1}{l|}{} &                                 \\ \hline
-JTI & this keyword insertion                                    & \multicolumn{1}{l|}{} &                                 \\ \hline
-JTD & this keyword deletion                                     & \multicolumn{1}{l|}{} &                                 \\ \hline
-JSI & static modifier   insertion                               & \multicolumn{1}{l|}{} &                                 \\ \hline
-JSD & static modifier   deletion                                & \multicolumn{1}{l|}{} &                                 \\ \hline
+JTI & This keyword insertion                                    & \multicolumn{1}{l|}{} &                                 \\ \hline
+JTD & This keyword deletion                                     & \multicolumn{1}{l|}{} &                                 \\ \hline
+JSI & Static modifier   insertion                               & \multicolumn{1}{l|}{} &                                 \\ \hline
+JSD & Static modifier   deletion                                & \multicolumn{1}{l|}{} &                                 \\ \hline
 JID & Member variable   initialization deletion                 & \multicolumn{1}{l|}{} &                                 \\ \hline
@@
 Table \ref{tab:mutators-overlap} shows an overview over the relations between the mutators of MuJava and PIT.
-This overview does include the mutators that have been reviewed in the studies of Kintis et al., and Marki and Lindstrom.
-However some of the new mutators map to the existing mutators.
+Some of the new mutators from the tools map to the already existing mutators of the other tools.
+To get a complete overview we include the mutators that have been reviewed in the studies of Kintis et al., and Marki and Lindstrom.
 
@@
 This difference can partially be explained by the fact that MuJava does not support source projects with Java version 1.7 or higher\cite{mujava}.
-All the the source code that uses features from this version of Java and higher resulted in an error during mutant generation.
+All the code that uses the of features from this version of Java and above resulted in an error during mutant generation.
 
@@
 \newpage
-\section{Mutation tool}
-We have studied literature on three different tools and extended the research with more recent data.
+\section{Selecting mutation testing tool}
+We have studied literature on three different tools and extended the research with the current state of the art.
 MuJava is not actively maintained and has not been updated in the last few years.
 It does not support JUnit 4 and all versions of TestNG\cite{mujava}.
-These are test engines\todo{add to vocabulary} and are crucial for executing tests.
+These are test engines and are crucial for executing tests.
 MuJava also does not support source projects with java version 1.7 or higher\cite{mujava}.
-Conforming to these requirements would recude the set of projects we could use in or experiments.
+Conforming to these requirements would reduce the set of projects we could use in or experiments.
 \newline
@@
 It supports Maven, Gradle, has a \acrfull{cli} and has a faster execution time than the other tools.
-For example PIT provides a plugin system in which you can inject your own code in various stages of mutation testing\cite{pit}.
+For example PIT provides a plugin system in which you can inject your own code in various stages of mutation testing process\cite{pit}.
 PIT also generated significantly more mutants in every project we have tested.
-Based on the information presented in this chapter we dcecided that PIT would be the best choice for generating and executing mutants.
+Based on the information presented in this chapter we decided that PIT would be the best choice for generating and executing mutants.
 \end{document}

### 2021-07-14 chapters/rq1/1_identifying_characteristics.tex
redo structure
latest changes
fixed our tool to the tool
changes

--- 2021-06-07 chapters/5_identifying_characteristics.tex
@@
-\documentclass[../main]{subfiles}
+\documentclass[../../main]{subfiles}
 \begin{document}
-
 \section{Identifying characteristics}
 \label{ch:clustering_characteristics}
-Our goal is to cluster mutants based on characteristics. 
+The goal of this experiment is to cluster mutants based on characteristics. 
 To achieve this we need to identify characteristics that the mutants contain. 
@@
 Zhang et al., identifies several mutant characteristics\cite{Zhang2019PredictiveTesting}. 
-Their research focuses on detecting if a mutant survives or not before executing the mutant.
+The goal of their research is detecting as accurate as possible whether a mutant survives or not before executing the mutant.
 The mutant characteristics identified by Zhang et al., have been used in other research with promising results\cite{Oonk2021}.
 While the goal of their research is different than ours, we can still use the characteristics they identified.
-We select characteristics relevant to our research from the research of Zhang et al and Oonk.
-We extend the list of characteristics with data we can extract from our selected mutation testing tool. 
+We select characteristics relevant to our research from the research of Zhang et al., and Oonk\cite{Oonk2021}.
+We extend the list of characteristics with data we can extract from the selected mutation testing tool. 
 \newline
@@
 \end{enumerate}
-In the following sections we will elaborate more on these characteristics.
+In the following sections we will elaborate on these characteristics.
 
@@
 There are existing similarities measures that base their similarity in different ways\cite{Gomma2013AApproaches}. 
-We select the Levenshtein distance as similarity measure for our research\cite{Gomma2013AApproaches}.
+\newline
 The Levenshtein distance defines the distance between two strings by counting the minimum number of operations(create/edit/delete) needed to transform one string into the other\cite{Gomma2013AApproaches}.
@@
 The Levenshtein distance measures this difference.
-This could be a useful characteristic because mutants that have the same similarity could be grouped together.
+We select the Levenshtein distance as similarity measure for our research\cite{Gomma2013AApproaches}.
+This could be a useful characteristic because mutants that have the same similarity may be grouped together.
 
@@
 Every test the mutant is challenged by is an occasion for the mutant to be killed or not.
-This could be a useful characteristic because mutants that have more chances to be killed could be grouped together.
+This could be a useful characteristic because mutants that have more chances to be killed may be grouped together.
 
@@
 We can extract this data from PIT and use it as a characteristic.
-It could be a useful characteristic because mutants that are changed in the same predefined way could be grouped together.
+It could be a useful characteristic because mutants that are changed in the same predefined way may be grouped together.
 
@@
 We identify the opcode as a characteristic.
+The opcode is a characteristic that influences the end result of the mutant in byte code representation which in turn can influence other characteristics, for example the Levenshtein distance.
 The opcode could be a useful characteristic because mutants with the same opcode may be grouped together.
@@
 The location can be identified as a characteristic of a mutant.
-PIT uses this location in the mutation testing and reporting processes.
+PIT uses this location during the mutation testing process and for reporting purposes.
 We can extract the location as defined by the properties in PIT.
-Each property is identified as a characters
+Each property is identified as a characteristic.
 The location of a piece of code is a distinctive property. 
 The same piece of code in a different location may have a different effect on the source code or unit tests.
+For example two mutants with the same byte code may have a different result in two different locations of a software project. 
+This is dependent on different factors one of which being the unit tests the mutant is challenged by.
 A mutant may not be unique if the location is not taken into account.
@@
 An addition to the try-catch blocks are finally blocks.
-A finally block is executed after a try and/or catch block\cite{javaDocs}.
+A finally block is executed after a try and catch block\cite{javaDocs}.
 There is a possibility that a mutant may be generated inside these blocks.
@@
 There is a chance a mutant modifies a local variable.
-A variable modified by a mutant increases the chance for a mutant to be killed.
+A variable modified by a mutant may increase the chance for a mutant to be killed.
 This is because a unit test may check for a specific value for a specific local variable.
@@
 This could be a useful characteristic as mutants with the same amount of local variables may be grouped together.
+It could also be possible that there is a relation between the amount of local variables and the result of a mutant(killed or survived).
 

### 2021-07-14 chapters/rq1/2_experiment_design.tex
changes

@@
 \section{Experiment design}
-In this chapter we will design an experiment to answer our first research question.
+In this chapter we design an experiment to answer the first research question.
 We hypothesise that we can cluster mutants with the set of characteristics we identified
 while maintaining effectiveness and reducing the amount executed when executing one mutant from each cluster that is randomly selected.
-Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.
+Our goal is to achieve a weighted mutation score that is as close as possible to the mutation score of a full set of executed mutants.
 \newline
@@
 Next we select and implement a clustering algorithm for which we can cluster mutants.
-We can then cluster the mutants and compare the results of the clusters with the results of the mutants.
+We can then cluster the mutants and compare the mutation scores to each other.
 
@@
 During the various stages we can extract characteristics by using PIT's plugin system.
+The plugins are executed after each other and per class.
+In other words first plugin A will make changes or analyse all classes then plugin B etc.
 \newline
@@
 With the combination of location and mutator we can generate the byte code of the mutant.
-With both pieces of code available we can calculate the Levenshtein distance between them.
+With both the mutated byte code and original byte code available we can calculate the Levenshtein distance.
 \newline
@@
 \newline
-When the mutation testing process has finished we need to merge the characteristic gathered in the second plugin.
+When the mutation testing process has finished we need to merge the characteristic gathered in from the plugins.
 We add the characteristic; the number of tests the mutant is challenged by to the data file containing all the other characteristics.
-We can merge the data because both files contain the same unique identifier per mutant.
+All the files contain the same unique identifier per mutant on which we can merge the data.
 We do not add the information about the survival of a mutant.
-This is data we need to validate our results.
+We only need this data to validate our results.
 
@@
 \newline
-For our characteristic we will use the Levenshtein distance between a Java byte code mutant and its parent.
 By calculating the distance on byte code we filter out the context that is present in Java code.
 This gives us a Levenshtein distance that represents more of the semantic difference, between a mutant and parent, than calculating the distance for Java code.
+For this characteristic we will use the Levenshtein distance between a Java byte code mutant and its parent.
 
@@
 For our experiment we need to calculate the Levenshtein distance on the byte code generated in PIT.
-Since we make use of the plugin system it was natural to use an implementation that was written in Java.
+Since we make use of the plugin system we also choose to use an implementation that is written in Java.
 The Levenshtein distance can be calculated in different ways for example with recursion.
@@
 This last implementation is the implementation we selected for our experiment.
+The implementation described above is also offered by a library.
+This library is the Apache Commons Text library\cite{commonsText}, in our experiment we make use of this implementation for calculating the Levenshtein distance.
 
@@
 It also states that varying the parameters of hierarchical clustering improves the performance compared to that of the default settings of the algorithm\cite{Rodriguez2019}.
-We selected hierarchical clustering to cluster the mutants.
 We have more than ten characteristics which we can use as features and adjust the parameters of the algorithm based on the characteristics in our data set.
+Bases on these observations we select hierarchical clustering to cluster the mutants.
 Next we will explain the configuration we use for clustering the mutants. 
@@
 The linkage criteria specifies the dissimilarity in the sets as a function of the pair-wise distances of observations in those sets\cite{Rajalingam2011}.
+We reviewed the different linkage criteria and concluded the following.
 \newline
@@
 Since we cannot assume that our clusters are well separated we chose to use the ward linkage method for our clustering algorithm.
-The agglomerative hierarchical clustering algorithm with ward linkage will cluster our mutants represented by the characteristics we gathered.
+In summary, the agglomerative hierarchical clustering algorithm with ward linkage will cluster our mutants represented by the characteristics we gathered.
 
-\subsection{weighted mutation score}
+\subsection{Weighted mutation score}
 \label{ch:weighted_score}
@@
 \newline
-For example, take a full set with a score of 75/100 killed mutants. \todo{example too much?}
+For example, take a full set with a score of 75/100 killed mutants.
 This gives us a mutation score of 75\%. 
@@
 \subsection{Categorical data}
-The hierarchical clustering algorithm needs all characteristics in a numerical form\cite{Vijaya2019ComparativeClustering}.
-The characteristics mutator identifier, class name, method name and return type are non numerical.
-To deal with this problem we need to apply categorical variable encoding to these specific features.
 There are different categorical variable encoding techniques available\cite{Potdar2017AClassifiers}.
@@
 I.e. a return type void is not better or worse than a return type string.
-The same goes for the location characteristics, there is no location that should weigh more during the clustering than the other locations.
+The same goes for the location characteristics, there is no location that should have a bigger weight than the other locations.
 The individual characteristics do contain a finite set of values.
 For example multiple mutants may contain the same class name.
-Taking into account the properties of our categorical characteristics we chose to use a nominal variable encoding.
+The hierarchical clustering algorithm needs all characteristics in a numerical form\cite{Vijaya2019ComparativeClustering}.
+The characteristics mutator identifier, class name, method name and return type are non numerical.
+To deal with this problem we apply categorical variable encoding to these specific features.
+Taking into account the properties of our categorical characteristics the nominal variable encoding fits our requirements.
 Nominal encoding comprises a finite set of discrete values with no relationship between values\cite{Potdar2017AClassifiers}.
+Therefore we implement this type of encoding in the experiment.
 
@@
 We decide on the number of clusters based on the amount of mutants generated in a full set.
-As a baseline we will cut off clustering in steps of 25\%.
-The steps of 25\% will give a good global indication in reduction.
-If necessary we can decide to cut of the clustering at different points.
+The reasoning for the selecting the number of clusters will be from a performance reducing perspective.
+Mutation clustering increases the performance by reducing the amount of mutants executed\cite{Pizzoleto2019}.
+Reducing the performance by a quarter, half or three quarters will give an indication of performance increases.
+To translate this to cluster size we will reduce the number of clusters in steps op 25\%.
+In other words by reducing the amount of mutants, for example by half, we increase the performance by 50\%\cite{Just2014}.
+If necessary we can decide to cut off the clustering at different points.
 The baseline will consist of three experiments where the cluster size is 25\%, 50\% and 75\% of the total amount of mutants respectively.
@@
 \label{ch:exp1_validation}
-The most efficient way to measure test effectiveness with mutation testing is by running all mutants that a tool can possibly generate.
+The most efficient way to measure test effectiveness with mutation testing is by running all mutants that a mutation testing tool can possibly generate.
 The goal of this experiment is to reduce the amount of mutants executed while maintaining effectiveness.
@@
 Depending on the effectiveness of our clustering algorithm we may loose accuracy.
-This can happen if a cluster contains mutants both types. 
-We can measure this accuracy by calculating a percentage of all mutants that have survived against the ones that have been killed in a cluster.
+This can happen if a cluster contains mutants of both results. 
+We can measure the accuracy inside a cluster by calculating a percentage of all mutants that have survived against the ones that have been killed in a cluster.
 If the majority of the mutants in a cluster is killed then we consider that cluster to represent a killed cluster and the other way around for survived mutants.
@@
 If the weighted mutation score of our clustered set deviates more than 5\% from the score of a full set we will reject our hypothesis.
-We also compare the inaccuracy to the significance level to decide if our inaccuracy is statistical significant.
 
@@
 As stated in our hypothesis we will randomly select a mutant from each cluster to be executed.
-To make our results reproducible we will select a random mutant based on a generated seed.
+To make our results reproducible we will select the random mutant based on a generated seed.
 The seed generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.
@@
 Achieving consistent results while applying random selection will contribute to the validity of the experiment.
-We will use the same statistical significance threshold to decide if the variation by the random sampling is acceptable or not.
 

### 2021-07-14 chapters/rq2/1_experiment_design.tex
changes

@@
 \begin{document}
-% Because we use sections but these sections should be interpreted as chapters for rq1 and rq2 sectioning.
-\clearpage
 \section{Experiment design}
-In this chapter we will design an experiment to answer our second research question.
+In this chapter we will design an experiment to answer the second research question.
 We hypothesise that we can train a machine learning model with the set of characteristics we identified
@@
 Subtractive clustering is designed for learning fuzzy systems models from data\cite{Chiu1994FuzzyEstimation}. 
-To find out the number of clusters we need for our model we can make use of these models.
+If it is required to find out the number of clusters we need for our model we can make use of these models.
 
@@
 While the ART models are very good at what they do, they do not fit our purpose of clustering mutants.
-As described above the ART models expect rapidly changing sequences of input patterns, our data set does not constists of this kind of data.
+As described above the ART models expect rapidly changing sequences of input patterns.
+Our data does not contain binary input patterns and it is also not rapidly changing.
 Therefore we will not select an ART model.
@@
 To find its neighborhood rank, each neuron compares its distance to the input vector with those of all the other neurons to the input vector\cite{Du2010Clustering:Approach}. 
-Unlike the self organizing map, which uses predefined static neighborhood relations, the neural gas network determines a dynamical neighborhood relation as learning proceeds\cite{Du2010Clustering:Approach}.
-A big pro of the neural gas network is that it is able to spawn neurons.
-This means that we do not need to know the number of clusters in advance.
+Unlike the \acrfull{som}(see Chapter \ref{ch:topThreeModels}, which uses predefined static neighborhood relations, the neural gas network determines a dynamical neighborhood relation as learning proceeds\cite{Du2010Clustering:Approach}.
+A benefit of the neural gas network is that it is able to spawn neurons.
+This means that it is not necessary to know the number of clusters in advance.
 Unfortunately the neural gas network is a learning network.
@@
 \subsubsection{Self organizing map, C-means and fuzzy C-means}
-The self organizing map, C-means and fuzzy C-means models are all models that would fit our purpose. 
+\label{ch:topThreeModels}
+The \acrfull{som}, C-means and fuzzy C-means models are all models that would fit our purpose. 
 To choose between these models we will look at the performance of each model.
@@
 The C-means method was very affected by the presence of a large amount of outliers.
-They conclude that the self organizing map did not perform well in many cases being very affected by the amount of variables and clusters even for the non overlapping cases\cite{Mingoti2006ComparingAlgorithms}.
+They conclude that the \acrshort{som} did not perform well in many cases being very affected by the amount of variables and clusters even for the non overlapping cases\cite{Mingoti2006ComparingAlgorithms}.
 They also conclude that their results partially overlap with other research in the same domain.

### 2021-07-14 chapters/1_abstract.tex
acr fuzzy c-means

@@
 We started our experiment by identifying characteristics to represent the mutants.
-For the first experiment we used hierarchical clustering and in the second experiment we trained the fuzzy c-means model.
+For the first experiment we used hierarchical clustering and in the second experiment we trained the \acrlong{fcm} model.
 We calculated a weighted mutation score and compared this with the mutation score of a full set of mutants executed.
@@
 The machine learning model didn't perform as well as the hierarchical clustering.
-However, more work is needed in finding the right parameters for the fuzzy c-means algorithm. 
+However, more work is needed in finding the right parameters for the \acrlong{fcm} algorithm. 
 \end{document}

### 2021-07-14 chapters/rq2/1_experiment_design.tex
acr fuzzy c-means

@@
 
-\subsubsection{Self organizing map, C-means and fuzzy C-means}
+\subsubsection{Self organizing map, C-means and \acrlong{fcm}}
 \label{ch:topThreeModels}
-The \acrfull{som}, C-means and fuzzy C-means models are all models that would fit our purpose. 
+The \acrfull{som}, C-means and \acrfull{fcm} models are all models that would fit our purpose. 
 To choose between these models we will look at the performance of each model.
@@
 Other variables such as cluster boundaries and overlap were also controlled in the experiment.
-They observe that the C-means and fuzzy C-means had good performance for non overlapping situations\cite{Mingoti2006ComparingAlgorithms}.
-The best results for average recovery and internal dispersion rates were found for Fuzzy C-means which was very stable in all situations achieving recovery averages over 90\%\cite{Mingoti2006ComparingAlgorithms}
+They observe that the C-means and \acrlong{fcm} had good performance for non overlapping situations\cite{Mingoti2006ComparingAlgorithms}.
+The best results for average recovery and internal dispersion rates were found for \acrlong{fcm} which was very stable in all situations achieving recovery averages over 90\%\cite{Mingoti2006ComparingAlgorithms}
 The C-means method was very affected by the presence of a large amount of outliers.
@@
 \newline
-Based on these results we can conclude that the fuzzy c-means model has the best performance overall.
-For this experiment we select the fuzzy c-means model.
+Based on these results we can conclude that the \acrshort{fcm} model has the best performance overall.
+For this experiment we select the \acrshort{fcm} model.
 
@@
 \subsection{Implementation of algorithm}
-The research paper of the fuzzy c-means algorithm is referenced in the survey paper of Du et al\cite{Du2010Clustering:Approach}.
-The fuzzy c-means algorithm is developed by Bezdek et al.\cite{Bezdek1984FCM:Algorithm}.
+The research paper of the \acrshort{fcm} algorithm is referenced in the survey paper of Du et al\cite{Du2010Clustering:Approach}.
+The \acrshort{fcm} algorithm is developed by Bezdek et al.\cite{Bezdek1984FCM:Algorithm}.
 Based on their research an implementation can be developed.
-There are existing libraries that contain the implementation of the fuzzy c-means algorithm.
-Developing our own version of the fuzzy c-means algorithm is out of scope for this research.
-For this reason we choose to use a library that implements the fuzzy c-means algorithm.
-The library we will use for this experiment is the fuzzy c-means implementation build by Dias et al.\cite{dias2019fuzzy}.
+There are existing libraries that contain the implementation of the \acrshort{fcm} algorithm.
+Developing our own version of the \acrshort{fcm} algorithm is out of scope for this research.
+For this reason we choose to use a library that implements the \acrshort{fcm} algorithm.
+The library we will use for this experiment is the \acrlong{fcm} implementation build by Dias et al.\cite{dias2019fuzzy}.
 The theory developed by Bezdek et al., is implemented in this library.
@@
 \subsection{Algorithm parameters}
-The fuzzy c-means algorithm has to initialize the centroids of the clusters.
+The \acrshort{fcm} algorithm has to initialize the centroids of the clusters.
 Our selected library does this based on a random number.
@@
 \newline
-Bezdek et al., also researched cluster validity for the fuzzy c-means model\cite{Bezdek1995OnModel}.
+Bezdek et al., also researched cluster validity for the \acrshort{fcm} model\cite{Bezdek1995OnModel}.
 They researched what effect which parameters have on the validity of the clusters.
@@
 Wu proposed a new guideline for selecting parameter \textit{m}\cite{Wu2012}.
-His point of view was that a large \textit{m} value will make the fuzzy c-means more robust to noise and outliers.
-He suggests implementing fuzzy c-means with \textit{m}=[1.5, 4].
+His point of view was that a large \textit{m} value will make the \acrlong{fcm} more robust to noise and outliers.
+He suggests implementing \acrlong{fcm} with \textit{m}=[1.5, 4].
 When the data set contains noise and outliers, the fuzzifier\textit{m}=4 is recommended in a theoretical upper bound case.
@@
 We will select and implement the subtractive clustering model to estimate the number of clusters.
-We can use this number for the number of clusters parameter in the fuzzy c-means algorithm.
+We can use this number for the number of clusters parameter in the \acrshort{fcm} algorithm.
 

### 2021-07-14 chapters/rq2/1_experiment_design.tex
Updates from Overleaf

--- 2021-07-14 chapters/rq2/1_experiment_design.tex
@@
 To find its neighborhood rank, each neuron compares its distance to the input vector with those of all the other neurons to the input vector\cite{Du2010Clustering:Approach}. 
-Unlike the \acrfull{som}(see Chapter \ref{ch:topThreeModels}, which uses predefined static neighborhood relations, the neural gas network determines a dynamical neighborhood relation as learning proceeds\cite{Du2010Clustering:Approach}.
+Unlike the \acrlong{som}(see Chapter \ref{ch:topThreeModels}, which uses predefined static neighborhood relations, the neural gas network determines a dynamical neighborhood relation as learning proceeds\cite{Du2010Clustering:Approach}.
 A benefit of the neural gas network is that it is able to spawn neurons.

### 2021-07-14 chapters/rq2/1_experiment_design.tex
minor typo
Updates from Overleaf

@@
 They did a Monte Carlo simulation where the cluster sizes and amount of random numbers in the data set varied each simulation\cite{Mingoti2006ComparingAlgorithms}.
-Other variables such as cluster boundaries and overlap were also controlled in the experiment.
+Other variables such as cluster boundaries and overlap were also controlled variables in the experiment.
 They observe that the C-means and \acrlong{fcm} had good performance for non overlapping situations\cite{Mingoti2006ComparingAlgorithms}.
 The best results for average recovery and internal dispersion rates were found for \acrlong{fcm} which was very stable in all situations achieving recovery averages over 90\%\cite{Mingoti2006ComparingAlgorithms}
-The C-means method was very affected by the presence of a large amount of outliers.
-They conclude that the \acrshort{som} did not perform well in many cases being very affected by the amount of variables and clusters even for the non overlapping cases\cite{Mingoti2006ComparingAlgorithms}.
-They also conclude that their results partially overlap with other research in the same domain.
+The C-means method was affected by the presence of a large amount of outliers.
+They conclude that the \acrshort{som} did not perform well in many cases being affected by the amount of variables and clusters even for the non overlapping cases\cite{Mingoti2006ComparingAlgorithms}.
+They also observe  that their results partially overlap with other research in the same domain.
 \newline
-Based on these results we can conclude that the \acrshort{fcm} model has the best performance overall.
-For this experiment we select the \acrshort{fcm} model.
+Based the conclusions made by Mingoti et al., we can conclude that the \acrshort{fcm} model has the best performance overall.
+Thus for this experiment we select the \acrshort{fcm} model.
 
@@
 The characteristics of the mutants are generated per project and have a project specific context.
-The relations between the different mutants are also specific per project.
-For example mutant with a mutator in a certain project can have a different effect on the source code and can give other results than a mutator in a different project.
+The relations between the mutants are scoped to their own project. 
+In other words if a specific mutant has a relation to another mutant in the same project, this relation may not exist between the same mutant and a mutant from a different project.
+For example, a mutant with a mutator in a certain project can have a different effect on the source code and can give other results than a mutator in a different project.
 To keep the relations and context we decided to train a separate model per project.
@@
 To make our results reproducible we will select a random division based on a generated seed.
-The seed generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.
+The seeds generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.
 
 \subsection{Implementation of algorithm}
-The research paper of the \acrshort{fcm} algorithm is referenced in the survey paper of Du et al\cite{Du2010Clustering:Approach}.
+The research paper of the \acrshort{fcm} algorithm is referenced in the survey of Du et al\cite{Du2010Clustering:Approach}.
 The \acrshort{fcm} algorithm is developed by Bezdek et al.\cite{Bezdek1984FCM:Algorithm}.
@@
 There are existing libraries that contain the implementation of the \acrshort{fcm} algorithm.
-Developing our own version of the \acrshort{fcm} algorithm is out of scope for this research.
-For this reason we choose to use a library that implements the \acrshort{fcm} algorithm.
+Developing our own version of the \acrshort{fcm} algorithm would be valuable if we want to make adjustments to the algorithm itself.
+For this experiment we do not require to make adjustments to the \acrshort{fcm} algorithm.
+Therefore we choose to use a library that implements the \acrshort{fcm} algorithm.
 The library we will use for this experiment is the \acrlong{fcm} implementation build by Dias et al.\cite{dias2019fuzzy}.
 The theory developed by Bezdek et al., is implemented in this library.
-The library has been used in other research\cite{DeAlmeidaNeto2020, Kopf2019, Nwadiugwu2020} and is actively maintained.
+The library is actively maintained and has been used before in other research\cite{DeAlmeidaNeto2020, Kopf2019, Nwadiugwu2020}.
 
@@
 The \acrshort{fcm} algorithm has to initialize the centroids of the clusters.
-Our selected library does this based on a random number.
+The library we selected does this based on a random number.
 It also allows us to control this random number.
-To make the initialization of the centroid reproducible and deterministic we will use a generated seed.
-The seed generated is included in the Github repository\cite{rbasarat-repo}.
+To make the initialization of the centroid reproducible we will use the same generated seeds as in our first experiment.
+The seeds generated are included in the Github repository\cite{rbasarat-repo}.
 \newline
@@
 He suggests implementing \acrlong{fcm} with \textit{m}=[1.5, 4].
-When the data set contains noise and outliers, the fuzzifier\textit{m}=4 is recommended in a theoretical upper bound case.
+When the data set contains noise and outliers, the fuzzifier \textit{m}=4 is recommended in a theoretical upper bound case.
 Since our data set does not contain noise and outliers we choose to select the midpoint concluded in the research of Bezdek et al.\cite{Bezdek1995OnModel}.

### 2021-07-15 chapters/1_abstract.tex
chagnes

@@
 Maintaining and monitoring these test suites is important because they influence the cost of maintenance.
-Monitoring the quality of test code has shown that tests with lower quality leads to more defect-prone production code.
+Monitoring the quality of test code has shown that tests with lower quality lead to more defect-prone production code.
 \newline
@@
 \newline
-We conducted two experiment in which we cluster the mutants with two different techniques.
+We conducted two experiments in which we cluster the mutants with two different techniques.
 Before conducting the experiment we conducted a state of the art comparison between three mutation testing tools for Java.
@@
 The clusters generated by the machine learning model were less accurate and showed a significant decrease in effectiveness.
-The machine learning model didn't perform as well as the hierarchical clustering.
+The machine learning model did not perform as well as the hierarchical clustering.
 However, more work is needed in finding the right parameters for the \acrlong{fcm} algorithm. 

### 2021-07-15 chapters/2_introduction.tex
chagnes

@@
 \newline
+\todo[inline]{is this your approach? the current general approach?}
 Mutation testing tools create small faulty versions of the program.
@@
 If a large amount of mutants survives, it might be an indication that the quality of the test suite is poor as programming errors remain undetected.
-Mutation testing has considerable drawbacks, such as the equivalent mutant problem and being resource expensive.
+Mutation testing has considerable drawbacks\todo{add ref}, such as the equivalent mutant problem and being resource expensive.
 
@@
 To reduce the cost of mutation testing, we try to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants. 
-Our goal is to cluster any mutant that is generated.
-We will do this with two different approaches.
-Our first approach will be a qualitative approach and second a quantitative approach. 
-We structure our research on the following research questions:
+Our goal is to cluster any mutant that is generated \todo{make the goal clearer: no requirements on which mutants get clustered: aim at generic mutant clustering}.
+We will do this with two different approaches: qualitative and quantitative. 
+We structure our research as follows:
 \newline
@@
 \begin{enumerate}
- \item A methodology to cluster generated mutants based on chosen characteristics to reduce the cost of mutation testing.
+ \item A methodology to cluster generated mutants based on chosen characteristics to reduce the cost of mutation testing.\todo{both white box and black box?}
  \item A \acrfull{poc} which implements the qualitative methodology chosen and elaborated within the thesis.

### 2021-07-15 chapters/rq2/1_experiment_design.tex
chagnes

@@
 \subsection{Implementation of algorithm}
+\label{ch:fcm_implementation}
 The research paper of the \acrshort{fcm} algorithm is referenced in the survey of Du et al\cite{Du2010Clustering:Approach}.
@@
 Bezdek et al., also researched cluster validity for the \acrshort{fcm} model\cite{Bezdek1995OnModel}.
-They researched what effect which parameters have on the validity of the clusters.
+They researched what effect which parameters could have on the validity of the clusters.
 They specifically analyzed the role of weighting exponent \textit{m}(fuzziness parameter).
-They concluded that the best choice for \textit{m} is probably in the interval [1.5, 2.5], whose mean and midpoint is \textit{m}=2.
+They concluded that the best choice for \textit{m} is in the interval [1.5, 2.5], whose mean and midpoint is \textit{m}=2.
 Wu proposed a new guideline for selecting parameter \textit{m}\cite{Wu2012}.
@@
 When the data set contains noise and outliers, the fuzzifier \textit{m}=4 is recommended in a theoretical upper bound case.
-Since our data set does not contain noise and outliers we choose to select the midpoint concluded in the research of Bezdek et al.\cite{Bezdek1995OnModel}.
+Since our data set does not contain noise we choose to select the midpoint concluded in the research of Bezdek et al.\cite{Bezdek1995OnModel}.
 The value of \textit{m} in this experiment is 2.
@@
 Finding the correct amount of clusters has been an ongoing problem for clustering algorithms\cite{Du2010Clustering:Approach}.
-Fortunately there are some techniques available to estimate the number of clusters.
+Fortunately there are techniques available to estimate the number of clusters.
 During model selection we discussed the mountain and subtractive models. 
@@
 We will select and implement the subtractive clustering model to estimate the number of clusters.
-We can use this number for the number of clusters parameter in the \acrshort{fcm} algorithm.
+We will use this number as the value for the number of clusters parameter in the \acrshort{fcm} algorithm.
 
 \subsection{Implementation of subtractive clustering}
+For the same reasons as in Chapter \ref{ch:fcm_implementation} we will select a library that implements the subtractive clsutering algorithm.
 The research paper of the mountain and subtractive algorithms are referenced in the survey paper of Du et al\cite{Du2010Clustering:Approach}.
@@
 To validate our machine learning model we will calculate the weighted mutation score and cluster accuracy based on the clusters calculated by the model.
-This is the same validation method as described in Chapter \ref{ch:exp1_validation}.
+We use the same validation method as described in Chapter \ref{ch:exp1_validation}.
 We will use the mutants of projects that are not selected for training as input for validating the model.
-As described in Chapter \ref{ch:training_strategy} the projects should represent approximately 20\% of the total amount of mutants in our data set.
+As described in Chapter \ref{ch:training_strategy} the mutants should represent approximately 20\% of the total amount of mutants per project.
 It may be possible the model has to cluster mutants it has never seen before.

### 2021-07-15 chapters/9_conclusion.tex
acr fuzzy c-means
styling issues

@@
 While the minimum clustering accuracy was higher than that of the hierarchical clustering, the maximum cluster accuracy was lower than that of the hierarchical clustering.
-The cluster accuracy had a greater impact on the clusters of the machine learning model because there was a relatively small number of clusters.
+The cluster accuracy had a larger impact on the clusters of the machine learning model because there was a relatively small number of clusters.
 Our research proves that it is possible to determine the means of domains as the centre of a cluster within a programming language without narrowing down the scope to specific pieces of code.
@@
 These proved to be non optimal for our purpose.
-This research can be extended by finding the optimal parameters for the fuzzy c-means clustering algorithm.
+This research can be extended by finding the optimal parameters for the \acrshort{som} clustering algorithm.
 The results showed a higher minimum accuracy which can indicate that there is potential for clustering the mutants with the machine learning model.

### 2021-07-15 chapters/rq2/1_experiment_design.tex
styling issues

@@
 The stability and plasticity properties as well as the ability to efficiently process dynamic data make the ART attractive for clustering large, rapidly changing sequences of input patterns\cite{Du2010Clustering:Approach}.
-While the ART models are very good at what they do, they do not fit our purpose of clustering mutants.
+While the ART models are good at what they do, they do not fit our purpose of clustering mutants.
 As described above the ART models expect rapidly changing sequences of input patterns.
@@
 They observe that the C-means and \acrlong{fcm} had good performance for non overlapping situations\cite{Mingoti2006ComparingAlgorithms}.
-The best results for average recovery and internal dispersion rates were found for \acrlong{fcm} which was very stable in all situations achieving recovery averages over 90\%\cite{Mingoti2006ComparingAlgorithms}
+The best results for average recovery and internal dispersion rates were found for \acrlong{fcm} which was stable in all situations achieving recovery averages over 90\%\cite{Mingoti2006ComparingAlgorithms}
 The C-means method was affected by the presence of a large amount of outliers.

### 2021-07-15 chapters/3_background.tex
changes
chagnes
minor style change

@@
 \label{ch:background}
-This chapter will present the necessary background information for this thesis. First, we define some basic terminology that will be used throughout this thesis.
+This chapter presents the necessary background information for this thesis. First, we define some basic terminology that will be used throughout this thesis.
 Next, we will compare existing mutation testing tools. We start of by reviewing existing literature and then extend the comparison. 
@@
 Mutator.
-test engine??
+test engine.
+weighted mutation score.
 \section{Project selection}
@@
 The unfiltered sample contains 50+ projects.
-From there we filter out all the projects that contain failing tests and that aren't libraries or applications.
+From there we filter out all the projects that contain failing tests and that are not libraries or applications.
 The result consists of a sample with fifteen projects in total.

### 2021-07-15 chapters/7_discussion.tex
finish manual reading notes

@@
 \label{ch:discussion}
-In this chapter, we discuss the results of our experiments on clustering the mutants.
-First we discuss the results of the first experiment. 
-We will discuss the all the variant results together and then compare the results between the variants.
+In this chapter, we discuss the results of our experiments on clustering mutants.
+First, we discuss the results of the first experiment. 
+We will discuss the all results together and then compare the results between the experiments including and excluding the Levenshtein distance characteristic.
 Second we will discuss the results of the second experiment and compare them to those of the first.
@@
 The hypothesis holds for all the projects and variants of the experiment.
-66\% of the displayed results had a deviation of less than 1\%. 
+66\% of the displayed results have a deviation of less than 1\%. 
 This includes both variations of the experiment.
 \newline
-The results prove that the relations between the characteristics we identified can be used to cluster mutants efficiently. 
-This can be explained by the correlation between the result of a mutant and the euclidean distance used to decide whether a mutants should be clustered together.
+The results prove that the relations between the mutants represented by the characteristics we identified can be used to cluster mutants efficiently. 
+This can be explained by the correlation between the result of a mutant, the calculated euclidean distance and the distance threshold(linkage criteria) used to decide whether a mutants should be clustered together.
+\newline
 If two mutants are grouped it is \textit{n} percent likely that the mutants have the same result where \textit{n} is the cluster accuracy.
 \begin{finding}
-    With a reduction up to 75\% of the total amount of mutants we can still maintain effectiveness.
+    When executing 25\% of the total amount of mutants we can still maintain effectiveness.
 \end{finding}
-In both variants we observe a reduction in average accuracy when reducing the number of clusters.
+In both variations of the experiment we observe a reduction in average accuracy when increasing the performance by reducing the number of clusters.
 These reductions are 0.1\% and 0.04\% for clustering with and without Levenshtein distance respectively.
@@
 \begin{finding}
-    The more you reduce the amount of mutants executed the more accuracy is lost.
+    The more you reduce the number of mutants executed the more accuracy is lost.
 \end{finding}
@@
 \begin{finding}
-    We can reduce the amount of mutants executed in our data up to half of the original amount before reducing the accuracy even more.
+    We can reduce the number of mutants executed in our data set up to 50\% before losing more accuracy.
 \end{finding}
@@
 \end{itemize}
-With the exception of Google Auto Service and Commons Text, with a reduction of 25\% and all characteristics, all projects showed an average accuracy above 90\%.
+With the exception of Google Auto Service and Commons Text, a reduction of 25\% and all characteristics, all projects showed an average accuracy above 90\%.
 The combination of hierarchical clustering and the identified characteristics results in high accuracy clusters.
-As logic dictates we expect that a higher accuracy means a score closer to that of the original.
+\newline
+According the rules of logic, a higher cluster accuracy means a score closer to that of the original.
+In other words if all clusters had an accuracy of 100\% we will achieve the same mutation score as that of the full set.
 However we do not observe this behaviour in our results.
@@
 \end{finding}
-As stated in Chapter \ref{ch:results_rq1} the calculation of the Levenshtein distance took a long time.
-For this reason we repeated the experiment without the Levenshtein distance.
-Comparing the results between these variants we can observe a max deviation of 0.05\% between average difference of the mutation score of.
+As stated in Chapter \ref{ch:results_rq1} the calculation of the Levenshtein distance characteristic took a long time.
+For this reason we repeated the experiment without the Levenshtein distance characteristic.
+Comparing the results between these variants we can observe a max deviation of 0.05\% between average difference of the mutation scores.
 In 66\% of the results displayed in Chapter \ref{ch:results_rq1} clustering without the Levenshtein distance is more accurate.
 The biggest difference in mutation score measured between all the projects and their reductions is 3.63\%.
-Clustering without Levenshtein distance is on average more accurate.
+Clustering without Levenshtein distance characteristic is on average more accurate.
 \begin{finding}
-    While clustering without the Levenshtein distance was more accurate in our results, the Levenshtein distance characteristic does not significantly increase or decrease the cluster accuracy.
+    While clustering without the Levenshtein distance characteristic was more accurate in our results, the Levenshtein distance characteristic does not significantly increase or decrease the cluster accuracy.
 \end{finding}
@@
 \section{RQ 2: Training a machine learning model and clustering mutant characteristics}
-The maximum deviation from the original mutation score, for all variants(with and without Levenshtein distance) of the experiment, is above the statistical threshold.
+The maximum deviation from the original mutation score, for all variants(with and without Levenshtein distance characteristic) of the experiment, is above the statistical threshold.
 The biggest deviation measured is 22.39\%.
@@
 \begin{finding}
-Our hypothesis does not hold for the clustering the mutants with the machine learning model trained in our experiment.
+Our hypothesis does not hold for the clustering the mutants with the \acrshort{fcm} model trained in our experiment.
 \end{finding}
 We observe that the number of clusters does not have an obvious relation to the total amount of mutants. 
-This is expected as the subtractive clustering does not look at the total amount of data points(mutants) but to the contents of the data points.
+This is expected as the subtractive clustering does not look at the total amount of data points(mutants) but to the contents(characteristics) of the data points.
 The number of clusters compared to the that of the hierarchical clustering is significantly lower.
@@
 \begin{finding}
-The overall accuracy of the clusters generated by the machine learning model in both variants is lower compared to that of the hierarchical clustering.
+The overall accuracy of the clusters generated by the \acrshort{fcm} model in both variants is lower compared to that of the hierarchical clustering.
 \end{finding}
-When we compare the accuracy per project between the machine learning model and hierarchical clustering we can observe that the lower accuracy of the machine learning model clusters have a higher impact on the deviation in mutation score.
+When we compare the accuracy per project between the \acrshort{fcm} model and hierarchical clustering we can observe that the lower accuracy of the \acrshort{fcm} model clusters have a higher impact on the deviation in mutation score.
 This is caused by the relatively low number of clusters.
@@
 \section{Application}
-The application of our research is limited to all software projects written in the language Java and the limitations of the selected mutation testing tool PIT\cite{pit}.
+The application of our research is limited to all software projects written in the language Java and the requirements of PIT\cite{pit}.
 We can repeat our experiment for every java project that can be run by PIT.
-This scope is significantly larger than most of the research done in the domain of mutation clustering(Chapter \ref{ch:related_work}).
+The scope of the research done in the domain of mutation clustering(Chapter \ref{ch:related_work}) is a subset of the scope in our research.
+With our research it is possible to cluster all the mutants used in other related research.
 The extraction of mutants is build within the PIT plugin system which is a plug and play system.
@@
 The first step is to run PIT with the plugin to gather the characteristics.
-This plugin makes sure PIT exits without executing mutants after gathering the characteristics.
-The second step is to run the script for either the machine learning model or the hierarchical clustering.
+Gathering the characteristics does not take more than a few minutes depending on the number of mutants generated.
+The plugin also makes sure PIT exits without executing mutants after gathering the characteristics.
+The second step is to run the script for either the \acrshort{fcm} model or the hierarchical clustering.
 The last step is to run PIT again with the plugin to filter the mutants based on the results of the clustering.
 It is possible to build all these steps into a DevOps pipeline to automate the process. 
-The plugins and scripts needed for these steps are available on Github\cite{rbasarat-repo}.
+The plugins and scripts needed for these steps are available in the code repository on Github\cite{rbasarat-repo}.
 
@@
 \subsubsection{Equivalent mutants}
-Equivalent mutants are mutants that are syntactically different but functionally the same.
+Equivalent mutants are mutants that are syntactically different but functionally identical.
 These mutants influence the mutation score while giving the same result as their equivalent counterpart.
@@
 Chapter \ref{ch:project_selection} provides an indication but not a wide representation of industrial software projects.
-The conclusions drawn regarding the increase performance by reducing the mutants are based on the comparison between our own baseline measurements.
+The conclusions drawn regarding the increase in performance by reducing the mutants are based on the comparison between our own baseline measurements.
+\newline
 Another threat which is the result of the representation of our test set is the representation of the mutants.
 While our data set contains around 1 million mutants we do not know if we have seen all mutants that exist.
-This will always be a problem as it would be nearly impossible to cluster and mutation test every piece of Java code, that conforms to the requirements of PIT, ever written.
+This will always be a problem as it would be nearly impossible to cluster and mutation test every piece of Java code, that conforms to the requirements of our research, ever written.
 
@@
 Some tests may not be recognized by PIT due to some external data required for the test.
-Another reason may be that the test uses global variables that are not available to PIT.
+Another reason may be that the tests uses global variables that are not available to PIT.
 We encountered some of these tests in our data set.
-While these tests did pass when run with only the unit test framework they would not when run by PIT.
+While these tests did pass when run with only the unit test framework they would fail when run by PIT.
 These projects still met the requirements stated in Chapter \ref{ch:project_selection}.

### 2021-07-15 chapters/8_related_work.tex
finish manual reading notes

@@
 \subsection{Clustering by scope}
-Yu et al., extended the research about overlapping mutants\cite{Yu2019PossibilityScope}.
-They added more mutators to the scope.
+Yu et al., extended the research about overlapping mutants(See chapter \ref{ch:overlapping_mutants})\cite{Yu2019PossibilityScope}.
+They extended the scope by adding more mutators.
 The also extended the research to include statements and blocks.
 Their results shows an increase in clustered mutants.
-The problems mentioned in Chapter \ref{ch:overlapping_mutants} are not addressed in this research.
+The research of Yu et al., suffers from the same limitations and threats  mentioned in Chapter \ref{ch:overlapping_mutants}.
 
@@
 Ji et al., did a qualitative domain analysis on mutants\cite{Ji2009}.
-As a result they identified the Hamming distance for to cluster with. 
+As a result they identified the Hamming distance to cluster with. 
 The Hamming distance is, like the Levenshtein distance, a similarity measure.
-They use the k-means to cluster the mutants with the Hamming distance.
+They use the k-means algorithm to cluster mutants represented by Hamming distance.
 They write that the reduced test set in their experiment is still as strong as the original test set\cite{Ji2009}.
+\newline
 Ji et al., were successful in clustering mutants with Hamming distance and the k-means algorithm.
-But they do remark that their research still has problems with rationally determining the domains of their clusters.
-Our research does not have this problem.
-In our reserach the centre of of the domains are decided by the characteristics of the mutants.
+They do acknowledge that their research still has problems with rationally determining the domains of their clusters.
+Our research does not have this problem, as the centre of the domains are decided by the characteristics of the mutants.
 
@@
 They defined multiple definitions for the mutants; similarity distance, distance between the mutants and a killing matrix of the mutants.
-With these definitions they reduce the mutants based on their proposed reduction method.
-This method is based on spectral clustering (SCMT), including the determination method of the number of clusters, spectral clustering of mutants, and selection of representative mutants.
-The reduced were then clustered with a classical clustering algorithm.
-Their results are promising and show good cluster accuracy.
+With these definitions they reduced the mutants based on their proposed reduction method.
+This method is based on spectral clustering (SCMT), the determination method of the number of clusters, spectral clustering of mutants, and selection of representative mutants.
+The reduced mutants were then clustered with a classical clustering algorithm.
+Their results are promising and show high cluster accuracy.
 They write that there is still work left to do in optimizing the matrices and clustering algorithms.
 Our goal overlaps with that of Wei, the difference is that we used a different methodology to achieve that goal.
+Another difference is that our research does not require to know the result of a mutant(killed or survived).
 
@@
 Hussain et al., used the k-means and agglomerative clustering algorithm to cluster the mutants according to a similarity measure. 
-They used the Hamming distance a similarity measure.
-They calculated the distance and used this as data te feed into the clustering algorithms.
+They used the Hamming distance as similarity measure.
+They calculated the distance and used this as data to feed into the clustering algorithms.
 The number of clusters and the initial position of the cluster center in the k-means algorithm are difficult to determine, and the process of the agglomerative clustering algorithm makes it difficult to correct the existing cluster formation\cite{Hussain2008}.
@@
 The difference is that our research used different and more elaborated methods to cluster the mutants.
+Hussain et al., calculated the mutation score by counting one cluster as one mutant.
+Our research calculated a weighted mutation score to reflect the clusters more accurately.
 
@@
 The three metrics are usefulness, frequency and dependency.
-Combining the results of usefulness and frequency metrics, they observed that reducing the number of generated mutants gives noticeable mutant cost reduction without a loss of the mutation score accuracy.
-Wilinski et al., their research is very narrowly scoped to the specific mutators they decided to do research on.
+Combining the results of usefulness and frequency metrics, they observed that reducing the number of generated mutants gives noticeable cost reduction without a loss of the mutation score accuracy.
+Wilinski et al., their research is narrowly scoped to the specific mutators they decided to do research on.
 Our research does not contain this limitation.

### 2021-07-15 chapters/1_abstract.tex
changed tense in introductions

@@
 Software testing is a crucial part of the software engineering process. 
-It is widely used in the industry.
 A part of software testing is building test suites which contain unit tests.
@@
 As projects grow the test suite grows along.
-Maintaining and monitoring these test suites is important because they influence the cost of maintenance.
+Maintaining and monitoring these test suites is important as they influence the cost of maintenance.
+For example, a project with a smaller test suite may have a higher maintenance cost as the effect of change has to be measured manually, which can be a time consuming process.
 Monitoring the quality of test code has shown that tests with lower quality lead to more defect-prone production code.
+A technique to measure the quality of test code is by measuring the test effectiveness.
 \newline
@@
 \newline
-We conducted two experiments in which we cluster the mutants with two different techniques.
-Before conducting the experiment we conducted a state of the art comparison between three mutation testing tools for Java.
+Our research consists of two parts in which we cluster the mutants.
+The first part consists with a white box approach and the second part of a black box approach.
+We conducted a state of the art comparison between three mutation testing tools for Java.
 Based on this research we selected a mutation testing tool to use in our experiments.
 \newline
-We started our experiment by identifying characteristics to represent the mutants.
-For the first experiment we used hierarchical clustering and in the second experiment we trained the \acrlong{fcm} model.
+We identified characteristics to represent the mutants.
+For our white box approach we used hierarchical clustering and in for the black box approach we trained the \acrlong{fcm} model.
 We calculated a weighted mutation score and compared this with the mutation score of a full set of mutants executed.
@@
 The machine learning model did not perform as well as the hierarchical clustering.
-However, more work is needed in finding the right parameters for the \acrlong{fcm} algorithm. 
+However, more can be done by experimenting with different approaches in finding the right parameters for the \acrlong{fcm} algorithm. 
 \end{document}

### 2021-07-15 chapters/2_introduction.tex
changed tense in introductions

@@
 If a large amount of mutants survives, it might be an indication that the quality of the test suite is poor as programming errors remain undetected.
-Mutation testing has considerable drawbacks\todo{add ref}, such as the equivalent mutant problem and being resource expensive.
+Mutation testing has considerable drawbacks\cite{Pizzoleto2019}, such as the equivalent mutant problem and being resource expensive.
 
@@
 \begin{enumerate}
- \item A methodology to cluster generated mutants based on chosen characteristics to reduce the cost of mutation testing.\todo{both white box and black box?}
+ \item A white box methodology to cluster generated mutants based on chosen characteristics to reduce the cost of mutation testing.
+  \item A black box methodology to cluster generated mutants based on chosen characteristics to reduce the cost of mutation testing.
  \item A \acrfull{poc} which implements the qualitative methodology chosen and elaborated within the thesis.

### 2021-07-15 chapters/7_discussion.tex
changed tense in introductions

@@
 In this chapter, we discuss the results of our experiments on clustering mutants.
-First, we discuss the results of the first experiment. 
-We will discuss the all results together and then compare the results between the experiments including and excluding the Levenshtein distance characteristic.
-Second we will discuss the results of the second experiment and compare them to those of the first.
-And last we will discuss the practicality of our research.
+First, we discuss the results of the white box approach. 
+The results are discussed as one and are compared between the variations(with Levenshtein distance characteristic and without).
+Second we discuss the results of the black box approach and compare them to those of the white box approach.
+And last we discuss the practicality of our research.
 

### 2021-07-15 chapters/rq1/2_experiment_design.tex
changed tense in introductions

@@
 \newline
-We will start with extracting the characteristics from the selected mutation testing tool. 
+We start with extracting the characteristics from the selected mutation testing tool. 
 Next we select and implement a clustering algorithm for which we can cluster mutants.
-We can then cluster the mutants and compare the mutation scores to each other.
+Last we cluster the mutants and compare the mutation scores to each other.
 

### 2021-07-15 chapters/rq2/1_experiment_design.tex
changed tense in introductions

@@
 \section{Experiment design}
-In this chapter we will design an experiment to answer the second research question.
+In this chapter we design an experiment to validate our black box approach.
 We hypothesise that we can train a machine learning model with the set of characteristics we identified
@@
 Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.
-We will select a machine learning model, decide on a training strategy, train the model and validate the results.
+We select a machine learning model, decide on a training strategy, train the model and validate the results.
 

### 2021-07-15 chapters/2_introduction.tex
feedback processing

@@
 To reduce the cost of mutation testing, we try to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants. 
-Our goal is to cluster any mutant that is generated \todo{make the goal clearer: no requirements on which mutants get clustered: aim at generic mutant clustering}.
+Our goal is to cluster any mutant that is generated within the Java programming language.
+This research aims to remove the scoping limitations present in existing research.
+There should be no requirements for which mutants can be clustered.
 We will do this with two different approaches: qualitative and quantitative. 

### 2021-07-15 chapters/3_background.tex
feedback processing

@@
 This chapter presents the necessary background information for this thesis. First, we define some basic terminology that will be used throughout this thesis.
-Next, we will compare existing mutation testing tools. We start of by reviewing existing literature and then extend the comparison. 
 This thesis relies on clustering techniques.
@@
 \section{Terminology}
-Subsuming mutants.
-Mutation testing.
-Levenshtein distance.
-Maven.
-Gradle.
-Mutator.
-test engine.
-weighted mutation score.
+\textbf{Mutation testing}
+a form of white box testing in which testers change specific components of an application's source code to ensure a software test suite will be able to detect the changes.
+\textbf{Levenshtein distance} the minimum number of single-character edits (insertions, deletions or substitutions) required to change one piece of text into the other.
+\textbf{Maven} build tooling for Java software projects.
+\textbf{Gradle} build tooling for Java and Android software projects.
+\textbf{Mutator} definition of how the original source code is transformed into a mutant.
+\textbf{test engine} a test engine facilitates discovery and execution of tests. Examples for Java are TestNG and JUnit.
+\textbf{weighted mutation score} mutation score calculation designed to be comparable with the mutation score of a full set(see Chapter \ref{ch:weighted_score} for a detailed explanation).
+
 \section{Project selection}

### 2021-07-15 chapters/4_mutation_tool.tex
feedback processing

@@
 \label{ch:mutation-tool}
-
-Before we can start clustering mutants we need to select a tool that helps us generate mutants. 
+To cluster mutants we need to select a tool that generate mutants and executes mutants.
+Such tools are called mutation testing tools.
 Our goal is to find the tool that generates the most amount of mutants. 
@@
 They found that the mutation tools do not subsume each other. 
+"one mutant subsumes a second mutant if every test that kills the first mutant is guaranteed also to kill the second \cite{Ammann2014EstablishingMutants}"
 According to the research a ranking of the tools would be as follows; MuJava generated the strongest mutants followed by Major and PIT.

### 2021-07-17 chapters/rq1/1_identifying_characteristics.tex
styling issues
tmp

@@
 \label{ch:clustering_characteristics}
-The goal of this experiment is to cluster mutants based on characteristics. 
+The goal of this research is to cluster mutants based on characteristics. 
 To achieve this we need to identify characteristics that the mutants contain. 
@@
 A unit test tests a small piece of code.
-If this piece of codes differs too much from what is being tested the test is more likely fail. 
+The more this piece of code differs from its original the more likely it is the test will fail.
 The Levenshtein distance measures this difference.

### 2021-07-17 chapters/rq1/2_experiment_design.tex
tmp

@@
 \section{Experiment design}
-In this chapter we design an experiment to answer the first research question.
+In this chapter we design an experiment to validate our research.
 We hypothesise that we can cluster mutants with the set of characteristics we identified
@@
 We only need this data to validate our results.
-
-\subsection{Levenshtein distance on Java byte code}
-A mutant is a piece of code that differs in a predefined way from its parent.
-Java source code compiles into Java byte code.
-During the compilation of Java code certain context is abstracted away\cite{byteCodeEngineering}.
-Optimisations are applied which also changes the Java byte code\cite{byteCodeEngineering}.
-As a result the textual similarity between a Java code mutant and its parent is different from the similarity between a Java byte code mutant and its parent.
-While the textual similarity may be different between the Java code and Java byte code the functionality remains the same.
-Java byte code reflects more of the semantic nature of the source code than the source code itself does. 
-PIT generates and executes mutants on byte code level\cite{pitestBytecode}. 
-In other words PIT executes the unit tests of a source against the byte code of a mutant.
-\newline
-By calculating the distance on byte code we filter out the context that is present in Java code.
-This gives us a Levenshtein distance that represents more of the semantic difference, between a mutant and parent, than calculating the distance for Java code.
-For this characteristic we will use the Levenshtein distance between a Java byte code mutant and its parent.
 
@@
 
-\subsection{Mutant selection}
+\subsection{Mutant selection and random sampling}
+\label{ch:mutant_selection}
 As stated in our hypothesis we will randomly select a mutant from each cluster to be executed.
+To validate our sample we make use of statistical hypothesis testing\cite{Emmert-Streib2019UnderstandingInference}. 
+Our null hypothesis states that there is no relation between the characteristics identified and the results of a mutant.
+To test this hypothesis we select an alpha value of 0.05.
+To translate; if more than five percent of the values in a sample deviates more than the significance level(see Chapter \ref{ch:exp1_validation}) we cannot reject the null hypothesis.
 To make our results reproducible we will select the random mutant based on a generated seed.

### 2021-07-17 chapters/rq1/clustering_mutants.tex
tmp

--- 2021-06-04 chapters/6_experiment_design.tex
@@
-\documentclass[../main]{subfiles}
+\documentclass[../../main]{subfiles}
 \begin{document}
-\chapter{Experiment design rq1}\todo{discuss how to split this with exp design for rq2}
-In this chapter we will design an experiment to answer our first research question.
-We hypothesise that we can cluster mutants with the identified set of characteristics 
-while maintaining effectiveness when executing one mutant from each cluster that is in the majority of that cluster.
-If a cluster contains ten mutants with eight mutants killed and two mutants survived, all the mutants that are killed are in the majority of that cluster.
-Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.
-\newline
-We will start with extracting the characteristics from the selected mutation testing tool. 
-Next we select and implement a clustering algorithm for which we can cluster mutants.
-We can then cluster the mutants and compare the results of the clusters with the results of the mutants.
+\section{Clustering mutants, a white box approach}
+\label{ch:clustering_characteristics}
+We identified characteristics to represent mutants in Chapter \ref{ch
 
-\section{Extracting characteristics}
-To extract all the characteristics we identified we need to run PIT and configure it to generate all mutants that are possible within the tool. 
-PIT works in phases, the first phase is the generation phase where mutants are generated.
-The second phase is the execution phase where mutants are challenged by the test suite.
-The last phase is the reporting phase where a report is generated based on the users preferences and data generated by PIT.
-During the various stages we can extract characteristics by using PIT's plugin system.
-\newline
-PIT offers a plugin system in which developers can inject their own code in PIT\cite{pitestPlugin}.
-There are two main types of plugins; a Mutation Result Listener and a Mutation interceptor\cite{pitestPlugin}.
-A mutation result listener receives the details of analysed mutations as they arrive\cite{pitestPlugin}.
-A Mutation interceptor is passed a complete list of all mutation that will be generated to each class before the mutation are challenged by tests\cite{pitestPlugin}.
-The implementation of our experiment makes use of this plugins system.
-\newline
-We developed two plugins to extract characteristics.
-The first plugin is of type Mutation interceptor.
-It is provided a list of details per class. 
-We can extract all characteristics but the Levensthein distance and the amount of tests a mutant is challenged by.
-To calculate the Levensthein distance we need the byte code of the original code and the mutated code.
-We generate the byte code of the original code based on the location provided in the list of details.
-We make use of existing functionality in PIT for generating byte code.
-The mutator is also provided in the list of details.
-With the combination of location and mutator we can generate the byte code of the mutant.
-With both pieces of code available we can calculate the Levenshtein distance between them.
-\newline
-The characteristics gathered until this phase of the mutation testing process are written, together with a unique identifier, to a \acrshort{csv} file.
-\newline
-The second plugin is a mutation result listener. 
-This plugin gets passed a list of the results per mutant.
-This list contains data about the mutant survival status.
-It also contains data about the number of tests the mutant is challenged by.
-We store this data together with the identifier to a different \acrshort{csv} file.
-\newline
-When the mutation testing process has finished we need to merge the characteristic gathered in the second plugin.
-We add the characteristic; the number of tests the mutant is challenged by to the data file containing all the other characteristics.
-We can merge the data because both files contain the same unique identifier per mutant.
-
-\section{Levenshtein distance on Java byte code}
+\subsection{Levenshtein distance on Java byte code}
 A mutant is a piece of code that differs in a predefined way from its parent.
@@
 \newline
-For our characteristic we will use the Levenshtein distance between a Java byte code mutant and its parent.
 By calculating the distance on byte code we filter out the context that is present in Java code.
 This gives us a Levenshtein distance that represents more of the semantic difference, between a mutant and parent, than calculating the distance for Java code.
-
-\subsection{Levenshtein distance implementation}
-write here that we use a memory efficient algorithm
-
-\section{Hierarchical clustering}\todo{elaborate on why hierarchical clustering}
-Clustering Levenshtein distances has been done before with hierarchical clustering\cite{Rajalingam2011HierarchicalStudy, Gothai2010PerformanceAlgorithms}. 
-Hierarchical clustering is subdivided into agglomerative and divisive. 
-The agglomerative hierarchical technique follows bottom up approach whereas divisive follows top-down approaches.
-Hierarchical clustering uses different metrics which measures the the distance between two clusters and the linkage criteria\cite{Rajalingam2011HierarchicalStudy}. 
-The linkage criteria specifies the dissimilarity in the sets as a function of the pair-wise distances of observations in those sets\cite{Rajalingam2011HierarchicalStudy}.
-\newline
-Research shows that the complete linkage outperforms the single linkage method\cite{Vijaya2019ComparativeClustering}.
-The ward linkage and complete linkage methods perform the same when clusters are well separated\cite{Vijaya2019ComparativeClustering}.
-However if the clusters overlap the ward linkage outperforms the complete linkage\cite{Vijaya2019ComparativeClustering}.
-\newline
-We identify all characteristics per mutant as a separate cluster.
-Starting out with each mutant as a separate cluster we can use the agglomerative form of hierarchical clustering.
-Since we cannot assume that our clusters are well separated we chose to use the ward linkage method for our clustering algorithm.
-The agglormorative hierarchical clustering algorithm with ward linkage will cluster our mutants represented as Levenshtein distances.
-\todo{write about implementation}
-
-\section{Weighed mutation score}
-Our hypothesis states that each mutants executed should be in the majority of mutants in its cluster. 
-With a mutant executed from each cluster we can calculate a mutation score.
-This mutation score is a weighed mutation score.
-This weighed mutation score is the product of the result of a mutant(1 for killed and 0 for survived) and the amount of mutants in the cluster it represents.
-The weighed mutation score is then comparable to the score of a full set as the total number of mutants will be the same.
-\newline
-For example, take a full set with a score of 75/100 killed mutants. \todo{example too much?}
-This gives us a mutation score of 75\%. 
-We then cluster the mutants in four clusters consisting of 12, 30, 38 and 20 mutants, respectively.
-We randomly select four mutants of each cluster and execute them.
-The mutants representing cluster one and four survive.
-If we calculate the weighed score we get 68/100 which is 68\%.
-We can then compare this to the score of a full set because the amount of mutants executed is the same: 75/100(75\%) and 68/100(68\%).
-
+For this characteristic we will use the Levenshtein distance between a Java byte code mutant and its parent.
 \end{document}

### 2021-07-17 appendix/appendix.tex
WIP structure

\begin{appendices}

\chapter{PIT maven code}
\label{ap:PIT_maven_code}
Example of code used to execute a full set of mutants with PIT.
\begin{lstlisting}
    <plugins>
      <plugin>
        <groupId>org.pitest</groupId>
        <artifactId>pitest-maven</artifactId>
        <version>1.6.4</version>
        <dependencies>
          <dependency>
            <groupId>com.niverhawk</groupId>
            <artifactId>pitest-clustering-plugin</artifactId>
            <version>1.0-SNAPSHOT</version>
          </dependency>
        </dependencies>
        <configuration>
          <exportLineCoverage>true</exportLineCoverage>
          <mutators>
            <mutator>ALL</mutator>
          </mutators>
          <threads>6</threads>
        </configuration>
      </plugin>
\end{lstlisting}

\end{appendices}

### 2021-07-17 chapters/1_abstract.tex
feedback processing
WIP structure

@@
 Our research consists of two parts in which we cluster the mutants.
-The first part consists with a white box approach and the second part of a black box approach.
+The first part consists of a white box approach and the second part of a black box approach.
 We conducted a state of the art comparison between three mutation testing tools for Java.
@@
 \newline
-We identified characteristics to represent the mutants.
-For our white box approach we used hierarchical clustering and in for the black box approach we trained the \acrlong{fcm} model.
+We identified characteristics to represent the mutants so that we can cluster them.
+For our white box approach we used hierarchical clustering and for the black box approach we trained the \acrlong{fcm} model.
 We calculated a weighted mutation score and compared this with the mutation score of a full set of mutants executed.

### 2021-07-17 chapters/3_background.tex
WIP structure

@@
 \textbf{weighted mutation score} mutation score calculation designed to be comparable with the mutation score of a full set(see Chapter \ref{ch:weighted_score} for a detailed explanation).
-
-\section{Project selection}
-\label{ch:project_selection}
-We choose three main requirements for selecting software projects; the projects should have a test suite, the test suite should not contain failing tests and the mutation testing tool should be able to execute mutants for the sample project.
-We selected projects that were also used in other research within the mutation testing and testing domain\cite{Pizzoleto2019,Yu2019PossibilityScope,Wei2021SpectralTesting, Zhang2019PredictiveTesting, Chen2018SpeedingStudy, Laurent2017AssessingPIT}.
-\newline
-We extend our sample by selecting projects from the first six pages of the most popular Java projects on GitHub\footnote{\url{https://github.com/search?p=7&q=language\%3Ajava+stars\%3A\%3E10000&type=Repositories}}.
-The unfiltered sample contains 50+ projects.
-From there we filter out all the projects that contain failing tests and that are not libraries or applications.
-The result consists of a sample with fifteen projects in total.
 

### 2021-07-17 chapters/5_research_question_1.tex
changes
WIP structure

@@
 \begin{document}
-\chapter{Research question 1}
+\chapter{White box approach}
 \label{ch:reserach_question1}
+The following sections describe the white box approach for clustering mutants.
+We do a qualitative analysis on the mutants to identify characteristics. 
+These characteristics represent the mutants.
+With the mutants represented we devise a methodology for clustering these mutants.
+Our goal is to cluster every mutant that is generated while main tainting effectiveness.
 \subfile{rq1/1_identifying_characteristics}
-\subfile{rq1/2_experiment_design}
-\subfile{rq1/3_results}
+\subfile{rq1/2_clustering_mutants}
+\subfile{rq1/3_experiment_design}
+\subfile{rq1/4_results}
 \end{document}

### 2021-07-17 chapters/rq1/1_identifying_characteristics.tex
WIP structure

@@
 \section{Identifying characteristics}
-\label{ch:clustering_characteristics}
-The goal of this research is to cluster mutants based on characteristics. 
-To achieve this we need to identify characteristics that the mutants contain. 
-In this chapter we identify characteristics that are relevant for defining clusters.
+\label{ch:identifying_characteristics}
+Our goal to cluster any mutant.
+We want to solve the problem of scoping in existing research\cite{Ji2009,Wilinski2015,Ma2016,Wei2021SpectralTesting}.
+To achieve this we need to identify characteristics that every mutant contains. 
+In this chapter we will do a qualitative analysis on mutants and identify characteristics that are relevant for defining the means of domains as the centre of clusters.
 
@@
 
+\subsection{Extracting characteristics}
+To extract all the characteristics we identified we need to run PIT and configure it to generate all mutants that are possible within the tool. 
+PIT works in phases, the first phase is the generation phase where mutants are generated.
+The second phase is the execution phase where mutants are challenged by the test suite.
+The last phase is the reporting phase where a report is generated based on the users preferences and data generated by PIT.
+During the various stages we can extract characteristics by using PIT's plugin system.
+The plugins are executed after each other and per class.
+In other words first plugin A will make changes or analyse all classes then plugin B etc.
+\newline
+PIT offers a plugin system in which developers can inject their own code in PIT\cite{pitestPlugin}.
+There are two main types of plugins; a Mutation Result Listener and a Mutation interceptor\cite{pitestPlugin}.
+A mutation result listener receives the details of analysed mutations as they arrive\cite{pitestPlugin}.
+A Mutation interceptor is passed a complete list of all mutation that will be generated to each class before the mutation are challenged by tests\cite{pitestPlugin}.
+The implementation of our experiment makes use of this plugins system.
+\newline
+We developed two plugins to extract characteristics.
+The first plugin is of type Mutation interceptor.
+It is provided a list of details per class. 
+We can extract all characteristics but the Levensthein distance and the amount of tests a mutant is challenged by.
+To calculate the Levensthein distance we need the byte code of the original code and the mutated code.
+We generate the byte code of the original code based on the location provided in the list of details.
+We make use of existing functionality in PIT for generating byte code.
+The mutator is also provided in the list of details.
+With the combination of location and mutator we can generate the byte code of the mutant.
+With both the mutated byte code and original byte code available we can calculate the Levenshtein distance.
+\newline
+The characteristics gathered until this phase of the mutation testing process are written, together with a unique identifier, to a \acrshort{csv} file.
+\newline
+The second plugin is a mutation result listener. 
+This plugin gets passed a list of the results per mutant.
+This list contains data about the mutant survival status.
+It also contains data about the number of tests the mutant is challenged by.
+We store this data together with the identifier to a different \acrshort{csv} file.
+\newline
+When the mutation testing process has finished we need to merge the characteristic gathered in from the plugins.
+We add the characteristic; the number of tests the mutant is challenged by to the data file containing all the other characteristics.
+All the files contain the same unique identifier per mutant on which we can merge the data.
+We do not add the information about the survival of a mutant.
+We only need this data to validate our results.
+
 \end{document}

### 2021-07-17 chapters/rq1/2_clustering_mutants.tex
WIP structure

--- 2021-07-15 chapters/rq1/2_experiment_design.tex
@@
 \begin{document}
-% Because we use sections but these sections should be interpreted as chapters for rq1 and rq2 sectioning.
-\clearpage
-\section{Experiment design}
-In this chapter we design an experiment to answer the first research question.
-We hypothesise that we can cluster mutants with the set of characteristics we identified
-while maintaining effectiveness and reducing the amount executed when executing one mutant from each cluster that is randomly selected.
-Our goal is to achieve a weighted mutation score that is as close as possible to the mutation score of a full set of executed mutants.
-\newline
-We start with extracting the characteristics from the selected mutation testing tool. 
-Next we select and implement a clustering algorithm for which we can cluster mutants.
-Last we cluster the mutants and compare the mutation scores to each other.
-
-\subsection{Extracting characteristics}
-To extract all the characteristics we identified we need to run PIT and configure it to generate all mutants that are possible within the tool. 
-PIT works in phases, the first phase is the generation phase where mutants are generated.
-The second phase is the execution phase where mutants are challenged by the test suite.
-The last phase is the reporting phase where a report is generated based on the users preferences and data generated by PIT.
-During the various stages we can extract characteristics by using PIT's plugin system.
-The plugins are executed after each other and per class.
-In other words first plugin A will make changes or analyse all classes then plugin B etc.
-\newline
-PIT offers a plugin system in which developers can inject their own code in PIT\cite{pitestPlugin}.
-There are two main types of plugins; a Mutation Result Listener and a Mutation interceptor\cite{pitestPlugin}.
-A mutation result listener receives the details of analysed mutations as they arrive\cite{pitestPlugin}.
-A Mutation interceptor is passed a complete list of all mutation that will be generated to each class before the mutation are challenged by tests\cite{pitestPlugin}.
-The implementation of our experiment makes use of this plugins system.
-\newline
-We developed two plugins to extract characteristics.
-The first plugin is of type Mutation interceptor.
-It is provided a list of details per class. 
-We can extract all characteristics but the Levensthein distance and the amount of tests a mutant is challenged by.
-To calculate the Levensthein distance we need the byte code of the original code and the mutated code.
-We generate the byte code of the original code based on the location provided in the list of details.
-We make use of existing functionality in PIT for generating byte code.
-The mutator is also provided in the list of details.
-With the combination of location and mutator we can generate the byte code of the mutant.
-With both the mutated byte code and original byte code available we can calculate the Levenshtein distance.
-\newline
-The characteristics gathered until this phase of the mutation testing process are written, together with a unique identifier, to a \acrshort{csv} file.
-\newline
-The second plugin is a mutation result listener. 
-This plugin gets passed a list of the results per mutant.
-This list contains data about the mutant survival status.
-It also contains data about the number of tests the mutant is challenged by.
-We store this data together with the identifier to a different \acrshort{csv} file.
-\newline
-When the mutation testing process has finished we need to merge the characteristic gathered in from the plugins.
-We add the characteristic; the number of tests the mutant is challenged by to the data file containing all the other characteristics.
-All the files contain the same unique identifier per mutant on which we can merge the data.
-We do not add the information about the survival of a mutant.
-We only need this data to validate our results.
+\section{Clustering mutants}
+\label{ch:clustering_characteristics}
+We identified characteristics to represent mutants in Chapter \ref{identifying_characteristics}.
+With the mutants represented it is possible to cluster them.
+The following sections describe the methodology for our white box approach of clustering mutants.
 
@@
 The amount of mutants inside a cluster will be decided by the clustering algorithm.
-
-\subsection{Validation}
-\label{ch:exp1_validation}
-The most efficient way to measure test effectiveness with mutation testing is by running all mutants that a mutation testing tool can possibly generate.
-The goal of this experiment is to reduce the amount of mutants executed while maintaining effectiveness.
-To reduce the amounts executed we cluster the mutants. 
-We can measure this by counting the number of clusters we generate and compare it to the number of total mutants generated by the selected tool.
-\newline
-The baseline for our experiment is the mutation score of the set of all mutants generated by the selected mutation testing tool.
-\newline
-To validate how effective our clustering is we can compare the weighted mutation score(see Chapter \ref{ch:weighted_score}) of our clustered set to the mutation score of the full set.
-The closer the weighted score is to that of a full set the more effective our set of characteristics and clustering algorithm proves to be.
-In other words we want achieve a mutation score that is as close as possible as to that of a full set.
-We select a statistical significance level of $\leq 0.05$.
-This is the conventional threshold for declaring statistical significance\cite{Kirk1996PracticalCome}.
-\newline
-Depending on the effectiveness of our clustering algorithm we may loose accuracy.
-This can happen if a cluster contains mutants of both results. 
-We can measure the accuracy inside a cluster by calculating a percentage of all mutants that have survived against the ones that have been killed in a cluster.
-If the majority of the mutants in a cluster is killed then we consider that cluster to represent a killed cluster and the other way around for survived mutants.
-We consider the mutants that are not in the majority of the cluster as inaccuracy.
-\newline
-If the weighted mutation score of our clustered set deviates more than 5\% from the score of a full set we will reject our hypothesis.
-
-\subsection{Mutant selection}
-As stated in our hypothesis we will randomly select a mutant from each cluster to be executed.
-To make our results reproducible we will select the random mutant based on a generated seed.
-The seed generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.
-We will repeat our experiment 30 times with 30 different seeds.
-Achieving consistent results while applying random selection will contribute to the validity of the experiment.
-
 \end{document}

### 2021-07-17 chapters/rq1/3_experiment_design.tex
WIP structure

--- 2021-07-17 chapters/rq1/2_experiment_design.tex
@@
 Our goal is to achieve a weighted mutation score that is as close as possible to the mutation score of a full set of executed mutants.
+
+\subsection{Project selection}
+\label{ch:project_selection}
+We choose three main requirements for selecting software projects; the projects should have a test suite, the test suite should not contain failing tests and the mutation testing tool should be able to execute mutants for the sample project.
+We selected projects that were also used in other research within the mutation testing and testing domain\cite{Pizzoleto2019,Yu2019PossibilityScope,Wei2021SpectralTesting, Zhang2019PredictiveTesting, Chen2018SpeedingStudy, Laurent2017AssessingPIT}.
 \newline
-We start with extracting the characteristics from the selected mutation testing tool. 
-Next we select and implement a clustering algorithm for which we can cluster mutants.
-Last we cluster the mutants and compare the mutation scores to each other.
+We extend our sample by selecting projects from the first six pages of the most popular Java projects on GitHub\footnote{\url{https://github.com/search?p=7&q=language\%3Ajava+stars\%3A\%3E10000&type=Repositories}}.
+The unfiltered sample contains 50+ projects.
+From there we filter out all the projects that contain failing tests and that are not libraries or applications.
+The result consists of a sample with fifteen projects in total.
 
-\subsection{Extracting characteristics}
-To extract all the characteristics we identified we need to run PIT and configure it to generate all mutants that are possible within the tool. 
-PIT works in phases, the first phase is the generation phase where mutants are generated.
-The second phase is the execution phase where mutants are challenged by the test suite.
-The last phase is the reporting phase where a report is generated based on the users preferences and data generated by PIT.
-During the various stages we can extract characteristics by using PIT's plugin system.
-The plugins are executed after each other and per class.
-In other words first plugin A will make changes or analyse all classes then plugin B etc.
-\newline
-PIT offers a plugin system in which developers can inject their own code in PIT\cite{pitestPlugin}.
-There are two main types of plugins; a Mutation Result Listener and a Mutation interceptor\cite{pitestPlugin}.
-A mutation result listener receives the details of analysed mutations as they arrive\cite{pitestPlugin}.
-A Mutation interceptor is passed a complete list of all mutation that will be generated to each class before the mutation are challenged by tests\cite{pitestPlugin}.
-The implementation of our experiment makes use of this plugins system.
-\newline
-We developed two plugins to extract characteristics.
-The first plugin is of type Mutation interceptor.
-It is provided a list of details per class. 
-We can extract all characteristics but the Levensthein distance and the amount of tests a mutant is challenged by.
-To calculate the Levensthein distance we need the byte code of the original code and the mutated code.
-We generate the byte code of the original code based on the location provided in the list of details.
-We make use of existing functionality in PIT for generating byte code.
-The mutator is also provided in the list of details.
-With the combination of location and mutator we can generate the byte code of the mutant.
-With both the mutated byte code and original byte code available we can calculate the Levenshtein distance.
-\newline
-The characteristics gathered until this phase of the mutation testing process are written, together with a unique identifier, to a \acrshort{csv} file.
-\newline
-The second plugin is a mutation result listener. 
-This plugin gets passed a list of the results per mutant.
-This list contains data about the mutant survival status.
-It also contains data about the number of tests the mutant is challenged by.
-We store this data together with the identifier to a different \acrshort{csv} file.
-\newline
-When the mutation testing process has finished we need to merge the characteristic gathered in from the plugins.
-We add the characteristic; the number of tests the mutant is challenged by to the data file containing all the other characteristics.
-All the files contain the same unique identifier per mutant on which we can merge the data.
-We do not add the information about the survival of a mutant.
-We only need this data to validate our results.
-
-\subsection{Levenshtein distance implementation}
-There are different implementations of the Levensthein distance.
-For our experiment we need to calculate the Levenshtein distance on the byte code generated in PIT.
-Since we make use of the plugin system we also choose to use an implementation that is written in Java.
-The Levenshtein distance can be calculated in different ways for example with recursion.
-The implementation with recursion is has a complexity of O(N\^{}2) and a memory need of O(N\^{}2).
-There is also an improved implementation that does not use recursion.
-While the complexity is the same the memory need is reduced to O(N).
-This last implementation is the implementation we selected for our experiment.
-The implementation described above is also offered by a library.
-This library is the Apache Commons Text library\cite{commonsText}, in our experiment we make use of this implementation for calculating the Levenshtein distance.
-
-\subsection{Hierarchical clustering}
-Clustering Levenshtein distances has been done before with hierarchical clustering\cite{Rajalingam2011, Gothai2010PerformanceAlgorithms}. 
-Research shows that hierarchical clustering performs better when clustering with at least ten features\cite{Rodriguez2019}.
-It also states that varying the parameters of hierarchical clustering improves the performance compared to that of the default settings of the algorithm\cite{Rodriguez2019}.
-We have more than ten characteristics which we can use as features and adjust the parameters of the algorithm based on the characteristics in our data set.
-Bases on these observations we select hierarchical clustering to cluster the mutants.
-Next we will explain the configuration we use for clustering the mutants. 
-\newline
-Hierarchical clustering is subdivided into agglomerative and divisive. 
-The agglomerative hierarchical technique follows bottom up approach whereas divisive follows top-down approaches.
-Hierarchical clustering uses different metrics which calculates the euclidean distance between two clusters and the linkage criteria\cite{Rajalingam2011}. 
-The linkage criteria specifies the dissimilarity in the sets as a function of the pair-wise distances of observations in those sets\cite{Rajalingam2011}.
-We reviewed the different linkage criteria and concluded the following.
-\newline
-Research shows that the complete linkage outperforms the single linkage method\cite{Vijaya2019ComparativeClustering}.
-The ward linkage and complete linkage methods perform the same when clusters are well separated\cite{Vijaya2019ComparativeClustering}.
-However if the clusters overlap the ward linkage outperforms the complete linkage\cite{Vijaya2019ComparativeClustering}.
-\newline
-We identify all characteristics per mutant as a separate cluster.
-Starting out with each mutant as a separate cluster we can use the agglomerative form of hierarchical clustering.
-Since we cannot assume that our clusters are well separated we chose to use the ward linkage method for our clustering algorithm.
-In summary, the agglomerative hierarchical clustering algorithm with ward linkage will cluster our mutants represented by the characteristics we gathered.
-
-\subsection{Weighted mutation score}
-\label{ch:weighted_score}
-Our hypothesis states that each mutants executed should represent that whole cluster.
-With a mutant executed from each cluster we can calculate a mutation score.
-This mutation score is a weighted mutation score.
-This weighted mutation score is the product of the result of a mutant(1 for killed and 0 for survived) and the amount of mutants in the cluster it represents.
-The weighted mutation score is then comparable to the score of a full set as the total number of mutants will be the same.
-\newline
-For example, take a full set with a score of 75/100 killed mutants.
-This gives us a mutation score of 75\%. 
-We then cluster the mutants in four clusters consisting of 12, 30, 38 and 20 mutants, respectively.
-We randomly select four mutants of each cluster and execute them.
-The mutants representing cluster one and four survive and two and three are killed.
-If we calculate the weighted score we get 68/100 which is 68\%.
-We can then compare this to the score of a full set because the amount of mutants executed is the same: 75/100(75\%) and 68/100(68\%).
-
-\subsection{Categorical data}
-There are different categorical variable encoding techniques available\cite{Potdar2017AClassifiers}.
-The categorical characteristics we use have no particular ranking compared to each other.
-There is also no specific order to the characteristics.
-I.e. a return type void is not better or worse than a return type string.
-The same goes for the location characteristics, there is no location that should have a bigger weight than the other locations.
-The individual characteristics do contain a finite set of values.
-For example multiple mutants may contain the same class name.
-The hierarchical clustering algorithm needs all characteristics in a numerical form\cite{Vijaya2019ComparativeClustering}.
-The characteristics mutator identifier, class name, method name and return type are non numerical.
-To deal with this problem we apply categorical variable encoding to these specific features.
-Taking into account the properties of our categorical characteristics the nominal variable encoding fits our requirements.
-Nominal encoding comprises a finite set of discrete values with no relationship between values\cite{Potdar2017AClassifiers}.
-Therefore we implement this type of encoding in the experiment.
-
-\subsection{Cluster size}
-Agglomerative hierarchical clustering will continue to cluster until there is one cluster left. 
-Naturally this cluster will contain all the mutants.
-We can cut off the clustering algorithm at any point.
-We decide on the number of clusters based on the amount of mutants generated in a full set.
-The reasoning for the selecting the number of clusters will be from a performance reducing perspective.
-Mutation clustering increases the performance by reducing the amount of mutants executed\cite{Pizzoleto2019}.
-Reducing the performance by a quarter, half or three quarters will give an indication of performance increases.
-To translate this to cluster size we will reduce the number of clusters in steps op 25\%.
-In other words by reducing the amount of mutants, for example by half, we increase the performance by 50\%\cite{Just2014}.
-If necessary we can decide to cut off the clustering at different points.
-The baseline will consist of three experiments where the cluster size is 25\%, 50\% and 75\% of the total amount of mutants respectively.
-The amount of mutants inside a cluster will be decided by the clustering algorithm.
+\subsection{Baseline measurement}
+The baseline for our experiment is the mutation score of the set of all mutants generated by the selected mutation testing tool.
+To extract this score we need to execute a full set of mutants within PIT.
+For each source we have added the PIT plugin and required configuration in the \textit{pom.xml}, this code instructs PIT to generate and execute mutants with all mutators available, an example of such code can be found in Appendix \ref{ap:PIT_maven_code}.
+We can then use the \acrfull{cli} and a maven command(see Listing \ref{lst:mvn-full}) to execute PIT.
+\begin{lstlisting}[label=lst:mvn-full,caption=Command used to execute full set of mutants with PIT.]
+mvn -U org.pitest:pitest-maven:mutationCoverage
+\end{lstlisting}
 
@@
 \label{ch:exp1_validation}
-The most efficient way to measure test effectiveness with mutation testing is by running all mutants that a mutation testing tool can possibly generate.
+The most efficient way to measure test effectiveness with mutation testing is by executing all mutants that a mutation testing tool can possibly generate.
 The goal of this experiment is to reduce the amount of mutants executed while maintaining effectiveness.
@@
 \newline
-The baseline for our experiment is the mutation score of the set of all mutants generated by the selected mutation testing tool.
-\newline
-To validate how effective our clustering is we can compare the weighted mutation score(see Chapter \ref{ch:weighted_score}) of our clustered set to the mutation score of the full set.
+To validate how effective our method of clustering is we can compare the weighted mutation score(see Chapter \ref{ch:weighted_score}) of the clustered set to the mutation score of the full set.
 The closer the weighted score is to that of a full set the more effective our set of characteristics and clustering algorithm proves to be.
@@
 
-\subsection{Mutant selection and random sampling}
+\subsection{Random mutant selection}
 \label{ch:mutant_selection}
-As stated in our hypothesis we will randomly select a mutant from each cluster to be executed.
+As stated in our hypothesis we randomly select a mutant from each cluster to be executed.
 To validate our sample we make use of statistical hypothesis testing\cite{Emmert-Streib2019UnderstandingInference}. 

### 2021-07-17 chapters/11_conclusion.tex
wip restructure

--- 2021-07-15 chapters/9_conclusion.tex
@@
 \newline
-We created a list of characteristics based on existing research. 
+We created a list of identified characteristics based on existing research. 
 We then extended this list by identifying characteristic with logic and data available.
@@
 \newline
-We clustered the mutants with two different clustering techniques; hierarchical clustering and by training a machine learning clustering model.
+We clustered the mutants with white box approach and a black box approach.
+For our white box approach we selected the hierarchical clustering algorithm.
 The hierarchical clustering showed that we can cluster mutants while maintaining effectiveness.
@@
 \newline
-The machine learning model showed a significant reduction in effectiveness.
-We trained the machine learning model with parameters that were optimal according to existing research. 
-We also used a machine learning model, specialized in calculating the number of clusters for fuzzy clustering, to calculate the number of clusters.
+For the black box approach we selected a machine learning model, specifically the \acrlong{fcm} model.
+
+The \acrshort{fcm} model showed a significant reduction in effectiveness.
+We trained the\acrshort{fcm} model with parameters that were optimal according to existing research. 
+We used a different machine learning model, specialized in calculating the number of clusters for fuzzy clustering, to calculate the number of clusters.
 While the minimum clustering accuracy was higher than that of the hierarchical clustering, the maximum cluster accuracy was lower than that of the hierarchical clustering.
@@
 Our research proves that it is possible to determine the means of domains as the centre of a cluster within a programming language without narrowing down the scope to specific pieces of code.
+The centroids our research identified is decided by the characteristics of the mutants.
 It can be applied to all Java projects that can be run with PIT. 
@@
 \subsubsection{Optimizing parameters of machine learning model}
-We trained the machine learning model with specific parameters guided by science. 
+We trained the machine learning model with specific parameters guided by relevant research. 
 These proved to be non optimal for our purpose.
 This research can be extended by finding the optimal parameters for the \acrshort{som} clustering algorithm.
-The results showed a higher minimum accuracy which can indicate that there is potential for clustering the mutants with the machine learning model.
+The results showed a higher minimum accuracy which may indicate that there is potential for clustering the mutants with the machine learning model.
 Further research and fine-tuning is required to find out if the model can be trained to achieve more accurate scores.
@@
 For this researched we trained a model for each source project. 
-A different way is to train one machine learning model with the total set of mutants.
+A different way is to train a machine learning model with the combined set of mutants.
 The model could be continuously trained each time a new source is added.
@@
 It would be valuable to find out if it possible to reduce or expand the set of characteristics and still maintain accurate results.
-Further experimentation can be done with different sets of characteristics.
+Further experimentation may be done with different sets of characteristics.
 \end{document}

### 2021-07-17 chapters/7_experiment_design.tex
wip restructure

--- 2021-07-17 chapters/rq1/3_experiment_design.tex
@@
-\documentclass[../../main]{subfiles}
+\documentclass[../main]{subfiles}
 \begin{document}
@@
 \clearpage
-\section{Experiment design}
+\chapter{Experiment design}
 In this chapter we design an experiment to validate our research.
@@
 
-\subsection{Project selection}
+\section{Project selection}
 \label{ch:project_selection}
@@
 
-\subsection{Baseline measurement}
+\section{Baseline measurement}
 The baseline for our experiment is the mutation score of the set of all mutants generated by the selected mutation testing tool.
@@
 
-\subsection{Validation}
+\section{Validation}
 \label{ch:exp1_validation}
@@
 
-\subsection{Random mutant selection}
+\section{Random mutant selection}
 \label{ch:mutant_selection}

### 2021-07-17 chapters/8_results.tex
wip restructure

\documentclass[../main]{subfiles}
\begin{document}
\chapter{Results}
\label{ch:results}
In this chapter we present the results of our experiments.
We divide the results in two sections. 
The first section displays the results of clustering with hierarchical clustering.
The second section displays the results of clustering with the \acrshort{fcm} model.
\newline
During the process of extracting the characteristics we found that calculating the Levenshtein distance took a long time.
Calculating the Levenshtein distance for the four biggest projects in the data set cost days.
Even for the smaller projects the extraction of the characteristics cost hours.
Including the Levenshtein distance defeats the purpose of increasing performance as every bit of performance we increase would be replaced by the cost of calculating the Levenshtein distance.
To see if we could mitigate this risk we decided to repeat the experiment without the Levensthein distance characteristic.

\subfile{results/1_results_rq1}
\subfile{results/2_results_rq2}
\end{document}

### 2021-07-17 chapters/results/1_results_rq1.tex
wip restructure

\documentclass[../../main]{subfiles}
\begin{document}
\section{Hierarchical clustering}
\label{ch:results_rq1}
In the this section we present the results of the experiment with hierarchical clustering.
We divide the results in two variants. 
The first variant displays the results of clustering with all characteristics.
The second variant displays the results of clustering without the Levensthein distance.
\newline
Due to the limitation of resources and time we had to omit four source projects from these results. 
The hardware was not equipped to store the amount of generated mutants for these projects, we did not have enough memory.
Trying to cluster these mutants this resulted in a out of memory error.

\subsection{Clustering with all characteristics}
This section displays the results of our clustering algorithm when executed with data that consists of all characteristics.
First we show the results to test our null hypothesis discussed in Chapter \ref{ch:mutant_selection}.
The results are be grouped based on the percentage in reduction of mutants executed.
We also display the accuracy of the clusters per reduction.
The results displayed are the average of the results of the 30 repetitions of the experiment.
The result of each single repetition can be found in our Github repository\cite{rbasarat-repo}.

\begin{table}[!htb]
\begin{tabular}{|l|c|c|c|c|c|}
\hline
\multicolumn{1}{|c|}{Project} & \begin{tabular}[c]{@{}c@{}}Mutation score\\ full set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Mutation score \\ clustered set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg. cluster\\  accuracy\end{tabular} & \begin{tabular}[c]{@{}c@{}}Min. cluster\\ accuracy\end{tabular} & \begin{tabular}[c]{@{}c@{}}Max. cluster\\ accuracy\end{tabular} \\ \hline
Google Auto Service           & 0.00\%                                                                                  & 0.00\%                                                                                      & 0.00\%                                                                               & 0.00\%                                                                                & 0.00\%                                                                               \\ \hline
ScribeJava-Core               & 33.52\%                                                                                 & 33.63\%                                                                                     & 93.29\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Auto Factory           & 48.41\%                                                                                 & 48.58\%                                                                                     & 94.18\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Auto Common            & 59.59\%                                                                                 & 60.14\%                                                                                     & 91.12\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Auto Value             & 61.42\%                                                                                 & 62.89\%                                                                                     & 94.33\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Gson                   & 66.83\%                                                                                 & 66.87\%                                                                                     & 91.61\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-io                    & 69.70\%                                                                                 & 66.18\%                                                                                     & 90.36\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-cli                   & 79.07\%                                                                                 & 79.01\%                                                                                     & 91.35\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-text                  & 79.29\%                                                                                 & 79.33\%                                                                                     & 89.95\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-codec                 & 84.56\%                                                                                 & 84.71\%                                                                                     & 94.44\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-csv                   & 85.16\%                                                                                 & 85.42\%                                                                                     & 92.08\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline

\end{tabular}
\caption{\label{tab:clustering_all_25}Results of clustering mutants with all characteristics and cluster size of n*0.25}
\end{table}
\FloatBarrier

Table \ref{tab:clustering_all_25} displays the results of clustering mutants with all characteristics and the number of clusters equal to the total amount of mutants* 0.25.  
The maximum and minimum differences between the score of a full set and clustered are is 3.52\% and 0.04\% respectively.  
The average of the differences between full set score and clustered score is 0.58\%.
It is also noticeable that the minimum and maximum accuracy are 50\% and 100\% respectively.
This means that there was at least one cluster that consisted of only mutants that were killed or survived.
The minimum accuracy cannot go below 50\%.
If the percentage does go below 50\% the representation of a cluster will flip from survived to killed or the other way around.

\begin{table}[!htb]
\begin{tabular}{|l|c|c|c|c|c|}
\hline
\multicolumn{1}{|c|}{Project} & \begin{tabular}[c]{@{}c@{}}Mutation score\\ full set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Mutation score \\ clustered set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg. cluster\\  accuracy\end{tabular} & \begin{tabular}[c]{@{}c@{}}Min. cluster\\ accuracy\end{tabular} & \begin{tabular}[c]{@{}c@{}}Max. cluster\\ accuracy\end{tabular} \\ \hline
Google Auto Service           & 0.00\%                                                                                  & 0.00\%                                                                                      & 0.00\%                                                                               & 0.00\%                                                                                & 0.00\%                                                                             \\ \hline
ScribeJava-Core               & 33.52\%                                                                                 & 33.54\%                                                                                     & 96.94\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Auto Factory           & 48.41\%                                                                                 & 48.39\%                                                                                     & 96.12\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Auto Common            & 59.59\%                                                                                 & 59.76\%                                                                                     & 94.49\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Auto Value             & 61.42\%                                                                                 & 62.71\%                                                                                     & 96.09\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Gson                   & 66.83\%                                                                                 & 66.81\%                                                                                     & 94.75\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-io                    & 69.70\%                                                                                 & 66.06\%                                                                                     & 94.03\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-cli                   & 79.07\%                                                                                 & 79.17\%                                                                                     & 94.19\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-text                  & 79.29\%                                                                                 & 79.29\%                                                                                     & 93.21\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-codec                 & 84.56\%                                                                                 & 84.60\%                                                                                     & 95.92\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-csv                   & 85.16\%                                                                                 & 85.13\%                                                                                     & 94.65\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
\end{tabular}
\caption{\label{tab:clustering_all_50}Results of clustering mutants with all characteristics and size n*0.50}
\end{table}
\FloatBarrier

Table \ref{tab:clustering_all_50} displays the results of clustering mutants with all characteristics and the number of clusters equal to the total amount of mutants* 0.50.  
The maximum and minimum differences between the score of a full set and clustered set are 3.63\% and 0.02\% respectively.
The average of the differences between full set score and clustered score is 0.48\% which is smaller than the average difference than that of the 25\% set.

\begin{table}[!htb]
\begin{tabular}{|l|c|c|c|c|c|}
\hline
\multicolumn{1}{|c|}{Project} & \begin{tabular}[c]{@{}c@{}}Mutation score\\ full set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Mutation score \\ clustered set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg. cluster\\  accuracy\end{tabular} & \begin{tabular}[c]{@{}c@{}}Min. cluster\\ accuracy\end{tabular} & \begin{tabular}[c]{@{}c@{}}Max. cluster\\ accuracy\end{tabular} \\ \hline
Google Auto Service           & 0.00\%                                                                                  & 0.00\%                                                                                      & 0.00\%                                                                               & 0.00\%                                                                                & 0.00\%                                                                               \\ \hline
ScribeJava-Core               & 33.52\%                                                                                 & 33.52\%                                                                                     & 97.99\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Auto Factory           & 48.41\%                                                                                 & 48.41\%                                                                                     & 97.46\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Auto Common            & 59.59\%                                                                                 & 59.76\%                                                                                     & 96.16\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Auto Value             & 61.42\%                                                                                 & 62.68\%                                                                                     & 97.31\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Gson                   & 66.83\%                                                                                 & 66.84\%                                                                                     & 97.59\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-io                    & 69.70\%                                                                                 & 66.07\%                                                                                     & 97.19\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-cli                   & 79.07\%                                                                                 & 79.10\%                                                                                     & 96.68\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-text                  & 79.29\%                                                                                 & 79.33\%                                                                                     & 97.61\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-codec                 & 84.56\%                                                                                 & 84.61\%                                                                                     & 98.84\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-csv                   & 85.16\%                                                                                 & 85.07\%                                                                                     & 97.64\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
\end{tabular}
\caption{\label{tab:clustering_all_75}Results of clustering mutants with a single characteristic and size n*0.75}
\end{table}
\FloatBarrier

Table \ref{tab:clustering_all_75} displays the results of clustering mutants with all characteristics and the number of clusters equal to the total amount of mutants* 0.75.  
The maximum and minimum differences between the score of a full set and clustered set are 3.63\% and 0\% respectively.  
The average of the differences between full set score and clustered score is 0.48\% which is higher than the average difference than that of the 25\% and the same of that of the 50\% set.

\subsection{Clustering without Levenshtein distance}
This section will display the results of our clustering algorithm when executed without the Levenshtein distance.
The results will be grouped based on the reduced percentage of mutants executed.
We will also display the accuracy of the clusters per reduction amount.
The results displayed are the average of the results of the 30 repetitions of the experiment.
The result of each single repetition can be found in our Github repository\cite{rbasarat-repo}.
\begin{table}[htb]
\begin{tabular}{|l|c|c|c|c|c|}
\hline
\multicolumn{1}{|c|}{Project} & \begin{tabular}[c]{@{}c@{}}Mutation score\\ full set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Mutation score \\ clustered set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg. cluster\\  accuracy\end{tabular} & \begin{tabular}[c]{@{}c@{}}Min. cluster\\ accuracy\end{tabular} & \begin{tabular}[c]{@{}c@{}}Max. cluster\\ accuracy\end{tabular} \\ \hline
Google Auto Service           & 0.00\%                                                                                  & 0.00\%                                                                                      & 0.00\%                                                                               & 0.00\%                                                                                & 0.00\%                                                                               \\ \hline
ScribeJava-Core               & 33.52\%                                                                                 & 33.72\%                                                                                     & 92.24\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Auto Factory           & 48.41\%                                                                                 & 48.32\%                                                                                     & 94.21\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Auto Common            & 59.59\%                                                                                 & 60.03\%                                                                                     & 91.19\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Auto Value             & 61.42\%                                                                                 & 62.58\%                                                                                     & 93.90\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Gson                   & 66.83\%                                                                                 & 66.89\%                                                                                     & 91.75\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-io                    & 69.70\%                                                                                 & 66.08\%                                                                                     & 90.20\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-cli                   & 79.07\%                                                                                 & 78.97\%                                                                                     & 91.11\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-text                  & 79.29\%                                                                                 & 79.30\%                                                                                     & 90.95\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-codec                 & 84.56\%                                                                                 & 84.59\%                                                                                     & 94.56\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-csv                   & 85.16\%                                                                                 & 85.24\%                                                                                     & 92.76\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
\end{tabular}
\caption{\label{tab:clustering_no_distance_25}Results of clustering mutants without Levensthein distance and cluster size n*0.25}
\end{table}
\FloatBarrier

Table \ref{tab:clustering_no_distance_25} displays the results of clustering mutants without Levensthein distance and the number of clusters equal to the total amount of mutants * 0.25. 
We observe that the maximum and minimum differences in mutation score between a full set and a clustered set  3.61\% and 0.09\% respectively.
The average of the differences between full set score and clustered score is 0.53\%.

\begin{table}[htb]
\begin{tabular}{|l|c|c|c|c|c|}
\hline
\multicolumn{1}{|c|}{Project} & \begin{tabular}[c]{@{}c@{}}Mutation score\\ full set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Mutation score \\ clustered set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg. cluster\\  accuracy\end{tabular} & \begin{tabular}[c]{@{}c@{}}Min. cluster\\ accuracy\end{tabular} & \begin{tabular}[c]{@{}c@{}}Max. cluster\\ accuracy\end{tabular} \\ \hline
Google Auto Service           & 0.00\%                                                                                  & 0.00\%                                                                                      & 0.00\%                                                                               & 0.00\%                                                                                & 0.00\%                                                                               \\ \hline
ScribeJava-Core               & 33.52\%                                                                                 & 33.55\%                                                                                     & 97.16\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Auto Factory           & 48.41\%                                                                                 & 48.44\%                                                                                     & 97.71\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Auto Common            & 59.59\%                                                                                 & 59.86\%                                                                                     & 96.35\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Auto Value             & 61.42\%                                                                                 & 62.62\%                                                                                     & 97.31\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Gson                   & 66.83\%                                                                                 & 66.81\%                                                                                     & 95.29\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-io                    & 69.70\%                                                                                 & 66.05\%                                                                                     & 95.08\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-cli                   & 79.07\%                                                                                 & 78.97\%                                                                                     & 95.55\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-text                  & 79.29\%                                                                                 & 79.30\%                                                                                     & 95.34\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-codec                 & 84.56\%                                                                                 & 84.59\%                                                                                     & 96.95\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-csv                   & 85.16\%                                                                                 & 85.11\%                                                                                     & 95.87\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
\end{tabular}
\caption{\label{tab:clustering_no_distance_50}Results of clustering mutants without Levensthein distance and cluster size n*0.50}
\end{table}
\FloatBarrier

Table \ref{tab:clustering_no_distance_50} displays the results of clustering mutants without Levensthein distance and the number of clusters equal to the total amount of mutants * 0.50.
The maximum and minimum differences between the score of a full set and clustered set are 3.65\% and 0.03\% respectively.
We observe that the average cluster accuracy is higher than the average cluster accuracy of the 25\% set. 
The average of the differences between full set score and clustered score is 0.49\% which is smaller than the average difference than that of the 25\% set.

\begin{table}[htb]
\begin{tabular}{|l|c|c|c|c|c|}
\hline
\multicolumn{1}{|c|}{Project} & \begin{tabular}[c]{@{}c@{}}Mutation score\\ full set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Mutation score \\ clustered set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg. cluster\\  accuracy\end{tabular} & \begin{tabular}[c]{@{}c@{}}Min. cluster\\ accuracy\end{tabular} & \begin{tabular}[c]{@{}c@{}}Max. cluster\\ accuracy\end{tabular} \\ \hline
Google Auto Service           & 0.00\%                                                                                  & 0.00\%                                                                                      & 0.00\%                                                                               & 0.00\%                                                                                & 0.00\%                                                                               \\ \hline
ScribeJava-Core               & 33.52\%                                                                                 & 33.48\%                                                                                     & 98.63\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Auto Factory           & 48.41\%                                                                                 & 48.34\%                                                                                     & 98.41\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Auto Common            & 59.59\%                                                                                 & 59.71\%                                                                                     & 97.08\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Auto Value             & 61.42\%                                                                                 & 62.66\%                                                                                     & 97.99\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
Google Gson                   & 66.83\%                                                                                 & 66.84\%                                                                                     & 97.37\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-io                    & 69.70\%                                                                                 & 66.07\%                                                                                     & 97.34\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-cli                   & 79.07\%                                                                                 & 79.12\%                                                                                     & 96.83\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-text                  & 79.29\%                                                                                 & 79.33\%                                                                                     & 97.60\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-codec                 & 84.56\%                                                                                 & 84.61\%                                                                                     & 98.43\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
commons-csv                   & 85.16\%                                                                                 & 85.07\%                                                                                     & 97.49\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
\end{tabular}
\caption{\label{tab:clustering_no_distance_75}Results of clustering mutant without Levensthein distance and cluster size n*0.75}
\end{table}
\FloatBarrier

Table \ref{tab:clustering_no_distance_75} displays the results of clustering mutants without Levensthein distance and the number of clusters equal to the total amount of mutants * 0.50.
The maximum and minimum differences between the score of a full set and clustered set are 3.63\% and 0.04\% respectively.
We observe that the average accuracy is higher than the average accuracy of the other sets.
The average of the differences between full set score and clustered score is  0.49\% which is higher than the average difference than that of the 25\% and the same of that of the 50\% set.
\end{document}

### 2021-07-17 chapters/rq2/1_clustering_mutants.tex
wip restructure

--- 2021-07-15 chapters/rq2/1_experiment_design.tex
@@
 \begin{document}
-\section{Experiment design}
-In this chapter we design an experiment to validate our black box approach.
-We hypothesise that we can train a machine learning model with the set of characteristics we identified
-while maintaining effectiveness and reducing the amount executed when executing one mutant from each cluster that is randomly selected.
+\section{Clustering mutants}
+In this chapter devise a methodology for our black box approach.
 Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.
-We select a machine learning model, decide on a training strategy, train the model and validate the results.
+We select a machine learning model and decide on a training strategy.
 

### 2021-07-17 chapters/5_research_question_1.tex
wip restructure
restructure done

--- 2021-06-23 chapters/5_research_question_1.tex
@@
 \begin{document}
-\chapter{Research question 1}
+\chapter{Qualitative clustering approach}
 \label{ch:reserach_question1}
+The following sections describe the white box approach for clustering mutants.
+We perform a qualitative analysis on mutants to identify characteristics. 
+These characteristics represent the mutants.
+With the mutants represented we devise a qualitative methodology for clustering mutants.
+Our goal is to cluster every mutant that is generated while maintaining effectiveness.
 \subfile{rq1/1_identifying_characteristics}
-\subfile{rq1/2_experiment_design}
-\subfile{rq1/3_results}
+\subfile{rq1/2_clustering_mutants}
 \end{document}

### 2021-07-17 chapters/6_research_question_2.tex
redo structure
minor tweaks exp2 design + results
WIP structure
wip restructure
restructure done

--- 2021-06-23 chapters/5_research_question_1.tex
@@
 \begin{document}
-\chapter{Research question 1}
-\label{ch:reserach_question1}
-\subfile{rq1/1_identifying_characteristics}
-\subfile{rq1/2_experiment_design}
-\subfile{rq1/3_results}
+\chapter{Quantitative clustering approach}
+\label{ch:reserach_question2}
+In the following sections we will devise a black box approach for clustering mutants.
+We make use of the characteristics identified in Chapter \ref{ch:identifying_characteristics}.
+While our goal remains the same as in Chapter \ref{ch:reserach_question1}, the way to achieve it is different.
+We start by selecting a machine learning algorithm.
+Next we select a training strategy and an implementation.
+Last we select the parameters required by the algorithm.
+
+\subfile{rq2/1_clustering_mutants}
 \end{document}

### 2021-07-17 chapters/7_experiment_design.tex
remove clearpage
restructure done

@@
 \begin{document}
-% Because we use sections but these sections should be interpreted as chapters for rq1 and rq2 sectioning.
-\clearpage
 \chapter{Experiment design}
@@
 Our goal is to achieve a weighted mutation score that is as close as possible to the mutation score of a full set of executed mutants.
+We will repeat the experiment for each approach we devised in Chapters \ref{ch:reserach_question1} and \ref{ch:reserach_question2}.
 
@@
 \end{lstlisting}
+
+\subsection{Weighted mutation score}
+\label{ch:weighted_score}
+Our hypothesis states that each mutants executed should represent that whole cluster.
+With a mutant executed from each cluster we can calculate a mutation score.
+This mutation score is a weighted mutation score.
+This weighted mutation score is the product of the result of a mutant(1 for killed and 0 for survived) and the amount of mutants in the cluster it represents.
+The weighted mutation score is then comparable to the score of a full set as the total number of mutants will be the same.
+\newline
+For example, take a full set with a score of 75/100 killed mutants.
+This gives us a mutation score of 75\%. 
+We then cluster the mutants in four clusters consisting of 12, 30, 38 and 20 mutants, respectively.
+We randomly select four mutants of each cluster and execute them.
+The mutants representing cluster one and four survive and two and three are killed.
+If we calculate the weighted score we get 68/100 which is 68\%.
+We can then compare this to the score of a full set because the amount of mutants executed is the same: 75/100(75\%) and 68/100(68\%).
 

### 2021-07-17 chapters/rq1/1_identifying_characteristics.tex
restructure done

@@
 \label{ch:identifying_characteristics}
-Our goal to cluster any mutant.
-We want to solve the problem of scoping in existing research\cite{Ji2009,Wilinski2015,Ma2016,Wei2021SpectralTesting}.
+Our goal is to cluster any mutant that fit the requirements of the selected mutation testing tools.
+We want to solve the scoping problem in existing research\cite{Ji2009,Wilinski2015,Ma2016,Wei2021SpectralTesting}.
 To achieve this we need to identify characteristics that every mutant contains. 

### 2021-07-17 chapters/rq1/2_clustering_mutants.tex
restructure done

@@
 \label{ch:clustering_characteristics}
-We identified characteristics to represent mutants in Chapter \ref{identifying_characteristics}.
+We identified characteristics to represent mutants in Chapter \ref{ch:identifying_characteristics}.
 With the mutants represented it is possible to cluster them.
-The following sections describe the methodology for our white box approach of clustering mutants.
+The following sections describes the methodology for our white box approach of clustering mutants.
+
+\subsection{Hierarchical clustering}
+Clustering Levenshtein distances has been done before with hierarchical clustering\cite{Rajalingam2011, Gothai2010PerformanceAlgorithms}. 
+Research shows that hierarchical clustering performs better when clustering with at least ten features\cite{Rodriguez2019}.
+It also states that varying the parameters of hierarchical clustering improves the performance compared to that of the default settings of the algorithm\cite{Rodriguez2019}.
+We identified more than ten characteristics and can use them as features.
+We adjust the parameters of the algorithm based on the characteristics in our data set.
+Bases on these observations we select hierarchical clustering to cluster the mutants.
+Next we will explain the configuration we use for clustering the mutants. 
+\newline
+Hierarchical clustering is subdivided into agglomerative and divisive. 
+The agglomerative hierarchical technique follows bottom up approach whereas divisive follows top-down approaches.
+Hierarchical clustering uses different metrics which calculates the euclidean distance between two clusters and the linkage criteria\cite{Rajalingam2011}. 
+The linkage criteria specifies the dissimilarity in the sets as a function of the pair-wise distances of observations in those sets\cite{Rajalingam2011}.
+We reviewed the different linkage criteria and concluded the following.
+\newline
+Research shows that the complete linkage outperforms the single linkage method\cite{Vijaya2019ComparativeClustering}.
+The ward linkage and complete linkage methods perform the same when clusters are well separated\cite{Vijaya2019ComparativeClustering}.
+However if the clusters overlap the ward linkage outperforms the complete linkage\cite{Vijaya2019ComparativeClustering}.
+\newline
+We identify all characteristics per mutant as a separate cluster.
+Starting out with each mutant as a separate cluster we can use the agglomerative form of hierarchical clustering.
+Since we cannot assume that our clusters are well separated we chose to use the ward linkage method for our clustering algorithm.
+In summary, the agglomerative hierarchical clustering algorithm with ward linkage will cluster our mutants represented by the characteristics we gathered.
 
@@
 This library is the Apache Commons Text library\cite{commonsText}, in our experiment we make use of this implementation for calculating the Levenshtein distance.
-
-\subsection{Hierarchical clustering}
-Clustering Levenshtein distances has been done before with hierarchical clustering\cite{Rajalingam2011, Gothai2010PerformanceAlgorithms}. 
-Research shows that hierarchical clustering performs better when clustering with at least ten features\cite{Rodriguez2019}.
-It also states that varying the parameters of hierarchical clustering improves the performance compared to that of the default settings of the algorithm\cite{Rodriguez2019}.
-We have more than ten characteristics which we can use as features and adjust the parameters of the algorithm based on the characteristics in our data set.
-Bases on these observations we select hierarchical clustering to cluster the mutants.
-Next we will explain the configuration we use for clustering the mutants. 
-\newline
-Hierarchical clustering is subdivided into agglomerative and divisive. 
-The agglomerative hierarchical technique follows bottom up approach whereas divisive follows top-down approaches.
-Hierarchical clustering uses different metrics which calculates the euclidean distance between two clusters and the linkage criteria\cite{Rajalingam2011}. 
-The linkage criteria specifies the dissimilarity in the sets as a function of the pair-wise distances of observations in those sets\cite{Rajalingam2011}.
-We reviewed the different linkage criteria and concluded the following.
-\newline
-Research shows that the complete linkage outperforms the single linkage method\cite{Vijaya2019ComparativeClustering}.
-The ward linkage and complete linkage methods perform the same when clusters are well separated\cite{Vijaya2019ComparativeClustering}.
-However if the clusters overlap the ward linkage outperforms the complete linkage\cite{Vijaya2019ComparativeClustering}.
-\newline
-We identify all characteristics per mutant as a separate cluster.
-Starting out with each mutant as a separate cluster we can use the agglomerative form of hierarchical clustering.
-Since we cannot assume that our clusters are well separated we chose to use the ward linkage method for our clustering algorithm.
-In summary, the agglomerative hierarchical clustering algorithm with ward linkage will cluster our mutants represented by the characteristics we gathered.
-
-\subsection{Weighted mutation score}
-\label{ch:weighted_score}
-Our hypothesis states that each mutants executed should represent that whole cluster.
-With a mutant executed from each cluster we can calculate a mutation score.
-This mutation score is a weighted mutation score.
-This weighted mutation score is the product of the result of a mutant(1 for killed and 0 for survived) and the amount of mutants in the cluster it represents.
-The weighted mutation score is then comparable to the score of a full set as the total number of mutants will be the same.
-\newline
-For example, take a full set with a score of 75/100 killed mutants.
-This gives us a mutation score of 75\%. 
-We then cluster the mutants in four clusters consisting of 12, 30, 38 and 20 mutants, respectively.
-We randomly select four mutants of each cluster and execute them.
-The mutants representing cluster one and four survive and two and three are killed.
-If we calculate the weighted score we get 68/100 which is 68\%.
-We can then compare this to the score of a full set because the amount of mutants executed is the same: 75/100(75\%) and 68/100(68\%).
 

### 2021-07-17 chapters/rq2/1_clustering_mutants.tex
restructure done

@@
 \begin{document}
-\section{Clustering mutants}
-In this chapter devise a methodology for our black box approach.
-Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.
-We select a machine learning model and decide on a training strategy.
 
-\subsection{Selecting a machine learning model}
+\section{Selecting a machine learning model}
+In this section we will select a machine learning model.
 We will first choose between supervised or unsupervised learning.
@@
 
-\subsubsection{Supervised or unsupervised learning}
+\subsection{Supervised or unsupervised learning}
 In this subsection we will decide on supervised or unsupervised learning.
@@
 
-\subsubsection{Mountain and subtractive clustering}
+\subsection{Mountain and subtractive clustering}
 The mountain and subtractive clustering models are an effective method for estimating the number of clusters\cite{Du2010Clustering:Approach}. 
@@
 
-\subsubsection{ART models}
+\subsection{ART models}
 ART models are characterized by systems of differential equations that formulate stable self-organizing learning methods\cite{Du2010Clustering:Approach}.
@@
 
-\subsubsection{Neural gas}
+\subsection{Neural gas}
 The neural gas is a topology-preserving network, and can be treated as an extension to the C-means\cite{Du2010Clustering:Approach}.
@@
 
-\subsubsection{Learning vector quantization}
+\subsection{Learning vector quantization}
 The learning vector quantization model is a supervised learning model.
@@
 
-\subsubsection{Self organizing map, C-means and \acrlong{fcm}}
+\subsection{Self organizing map, C-means and \acrlong{fcm}}
 \label{ch:topThreeModels}
@@
 
-\subsection{Training strategy}
+\section{Training strategy}
 \label{ch:training_strategy}
-Our data set contains of mutants represented by their characteristics. 
+Our data set consists of mutants represented by their characteristics. 
 The characteristics of the mutants are generated per project and have a project specific context.
+For example, a mutant with a mutator in a certain project can have a different effect on the source code and can give other results than a mutator in a different project.
 The relations between the mutants are scoped to their own project. 
 In other words if a specific mutant has a relation to another mutant in the same project, this relation may not exist between the same mutant and a mutant from a different project.
-For example, a mutant with a mutator in a certain project can have a different effect on the source code and can give other results than a mutator in a different project.
-To keep the relations and context we decided to train a separate model per project.
+To preserve these relations and context we decided to train a separate model per project.
 \newline
 To validate the model we need to split our data set into a training set and a validation set.
-Our total data set consists of approximately 1 million mutants generated over fourteen projects.
 Research has shown that if the data set is big enough a ratio of 80/20 is sufficient\cite{Guyon1997ARatio}.
@@
 The division of the mutants will be done randomly. 
-To make our results reproducible we will select a random division based on a generated seed.
+To make our results reproducible we select a random division based on a generated seed.
 The seeds generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.
 
-\subsection{Implementation of algorithm}
+\section{Implementation of algorithm}
 \label{ch:fcm_implementation}
-The research paper of the \acrshort{fcm} algorithm is referenced in the survey of Du et al\cite{Du2010Clustering:Approach}.
+The research paper of the \acrshort{fcm} algorithm is referenced in the survey of Du et al.\cite{Du2010Clustering:Approach}.
 The \acrshort{fcm} algorithm is developed by Bezdek et al.\cite{Bezdek1984FCM:Algorithm}.
@@
 
-\subsection{Algorithm parameters}
+\section{Algorithm parameters}
 The \acrshort{fcm} algorithm has to initialize the centroids of the clusters.
@@
 
-\subsection{Implementation of subtractive clustering}
+\section{Implementation of subtractive clustering}
 For the same reasons as in Chapter \ref{ch:fcm_implementation} we will select a library that implements the subtractive clsutering algorithm.
@@
 
-\subsection{Validation}
-To validate our machine learning model we will calculate the weighted mutation score and cluster accuracy based on the clusters calculated by the model.
-We use the same validation method as described in Chapter \ref{ch:exp1_validation}.
-We will use the mutants of projects that are not selected for training as input for validating the model.
-As described in Chapter \ref{ch:training_strategy} the mutants should represent approximately 20\% of the total amount of mutants per project.
-It may be possible the model has to cluster mutants it has never seen before.
-We will use the same statistical significance threshold as in our firsts experiment to decide whether we will accept or reject or hypothesis.
-To make our results reproducible we will select a random mutant based on a generated seed.
-We will repeat our experiment 30 times with 30 different seeds.
-Achieving consistent results while applying random selection and cluster initialization will contribute to the validity of the experiment.
-The seed generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.
-
 \end{document}

### 2021-07-17 chapters/10_related_work.tex
minor changes

--- 2021-07-15 chapters/8_related_work.tex
@@
 
-\subsection{Clustering overlapped mutants}
+\section{Clustering overlapped mutants}
 \label{ch:overlapping_mutants}
@@
 
-\subsection{Clustering by scope}
+\section{Clustering by scope}
 Yu et al., extended the research about overlapping mutants(See chapter \ref{ch:overlapping_mutants})\cite{Yu2019PossibilityScope}.
@@
 
-\subsection{Clustering Hamming distance}
+\section{Clustering Hamming distance}
 Ji et al., did a qualitative domain analysis on mutants\cite{Ji2009}.
@@
 
-\subsection{Spectral clustering}
+\section{Spectral clustering}
 Wei et al., makes use of an intelligent technique, namely spectral clustering, to improve the efficacy of mutant reduction\cite{Wei2021SpectralTesting}.
@@
 
-\subsection{Clustering similarity}
+\section{Clustering similarity}
 Hussain et al., used the k-means and agglomerative clustering algorithm to cluster the mutants according to a similarity measure. 
@@
 
-\subsection{Generalizing mutants}
+\section{Generalizing mutants}
 Wilinski et al., tried to generalize the mutants by defining metrics.
@@
 Our research does not contain this limitation.
-The mutator is used as a characteristic in our research instead of limiting our research it helps define the the mutant.
+The mutator is used as a characteristic in our research instead of limiting our research it helps define a mutant.
 

### 2021-07-17 chapters/7_experiment_design.tex
minor changes

@@
 
-\subsection{Weighted mutation score}
+\section{Weighted mutation score}
 \label{ch:weighted_score}
-Our hypothesis states that each mutants executed should represent that whole cluster.
+Our hypothesis states that each mutant executed should represent that whole cluster.
 With a mutant executed from each cluster we can calculate a mutation score.
@@
 The most efficient way to measure test effectiveness with mutation testing is by executing all mutants that a mutation testing tool can possibly generate.
-The goal of this experiment is to reduce the amount of mutants executed while maintaining effectiveness.
+The goal of this research is to reduce the amount of mutants executed while maintaining effectiveness.
 To reduce the amounts executed we cluster the mutants. 

### 2021-07-17 chapters/rq2/1_clustering_mutants.tex
minor changes

@@
 
-\subsection{Self organizing map, C-means and \acrlong{fcm}}
+\subsection{Self organizing map, C-means and fuzzy c-means}
 \label{ch:topThreeModels}
@@
 When the data set contains noise and outliers, the fuzzifier \textit{m}=4 is recommended in a theoretical upper bound case.
-Since our data set does not contain noise we choose to select the midpoint concluded in the research of Bezdek et al.\cite{Bezdek1995OnModel}.
+Since our data set does not contain noise(all mutants consists of every characteristic) we choose to select the midpoint concluded in the research of Bezdek et al.\cite{Bezdek1995OnModel}.
 The value of \textit{m} in this experiment is 2.

### 2021-07-17 chapters/1_abstract.tex
wip until now

@@
 We conducted a state of the art comparison between three mutation testing tools for Java.
-Based on this research we selected a mutation testing tool to use in our experiments.
+Based on this research we selected a mutation testing tool to use in our research.
 \newline
-We identified characteristics to represent the mutants so that we can cluster them.
-For our white box approach we used hierarchical clustering and for the black box approach we trained the \acrlong{fcm} model.
+We identified characteristics to represent the mutants such that we can cluster them.
+For our white box approach we used hierarchical clustering and for our black box approach we trained the \acrlong{fcm} model.
 We calculated a weighted mutation score and compared this with the mutation score of a full set of mutants executed.
@@
 The machine learning model did not perform as well as the hierarchical clustering.
-However, more can be done by experimenting with different approaches in finding the right parameters for the \acrlong{fcm} algorithm. 
+However, more can be done by researching different approaches for training the \acrshort{fcm} model and finding the right parameters. 
 \end{document}

### 2021-07-17 chapters/9_discussion.tex
wip until now

--- 2021-07-15 chapters/7_discussion.tex
@@
 \end{finding}
-
 \section{Application}

### 2021-07-19 chapters/3_background.tex
wip boxplots

@@
 \textbf{weighted mutation score} mutation score calculation designed to be comparable with the mutation score of a full set(see Chapter \ref{ch:weighted_score} for a detailed explanation).
+
+\section{Statistics}
+The terminology of statistics may lead to confusion if not understood correctly.
+Thus we discuss some basic terminology and principles in statistics.
+When performing a statistical test we have two hypotheses; a null hypothesis and an alternative hypothesis.
+A test contains a specific null and alternative hypothesis.
+The goal of such a test is to reject the null hypothesis.
+Rejecting a null hypothesis can be done by looking at the resulting \textit{p-value}.
+The \textit{p-value} is the chance that the value of the statistical test occurs if the null hypothesis is true on a
+zero to one scale.
+We can reject a null hypothesis if we think that this chance is below a certain threshold.
+This threshold is determined beforehand and named the \textit{alpha-value}.
+A commonly used value for alpha is 5\%.
+We cannot reject the null hypothesis if the \textit{p-value} is below the \textit{alpha-value}.
+This does not necessarily mean that the null hypothesis is true.
+This can have different reasons, for example, the data set used to validate the hypothesis is too small.
 

### 2021-07-19 chapters/results/1_results_rq1.tex
restructure done
minor changes
wip boxplots

@@
 \newline
+For each variant and project we display the weighted mutation score to create box-plots.
+There are a total of eleven projects and thus also eleven box-plots per variant.
+In a box-plot the box is from the first quartile till the third quartile and a line at the median. 
+The dots are measurement points laying outside the fences, which are located at Q1  1.5(Q3  Q1) and Q3+1.5(Q3  Q1). 
+The whiskers show the range of points that are outside the box but inside the fences.
+\newline
 Due to the limitation of resources and time we had to omit four source projects from these results. 
@@
 Trying to cluster these mutants this resulted in a out of memory error.
+\newline
+The results displayed are the average of the results of the 30 repetitions of the experiment.
+The result of each single repetition can be found in our Github repository\cite{rbasarat-repo}.
 
@@
 This section displays the results of our clustering algorithm when executed with data that consists of all characteristics.
-First we show the results to test our null hypothesis discussed in Chapter \ref{ch:mutant_selection}.
 The results are be grouped based on the percentage in reduction of mutants executed.
@@
 The result of each single repetition can be found in our Github repository\cite{rbasarat-repo}.
+
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{images/boxplot_summary/boxplot_hc_full_0.25.png}
+\caption{\label{box:clustering_all_25}Box-plots containing weighted mutation score of clustering mutants with all characteristics and cluster size of n*0.25}
+\end{figure}
 
@@
 
+Figure \ref{box:clustering_all_25} displays the box-plots of the weighted mutation score obtained from each individual sample.
+We can observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
 Table \ref{tab:clustering_all_25} displays the results of clustering mutants with all characteristics and the number of clusters equal to the total amount of mutants* 0.25.  
@@
 If the percentage does go below 50\% the representation of a cluster will flip from survived to killed or the other way around.
+
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{images/boxplot_summary/boxplot_hc_full_0.5.png}
+\caption{\label{box:clustering_all_50}Box-plots containing weighted mutation score of clustering mutants with all characteristics and cluster size of n*0.5}
+\end{figure}
 
@@
 
+Figure \ref{box:clustering_all_50} displays the box-plots of the weighted mutation score obtained from each individual sample.
+We can observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
 Table \ref{tab:clustering_all_50} displays the results of clustering mutants with all characteristics and the number of clusters equal to the total amount of mutants* 0.50.  
@@
 The average of the differences between full set score and clustered score is 0.48\% which is smaller than the average difference than that of the 25\% set.
+
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{images/boxplot_summary/boxplot_hc_full_0.75.png}
+\caption{\label{box:clustering_all_75}Box-plots containing weighted mutation score of clustering mutants with all characteristics and cluster size of n*0.75}
+\end{figure}
 
@@
 
+Figure \ref{box:clustering_all_75} displays the box-plots of the weighted mutation score obtained from each individual sample.
+We can observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
 Table \ref{tab:clustering_all_75} displays the results of clustering mutants with all characteristics and the number of clusters equal to the total amount of mutants* 0.75.  
@@
 The average of the differences between full set score and clustered score is 0.48\% which is higher than the average difference than that of the 25\% and the same of that of the 50\% set.
+
+%---------------- from here no levenshtein ---------------------------%
 
@@
 The result of each single repetition can be found in our Github repository\cite{rbasarat-repo}.
+
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{images/boxplot_summary/boxplot_hc_no_distance_0.25.png}
+\caption{\label{box:clustering_no_distance_25}Box-plots containing weighted mutation score of clustering mutants without Levensthein distance and cluster size n*0.25}
+\end{figure}
+
 \begin{table}[htb]
@@
 
+Figure \ref{box:clustering_no_distance_25} displays the box-plots of the weighted mutation score obtained from each individual sample.
+We can observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
 Table \ref{tab:clustering_no_distance_25} displays the results of clustering mutants without Levensthein distance and the number of clusters equal to the total amount of mutants * 0.25. 
@@
 The average of the differences between full set score and clustered score is 0.53\%.
+
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{images/boxplot_summary/boxplot_hc_no_distance_0.5.png}
+\caption{\label{box:clustering_no_distance_50}Box-plots containing weighted mutation score of clustering mutants without Levensthein distance and cluster size n*0.50}
+\end{figure}
 
@@
 
+Figure \ref{box:clustering_no_distance_50} displays the box-plots of the weighted mutation score obtained from each individual sample.
+We can observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
 Table \ref{tab:clustering_no_distance_50} displays the results of clustering mutants without Levensthein distance and the number of clusters equal to the total amount of mutants * 0.50.
@@
 The average of the differences between full set score and clustered score is 0.49\% which is smaller than the average difference than that of the 25\% set.
+
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{images/boxplot_summary/boxplot_hc_no_distance_0.75.png}
+\caption{\label{box:clustering_no_distance_75}Box-plots containing weighted mutation score of clustering mutants without Levensthein distance and cluster size n*0.75}
+\end{figure}
 
@@
 
+Figure \ref{box:clustering_no_distance_75} displays the box-plots of the weighted mutation score obtained from each individual sample.
+We can observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
 Table \ref{tab:clustering_no_distance_75} displays the results of clustering mutants without Levensthein distance and the number of clusters equal to the total amount of mutants * 0.50.

### 2021-07-19 chapters/results/1_results_rq1.tex
wip statistics

@@
 We can observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
+To view the box-plots separately and in more detail see Appendix \ref{ap:full_25}.
 Table \ref{tab:clustering_all_25} displays the results of clustering mutants with all characteristics and the number of clusters equal to the total amount of mutants* 0.25.  
@@
 We can observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
+To view the box-plots separately and in more detail see Appendix \ref{ap:full_50}.
 Table \ref{tab:clustering_all_50} displays the results of clustering mutants with all characteristics and the number of clusters equal to the total amount of mutants* 0.50.  
@@
 We can observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
+To view the box-plots separately and in more detail see Appendix \ref{ap:full_75}.
 Table \ref{tab:clustering_all_75} displays the results of clustering mutants with all characteristics and the number of clusters equal to the total amount of mutants* 0.75.  
@@
 We can observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
+To view the box-plots separately and in more detail see Appendix \ref{ap:no_distance_25}.
 Table \ref{tab:clustering_no_distance_25} displays the results of clustering mutants without Levensthein distance and the number of clusters equal to the total amount of mutants * 0.25. 
@@
 We can observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
+To view the box-plots separately and in more detail see Appendix \ref{ap:no_distance_50}.
 Table \ref{tab:clustering_no_distance_50} displays the results of clustering mutants without Levensthein distance and the number of clusters equal to the total amount of mutants * 0.50.
@@
 We can observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
+To view the box-plots separately and in more detail see Appendix \ref{ap:no_distance_75}.
 Table \ref{tab:clustering_no_distance_75} displays the results of clustering mutants without Levensthein distance and the number of clusters equal to the total amount of mutants * 0.50.

### 2021-07-19 appendix/appendix.tex

@@
 
+\chapter{Box-plots of samples per project with all characteristics n*0.25 reduction}
+\label{ap:full_25}
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.32]{images/full_25/boxplot_google-auto-service.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.32]{images/full_25/boxplot_scribejava-core.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/full_25/boxplot_google-auto-factory.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/full_25/boxplot_google-auto-common.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/full_25/boxplot_google-auto-value.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/full_25/boxplot_gson.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/full_25/boxplot_commons-io.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/full_25/boxplot_commons-cli.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/full_25/boxplot_commons-text.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/full_25/boxplot_commonc-codec.png}}}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.3]{images/full_25/boxplot_commons-csv.png}
+\end{figure}
+
+\chapter{Box-plots of samples per project with all characteristics n*0.50 reduction}
+\label{ap:full_50}
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.32]{images/full_50/boxplot_google-auto-service.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.32]{images/full_50/boxplot_scribejava-core.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/full_50/boxplot_google-auto-factory.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/full_50/boxplot_google-auto-common.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/full_50/boxplot_google-auto-value.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/full_50/boxplot_gson.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/full_50/boxplot_commons-io.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/full_50/boxplot_commons-cli.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/full_50/boxplot_commons-text.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/full_50/boxplot_commonc-codec.png}}}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.3]{images/full_50/boxplot_commons-csv.png}
+\end{figure}
+
+\chapter{Box-plots of samples per project with all characteristics n*0.75 reduction}
+\label{ap:full_75}
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.32]{images/full_75/boxplot_google-auto-service.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.32]{images/full_75/boxplot_scribejava-core.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/full_75/boxplot_google-auto-factory.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/full_75/boxplot_google-auto-common.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/full_75/boxplot_google-auto-value.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/full_75/boxplot_gson.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/full_75/boxplot_commons-io.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/full_75/boxplot_commons-cli.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/full_75/boxplot_commons-text.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/full_75/boxplot_commonc-codec.png}}}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.3]{images/full_75/boxplot_commons-csv.png}
+\end{figure}
+
+\chapter{Box-plots of samples per project without Levenshtein distance characteristic n*0.25 reduction}
+\label{ap:no_distance_25}
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.32]{images/no_distance_25/boxplot_google-auto-service.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.32]{images/no_distance_25/boxplot_scribejava-core.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_25/boxplot_google-auto-factory.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_25/boxplot_google-auto-common.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_25/boxplot_google-auto-value.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_25/boxplot_gson.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_25/boxplot_commons-io.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_25/boxplot_commons-cli.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_25/boxplot_commons-text.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_25/boxplot_commonc-codec.png}}}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.3]{images/no_distance_25/boxplot_commons-csv.png}
+\end{figure}
+
+\chapter{Box-plots of samples per project without Levenshtein distance characteristic n*0.50 reduction}
+\label{ap:no_distance_50}
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.32]{images/no_distance_50/boxplot_google-auto-service.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.32]{images/no_distance_50/boxplot_scribejava-core.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_50/boxplot_google-auto-factory.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_50/boxplot_google-auto-common.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_50/boxplot_google-auto-value.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_50/boxplot_gson.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_50/boxplot_commons-io.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_50/boxplot_commons-cli.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_50/boxplot_commons-text.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_50/boxplot_commonc-codec.png}}}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.3]{images/no_distance_50/boxplot_commons-csv.png}
+\end{figure}
+
+\chapter{Box-plots of samples per project without Levenshtein distance characteristic n*0.75 reduction}
+\label{ap:no_distance_75}
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.32]{images/no_distance_75/boxplot_google-auto-service.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.32]{images/no_distance_75/boxplot_scribejava-core.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_75/boxplot_google-auto-factory.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_75/boxplot_google-auto-common.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_75/boxplot_google-auto-value.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_75/boxplot_gson.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_75/boxplot_commons-io.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_75/boxplot_commons-cli.png}}}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_75/boxplot_commons-text.png}}}
+    \qquad
+    \subfloat{{\includegraphics[scale=0.3]{images/no_distance_75/boxplot_commonc-codec.png}}}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.3]{images/no_distance_75/boxplot_commons-csv.png}
+\end{figure}
+
 \end{appendices}

### 2021-07-19 chapters/9_discussion.tex
wip until now

@@
 The maximum deviation from the original mutation score, for all variants(with and without Levenshtein distance) of the experiment, is below the statistical threshold.
+We can observe that the distance between the whiskers of the box-plots do not stretch more than the \textit{alpha-value} of 0.05.
+This means that there is not one sample that deviated more than 5\% from the original mutation score.
+Thus we can reject the null hypothesis and accept our alternative hypothesis.
 As hypothesised we observe that it is possible to cluster mutants with the set of characteristics we identified while maintaining effectiveness and reducing the amount executed.
@@
 The maximum deviation from the original mutation score, for all variants(with and without Levenshtein distance characteristic) of the experiment, is above the statistical threshold.
+We can observe that the distance between whiskers of the box-plots is more than the \textit{alpha-value} of 0.05.
+Thus we cannot reject the null hypothesis and accept our alternative hypothesis.
 The biggest deviation measured is 22.39\%.

### 2021-07-20 appendix/appendix.tex
background

@@
 
-\chapter{Box-plots of samples per project with all characteristics n*0.25 reduction}
+\chapter{Box plots}
+\section{Box-plots of samples per project with all characteristics with n*0.25 reduction}
 \label{ap:full_25}
@@
 
-\chapter{Box-plots of samples per project with all characteristics n*0.50 reduction}
+\section{Box-plots of samples per project with all characteristics with n*0.50 reduction}
 \label{ap:full_50}
@@
 
-\chapter{Box-plots of samples per project with all characteristics n*0.75 reduction}
+\section{Box-plots of samples per project with all characteristics with n*0.75 reduction}
 \label{ap:full_75}
@@
 
-\chapter{Box-plots of samples per project without Levenshtein distance characteristic n*0.25 reduction}
+\section{Box-plots of samples per project without Levenshtein distance characteristic with n*0.25 reduction}
 \label{ap:no_distance_25}
@@
 
-\chapter{Box-plots of samples per project without Levenshtein distance characteristic n*0.50 reduction}
+\section{Box-plots of samples per project without Levenshtein distance characteristic with n*0.50 reduction}
 \label{ap:no_distance_50}
@@
 
-\chapter{Box-plots of samples per project without Levenshtein distance characteristic n*0.75 reduction}
+\section{Box-plots of samples per project without Levenshtein distance characteristic with n*0.75 reduction}
 \label{ap:no_distance_75}

### 2021-07-20 chapters/3_background.tex
background

@@
 Rodriguez et al., compared the performance of nine different clustering algorithms:
-\begin{itemize}
-    \item k-means
-    \item CLARA
-    \item hierarchical
-    \item EM
-    \item hcmodel
-    \item spectral
-    \item subscpace
-    \item optics
-    \item dbscan
-\end{itemize}
+
+\subsubsection{K-means} 
+The K-means algorithm identifies \textit{k} number of centroids, and then allocates every data point to the nearest cluster, while keeping the centroids as small as possible\cite{Hartigan1979AlgorithmAlgorithm}.
+\subsubsection{CLARA} 
+CLARA uses multiple fixed samples of the data set to minimize sampling bias and, select the best medoids among those samples\cite{Sarle1991FindingAnalysis.}.
+A medoid is defined as the object \textit{i} for which the average dissimilarity to all other objects in its cluster is minimal\cite{Sarle1991FindingAnalysis.}.
+\subsubsection{Hierarchical clustering} 
+Hierarchical clustering groups similar objects into groups called clusters\cite{Fraley1998HowAnalysis}. 
+The endpoint is a set of clusters or groups, where each cluster is distinct from each other cluster, and the objects within each cluster are broadly similar to each other\cite{Fraley1998HowAnalysis}.
+\subsubsection{EM} 
+EM(expectation maximization) clustering technique is similar to the K-Means technique\cite{Fv1996TheAlgorithm}.
+Instead of assigning examples to clusters to maximize the differences in means for continuous variables, the EM clustering algorithm computes probabilities of cluster memberships based on one or more probability distributions\cite{Fv1996TheAlgorithm}.
+\subsubsection{Spectral clustering}
+Spectral clustering is a technique with roots in graph theory.
+It identifies communities of nodes in a graph based on the edges connecting them\cite{Filippone2008AClustering}.
+Spectral clustering uses information from the eigenvalues of special matrices built from the graph or the data set  to perform dimensionality reduction before clustering in fewer dimensions\cite{Filippone2008AClustering}.
+The similarity matrix is provided as an input and consists of a quantitative assessment of the relative similarity of each pair of points in the data set\cite{Filippone2008AClustering}.
+\subsubsection{Subspace}
+Subspace clustering algorithms consider the similarity between objects with respect to distinct subsets of the attributes\cite{Hartigan1979AlgorithmAlgorithm}. 
+Different subsets of the attributes might define distinctions between each other. 
+The algorithm can identify clusters that exist in multiple, possibly overlapping, sub spaces\cite{M2016SUBSPACEDATA}.
+\subsubsection{Dbscan}
+Dbscan is a density-based clustering non-parametric algorithm\cite{Daszykowski2009}.
+Given a set of points in some space, it groups together points that are closely packed together, marking outliers as points that lie alone in low-density regions\cite{Daszykowski2009}.
+\subsubsection{OPTICS}
+The OPTICS(Ordering Points To Identify the Clustering Structure) algorithm starts with a data point and expands its neighborhood using a similar procedure as in the Dbscan algorithm\cite{Ankerst1999OPTICS:Structure}.
+The difference is that the neighborhood is first expanded to points with low core-distance\cite{Ankerst1999OPTICS:Structure}. 
+The optics algorithm can detect clusters having large density variations and irregular shapes\cite{Ankerst1999OPTICS:Structure}.
+\newline
+
 The goal of their study was to guide researchers, who have little experience in data mining techniques, to the application of clustering methods.

### 2021-07-20 chapters/rq1/2_clustering_mutants.tex
minor changes
background

@@
 \subsection{Hierarchical clustering}
-Clustering Levenshtein distances has been done before with hierarchical clustering\cite{Rajalingam2011, Gothai2010PerformanceAlgorithms}. 
+Clustering Levenshtein distances has been done before with hierarchical clustering with promising results.\cite{Rajalingam2011, Gothai2010PerformanceAlgorithms}. 
 Research shows that hierarchical clustering performs better when clustering with at least ten features\cite{Rodriguez2019}.
@@
 We adjust the parameters of the algorithm based on the characteristics in our data set.
-Bases on these observations we select hierarchical clustering to cluster the mutants.
+Bases on these observations and the research of Rodriguez et al., we select hierarchical clustering as it gives us the best performance in terms of partitioning.
 Next we will explain the configuration we use for clustering the mutants. 
@@
 To translate this to cluster size we will reduce the number of clusters in steps op 25\%.
-In other words by reducing the amount of mutants, for example by half, we increase the performance by 50\%\cite{Just2014}.
+In other words by reducing the amount of mutants, for example by half, we increase the performance by 50\%\cite{Just2014AreTesting}.
 If necessary we can decide to cut off the clustering at different points.

### 2021-07-20 chapters/2_introduction.tex
wip restructure
dflakshdasgiudjsa

@@
 \label{ch:introduction}
-Software testing is a critical part in the software engineering process. It is widely used and accepted in the industry for quality assurance. 
+Software testing is a critical part in the software engineering process. 
+It is a standardized in the industry and is used for quality assurance\cite{ISOtesting}. 
 Software testing can detect software bugs during the development process and can also serve for regression purposes\cite{Bertolino2007SoftwareDreams}. 
@@
 We will do this with two different approaches: qualitative and quantitative. 
-We structure our research as follows:
+We structure our research on the following research questions:
 \newline
@@
 \section{Outline}
-In Chapter~\ref{ch:background} we describe the background of this thesis. 
+In Chapter~\ref{ch:background} we describe the background information of this thesis. 
 Chapter \ref{ch:mutation-tool} compares the different mutation testing tools to decide on a tool to use for this research.

### 2021-07-20 chapters/10_related_work.tex
writing style issue

@@
 Clustering mutants has been done before.
-The research that exists tries to cluster mutants by defining the centroid of a cluster and cluster the mutants on that definition.
+The research that exists tries to cluster mutants by defining the centroid of a cluster and cluster mutants on that definition.
 In our research extends mutation clustering by defining a centroid that is dynamic.
@@
 \section{Clustering similarity}
-Hussain et al., used the k-means and agglomerative clustering algorithm to cluster the mutants according to a similarity measure. 
+Hussain et al., used the k-means and agglomerative clustering algorithm to cluster mutants according to a similarity measure. 
 They used the Hamming distance as similarity measure.
@@
 Hussain et al, had the same goals as our research. 
-The difference is that our research used different and more elaborated methods to cluster the mutants.
+The difference is that our research used different and more elaborated methods to cluster mutants.
 Hussain et al., calculated the mutation score by counting one cluster as one mutant.

### 2021-07-20 chapters/2_introduction.tex
writing style issue

@@
 A technique to reduce the number of mutants or the number of test case executions is mutation clustering.
-Mutation clustering aims to reduce the amount of mutants to be executed by clustering the mutants\cite{Ma2016,Yu2019PossibilityScope}.
+Mutation clustering aims to reduce the amount of mutants to be executed by clustering mutants\cite{Ma2016,Yu2019PossibilityScope}.
 Clustering is a technique used for the grouping of data, for pattern recognition and classification, based on some similar characteristics\cite{Alsabti1997}.
@@
 In chapter \ref{ch:reserach_question1} we identify characteristics to represent mutants and design an experiment to cluster them.
-In chapter \ref{ch:reserach_question2} we select a machine learning model and design an experiment to cluster the mutants with this model.
+In chapter \ref{ch:reserach_question2} we select a machine learning model and design an experiment to cluster mutants with this model.
 Results are shown in Chapters \ref{ch:results_rq1} and \ref{ch:results_rq2} for the experiments described in Chapters \ref{ch:reserach_question1} and \ref{ch:reserach_question2} respectively. 

### 2021-07-20 chapters/7_experiment_design.tex
wip statistics

@@
 In this chapter we design an experiment to validate our research.
+We will execute the following steps:
+\begin{enumerate}
+    \item Gather all characteristics from PIT.
+    \item Execute full set of mutants to gather baseline measurement.
+    \item Cluster the mutants according the black and white box methodologies.
+    \item Execute one random mutant per cluster and gather result.
+    \item Calculate weighted mutation score.
+\end{enumerate}
 We hypothesise that we can cluster mutants with the set of characteristics we identified
@@
 To validate our sample we make use of statistical hypothesis testing\cite{Emmert-Streib2019UnderstandingInference}.
-Our null hypothesis states that there is no relation between the characteristics identified and the results of a mutant.
+Our null hypothesis states that there is no relation between the characteristics identified and the results of a mutant with the chosen methodology.
 To test this hypothesis we select an alpha value of 0.05.

### 2021-07-20 chapters/rq1/2_clustering_mutants.tex

@@
 
-\subsection{Cluster size}
-Agglomerative hierarchical clustering will continue to cluster until there is one cluster left. 
+\subsection{Number of clusters}
+Agglomerative hierarchical clustering can continue to cluster until there is one cluster left. 
 Naturally this cluster will contain all the mutants.
@@
 We decide on the number of clusters based on the amount of mutants generated in a full set.
-The reasoning for the selecting the number of clusters will be from a performance reducing perspective.
+We select the number of clusters from a performance increasing perspective.
 Mutation clustering increases the performance by reducing the amount of mutants executed\cite{Pizzoleto2019}.
-Reducing the performance by a quarter, half or three quarters will give an indication of performance increases.
-To translate this to cluster size we will reduce the number of clusters in steps op 25\%.
-In other words by reducing the amount of mutants, for example by half, we increase the performance by 50\%\cite{Just2014AreTesting}.
-If necessary we can decide to cut off the clustering at different points.
-The baseline will consist of three experiments where the cluster size is 25\%, 50\% and 75\% of the total amount of mutants respectively.
+By reducing the amount of mutants, for example by half, we increase the performance by 50\%\cite{Just2014AreTesting}.
+As a starting point we will perform three experiments where the cluster size is 25\%, 50\% and 75\% of the total amount of mutants respectively.
+We can evaluate the results and can decide to cut off the clustering at different points.
 The amount of mutants inside a cluster will be decided by the clustering algorithm.

### 2021-07-20 chapters/7_experiment_design.tex
ping

@@
 \begin{enumerate}
-    \item Gather all characteristics from PIT.
+    \item Gather all mutant characteristics from PIT.
     \item Execute full set of mutants to gather baseline measurement.
@@
 \newline
-If the weighted mutation score of our clustered set deviates more than 5\% from the score of a full set we will reject our hypothesis.
+If the weighted mutation score of our clustered set deviates more than 5\% from the score of a full set we reject our hypothesis.
 

### 2021-07-20 chapters/rq1/2_clustering_mutants.tex
pong

@@
 Bases on these observations and the research of Rodriguez et al., we select hierarchical clustering as it gives us the best performance in terms of partitioning.
-Next we will explain the configuration we use for clustering the mutants. 
+Next we explain the configuration we use for clustering the mutants. 
 \newline
@@
 Since we cannot assume that our clusters are well separated we chose to use the ward linkage method for our clustering algorithm.
-In summary, the agglomerative hierarchical clustering algorithm with ward linkage will cluster our mutants represented by the characteristics we gathered.
+In summary, we select the agglomerative hierarchical clustering algorithm with ward linkage to cluster mutants represented by the characteristics we gathered.
 
@@
 This gives us a Levenshtein distance that represents more of the semantic difference, between a mutant and parent, than calculating the distance for Java code.
-For this characteristic we will use the Levenshtein distance between a Java byte code mutant and its parent.
+For this characteristic we use the Levenshtein distance between a Java byte code mutant and its parent.
 
@@
 By reducing the amount of mutants, for example by half, we increase the performance by 50\%\cite{Just2014AreTesting}.
-As a starting point we will perform three experiments where the cluster size is 25\%, 50\% and 75\% of the total amount of mutants respectively.
+As a starting point we perform three experiments where the cluster size is 25\%, 50\% and 75\% of the total amount of mutants respectively.
 We can evaluate the results and can decide to cut off the clustering at different points.
-The amount of mutants inside a cluster will be decided by the clustering algorithm.
+The amount of mutants inside a cluster is be decided by the clustering algorithm.
 \end{document}

### 2021-07-20 chapters/rq2/1_clustering_mutants.tex
writing style issue
pong

@@
 \section{Selecting a machine learning model}
-In this section we will select a machine learning model.
-We will first choose between supervised or unsupervised learning.
-Our starting point for selecting a machine learning model are the models discussed in the survey done by K.L. Du.
-Based on the choice of supervised or unsupervised learning we will review the corresponding models and decide if they fit our use case by a process of elimination.
+In this section we select a machine learning model.
+We start by choosing between supervised or unsupervised learning.
+Then we look at the machine learning models that are discussed in the survey done by K.L. Du.
+Based on the choice of supervised or unsupervised learning we review the corresponding models and decide if they fit our use case by a process of elimination.
 
 \subsection{Supervised or unsupervised learning}
-In this subsection we will decide on supervised or unsupervised learning.
+In this subsection we decide on supervised or unsupervised learning.
 The application and description of these two types of learning are described in Chapter \ref{ch:background}.
@@
 Our data does not contain binary input patterns and it is also not rapidly changing.
-Therefore we will not select an ART model.
+Therefore we do not select an ART model.
 
@@
 For this reason we cannot use the neural gas network to predict in what cluster a new mutant should be placed.
-Therefore we will not select the neural gas network.
+Therefore we do not select the neural gas network.
 
@@
 The learning vector quantization model is a supervised learning model.
-We established that we need an unsupervised learning model to cluster the mutants.
-Therefore we will not select the learning vector quantization model.
+We established that we will use an unsupervised learning model to cluster mutants.
+Therefore we do not select the learning vector quantization model.
 
@@
 The \acrfull{som}, C-means and \acrfull{fcm} models are all models that would fit our purpose. 
-To choose between these models we will look at the performance of each model.
+To choose between these models we look at the performance of each model.
 Mingoti et al., compared the performance of these three models \cite{Mingoti2006ComparingAlgorithms}.
@@
 Research has shown that if the data set is big enough a ratio of 80/20 is sufficient\cite{Guyon1997ARatio}.
-We will divide each source projects into training and validation groups to reflect the 80/20 ratio as best as possible.
+We divide each source projects into training and validation groups to reflect the 80/20 ratio as best as possible.
 Depending on the amount of mutants per project we might not achieve a perfect 80/20 ratio.
-The division of the mutants will be done randomly. 
+The division of the mutants is be done randomly. 
 To make our results reproducible we select a random division based on a generated seed.
@@
 It approximates the centroids and amount of the clusters based on the data points in the data set\cite{Chiu1994FuzzyEstimation}.
-We will select and implement the subtractive clustering model to estimate the number of clusters.
-We will use this number as the value for the number of clusters parameter in the \acrshort{fcm} algorithm.
+We select and implement the subtractive clustering model to estimate the number of clusters.
+We use this number as the value for the number of clusters parameter in the \acrshort{fcm} algorithm.
 
 \section{Implementation of subtractive clustering}
-For the same reasons as in Chapter \ref{ch:fcm_implementation} we will select a library that implements the subtractive clsutering algorithm.
+For the same reasons as in Chapter \ref{ch:fcm_implementation} we select a library that implements the subtractive clsutering algorithm.
 The research paper of the mountain and subtractive algorithms are referenced in the survey paper of Du et al\cite{Du2010Clustering:Approach}.

### 2021-07-21 chapters/rq2/1_clustering_mutants.tex
minor change

@@
 The implementation is professionally build and maintained by Matlab\cite{matlabSubtractive}.
-Matlab is a tool commonly used in the academic world and is professionally maintained for this reason we decided to use their implementation of subtractive clustering.
+Matlab is a tool specifically designed for academic world and is professionally maintained for this reason we decided to use their implementation of subtractive clustering.
 

### 2021-07-21 chapters/rq2/1_clustering_mutants.tex

--- 2021-07-20 chapters/rq2/1_clustering_mutants.tex
@@
 \section{Implementation of subtractive clustering}
-For the same reasons as in Chapter \ref{ch:fcm_implementation} we select a library that implements the subtractive clsutering algorithm.
-The research paper of the mountain and subtractive algorithms are referenced in the survey paper of Du et al\cite{Du2010Clustering:Approach}.
+For the same reasons as in Chapter \ref{ch:fcm_implementation} we select a library that implements the subtractive clustering algorithm.
+The research paper of the mountain and subtractive algorithms are referenced in the survey paper of Du et al.\cite{Du2010Clustering:Approach}.
 Based on the paper of the theory of the subtractive algorithm\cite{Chiu1994FuzzyEstimation} we found an existing implementation\cite{matlabSubtractive}.
 The implementation is professionally build and maintained by Matlab\cite{matlabSubtractive}.
-Matlab is a tool commonly used in the academic world and is professionally maintained for this reason we decided to use their implementation of subtractive clustering.
+Matlab is a tool designed for the academic world and is professionally maintained for this reason we decided to use their implementation of subtractive clustering\cite{matlab}.
 

### 2021-07-21 chapters/3_background.tex
move model description to background

@@
     \item Mountain and subtractive clustering.
+    The mountain and subtractive clustering models are an effective method for estimating the number of clusters\cite{Du2010Clustering:Approach}. 
+The subtractive clustering model is an extension of the mountain model\cite{Du2010Clustering:Approach}. 
+\newline
+The method grids the data space and computes a potential value for each grid point based on its distance to the actual data points. Each grid point is a potential cluster center. 
+The potential for each grid is calculated based on the density of the surrounding data points. 
+The grid with the highest potential is selected as the first cluster center and then the potential values of all the other grids are reduced according to their distances to the first cluster center.
     \item Neural gas.
+    The neural gas is a topology-preserving network, and can be treated as an extension to the C-means\cite{Du2010Clustering:Approach}.
+A data optimal topological ordering is achieved by using neighborhood ranking within the input space at each training step\cite{Du2010Clustering:Approach}.
+To find its neighborhood rank, each neuron compares its distance to the input vector with those of all the other neurons to the input vector\cite{Du2010Clustering:Approach}. 
+Unlike the \acrlong{som}(see Chapter \ref{ch:topThreeModels}, which uses predefined static neighborhood relations, the neural gas network determines a dynamical neighborhood relation as learning proceeds\cite{Du2010Clustering:Approach}.
+A benefit of the neural gas network is that it is able to spawn neurons.
+This means that it is not necessary to know the number of clusters in advance.
     \item ART networks.
+    ART models are characterized by systems of differential equations that formulate stable self-organizing learning methods\cite{Du2010Clustering:Approach}.
+The stability and plasticity properties as well as the ability to efficiently process dynamic data make the ART attractive for clustering large, rapidly changing sequences of input patterns\cite{Du2010Clustering:Approach}.
     \item Fuzzy c-means.

### 2021-07-21 chapters/rq2/1_clustering_mutants.tex
move model description to background

@@
 \subsection{Mountain and subtractive clustering}
-The mountain and subtractive clustering models are an effective method for estimating the number of clusters\cite{Du2010Clustering:Approach}. 
-The subtractive clustering model is an extension of the mountain model\cite{Du2010Clustering:Approach}. 
-\newline
-The method grids the data space and computes a potential value for each grid point based on its distance to the actual data points.
-Each grid point is a potential cluster center.
-The potential for each grid is calculated based on the density of the surrounding data points.
-The grid with the highest potential is selected as the first cluster center and then the potential values of all the other grids are reduced according to their distances to the first cluster center.
-\newline
+As described in Chapter \ref{ch:neuralNetworkSurvey} the mountain and subtractive clustering models calculate the amount of clusters needed for a specific data set.
 Subtractive clustering is designed for learning fuzzy systems models from data\cite{Chiu1994FuzzyEstimation}. 
@@
 \subsection{ART models}
-ART models are characterized by systems of differential equations that formulate stable self-organizing learning methods\cite{Du2010Clustering:Approach}.
-The stability and plasticity properties as well as the ability to efficiently process dynamic data make the ART attractive for clustering large, rapidly changing sequences of input patterns\cite{Du2010Clustering:Approach}.
 While the ART models are good at what they do, they do not fit our purpose of clustering mutants.
-As described above the ART models expect rapidly changing sequences of input patterns.
+As described in Chapter \ref{ch:neuralNetworkSurvey} the ART models expect rapidly changing sequences of input patterns.
 Our data does not contain binary input patterns and it is also not rapidly changing.
@@
 \subsection{Neural gas}
-The neural gas is a topology-preserving network, and can be treated as an extension to the C-means\cite{Du2010Clustering:Approach}.
-A data optimal topological ordering is achieved by using neighborhood ranking within the input space at each training step\cite{Du2010Clustering:Approach}.
-To find its neighborhood rank, each neuron compares its distance to the input vector with those of all the other neurons to the input vector\cite{Du2010Clustering:Approach}. 
-Unlike the \acrlong{som}(see Chapter \ref{ch:topThreeModels}, which uses predefined static neighborhood relations, the neural gas network determines a dynamical neighborhood relation as learning proceeds\cite{Du2010Clustering:Approach}.
-A benefit of the neural gas network is that it is able to spawn neurons.
-This means that it is not necessary to know the number of clusters in advance.
-Unfortunately the neural gas network is a learning network.
+The neural gas network is a learning network.
 One of the properties of a learning network is that it is not designed to fit new data points\cite{supervisedUnsupervised}.

### 2021-08-04 chapters/7_experiment_design.tex
last feedback

@@
 In this chapter we design an experiment to validate our research.
-We will execute the following steps:
+We execute the following steps:
 \begin{enumerate}
@@
 \end{enumerate}
+Both research questions have the same goal, but with different approaches.
+We can devise one hypothesis for both research questions.
 We hypothesise that we can cluster mutants with the set of characteristics we identified
@@
 Our goal is to achieve a weighted mutation score that is as close as possible to the mutation score of a full set of executed mutants.
-We will repeat the experiment for each approach we devised in Chapters \ref{ch:reserach_question1} and \ref{ch:reserach_question2}.
+We repeat the experiment for each approach we devised in Chapters \ref{ch:reserach_question1} and \ref{ch:reserach_question2}.
 
@@
 To translate; if more than five percent of the values in a sample deviates more than the significance level(see Chapter \ref{ch:exp1_validation}) we cannot reject the null hypothesis.
-To make our results reproducible we will select the random mutant based on a generated seed.
+To make our results reproducible we select the random mutant based on a generated seed.
 The seed generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.
-We will repeat our experiment 30 times with 30 different seeds.
-Achieving consistent results while applying random selection will contribute to the validity of the experiment.
+We repeat our experiment 30 times with 30 different seeds.
+Achieving consistent results while applying random selection contributes to the validity of the experiment.
 

### 2021-08-04 chapters/results/1_results_rq1.tex
last feedback

@@
 The results displayed are the average of the results of the 30 repetitions of the experiment.
-The result of each single repetition can be found in our Github repository\cite{rbasarat-repo}.
+The result of each single repetition can be found in our Github repository in the folder \textit{experiment_results}\cite{rbasarat-repo}.
 
@@
 \subsection{Clustering without Levenshtein distance}
-This section will display the results of our clustering algorithm when executed without the Levenshtein distance.
-The results will be grouped based on the reduced percentage of mutants executed.
-We will also display the accuracy of the clusters per reduction amount.
+This section displays the results of our clustering algorithm when executed without the Levenshtein distance.
+The results are grouped based on the reduced percentage of mutants executed.
+We also display the accuracy of the clusters per reduction amount.
 The results displayed are the average of the results of the 30 repetitions of the experiment.
-The result of each single repetition can be found in our Github repository\cite{rbasarat-repo}.
+The result of each single repetition can be found in our Github repository in the folder \textit{experiment_results}\cite{rbasarat-repo}.
 

### 2021-08-04 chapters/results/1_results_rq1.tex
syntax error

@@
 The results displayed are the average of the results of the 30 repetitions of the experiment.
-The result of each single repetition can be found in our Github repository in the folder \textit{experiment_results}\cite{rbasarat-repo}.
+The result of each single repetition can be found in our Github repository in the folder \textit{experiment{_}results}\cite{rbasarat-repo}.
 
@@
 The results displayed are the average of the results of the 30 repetitions of the experiment.
-The result of each single repetition can be found in our Github repository in the folder \textit{experiment_results}\cite{rbasarat-repo}.
+The result of each single repetition can be found in our Github repository in the folder\textit{experiment{\_}results}\cite{rbasarat-repo}.
 

### 2021-11-24 mutation-workshop/paper.tex
Added paper

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
% \usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{makecell}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{hyperref}

\usepackage[flushleft]{threeparttable}
\usepackage{tabularx}

\usepackage[disable]{todonotes}
\usepackage{enumitem}

\usepackage[backend=biber, style=ieee, citestyle=numeric-comp,uniquelist=false, maxcitenames=2, maxbibnames=9, mincitenames=1]{biblatex}
\bibliography{../refs/main}

\usepackage{listings}
\lstset{
	basicstyle=\footnotesize,        % the size of the fonts that are used for the code
	breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
	breaklines=true,                 % sets automatic line breaking
	captionpos=b,                    % sets the caption-position to bottom
	frame=single,                    % adds a frame around the code
	language=Java,                 % the language of the code
	keywordstyle=\bf,
	tabsize=2                       % sets default tabsize to 2 spaces
}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\def\code#1{\texttt{#1}}
\begin{document}

\title{Mutation testing: Clustering the mutants}

\author{
\IEEEauthorblockN{Ana Oprescu}
\IEEEauthorblockA{\textit{Parallel Computing Systems} \\
\textit{University of Amsterdam}\\
Amsterdam, the Netherlands \\
a.m.oprescu@uva.nl}

\and
\IEEEauthorblockN{Rasjaad Basarat}
\IEEEauthorblockA{\textit{Parallel Computing Systems} \\ %% TODO: CHANGE THIS
\textit{University of Amsterdam}\\
Amsterdam, the Netherlands \\
rasjaad@nivero.io}
\and
\IEEEauthorblockN{Thomas Biesaart}
\IEEEauthorblockA{\textit{Parallel Computing Systems} \\ %% TODO: CHANGE THIS
Amsterdam, the Netherlands \\
thomas@emailhere.com}
}

\maketitle
% \todo[inline]{FROM REVIEWER: Make sure to be consistent with singular/plural (e.g. mutation operators will result in mutants, not just a mutant)}
% \todo[inline]{FROM REVIEWER:  I don't quite follow the discussion of the execution time problem. After all, even in a decentralized approach I can run multiple approaches on the same hardware and thus make a fair comparison. I suppose the point you're trying to make is comparison *across multiple papers/studies"? Please clarify.}
\begin{abstract}
Mutation testing is a computationally expensive technique that requires the ability to run tests.
Different techniques have been proposed to reduce the cost of mutation testing.
One of these techniques is the clustering of mutants.
By clustering mutants we can execute less mutants to reduce the cost.
Different techniques have been proposed to reduce the cost of mutation testing.
One of these techniques is the clustering of mutants.
Our research consists of two parts in which we cluster mutants.
To cluster mutants we need a mutation testing tools.
Three mutation testing tools for Java are compared and one is selected to use for this research
The first part consists of a white box approach and the second part of a black box approach.
Our research uses the mutation testing tool PIT.
We identified characteristics to represent the mutants such that we can cluster them.
For our white box approach we used hierarchical clustering and for our black box approach we trained the fuzzy c-means  model.
We calculated a weighted mutation score and compared this with the mutation score of a full set of mutants executed.
Results show that with hierarchical clustering we can reduce the amount of mutants executed while maintaining the effectiveness.
The clusters generated by the machine learning model were less accurate and showed a significant decrease in effectiveness.
\end{abstract}

\begin{IEEEkeywords}

\end{IEEEkeywords}

\section{Introduction}
\subsection{Contributions}
\subsection{Outline}

%%%%%% DO RESEARCH PART HERE %%%%%%%%

\section{Comparing mutation testing tooling}
\section{Qualitative clustering approach}
\section{Quantitative clustering approach}
\section{Experiment design}

\section{Discussion}

\section{Related work}
\label{sec:rw}

\section{Conclusion}
\label{sec:conclusion}

\printbibliography[heading=bibintoc]

\end{document}

### 2021-11-29 chapters/2_introduction.tex
Added introduction + selecting mutation

@@
-\documentclass[../main]{subfiles}
+wel e\documentclass[../main]{subfiles}
 \begin{document}
@@
 Real faults are unknown in advance, we can use mutation testing as a proxy measurement. 
-Research has found statistically significant evidence which correlates mutant detection with real fault detection\cite{Just2014AreTesting}.
+Existing research has found statistically significant evidence which correlates mutant detection with real fault detection\cite{Just2014AreTesting}.
 \newline
-\todo[inline]{is this your approach? the current general approach?}
 Mutation testing tools create small faulty versions of the program.
@@
 Mutation clustering aims to reduce the amount of mutants to be executed by clustering mutants\cite{Ma2016,Yu2019PossibilityScope}.
-Clustering is a technique used for the grouping of data, for pattern recognition and classification, based on some similar characteristics\cite{Alsabti1997}.
 The clustering of mutants has been researched with promising results\cite{Ji2009,Wilinski2015,Ma2016}. For example Ma et al., \cite{Ma2016} clustered the mutants for expressions. 
@@
 There should be no requirements for which mutants can be clustered.
-We will do this with two different approaches: qualitative and quantitative. 
+We do this with two different techniques: qualitative and quantitative.
+We devise a white and black box approach for clustering mutants.
+Our white box approach contains a qualitative analysis on mutants and a methodology to cluster them.
+The black box approach makes use of a machine learning model to cluster mutants.
 We structure our research on the following research questions:

### 2021-11-29 mutation-workshop/paper.tex
Added introduction + selecting mutation

@@
 \usepackage[backend=biber, style=ieee, citestyle=numeric-comp,uniquelist=false, maxcitenames=2, maxbibnames=9, mincitenames=1]{biblatex}
-\bibliography{../refs/main}
+\bibliography{../main}
+\bibliography{../custom}
+
+%% uncomment this and change compiler for dyslexic version
+%\usepackage{fontspec}
+%\setmainfont{OpenDyslexic}
 
@@
 Amsterdam, the Netherlands \\
-thomas@emailhere.com}
+thomas.biesaart@protonmail.com}
 }
@@
 To cluster mutants we need a mutation testing tools.
-Three mutation testing tools for Java are compared and one is selected to use for this research
+Three mutation testing tools for Java are compared and one is selected for further use in this research.
 The first part consists of a white box approach and the second part of a black box approach.
@@
 \section{Introduction}
+The high cost requirement is often a barrier for adopting mutation testing \cite{Pizzoleto2019}.
+A lot of techniques and methods have been developed to improve the performance, however, most of these approaches are not as effective as mutation testing a full set of mutants\cite{Pizzoleto2019,Yao2014}.
+\newline
+A technique to reduce the number of mutants or the number of test case executions is mutation clustering.
+Mutation clustering aims to reduce the amount of mutants to be executed by clustering mutants\cite{Ma2016,Yu2019PossibilityScope}.
+The clustering of mutants has been researched with promising results\cite{Ji2009,Wilinski2015,Ma2016}. For example Ma et al., \cite{Ma2016} clustered the mutants for expressions. 
+However in their paper they indicate that the choosing expressions as the domain of their cluster is a limitation.
+A recurring problem in the research of mutation clustering is to determine the means of domains as the centre of a cluster\cite{Ji2009,Wilinski2015,Ma2016,Wei2021SpectralTesting}.
+\newline
+To reduce the cost of mutation testing, we try to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants. 
+Our goal is to cluster any mutant that is generated within the Java programming language.
+This research aims to remove the scoping limitations present in existing research.
+There should be no requirements for which mutants can be clustered.
+We do this with two different techniques: qualitative and quantitative.
+We devise a white and black box approach for clustering mutants.
+Our white box approach contains a qualitative analysis on mutants and a methodology to cluster them.
+The black box approach makes use of a machine learning model to cluster mutants.
+We structure our research on the following research questions:
+\newline
+\textbf{Research Question 1}: What set of characteristics can we identify for clustering generated mutants to reduce the amount executed, while maintaining effectiveness?
+\newline
+\textbf{Research Question 1.1}: How do the existing mutation testing tools for Java compare to each other?
+\newline
+\textbf{Research Question 2}: How can we train a machine learning model to recognize and cluster generated mutants to reduce the amount executed, while maintaining effectiveness?
+
 \subsection{Contributions}
+Our research makes the following contributions:
+\begin{enumerate}
+ \item A white box methodology to cluster generated mutants based on chosen characteristics to reduce the cost of mutation testing.
+  \item A black box methodology to cluster generated mutants based on chosen characteristics to reduce the cost of mutation testing.
+ \item A proof of concept which implements the qualitative methodology chosen and elaborated within the thesis.
+ \item A proof of concept which implements the quantitative approach to cluster mutants and is elaborated within the thesis. 
+\end{enumerate}
 \subsection{Outline}
@@
 
-\section{Comparing mutation testing tooling}
+\section{Selecting a mutation testing tool}
+We reviewed literature on three different mutation testing tools[CITE OWN PAPER HERE].
+However the most recent study is from 2017.
+At the time of writing this paper these studies are at least four years old. 
+Some of the tools were updated or are still under active development\cite{pit-releases,Major}.
+There are three candidate tools; Major, MuJava and PIT.
+We extend the existing mutant operator comparison with the operators that have been added since the publishing date of the existing literature.
+Additionally we add an overview of the overlapping mutant operators for the new mutant operators.
+
+\todo{show tables here? maybe a summarized version of this?}
+\todo{the results can be found in ref to paper??}
+
+MuJava is not actively maintained and has not been updated in the last few years.
+It does not support JUnit 4 and all versions of TestNG\cite{mujava}.
+These are test engines and are crucial for executing tests.
+MuJava also does not support source projects with java version 1.7 or higher\cite{mujava}.
+Conforming to these requirements would reduce the set of projects we could use in or experiments.
+\newline
+While Major supports JUnit 4,
+we did not succeed in generating mutants with this tool[CITE OWN PAPER HERE].
+It would be too time consuming to customize all source projects to work with Major.
+Furthermore there is not much documentation available for this tool.
+\newline
+PIT targets the industry, is actively maintained and is open source\cite{Kintis2016AnalysingStudy}.
+It supports Maven, Gradle, has a command line interface and has a faster execution time than the other tools.
+For example PIT provides a plugin system in which you can inject your own code in various stages of mutation testing process\cite{pit}.
+PIT also generated significantly more mutants in every project we have tested[CITE OWN PAPER HERE].
+Based on the information presented in this chapter we decide that PIT is the best choice for generating and executing mutants for this research.
+
 \section{Qualitative clustering approach}
 \section{Quantitative clustering approach}
-\section{Experiment design}
-
-\section{Discussion}
+\section{validation}
 

### 2021-12-02 mutation-workshop/paper.tex
added template file for paper
Add "draft" mode for better review
Added draft for white box approach

@@
+\newif\ifdraft
+
+% Set the below value to \drafttrue to use draft formatting or
+% \draftfalse to enable "production" formatting
+\draftfalse
+
+\ifdraft
+\documentclass[conference,draftclsnofoot,onecolumn]{IEEEtran}
+\usepackage{fontspec}
+\setmainfont{OpenDyslexic}
+\else
 \documentclass[conference]{IEEEtran}
+\fi
+
 \IEEEoverridecommandlockouts
@@
 % \usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
 \usepackage{algorithmic}
@@
 \usepackage{hyperref}
-
+\usepackage{amsmath,amssymb}
 \usepackage[flushleft]{threeparttable}
 \usepackage{tabularx}
-
 \usepackage[disable]{todonotes}
 \usepackage{enumitem}
+\usepackage[backend=biber, style=ieee, citestyle=numeric-comp,uniquelist=false, maxcitenames=2, maxbibnames=9, mincitenames=1]{biblatex}
+\usepackage{multirow}
 
-\usepackage[backend=biber, style=ieee, citestyle=numeric-comp,uniquelist=false, maxcitenames=2, maxbibnames=9, mincitenames=1]{biblatex}
 \bibliography{../main}
 \bibliography{../custom}
-
-%% uncomment this and change compiler for dyslexic version
-%\usepackage{fontspec}
-%\setmainfont{OpenDyslexic}
 
@@
 \section{Selecting a mutation testing tool}
-We reviewed literature on three different mutation testing tools[CITE OWN PAPER HERE].
+To cluster mutants we need to select a tool that generate mutants and executes mutants.
+Such tools are called mutation testing tools.
+Our goal is to find the tool that generates the most amount of mutants. 
+\newline
+We reviewed literature on three different mutation testing tools\todo{CITE OWN PAPER HERE}.
 However the most recent study is from 2017.
@@
 We extend the existing mutant operator comparison with the operators that have been added since the publishing date of the existing literature.
-Additionally we add an overview of the overlapping mutant operators for the new mutant operators.
+In Table \ref{tab:mutators-current} shows the comparison of operators that were added.
+\begin{table*}[htb]
+\centering
+\begin{tabular}{|l|l|l|l|}
+\hline
+\multicolumn{2}{|c|}{\textbf{ MuJava}}                           & \multicolumn{2}{|c|}{\textbf{PIT}}                      \\ \hline
+IHD & Hiding variable deletion                                  & EM                    & Empty Returns                   \\ \hline
+IHI & Hiding variable   insertion                               & FR                    & False Returns                   \\ \hline
+IOD & Overriding method   deletion                              & NR                    & Null Returns                    \\ \hline
+IOP & Overriding method   calling position change               & TR                    & True Returns                    \\ \hline
+IOR & Overriding method   rename                                & PR                    & Primitive Returns               \\ \hline
+ISI & Super keyword insertion                                   & ER                    & Experimental Switch             \\ \hline
+ISD & Super keyword deletion                                    & BI                    & Big Integer                     \\ \hline
+IPC & Explicit call to a   parent’s constructor deletion        & NRC                   & Naked Receiver                  \\ \hline
+PNC & New method call with   child class type                   & N                     & Negation                        \\ \hline
+PMD & Member variable   declaration with parent class type      & AOR                   & Arithmic Operator Replacement   \\ \hline
+PPD & Parameter variable   declaration with child class type    & AOD                   & Arithmic Operator Deletion      \\ \hline
+PCI & Type cast operator   insertion                            & CR                    & Constant replacement            \\ \hline
+PCC & Cast type change                                          & BO                    & Bitwise Operator                \\ \hline
+PCD & Type cast operator   deletion                             & ROR                   & Relational Operator Replacement \\ \hline
+PRV & Reference assignment   with other comparable variable     & UOI                   & Unary Operation Insertion       \\ \hline
+OMR & Overloading method   contents replace                     & \multicolumn{1}{l|}{} &                                 \\ \hline
+OMD & Overloading method   deletion                             & \multicolumn{1}{l|}{} &                                 \\ \hline
+OAC & Arguments of   overloading method call change             & \multicolumn{1}{l|}{} &                                 \\ \hline
+JTI & This keyword insertion                                    & \multicolumn{1}{l|}{} &                                 \\ \hline
+JTD & This keyword deletion                                     & \multicolumn{1}{l|}{} &                                 \\ \hline
+JSI & Static modifier   insertion                               & \multicolumn{1}{l|}{} &                                 \\ \hline
+JSD & Static modifier   deletion                                & \multicolumn{1}{l|}{} &                                 \\ \hline
+JID & Member variable   initialization deletion                 & \multicolumn{1}{l|}{} &                                 \\ \hline
+JDC & Java-supported default   constructor deletion             & \multicolumn{1}{l|}{} &                                 \\ \hline
+EOA & Reference assignment   and content assignment replacement & \multicolumn{1}{l|}{} &                                 \\ \hline
+EOC & Reference comparison   and content comparison replacement & \multicolumn{1}{l|}{} &                                 \\ \hline
+EAM & Acessor method change                                     & \multicolumn{1}{l|}{} &                                 \\ \hline
+EMM & Modifier method change                                    & \multicolumn{1}{l|}{} &                                 \\ \hline
+\end{tabular}
+\caption{\label{tab:mutators-current}Current mutators supported.}
+\end{table*}
+Table \ref{tab:mutators-overlap} we show an overview of the overlapping mutant operators for the added mutant operators.
 
-\todo{show tables here? maybe a summarized version of this?}
-\todo{the results can be found in ref to paper??}
+\begin{table*}[htb]
+\centering
+\begin{tabular}{|l|l|l|l|}
+\hline
+\multicolumn{2}{|c|}{\textbf{ MuJava mutator}}      & \multicolumn{2}{c|}{\textbf{PIT mutator}}                         \\ \hline
+AORB & Arithmetic Operator Replacement Binary      & \multirow{4}{*}{M}   & \multirow{4}{*}{Math}                      \\ \cline{1-2}
+ASRS & Short-Cut Assignment   Operator Replacement &                      &                                            \\ \cline{1-2}
+SOR  & Shift Operator   Replacement                &                      &                                            \\ \cline{1-2}
+COR  & Conditional Operator   Replacement          &                      &                                            \\ \hline
+AOIU & Arithmetic Operator   Insertion Unary       & \multirow{2}{*}{UOI} & \multirow{2}{*}{Unary Operation Insertion} \\ \cline{1-2}
+AOIS & Arithmetic Operator   Insertion Short-cut   &                      &                                            \\ \hline
+ODL  & Operator Deletion                           & AOD                  & Arithmic Operator   Deletion               \\ \hline
+AODS & Arithmetic Operator   Deletion Short-cut    & RI                   & Remove Increments                          \\ \hline
+ROR  & Relational Operator   Replacement           & CB                   & Conditionals Boundary                      \\ \hline
+AORS & Arithmetic Operator   Replacement Short-Cut & I                    & Increments                                 \\ \hline
+COD  & Conditional Operator   Deletion             & RC                   & Remove Conditionials                       \\ \hline
+
+\end{tabular}
+\caption{\label{tab:mutators-overlap}Overview of mutator overlaps between  MuJavav and PIT}
+\end{table*}
 
@@
 While Major supports JUnit 4,
-we did not succeed in generating mutants with this tool[CITE OWN PAPER HERE].
+we did not succeed in generating mutants with this tool\todo{CITE OWN PAPER HERE}.
 It would be too time consuming to customize all source projects to work with Major.
@@
 For example PIT provides a plugin system in which you can inject your own code in various stages of mutation testing process\cite{pit}.
-PIT also generated significantly more mutants in every project we have tested[CITE OWN PAPER HERE].
+PIT also generated significantly more mutants in every project we have tested\todo{CITE OWN PAPER HERE}.
 Based on the information presented in this chapter we decide that PIT is the best choice for generating and executing mutants for this research.
@@
 \section{Qualitative clustering approach}
+\label{ch:reserach_question1}
+The following sections describe the white box approach for clustering mutants.
+We perform a qualitative analysis on mutants to identify characteristics. 
+These characteristics represent the mutants.
+With the mutants represented we devise a methodology for clustering mutants.
+Our goal is to cluster every mutant that is generated while maintaining effectiveness.
+
+\subsection{Identifying characteristics}
+\label{ch:identifying_characteristics}
+Zhang et al., identifies several mutant characteristics\cite{Zhang2019PredictiveTesting}. 
+The goal of their research is detecting as accurate as possible whether a mutant survives or not before executing the mutant.
+The mutant characteristics identified by Zhang et al., have been used in other research with promising results\cite{Oonk2021}.
+While the goal of their research is different than ours, we can still use the characteristics they identified.
+We select characteristics relevant to our research from the research of Zhang et al., and Oonk\cite{Oonk2021}.
+We extend the list of characteristics with data we can extract from the selected mutation testing tool. 
+\newline
+PIT generates and executes mutants on byte code level\cite{pitestBytecode}.
+By mutating in Java byte code we can identify characteristics specific to Java byte code.
+PIT also gathers mutant metadata.
+We can extract this mutant metadata and identify characteristics. 
+With the combination of information extracted from PIT and characteristics selected from existing research we identify the following characteristics\todo{CITE OWN PAPER HERE}:
+\begin{enumerate}
+    \item Mutant similarity.
+    \item Amount of tests the mutant is challenged by.
+    \item Mutator identifier.
+    \item Mutant opcode.
+    \item Mutant return type.
+    \item Mutant contains local variables.
+    \item Mutant is in try catch block.
+    \item Mutant is in finally block.
+    \item Name of class that contains the mutant.
+    \item Name of function that contains the mutant.
+    \item Line number of the start of the block that contains the mutant.
+    \item Line number on which the mutation occur.
+\end{enumerate}
+
+\subsection{Levenshtein distance}
+A piece of code is a textual representation of instructions for a computer.
+A mutant is a transformation of a piece of code.
+This transformation is predefined.
+The difference between a mutant and its parent can be defined as similarity.
+We identify the similarity between a mutant and parent as a characteristic.
+\newline
+To measure this characteristic we select a similarity measure.
+There are existing similarities measures that base their similarity in different ways\cite{Gomma2013AApproaches}. 
+\newline
+The Levenshtein distance defines the distance between two strings by counting the minimum number of operations(create/edit/delete) needed to transform one string into the other\cite{Gomma2013AApproaches}.
+The more the textual representation of a mutant differs from its parent the bigger the Levenshtein distance will be.
+A unit test tests a small piece of code.
+The more this piece of code differs from its original the more likely it is the test will fail.
+The Levenshtein distance measures this difference.
+We select the Levenshtein distance as similarity measure for our research\cite{Gomma2013AApproaches}.
+\newline
+A mutant is a piece of code that differs in a predefined way from its parent.
+Java source code compiles into Java byte code.
+During the compilation of Java code certain context is abstracted away\cite{byteCodeEngineering}.
+Optimisations are applied which also changes the Java byte code\cite{byteCodeEngineering}.
+As a result the textual similarity between a Java code mutant and its parent is different from the similarity between a Java byte code mutant and its parent.
+While the textual similarity may be different between the Java code and Java byte code the functionality remains the same.
+Java byte code reflects more of the semantic nature of the source code than the source code itself does. 
+PIT generates and executes mutants on byte code level\cite{pitestBytecode}. 
+In other words PIT executes the unit tests of a source against the byte code of a mutant.
+\newline
+By calculating the distance on byte code we filter out the context that is present in Java code.
+This gives us a Levenshtein distance that represents more of the semantic difference, between a mutant and parent, than calculating the distance for Java code.
+For this characteristic we use the Levenshtein distance between a Java byte code mutant and its parent.
+
+\subsection{Hierarchical clustering}
+Clustering Levenshtein distances has been done before with hierarchical clustering with promising results.\cite{Rajalingam2011, Gothai2010PerformanceAlgorithms}. 
+Research shows that hierarchical clustering performs better when clustering with at least ten features\cite{Rodriguez2019}.
+It also states that varying the parameters of hierarchical clustering improves the performance compared to that of the default settings of the algorithm\cite{Rodriguez2019}.
+We identified more than ten characteristics and can use them as features.
+We adjust the parameters of the algorithm based on the characteristics in our data set.
+Bases on these observations and the research of Rodriguez et al., we select hierarchical clustering as it gives us the best performance in terms of partitioning.
+Next we explain the configuration we use for clustering the mutants. 
+\newline
+Hierarchical clustering is subdivided into agglomerative and divisive. 
+The agglomerative hierarchical technique follows bottom up approach whereas divisive follows top-down approaches.
+Hierarchical clustering uses different metrics which calculates the euclidean distance between two clusters and the linkage criteria\cite{Rajalingam2011}. 
+The linkage criteria specifies the dissimilarity in the sets as a function of the pair-wise distances of observations in those sets\cite{Rajalingam2011}.
+We reviewed the different linkage criteria and concluded the following.
+\newline
+Research shows that the complete linkage outperforms the single linkage method\cite{Vijaya2019ComparativeClustering}.
+The ward linkage and complete linkage methods perform the same when clusters are well separated\cite{Vijaya2019ComparativeClustering}.
+However if the clusters overlap the ward linkage outperforms the complete linkage\cite{Vijaya2019ComparativeClustering}.
+\newline
+We identify all characteristics per mutant as a separate cluster.
+Starting out with each mutant as a separate cluster we can use the agglomerative form of hierarchical clustering.
+Since we cannot assume that our clusters are well separated we chose to use the ward linkage method for our clustering algorithm.
+In summary, we select the agglomerative hierarchical clustering algorithm with ward linkage to cluster mutants represented by the characteristics we gathered.
+
+\subsection{Categorical data}
+There are different categorical variable encoding techniques available\cite{Potdar2017AClassifiers}.
+The categorical characteristics we use have no particular ranking compared to each other.
+There is also no specific order to the characteristics.
+I.e. a return type void is not better or worse than a return type string.
+The same goes for the location characteristics, there is no location that should have a bigger weight than the other locations.
+The individual characteristics do contain a finite set of values.
+For example multiple mutants may contain the same class name.
+The hierarchical clustering algorithm needs all characteristics in a numerical form\cite{Vijaya2019ComparativeClustering}.
+The characteristics mutator identifier, class name, method name and return type are non numerical.
+To deal with this problem we apply categorical variable encoding to these specific features.
+Taking into account the properties of our categorical characteristics the nominal variable encoding fits our requirements.
+Nominal encoding comprises a finite set of discrete values with no relationship between values\cite{Potdar2017AClassifiers}.
+Therefore we implement this type of encoding in the experiment.
+
+\subsection{Number of clusters}
+Agglomerative hierarchical clustering can continue to cluster until there is one cluster left. 
+Naturally this cluster will contain all the mutants.
+We can cut off the clustering algorithm at any point.
+We decide on the number of clusters based on the amount of mutants generated in a full set.
+We select the number of clusters from a performance increasing perspective.
+Mutation clustering increases the performance by reducing the amount of mutants executed\cite{Pizzoleto2019}.
+By reducing the amount of mutants, for example by half, we increase the performance by 50\%\cite{Just2014AreTesting}.
+As a starting point we perform three experiments where the cluster size is 25\%, 50\% and 75\% of the total amount of mutants respectively.
+We can evaluate the results and can decide to cut off the clustering at different points.
+The amount of mutants inside a cluster is be decided by the clustering algorithm.
+
 \section{Quantitative clustering approach}
+In the following sections we devise a black box approach for clustering mutants.
+We make use of the characteristics identified in Chapter \ref{ch:identifying_characteristics}.
+While our goal remains the same as in Chapter \ref{ch:reserach_question1}, the method to achieve it is different.
+
 \section{validation}

### 2021-12-24 chapters/9_discussion.tex
background
draft paper 0.1

--- 2021-07-15 chapters/7_discussion.tex
@@
 First, we discuss the results of the white box approach. 
-The results are discussed as one and are compared between the variations(with Levenshtein distance characteristic and without).
+The results are discussed as one and are compared between the variants(with Levenshtein distance characteristic and without).
 Second we discuss the results of the black box approach and compare them to those of the white box approach.
@@
 66\% of the displayed results have a deviation of less than 1\%. 
-This includes both variations of the experiment.
+This includes both variants of the experiment.
 \newline
@@
 \end{finding}
-In both variations of the experiment we observe a reduction in average accuracy when increasing the performance by reducing the number of clusters.
+In both variants of the experiment we observe a reduction in average accuracy when increasing the performance by reducing the number of clusters.
 These reductions are 0.1\% and 0.04\% for clustering with and without Levenshtein distance respectively.
@@
 The plugins and scripts needed for these steps are available in the code repository on Github\cite{rbasarat-repo}.
+The plugins can be found in the directory \textit{pitest-clustering-plugin}.
+The scripts used can be found in the directories \textit{FCM-clustering} and \textit{hierarchical-clustering}.
 
 \section{Threats to validity}
-In this section, we discuss the threats to internal validity, external validity and reliability.
+In this section, we discuss the threats to internal validity and external validity.
 \subsection{Internal validity}

### 2021-12-24 mutation-workshop/paper.tex
draft paper 0.1

@@
 Three mutation testing tools for Java are compared and one is selected for further use in this research.
-The first part consists of a white box approach and the second part of a black box approach.
 Our research uses the mutation testing tool PIT.
 We identified characteristics to represent the mutants such that we can cluster them.
+These characteristics have shown that there exists a strong relationship between mutants. 
+We prove this by clustering them with two different approaches; a white box approach and a black box approach.
 For our white box approach we used hierarchical clustering and for our black box approach we trained the fuzzy c-means model.
@@
 The black box approach makes use of a machine learning model to cluster mutants.
+Before we can cluster mutants we need to select a mutation testing tool.
 We structure our research on the following research questions:
@@
 
-%%%%%% DO RESEARCH PART HERE %%%%%%%%
-
 \section{Selecting a mutation testing tool}
@@
 We extend the existing mutant operator comparison with the operators that have been added since the publishing date of the existing literature.
-In Table \ref{tab:mutators-current} shows the comparison of operators that were added.
-\begin{table*}[htb]
+Table \ref{tab:mutators-current} shows the comparison of operators that were added.
+
+\begin{table*}[ht]
 \centering
@@
 \end{tabular}
-\caption{\label{tab:mutators-current}Current mutators supported.}
+\caption{\label{tab:mutators-current}New mutators not taken into account in previous research.}
 \end{table*}
-Table \ref{tab:mutators-overlap} we show an overview of the overlapping mutant operators for the added mutant operators.
 
-\begin{table*}[htb]
+It is also notable that there is some overlap between the new mutators of MuJava and PIT. 
+Overlapping mutators have a different name but mutate the code in the same way. 
+An overview of overlapping mutators can be found in Table \ref{tab:mutators-overlap}.
+
+\begin{table*}[ht]
 \centering
@@
 
+We also compared the numbers of mutants generated by each tool.
+The goal of our thesis is to cluster as many mutants as possible. 
+Therefore it is valuable for this research to know how much mutants each tool generates on the same source.
+Table \ref{tab:mutants-generated} shows how much mutants the tools generate if used on the same source.
+
+\begin{table*}[ht]
+\centering
+\begin{tabular}{|l|l|c|c|c|}
+\hline
+                                & \textbf{version} & \textbf{\begin{tabular}[c]{@{}c@{}}LOC\\ (Lines of code)\end{tabular}} & \multicolumn{1}{l|}{\textbf{MuJava}} & \multicolumn{1}{l|}{\textbf{PIT}} \\ \hline
+Commons-numbers.core            & 1.0-beta1        & 450                                                                    & 988                                  & 4384                              \\ \hline
+Jodatime                        & 2.10.10          & 28811                                                                  & 52925                                & 112772                            \\ \hline
+Zxing 3.4.1                     & 3.4.1            & 24792                                                                  & 65983                                & 161409                            \\ \hline
+Google Auto Common              & 0.11.0           & 2338                                                                   & 20                                   & 5219                              \\ \hline
+Google Auto Factory             & 1.0-beta9        & 1507                                                                   & 69                                   & 5832                              \\ \hline
+Google Auto Value               & 1.7.5            & 7466                                                                   & 745                                  & 16746                             \\ \hline
+Google Auto Service             & 1.0-rc6          & 750                                                                    & 0                                    & 475                               \\ \hline
+ScribeJava-Core                 & 8.1.0            & 5709                                                                   & 1358                                 & 5746                              \\ \hline
+Checkstyle                      & 8.41.1           & 38491                                                                  & 4529                                 & 100952                            \\ \hline
+Fastjson                        & 1.2.75           & 43405                                                                  & 70577                                & 116188                            \\ \hline
+Jfreechart                      & 1.5.3            & 91876                                                                  & 0                                    & 350741                            \\ \hline
+Commons-lang                    & 3.12.0           & 29836                                                                  & 31145                                & 134764                            \\ \hline
+Commons-codec                   & 1.15.0           & 9656                                                                   & 21719                                & 54804                            \\ \hline
+Commons-text                   & 1.9.0             & 9781                                                                   & 10403                                & 48490                            \\ \hline
+Commons-io                     & 2.8.0             & 13947                                                                  & 13245                                & 44631                            \\ \hline
+Gson                           & 2.8.6             & 8078                                                                   & 9198                                 & 28485                            \\ \hline
+Commons-cli                    & 1.4.0             & 2782                                                                   & 831                                  & 7193                             \\ \hline
+Commons-csv                    & 1.8.0             & 1855                                                                   & 2081                                 & 6906                             \\ \hline
+\end{tabular}
+\caption{\label{tab:mutants-generated}Amount of mutants generated by MuJava and PIT}
+\end{table*}
+
 MuJava is not actively maintained and has not been updated in the last few years.
@@
 While Major supports JUnit 4,
-we did not succeed in generating mutants with this tool\todo{CITE OWN PAPER HERE}.
+we did not succeed in generating mutants with this tool\cite{thesis}.
 It would be too time consuming to customize all source projects to work with Major.
@@
 For example PIT provides a plugin system in which you can inject your own code in various stages of mutation testing process\cite{pit}.
-PIT also generated significantly more mutants in every project we have tested\todo{CITE OWN PAPER HERE}.
+PIT also generated significantly more mutants(see Table \ref{tab:mutants-generated}) in every project we have tested\cite{thesis}.
 Based on the information presented in this chapter we decide that PIT is the best choice for generating and executing mutants for this research.
@@
 
+In the next section we look at the Levenshtein distance as an additional characteristic.
+
 \subsection{Levenshtein distance}
@@
 For this characteristic we use the Levenshtein distance between a Java byte code mutant and its parent.
+\newline
+We select the Levenshtein distance as similarity measure for our research\cite{Gomma2013AApproaches}.
+This could be a useful characteristic because mutants that have the same similarity may be grouped together.
 
@@
 Bases on these observations and the research of Rodriguez et al., we select hierarchical clustering as it gives us the best performance in terms of partitioning.
+\newline
 Next we explain the configuration we use for clustering the mutants. 
-\newline
 Hierarchical clustering is subdivided into agglomerative and divisive. 
@@
 \section{Quantitative clustering approach}
+\label{ch:reserach_question2}
 In the following sections we devise a black box approach for clustering mutants.
+We select a machine learning model.
+We start by choosing between supervised or unsupervised learning.
+Then we look at the machine learning models that are discussed in the survey done by K.L. Du\cite{Du2010Clustering:Approach}.
 We make use of the characteristics identified in Chapter \ref{ch:identifying_characteristics}.
@@
 
-\section{validation}
+\subsection{Supervised or unsupervised learning}
+In this subsection we decide on supervised or unsupervised learning.
+\newline
+Supervised machine learning models require a labeled data set.
+We can argue that we can label our data set with the results of our first experiments.
+Each mutant would then be labeled with the identity of the cluster it is clustered in by the hierarchical clustering model.
+If we would train a supervised model with the data labeled as described, it may result in a model that can do the same as hierarchical clustering.
+Our research is not to prove if we can train a model to cluster mutants the same way as hierarchical clustering.
+We want to train a model to find the relations between the characteristics without having to define these relations manually.
+\newline
+Unsupervised learning models do not need a labeled data set and can recognize relations or discover hidden patterns in the characteristics we identified.
+This corresponds to what we want to achieve with this experiment.
+Therefore we choose unsupervised learning to continue with in our experiment.
 
+\subsection{Self organizing map, C-means and fuzzy c-means}
+\label{ch:topThreeModels}
+We reviewed the models discussed in the survey done by K.L. Du\cite{Du2010Clustering:Approach, thesis}. 
+Three models which best fit our use case are selected.
+The self-organizing map, C-means and fuzzy c-means models are all models that would fit our use case and purpose. 
+To choose between these models we look at the performance of each model.
+Mingoti et al., compared the performance of these three models \cite{Mingoti2006ComparingAlgorithms}.
+They did a Monte Carlo simulation where the cluster sizes and amount of random numbers in the data set varied each simulation\cite{Mingoti2006ComparingAlgorithms}.
+Other variables such as cluster boundaries and overlap were also controlled variables in the experiment.
+They observe that the C-means and fuzzy c-means had good performance for non overlapping situations\cite{Mingoti2006ComparingAlgorithms}.
+The best results for average recovery and internal dispersion rates were found for fuzzy c-means which was stable in all situations achieving recovery averages over 90\%\cite{Mingoti2006ComparingAlgorithms}
+The C-means method was affected by the presence of a large amount of outliers.
+They conclude that the self-organizing map did not perform well in many cases being affected by the amount of variables and clusters even for the non overlapping cases\cite{Mingoti2006ComparingAlgorithms}.
+They also observe  that their results partially overlap with other research in the same domain.
+\newline
+Based the conclusions made by Mingoti et al., we can conclude that the fuzzy c-means model has the best performance overall.
+Thus for this experiment we select the fuzzy c-means model.
+
+\subsection{Training strategy}
+\label{ch:training_strategy}
+Our data set consists of mutants represented by their characteristics. 
+The characteristics of the mutants are generated per project and have a project specific context.
+For example, a mutant with a mutator in a certain project can have a different effect on the source code and can give other results than a mutator in a different project.
+The relations between the mutants are scoped to their own project. 
+In other words if a specific mutant has a relation to another mutant in the same project, this relation may not exist between the same mutant and a mutant from a different project.
+To preserve these relations and context we decided to train a separate model per project.
+\newline
+To validate the model we need to split our data set into a training set and a validation set.
+Research has shown that if the data set is big enough a ratio of 80/20 is sufficient\cite{Guyon1997ARatio}.
+We divide each source projects into training and validation groups to reflect the 80/20 ratio as best as possible.
+Depending on the amount of mutants per project we might not achieve a perfect 80/20 ratio.
+The division of the mutants is be done randomly. 
+To make our results reproducible we select a random division based on a generated seed.
+The seeds generated for each run will be included in the results and can be found in the Github repository in the file \textit{hierarchical-clustering/main.py} at line 182\cite{rbasarat-repo}.
+
+\subsection{Algorithm parameters}
+The fuzzy c-means algorithm has to initialize the centroids of the clusters.
+The library we selected does this based on a random number.
+It also allows us to control this random number.
+To make the initialization of the centroid reproducible we will use the same generated seeds as in our first experiment.
+The seeds generated are included in the Github repository in the file \textit{hierarchical-clustering/main.py} at line 182\cite{rbasarat-repo}.
+\newline
+Bezdek et al., also researched cluster validity for the fuzzy c-means model\cite{Bezdek1995OnModel}.
+They researched what effect which parameters could have on the validity of the clusters.
+They specifically analyzed the role of weighting exponent \textit{m}(fuzziness parameter).
+They concluded that the best choice for \textit{m} is in the interval [1.5, 2.5], whose mean and midpoint is \textit{m}=2.
+Wu proposed a new guideline for selecting parameter \textit{m}\cite{Wu2012}.
+His point of view was that a large \textit{m} value will make the fuzzy c-means more robust to noise and outliers.
+He suggests implementing fuzzy c-means with \textit{m}=[1.5, 4].
+When the data set contains noise and outliers, the fuzzifier \textit{m}=4 is recommended in a theoretical upper bound case.
+Since our data set does not contain noise(all mutants consists of every characteristic) we choose to select the midpoint concluded in the research of Bezdek et al.\cite{Bezdek1995OnModel}.
+The value of \textit{m} in this experiment is 2.
+\newline
+Finding the correct amount of clusters has been an ongoing problem for clustering algorithms\cite{Du2010Clustering:Approach}.
+Fortunately there are techniques available to estimate the number of clusters.
+During model selection we discussed the mountain and subtractive models. 
+These models are designed to estimate the number of clusters in a data set.
+The subtractive clustering model is specifically designed to approximate the number of clusters for fuzzy system models\cite{Chiu1994FuzzyEstimation}.
+It approximates the centroids and amount of the clusters based on the data points in the data set\cite{Chiu1994FuzzyEstimation}.
+We select and implement the subtractive clustering model to estimate the number of clusters.
+We use this number as the value for the number of clusters parameter in the fuzzy c-means algorithm.
+
+\section{Experiment design}
+In this section we design an experiment to validate our research.
+We execute the following steps:
+\begin{enumerate}
+    \item Gather all mutant characteristics from PIT.
+    \item Execute full set of mutants to gather baseline measurement.
+    \item Cluster the mutants according the black and white box methodologies.
+    \item Execute one random mutant per cluster and gather result.
+    \item Calculate weighted mutation score.
+\end{enumerate}
+Both research questions have the same goal, but with different approaches.
+We can devise one hypothesis for both research questions.
+We hypothesise that we can cluster mutants with the set of characteristics we identified
+while maintaining effectiveness and reducing the amount executed when executing one mutant from each cluster that is randomly selected.
+Our goal is to achieve a weighted mutation score that is as close as possible to the mutation score of a full set of executed mutants.
+We repeat the experiment for each approach we devised in Chapters \ref{ch:reserach_question1} and \ref{ch:reserach_question2}.
+
+\subsection{Baseline measurement}
+The baseline for our experiment is the mutation score of the set of all mutants generated by the selected mutation testing tool.
+To extract this score we need to execute a full set of mutants within PIT.
+For each source we have added the PIT plugin and required configuration in the \textit{pom.xml}, this code instructs PIT to generate and execute mutants with all mutators available.
+We can then use the command line interface and a maven command(see Listing \ref{lst:mvn-full}) to execute PIT.
+\begin{lstlisting}[label=lst:mvn-full,caption=Command used to execute full set of mutants with PIT.]
+mvn -U org.pitest:pitest-maven:mutationCoverage
+\end{lstlisting}
+
+\subsection{Weighted mutation score}
+\label{ch:weighted_score}
+Our hypothesis states that each mutant executed should represent that whole cluster.
+With a mutant executed from each cluster we can calculate a mutation score.
+This mutation score is a weighted mutation score.
+This weighted mutation score is the product of the result of a mutant(1 for killed and 0 for survived) and the amount of mutants in the cluster it represents.
+The weighted mutation score is then comparable to the score of a full set as the total number of mutants will be the same.
+\newline
+For example, take a full set with a score of 75/100 killed mutants.
+This gives us a mutation score of 75\%. 
+We then cluster the mutants in four clusters consisting of 12, 30, 38 and 20 mutants, respectively.
+We randomly select four mutants of each cluster and execute them.
+The mutants representing cluster one and four survive and two and three are killed.
+If we calculate the weighted score we get 68/100 which is 68\%.
+We can then compare this to the score of a full set because the amount of mutants executed is the same: 75/100(75\%) and 68/100(68\%).
+
+\subsection{Random mutant selection}
+\label{ch:mutant_selection}
+As stated in our hypothesis we randomly select a mutant from each cluster to be executed.
+To validate our sample we make use of statistical hypothesis testing\cite{Emmert-Streib2019UnderstandingInference}.
+Our null hypothesis states that there is no relation between the characteristics identified and the results of a mutant with the chosen methodology.
+To test this hypothesis we select an alpha value of 0.05.
+To translate; if more than five percent of the values in a sample deviates more than the significance level(see Chapter \ref{ch:exp1_validation}) we cannot reject the null hypothesis.
+To make our results reproducible we select the random mutant based on a generated seed.
+The seed generated for each run will be included in the results and can be found in the Github repository in the file \textit{hierarchical-clustering/main.py} at line 182\cite{rbasarat-repo}.
+We repeat our experiment 30 times with 30 different seeds.
+Achieving consistent results while applying random selection contributes to the validity of the experiment.
+
+\section{Validation}
+\label{ch:exp1_validation}
+The most efficient way to measure test effectiveness with mutation testing is by executing all mutants that a mutation testing tool can possibly generate.
+The goal of this research is to reduce the amount of mutants executed while maintaining effectiveness.
+To reduce the amounts executed we cluster the mutants. 
+We can measure this by counting the number of clusters we generate and compare it to the number of total mutants generated by the selected tool.
+\newline
+To validate how effective our method of clustering is we can compare the weighted mutation score(see Chapter \ref{ch:weighted_score}) of the clustered set to the mutation score of the full set.
+The closer the weighted score is to that of a full set the more effective our set of characteristics and clustering algorithm proves to be.
+In other words we want achieve a mutation score that is as close as possible as to that of a full set.
+We select a statistical significance level of $\leq 0.05$.
+This is the conventional threshold for declaring statistical significance\cite{Kirk1996PracticalCome}.
+\newline
+Depending on the effectiveness of our clustering algorithm we may loose accuracy.
+This can happen if a cluster contains mutants of both results. 
+We can measure the accuracy inside a cluster by calculating a percentage of all mutants that have survived against the ones that have been killed in a cluster.
+If the majority of the mutants in a cluster is killed then we consider that cluster to represent a killed cluster and the other way around for survived mutants.
+We consider the mutants that are not in the majority of the cluster as inaccuracy.
+\newline
+If the weighted mutation score of our clustered set deviates more than 5\% from the score of a full set we reject our hypothesis.
+
+\section{Results}
+\label{ch:results}
+In this chapter we present the results of our experiments.
+We divide the results in two sections. 
+The first section displays the results of clustering with hierarchical clustering.
+The second section displays the results of clustering with the fuzzy c-means model.
+\newline
+During the process of extracting the characteristics we found that calculating the Levenshtein distance took a long time\cite{thesis}.
+Calculating the Levenshtein distance for the four biggest projects in the data set cost days\cite{thesis}.
+Even for the smaller projects the extraction of the characteristics cost hours\cite{thesis}.
+Including the Levenshtein distance defeats the purpose of increasing performance as every bit of performance we increase would be replaced by the cost of calculating the Levenshtein distance\cite{thesis}.
+
+\subsection{Hierarchical clustering}
+This section displays the results of our clustering algorithm when executed without the Levenshtein distance.
+The results are grouped based on the reduced percentage of mutants executed.
+We also display the accuracy of the clusters per reduction amount.
+The results displayed are the average of the results of the 30 repetitions of the experiment.
+The result of each single repetition can be found in our Github repository in the folder\textit{experiment{\_}results}\cite{rbasarat-repo}.
+
+\begin{figure}[ht]
+\includegraphics[width=0.5 \textwidth]{images/boxplot_summary/boxplot_hc_no_distance_0.25.png}
+\caption{\label{box:clustering_no_distance_25}Box-plots containing weighted mutation score of clustering mutants without Levensthein distance and cluster size n*0.25}
+\end{figure}
+
+\begin{table*}[ht]
+\centering
+\begin{tabular}{|l|c|c|c|c|c|}
+\hline
+\multicolumn{1}{|c|}{Project} & \begin{tabular}[c]{@{}c@{}}Mutation score\\ full set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Mutation score \\ clustered set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg.
+cluster\\  accuracy\end{tabular} & \begin{tabular}[c]{@{}c@{}}Min.
+cluster\\ accuracy\end{tabular} & \begin{tabular}[c]{@{}c@{}}Max.
+cluster\\ accuracy\end{tabular} \\ \hline
+Google Auto Service           & 0.00\%                                                                                  & 0.00\%                                                                                      & 0.00\%                                                                               & 0.00\%                                                                                & 0.00\%                                                                               \\ \hline
+ScribeJava-Core               & 33.52\%                                                                                 & 33.72\%                                                                                     & 92.24\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+Google Auto Factory           & 48.41\%                                                                                 & 48.32\%                                                                                     & 94.21\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+Google Auto Common            & 59.59\%                                                                                 & 60.03\%                                                                                     & 91.19\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+Google Auto Value             & 61.42\%                                                                                 & 62.58\%                                                                                     & 93.90\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+Google Gson                   & 66.83\%                                                                                 & 66.89\%                                                                                     & 91.75\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+commons-io                    & 69.70\%                                                                                 & 66.08\%                                                                                     & 90.20\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+commons-cli                   & 79.07\%                                                                                 & 78.97\%                                                                                     & 91.11\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+commons-text                  & 79.29\%                                                                                 & 79.30\%                                                                                     & 90.95\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+commons-codec                 & 84.56\%                                                                                 & 84.59\%                                                                                     & 94.56\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+commons-csv                   & 85.16\%                                                                                 & 85.24\%                                                                                     & 92.76\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+\end{tabular}
+\caption{\label{tab:clustering_no_distance_25}Results of clustering mutants without Levensthein distance and cluster size n*0.25}
+\end{table*}
+
+Figure \ref{box:clustering_no_distance_25} displays the box-plots of the weighted mutation score obtained from each individual sample.
+We can observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
+Table \ref{tab:clustering_no_distance_25} displays the results of clustering mutants without Levensthein distance and the number of clusters equal to the total amount of mutants * 0.25. 
+We observe that the maximum and minimum differences in mutation score between a full set and a clustered set  3.61\% and 0.09\% respectively.
+The average of the differences between full set score and clustered score is 0.53\%.
+
+\begin{figure}[ht]
+\includegraphics[width=0.5 \textwidth]{images/boxplot_summary/boxplot_hc_no_distance_0.5.png}
+\caption{\label{box:clustering_no_distance_50}Box-plots containing weighted mutation score of clustering mutants without Levensthein distance and cluster size n*0.50}
+\end{figure}
+
+\begin{table*}[ht]
+\centering
+\begin{tabular}{|l|c|c|c|c|c|}
+\hline
+\multicolumn{1}{|c|}{Project} & \begin{tabular}[c]{@{}c@{}}Mutation score\\ full set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Mutation score \\ clustered set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg.
+cluster\\  accuracy\end{tabular} & \begin{tabular}[c]{@{}c@{}}Min.
+cluster\\ accuracy\end{tabular} & \begin{tabular}[c]{@{}c@{}}Max.
+cluster\\ accuracy\end{tabular} \\ \hline
+Google Auto Service           & 0.00\%                                                                                  & 0.00\%                                                                                      & 0.00\%                                                                               & 0.00\%                                                                                & 0.00\%                                                                               \\ \hline
+ScribeJava-Core               & 33.52\%                                                                                 & 33.55\%                                                                                     & 97.16\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+Google Auto Factory           & 48.41\%                                                                                 & 48.44\%                                                                                     & 97.71\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+Google Auto Common            & 59.59\%                                                                                 & 59.86\%                                                                                     & 96.35\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+Google Auto Value             & 61.42\%                                                                                 & 62.62\%                                                                                     & 97.31\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+Google Gson                   & 66.83\%                                                                                 & 66.81\%                                                                                     & 95.29\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+commons-io                    & 69.70\%                                                                                 & 66.05\%                                                                                     & 95.08\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+commons-cli                   & 79.07\%                                                                                 & 78.97\%                                                                                     & 95.55\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+commons-text                  & 79.29\%                                                                                 & 79.30\%                                                                                     & 95.34\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+commons-codec                 & 84.56\%                                                                                 & 84.59\%                                                                                     & 96.95\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+commons-csv                   & 85.16\%                                                                                 & 85.11\%                                                                                     & 95.87\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+\end{tabular}
+\caption{\label{tab:clustering_no_distance_50}Results of clustering mutants without Levensthein distance and cluster size n*0.50}
+\end{table*}
+
+Figure \ref{box:clustering_no_distance_50} displays the box-plots of the weighted mutation score obtained from each individual sample. 
+We can observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
+Table \ref{tab:clustering_no_distance_50} displays the results of clustering mutants without Levensthein distance and the number of clusters equal to the total amount of mutants * 0.50.
+The maximum and minimum differences between the score of a full set and clustered set are 3.65\% and 0.03\% respectively.
+We observe that the average cluster accuracy is higher than the average cluster accuracy of the 25\% set. 
+The average of the differences between full set score and clustered score is 0.49\% which is smaller than the average difference than that of the 25\% set.
+
+\begin{figure}[ht]
+\includegraphics[width=0.5 \textwidth]{images/boxplot_summary/boxplot_hc_no_distance_0.75.png}
+\caption{\label{box:clustering_no_distance_75}Box-plots containing weighted mutation score of clustering mutants without Levensthein distance and cluster size n*0.75}
+\end{figure}
+
+\begin{table*}[ht]
+\centering
+\begin{tabular}{|l|c|c|c|c|c|}
+\hline
+\multicolumn{1}{|c|}{Project} & \begin{tabular}[c]{@{}c@{}}Mutation score\\ full set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Mutation score \\ clustered set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg.
+cluster\\  accuracy\end{tabular} & \begin{tabular}[c]{@{}c@{}}Min.
+cluster\\ accuracy\end{tabular} & \begin{tabular}[c]{@{}c@{}}Max.
+cluster\\ accuracy\end{tabular} \\ \hline
+Google Auto Service           & 0.00\%                                                                                  & 0.00\%                                                                                      & 0.00\%                                                                               & 0.00\%                                                                                & 0.00\%                                                                               \\ \hline
+ScribeJava-Core               & 33.52\%                                                                                 & 33.48\%                                                                                     & 98.63\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+Google Auto Factory           & 48.41\%                                                                                 & 48.34\%                                                                                     & 98.41\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+Google Auto Common            & 59.59\%                                                                                 & 59.71\%                                                                                     & 97.08\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+Google Auto Value             & 61.42\%                                                                                 & 62.66\%                                                                                     & 97.99\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+Google Gson                   & 66.83\%                                                                                 & 66.84\%                                                                                     & 97.37\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+commons-io                    & 69.70\%                                                                                 & 66.07\%                                                                                     & 97.34\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+commons-cli                   & 79.07\%                                                                                 & 79.12\%                                                                                     & 96.83\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+commons-text                  & 79.29\%                                                                                 & 79.33\%                                                                                     & 97.60\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+commons-codec                 & 84.56\%                                                                                 & 84.61\%                                                                                     & 98.43\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+commons-csv                   & 85.16\%                                                                                 & 85.07\%                                                                                     & 97.49\%                                                                              & 50.00\%                                                                               & 100.00\%                                                                             \\ \hline
+\end{tabular}
+\caption{\label{tab:clustering_no_distance_75}Results of clustering mutant without Levensthein distance and cluster size n*0.75}
+\end{table*}
+
+Figure \ref{box:clustering_no_distance_75} displays the box-plots of the weighted mutation score obtained from each individual sample. 
+We can observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
+Table \ref{tab:clustering_no_distance_75} displays the results of clustering mutants without Levensthein distance and the number of clusters equal to the total amount of mutants * 0.50.
+The maximum and minimum differences between the score of a full set and clustered set are 3.63\% and 0.04\% respectively.
+We observe that the average accuracy is higher than the average accuracy of the other sets.
+The average of the differences between full set score and clustered score is  0.49\% which is higher than the average difference than that of the 25\% and the same of that of the 50\% set.
+
+\subsection{Fuzzy c-means}
+\label{ch:results_rq2}
+In this section we present the results of our second experiment.
+We display the weighted mutation score to create box-plots.
+For the same reason as in Chapter \ref{ch:results} we removed the Levenshtein distance characteristic.
+\newline
+The results displayed are the average of the results of the 30 repetitions of the experiment.
+The result of each single repetition can be found in our Github repository in the folder \textit{experiment{\_}results}\cite{rbasarat-repo}.
+Instead of displaying the reduction amount, as we did in the results of experiment one, we display the number of clusters that is calculated by the subtractive clustering model.
+
+\begin{figure}[ht]
+\includegraphics[width=0.5 \textwidth]{images/fcm_summary/fcm_no_distance.png}
+\caption{\label{box:clustering_fcm_no_distance_25}Box-plots containing weighted mutation score of clustering mutants without Levenshtein distance characteristic.}
+\end{figure}
+
+\begin{table*}[ht]
+\centering
+\begin{tabular}{|l|l|l|l|l|l|l|}
+\hline
+Project               & \begin{tabular}[c]{@{}l@{}}Number of   \\ clusters\end{tabular} & \begin{tabular}[c]{@{}l@{}}Mutation score  \\  full set\end{tabular} & \begin{tabular}[c]{@{}l@{}}Mutation   score\\   clustered set\end{tabular} & \begin{tabular}[c]{@{}l@{}}Average cluster \\   accuracy\end{tabular} & \begin{tabular}[c]{@{}l@{}}Min.
+cluster   \\ accuracy\end{tabular} & \begin{tabular}[c]{@{}l@{}}Max.
+cluster  \\  accuracy\end{tabular} \\ \hline
+Google Auto   Service & 17                   & 0.00\%                            & 0.00\%                                                 & 0.00\%                             & 0.00\%                                                                                                                                     & 0.00\%                                                                                                                                     \\ \hline
+ScribeJava-Core       & 32                   & 33.52\%                           & 23.64\%                                                & 76.80\%                            & 58.77\%                                                                                                                                    & 96.60\%                                                                                                                                    \\ \hline
+Google Auto   Factory & 22                   & 48.41\%                           & 57.40\%                                                & 59.20\%                            & 50.67\%                                                                                                                                    & 73.96\%                                                                                                                                    \\ \hline
+Google Auto   Common  & 31                   & 59.59\%                           & 37.64\%                                                & 62.95\%                            & 50.59\%                                                                                                                                    & 83.53\%                                                                                                                                    \\ \hline
+Google Auto   Value   & 41                   & 61.42\%                           & 59.61\%                                                & 61.48\%                            & 50.60\%                                                                                                                                    & 83.53\%                                                                                                                                    \\ \hline
+Google Gson           & 40                   & 66.83\%                           & 57.61\%                                                & 59.11\%                            & 50.84\%                                                                                                                                    & 76.68\%                                                                                                                                    \\ \hline
+Commons-io            & 40                   & 69.70\%                           & 57.95\%                                                & 57.25\%                            & 50.62\%                                                                                                                                    & 72.70\%                                                                                                                                    \\ \hline
+Commons-cli           & 26                   & 79.07\%                           & 81.53\%                                                & 81.77\%                            & 67.70\%                                                                                                                                    & 95.75\%                                                                                                                                    \\ \hline
+Commons-text          & 55                   & 79.29\%                           & 77.92\%                                                & 78.22\%                            & 61.80\%                                                                                                                                    & 93.92\%                                                                                                                                    \\ \hline
+Commons-codec         & 17                   & 84.56\%                           & 79.46\%                                                & 77.73\%                            & 74.39\%                                                                                                                                    & 81.66\%                                                                                                                                    \\ \hline
+Commons-csv           & 32                   & 85.16\%                           & 82.84\%                                                & 83.47\%                            & 65.58\%                                                                                                                                    & 97.80\%                                                                                                                                    \\ \hline
+Commons-lang          & 28                   & 78.31\%                           & 76.83\%                                                & 76.99\%                            & 73.03\%                                                                                                                                    & 80.56\%                                                                                                                                    \\ \hline
+Jfreechart            & 28                   & 27.02\%                           & 26.41\%                                                & 74.11\%                            & 71.83\%                                                                                                                                    & 76.36\%                                                                                                                                    \\ \hline
+Jodatime              & 25                   & 70.30\%                           & 65.17\%                                                & 64.75\%                            & 60.82\%                                                                                                                                    & 69.81\%                                                                                                                                    \\ \hline
+Zxing                 & 28                   & 70.64\%                           & 80.03\%                                                & 83.11\%                            & 80.10\%                                                                                                                                    & 86.03\%                                                                                                                                    \\ \hline
+\end{tabular}
+\caption{\label{tab:ml_no_distance}Results of clustering mutants with the fuzzy c-means model without Levenshtein distance}
+\end{table*}
+
+Figure \ref{box:clustering_fcm_no_distance_25} displays the box-plots of the weighted mutation score obtained from each individual sample. 
+We can observe that for every box-plot the \textit{p-value} is more than 0.05 or 5\%.
+Table \ref{tab:ml_no_distance} displays the results of clustering mutants without Levenshtein distance.
+The maximum and minimum differences between the score of a full set and clustered set are 21.95\% and 0.61\% respectively.
+The average of the differences between full set score and clustered score is 4.44\% which is bigger than the average difference than that of the results with all characteristics.
+
+\section{Discussion}
+The maximum deviation from the original mutation score of the experiment with hierarchical clustering, is below the statistical threshold.
+We can observe that the distance between the whiskers of the box-plots do not stretch more than the \textit{alpha-value} of 0.05.
+This means that there is not one sample that deviated more than 5\% from the original mutation score.
+Thus we can reject the null hypothesis and accept our alternative hypothesis.
+As hypothesised we observe that it is possible to cluster mutants with the set of characteristics we identified while maintaining effectiveness and reducing the amount executed.
+The hypothesis holds for all the projects and variants of the experiment.
+66\% of the displayed results have a deviation of less than 1\%. 
+This includes both variants of the experiment.
+\newline
+The results prove that the relations between the mutants represented by the characteristics we identified can be used to cluster mutants efficiently. 
+This can be explained by the correlation between the result of a mutant, the calculated euclidean distance and the distance threshold(linkage criteria) used to decide whether a mutants should be clustered together.
+\newline
+According the rules of logic, a higher cluster accuracy means a score closer to that of the original.
+In other words if all clusters had an accuracy of 100\% we will achieve the same mutation score as that of the full set.
+However we do not observe this behaviour in our results.
+This is due to our random sampling method.
+If we would select a mutant in the majority of each cluster the accuracy per cluster would correlate to the deviation in mutation score.
+Unfortunately we do not know the majority of a cluster as we do now know if a mutant survives or is killed before execution.
+\newline
+The maximum deviation from the original mutation score of the experiment with the fuzzy c-means model, is above the statistical threshold.
+We can observe that the distance between whiskers of the box-plots is more than the \textit{alpha-value} of 0.05.
+Thus we cannot reject the null hypothesis and accept our alternative hypothesis.
+The biggest deviation measured is 22.39\%.
+This is significantly higher than the biggest deviation measured in the first experiment.
+56\% of the results of both variants had a deviation in mutation score bigger than 5\%.
+Our hypothesis does not hold for the clustering the mutants with the fuzzy c-means model trained in our experiment.
+\newline
+We observe that the number of clusters does not have an obvious relation to the total amount of mutants. 
+This is expected as the subtractive clustering does not look at the total amount of data points(mutants) but to the contents(characteristics) of the data points.
+The number of clusters compared to the that of the hierarchical clustering is significantly lower.
+While we do not know the exact clusters centroids the subtractive model calculated, a low number of clusters indicates close relationships between certain mutants.
+\newline
+We observe that the minimum lowest minimum cluster accuracy measured over all variants and projects is 50.54\% with and average of 62\%. 
+This is higher than the minimum of the hierarchical clustering.
+This means that in all project there was not one cluster that was evenly divided.
+While the minimum accuracy was higher, the maximum accuracy was lower compared to the maximum accuracy of the hierarchical clustering results.
+The highest accuracy measured over all variants and projects is 97.80\% with an average of 83.79\%
+This means that in all project there was not one cluster that consisted of only survived or killed mutants.
+This is caused by the relatively low number of clusters.
+One cluster contains more mutants which results in a bigger weight per cluster.
 \section{Related work}
-\label{sec:rw}
+\label{ch:related_work}
+Clustering mutants has been done before.
+The research that exists tries to cluster mutants by defining the centroid of a cluster and cluster mutants on that definition.
+Our research extends mutation clustering by defining a centroid that is dynamic.
+In other words the centroids of our clusters are defined by the mutants itself.
+
+\subsection{Clustering overlapped mutants}
+\label{ch:overlapping_mutants}
+Ma et al., clustered mutants by overlap\cite{Ma2016}.
+They defined the term overlap as as mutants that are functionally equivalent.
+This is close to an equivalent mutant but not the same.
+They explain that an equivalent mutants is functionally identical to the original source code while an overlapped mutant is functionally identical to at least one other mutant.
+If a mutant does not have an overlapped mutant that mutant would be a cluster on its own.
+The achieved a reduction, in mutants executed, of 10\%.
+They write in their research that it is limited to the mutator operators on expression level. 
+If a mutator did not generate two or more mutants they could not detect overlapping mutants.
+Their future work includes adding more mutators and widening the scope level to statements or blocks.
+\newline
+The research does not include efficiency.
+While they show that they can cluster mutants based on overlap they do not show how efficient the mutation testing process is compared to executing a full set of mutants.
+A threat to their validity, which is not addressed, is the fact that two overlapping mutants may not be challenged by the same tests and thus can result in different results.
+Our research solves this problem by taking into account the location of the mutants.
+
+\subsection{Clustering by scope}
+Yu et al., extended the research about overlapping mutants(see chapter \ref{ch:overlapping_mutants})\cite{Yu2019PossibilityScope}.
+They extended the scope by adding more mutators.
+The also extended the research to include statements and blocks.
+Their results shows an increase in clustered mutants.
+The research of Yu et al., suffers from the same limitations and threats  mentioned in Chapter \ref{ch:overlapping_mutants}.
+
+\subsection{Clustering Hamming distance}
+Ji et al., did a qualitative domain analysis on mutants\cite{Ji2009}.
+As a result they identified the Hamming distance to cluster with. 
+The Hamming distance is, like the Levenshtein distance, a similarity measure.
+They use the k-means algorithm to cluster mutants represented by Hamming distance.
+They write that the reduced test set in their experiment is still as strong as the original test set\cite{Ji2009}.
+\newline
+Ji et al., were successful in clustering mutants with Hamming distance and the k-means algorithm.
+They do acknowledge that their research still has problems with rationally determining the domains of their clusters.
+Our research does not have this problem, as the centre of the domains are decided by the characteristics of the mutants.
+
+\subsection{Spectral clustering}
+Wei et al., makes use of an intelligent technique, namely spectral clustering, to improve the efficacy of mutant reduction\cite{Wei2021SpectralTesting}.
+They defined multiple definitions for the mutants; similarity distance, distance between the mutants and a killing matrix of the mutants.
+With these definitions they reduced the mutants based on their proposed reduction method.
+This method is based on spectral clustering (SCMT), the determination method of the number of clusters, spectral clustering of mutants, and selection of representative mutants.
+The reduced mutants were then clustered with a classical clustering algorithm.
+Their results are promising and show high cluster accuracy.
+They write that there is still work left to do in optimizing the matrices and clustering algorithms.
+Our goal overlaps with that of Wei, the difference is that we used a different methodology to achieve that goal.
+Another difference is that our research does not require to know the result of a mutant(killed or survived).
+
+\subsection{Clustering similarity}
+Hussain et al., used the k-means and agglomerative clustering algorithm to cluster mutants according to a similarity measure. 
+They used the Hamming distance as similarity measure.
+They calculated the distance and used this as data to feed into the clustering algorithms.
+The number of clusters and the initial position of the cluster center in the k-means algorithm are difficult to determine, and the process of the agglomerative clustering algorithm makes it difficult to correct the existing cluster formation\cite{Hussain2008}.
+They did take into account efficiency.
+Hussain et al, had the same goals as our research. 
+The difference is that our research used different and more elaborated methods to cluster mutants.
+Hussain et al., calculated the mutation score by counting one cluster as one mutant.
+Our research calculated a weighted mutation score to reflect the clusters more accurately.
+
+\subsection{Generalizing mutants}
+Wilinski et al., tried to generalize the mutants by defining metrics.
+Each metric is calculated for a specific mutator.
+The three metrics are usefulness, frequency and dependency.
+Combining the results of usefulness and frequency metrics, they observed that reducing the number of generated mutants gives noticeable cost reduction without a loss of the mutation score accuracy.
+Wilinski et al., their research is narrowly scoped to the specific mutators they decided to do research on.
+Our research does not contain this limitation.
+The mutator is used as a characteristic in our research instead of limiting our research it helps define a mutant.
 
 \section{Conclusion}
-\label{sec:conclusion}
+\label{ch:conclusion}
+Our research started with comparing three mutation testing tools for Java and found that the state of the art was outdated.
+We replicated that research with data currently available on the mutation testing tools.
+This new information showed that two out of three mutation testing tools were not actively maintained anymore.
+The mutation testing tool PIT scored best on the metrics measured in our replication study, this is also the tool we decided to use for our research.
+\newline
+We performed a qualitative analysis on mutant characteristics.
+As a result we created a list of identified characteristics based on existing research. 
+We then extended this list by identifying characteristic with logic and data available.
+We used the identified characteristics to represent the mutants.
+\newline
+We clustered the mutants with a white box approach and a black box approach.
+For our white box approach we selected the hierarchical clustering algorithm.
+The hierarchical clustering showed that we can cluster mutants while maintaining effectiveness.
+We achieved reductions of up to 75\% of the total amount of mutants without significantly reducing the accuracy.
+\newline
+For the black box approach we selected a machine learning model, specifically the fuzzy c-means model.
+The fuzzy c-means model showed a significant reduction in effectiveness.
+We trained the fuzzy c-means model with parameters that were optimal according to existing research. 
+We used a different machine learning model, specialized in calculating the number of clusters for fuzzy clustering, to calculate the number of clusters.
+While the minimum clustering accuracy was higher than that of the hierarchical clustering, the maximum cluster accuracy was lower than that of the hierarchical clustering.
+The cluster accuracy had a larger impact on the clusters of the machine learning model because there was a relatively small number of clusters.
+Our research proves that it is possible to determine the means of domains as the centre of a cluster within a programming language without narrowing down the scope to specific pieces of code.
+The centroids our research identified is decided by the characteristics of the mutants.
+It can be applied to all Java projects that can be run with PIT. 
+Which is a larger scope than in existing mutation clustering research.
 

### 2022-01-25 mutation-workshop/paper.tex
wip until now

--- 2021-11-29 mutation-workshop/paper.tex
@@
 
-\title{Mutation testing: Clustering the mutants}
+\title{Towards clustering mutants to meet CI/CD timelines}
 
 \author{
+\IEEEauthorblockN{Rasjaad Basarat}
+\IEEEauthorblockA{\textit{Software Engineering Master} \\ %% TODO: CHANGE THIS
+\textit{University of Amsterdam}\\
+Amsterdam, the Netherlands \\
+rasjaad@nivero.io}
+
+\and
 \IEEEauthorblockN{Ana Oprescu}
-\IEEEauthorblockA{\textit{Parallel Computing Systems} \\
+\IEEEauthorblockA{\textit{Complex Cyber Infrastructure} \\
 \textit{University of Amsterdam}\\
@@
 a.m.oprescu@uva.nl}
-
-\and
-\IEEEauthorblockN{Rasjaad Basarat}
-\IEEEauthorblockA{\textit{Parallel Computing Systems} \\ %% TODO: CHANGE THIS
-\textit{University of Amsterdam}\\
-Amsterdam, the Netherlands \\
-rasjaad@nivero.io}
 \and
 \IEEEauthorblockN{Thomas Biesaart}
-\IEEEauthorblockA{\textit{Parallel Computing Systems} \\ %% TODO: CHANGE THIS
+\IEEEauthorblockA{\textit{} \\ %% TODO: CHANGE THIS
 Amsterdam, the Netherlands \\
@@
 \begin{abstract}
-Mutation testing is a computationally expensive technique that requires the ability to run tests.
-Different techniques have been proposed to reduce the cost of mutation testing.
+Mutation testing is a computationally expensive technique that requires the ability to run test suites.
+Mutation testing runs can take hours or days depending on the size of the project.
+Various techniques have been proposed to reduce the cost of mutation testing.
 One of these techniques is the clustering of mutants.
-By clustering mutants we can execute less mutants to reduce the cost.
-Different techniques have been proposed to reduce the cost of mutation testing.
-One of these techniques is the clustering of mutants.
-Our research consists of two parts in which we cluster mutants.
-To cluster mutants we need a mutation testing tools.
-Three mutation testing tools for Java are compared and one is selected for further use in this research.
-The first part consists of a white box approach and the second part of a black box approach.
-Our research uses the mutation testing tool PIT.
-We identified characteristics to represent the mutants such that we can cluster them.
-For our white box approach we used hierarchical clustering and for our black box approach we trained the fuzzy c-means  model.
-We calculated a weighted mutation score and compared this with the mutation score of a full set of mutants executed.
-Results show that with hierarchical clustering we can reduce the amount of mutants executed while maintaining the effectiveness.
-The clusters generated by the machine learning model were less accurate and showed a significant decrease in effectiveness.
+By clustering mutants we can execute fewer mutants to reduce the cost.
+Our goal is to reduce the cost as much as possible without significantly reducing the effectiveness.
+
+Our research identifies characteristics to represent the mutants such that we can cluster them. 
+The characteristics selected indicate a relationship between mutants.
+
+We use two clustering approaches: a qualitative, white box, approach and a quantitative, black box, approach.
+The white box approach uses hierarchical clustering and for the black box approach we train a fuzzy c-means model.
+
+We look at existing comparisons between mutation testing tools Mujava, Major and PIT.
+The tools are compared on cost, effectiveness and execution time\cite{Kintis2016AnalysingStudy}.
+Our goal is to generate as much mutants such that we can include as much different mutants as possible.
+For this reason we extend the existing research with a measurement of mutants generated per tool. 
+With the existing and extended research reviewed we selected PIT.
+
+We cluster mutants and calculate a weighted mutation score and compare this with the mutation score of a full set of executed mutants.
+Results show that with hierarchical clustering we can reduce the amount of mutants executed while maintaining the test suite effectiveness.
+The clusters generated by the machine learning model were less accurate and showed a significant decrease in test suite effectiveness.
+
+% Our research consists of two parts in which we cluster mutants. To cluster mutants we need a mutation testing tools.
+% Three mutation testing tools for Java are compared and one is selected for further use in this research.
+% Our research uses the mutation testing tool PIT.
+% We identified characteristics to represent the mutants such that we can cluster them.
+% These characteristics have shown that there exists a strong relationship between mutants. 
+% We prove this by clustering them with two different approaches; a white box approach and a black box approach.
+% For our white box approach we used hierarchical clustering and for our black box approach we trained the fuzzy c-means model.
+% We calculated a weighted mutation score and compared this with the mutation score of a full set of mutants executed.
+% Results show that with hierarchical clustering we can reduce the amount of mutants executed while maintaining the effectiveness.
+% The clusters generated by the machine learning model were less accurate and showed a significant decrease in effectiveness.
 \end{abstract}
@@
 \section{Introduction}
-The high cost requirement is often a barrier for adopting mutation testing \cite{Pizzoleto2019}.
-A lot of techniques and methods have been developed to improve the performance, however, most of these approaches are not as effective as mutation testing a full set of mutants\cite{Pizzoleto2019,Yao2014}.
-\newline
-A technique to reduce the number of mutants or the number of test case executions is mutation clustering.
-Mutation clustering aims to reduce the amount of mutants to be executed by clustering mutants\cite{Ma2016,Yu2019PossibilityScope}.
-The clustering of mutants has been researched with promising results\cite{Ji2009,Wilinski2015,Ma2016}. For example Ma et al., \cite{Ma2016} clustered the mutants for expressions. 
-However in their paper they indicate that the choosing expressions as the domain of their cluster is a limitation.
-A recurring problem in the research of mutation clustering is to determine the means of domains as the centre of a cluster\cite{Ji2009,Wilinski2015,Ma2016,Wei2021SpectralTesting}.
-\newline
+The high temporal cost requirement is often a barrier for adopting mutation testing~\cite{Pizzoleto2019}.
+Many techniques and methods have been developed to improve the performance, such as excluding certain mutants from a mutation testing run.
+Most of these approaches are not as effective as testing a full set of mutants, which, on the other hand, can take hours or days depending on the size of the project~\cite{Pizzoleto2019,Yao2014}.
+
+Another technique to reduce the number of mutants, and thus the number of test case executions is mutation clustering~\cite{Ma2016,Yu2019PossibilityScope}, which has been researched with promising results~\cite{Ji2009,Wilinski2015,Ma2016}.
+However, the reduction in temporal cost is still insufficient \todo{give an indication of how it would take compared to how much it should take to fit the CI/CD typical timeline} when considering CI/CD pipelines.
+When developing in a team waiting hours for a mutation test to complete is not practical, it will slow down development significantly. 
+Thus the cost outweighs the benefits.
+
 To reduce the cost of mutation testing, we try to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants. 
-Our goal is to cluster any mutant that is generated within the Java programming language.
+Our goal is to cluster every mutant that is generated within the Java programming language.
 This research aims to remove the scoping limitations present in existing research.
-There should be no requirements for which mutants can be clustered.
-We do this with two different techniques: qualitative and quantitative.
-We devise a white and black box approach for clustering mutants.
-Our white box approach contains a qualitative analysis on mutants and a methodology to cluster them.
-The black box approach makes use of a machine learning model to cluster mutants.
-We structure our research on the following research questions:
-\newline
+There should be no mutant excluded from clustering to ensure the validity of the mutation score.
+We achieve this using two different techniques: a qualitative, white box approach, and a quantitative, black box approach.
+The white box approach contains a qualitative analysis on mutants and a methodology to cluster them.
+The black box approach makes use of a machine learning model to cluster mutants. We also analyse several mutation testing tools before selecting one for conducting the research in this paper.
+
+We structure our research on the following questions:
+
 \textbf{Research Question 1}: What set of characteristics can we identify for clustering generated mutants to reduce the amount executed, while maintaining effectiveness?
-\newline
-\textbf{Research Question 1.1}: How do the existing mutation testing tools for Java compare to each other?
-\newline
+
 \textbf{Research Question 2}: How can we train a machine learning model to recognize and cluster generated mutants to reduce the amount executed, while maintaining effectiveness?
@@
 
-%%%%%% DO RESEARCH PART HERE %%%%%%%%
+\section{Selecting a mutation testing tool}
+For this research, we need a tool that generates and executes mutants.
+Without a tool to execute and generate mutants we cannot calculate mutation scores. 
+Comparing the mutation scores gives us a measurement on the effectiveness of the clusters that we will generate.
+Our goal is to find the tool that generates the most amount of mutants.
+By not excluding any mutants we can ensure the effectiveness of mutation testing.
 
-\section{Selecting a mutation testing tool}
-We reviewed literature on three different mutation testing tools[CITE OWN PAPER HERE].
-However the most recent study is from 2017.
-At the time of writing this paper these studies are at least four years old. 
-Some of the tools were updated or are still under active development\cite{pit-releases,Major}.
-There are three candidate tools; Major, MuJava and PIT.
-We extend the existing mutant operator comparison with the operators that have been added since the publishing date of the existing literature.
-Additionally we add an overview of the overlapping mutant operators for the new mutant operators.
+We reviewed literature, by searching for state of the art research on the comparison of mutation testing tools for Java\cite{thesis}.
+There are three candidate tools \cite{Kintis2016AnalysingStudy,Marki2017MutationJava}: Major\cite{Major}, MuJava\cite{mujava} and PIT\cite{pit}.
+The most recent study is from 2017. 
+Some of these tools were updated or are still under active development\cite{pit-releases,Major}.
+We extend the existing comparison with the operators that have been added after the publishing date of the existing literature.
+Table~\ref{tab:mutators-current} shows the comparison of operators that were added.
 
-\todo{show tables here? maybe a summarized version of this?}
-\todo{the results can be found in ref to paper??}
+\begin{table*}[ht]
+\centering
+\begin{tabular}{|l|l|l|l|}
+\hline
+\multicolumn{2}{|c|}{\textbf{MuJava}}                           & \multicolumn{2}{|c|}{\textbf{PIT}}                      \\ \hline
+IHD & Hiding variable deletion                                  & EM                    & Empty Returns                   \\ \hline
+IHI & Hiding variable   insertion                               & FR                    & False Returns                   \\ \hline
+IOD & Overriding method   deletion                              & NR                    & Null Returns                    \\ \hline
+IOP & Overriding method   calling position change               & TR                    & True Returns                    \\ \hline
+IOR & Overriding method   rename                                & PR                    & Primitive Returns               \\ \hline
+ISI & Super keyword insertion                                   & ER                    & Experimental Switch             \\ \hline
+ISD & Super keyword deletion                                    & BI                    & Big Integer                     \\ \hline
+IPC & Explicit call to a   parent’s constructor deletion        & NRC                   & Naked Receiver                  \\ \hline
+PNC & New method call with   child class type                   & N                     & Negation                        \\ \hline
+PMD & Member variable   declaration with parent class type      & AOR                   & Arithmic Operator Replacement   \\ \hline
+PPD & Parameter variable   declaration with child class type    & AOD                   & Arithmic Operator Deletion      \\ \hline
+PCI & Type cast operator   insertion                            & CR                    & Constant replacement            \\ \hline
+PCC & Cast type change                                          & BO                    & Bitwise Operator                \\ \hline
+PCD & Type cast operator   deletion                             & ROR                   & Relational Operator Replacement \\ \hline
+PRV & Reference assignment   with other comparable variable     & UOI                   & Unary Operation Insertion       \\ \hline
+OMR & Overloading method   contents replace                     & \multicolumn{1}{l|}{} &                                 \\ \hline
+OMD & Overloading method   deletion                             & \multicolumn{1}{l|}{} &                                 \\ \hline
+OAC & Arguments of   overloading method call change             & \multicolumn{1}{l|}{} &                                 \\ \hline
+JTI & This keyword insertion                                    & \multicolumn{1}{l|}{} &                                 \\ \hline
+JTD & This keyword deletion                                     & \multicolumn{1}{l|}{} &                                 \\ \hline
+JSI & Static modifier   insertion                               & \multicolumn{1}{l|}{} &                                 \\ \hline
+JSD & Static modifier   deletion                                & \multicolumn{1}{l|}{} &                                 \\ \hline
+JID & Member variable   initialization deletion                 & \multicolumn{1}{l|}{} &                                 \\ \hline
+JDC & Java-supported default   constructor deletion             & \multicolumn{1}{l|}{} &                                 \\ \hline
+EOA & Reference assignment   and content assignment replacement & \multicolumn{1}{l|}{} &                                 \\ \hline
+EOC & Reference comparison   and content comparison replacement & \multicolumn{1}{l|}{} &                                 \\ \hline
+EAM & Acessor method change                                     & \multicolumn{1}{l|}{} &                                 \\ \hline
+EMM & Modifier method change                                    & \multicolumn{1}{l|}{} &                                 \\ \hline
+\end{tabular}
+\caption{\label{tab:mutators-current}New mutators not taken into account in previous research.}
+\end{table*}
+There is some overlap between the new mutators of MuJava and PIT.
+These mutators have different names but mutate the code in the same way.
+Table \ref{tab:mutators-overlap} shows an overview of overlapping mutators.
+
+\begin{table*}[ht]
+\centering
+\begin{tabular}{|l|l|l|l|}
+\hline
+\multicolumn{2}{|c|}{\textbf{ MuJava mutator}}      & \multicolumn{2}{c|}{\textbf{PIT mutator}}                         \\ \hline
+AORB & Arithmetic Operator Replacement Binary      & \multirow{4}{*}{M}   & \multirow{4}{*}{Math}                      \\ \cline{1-2}
+ASRS & Short-Cut Assignment   Operator Replacement &                      &                                            \\ \cline{1-2}
+SOR  & Shift Operator   Replacement                &                      &                                            \\ \cline{1-2}
+COR  & Conditional Operator   Replacement          &                      &                                            \\ \hline
+AOIU & Arithmetic Operator   Insertion Unary       & \multirow{2}{*}{UOI} & \multirow{2}{*}{Unary Operation Insertion} \\ \cline{1-2}
+AOIS & Arithmetic Operator   Insertion Short-cut   &                      &                                            \\ \hline
+ODL  & Operator Deletion                           & AOD                  & Arithmic Operator   Deletion               \\ \hline
+AODS & Arithmetic Operator   Deletion Short-cut    & RI                   & Remove Increments                          \\ \hline
+ROR  & Relational Operator   Replacement           & CB                   & Conditionals Boundary                      \\ \hline
+AORS & Arithmetic Operator   Replacement Short-Cut & I                    & Increments                                 \\ \hline
+COD  & Conditional Operator   Deletion             & RC                   & Remove Conditionials                       \\ \hline
+
+\end{tabular}
+\caption{\label{tab:mutators-overlap}Overview of mutator overlaps between  MuJava and PIT}
+\end{table*}
+
+Our goal is to cluster as many mutants as possible, we note how many mutants each tool generates on the same source (numbers shown in Table~\ref{tab:mutants-generated}).
+\begin{table}[ht]
+\centering
+\begin{tabular}{|l|l|c|c|c|}
+\hline
+                                & \textbf{version} & \textbf{\begin{tabular}[c]{@{}c@{}}LOC\end{tabular}} & \multicolumn{1}{l|}{\textbf{MuJava}} & \multicolumn{1}{l|}{\textbf{PIT}} \\ \hline
+Commons-numbers.core            & 1.0-beta1        & 450                                                                    & 988                                  & 4384                              \\ \hline
+Jodatime                        & 2.10.10          & 28811                                                                  & 52925                                & 112772                            \\ \hline
+Zxing 3.4.1                     & 3.4.1            & 24792                                                                  & 65983                                & 161409                            \\ \hline
+Google Auto Common              & 0.11.0           & 2338                                                                   & 20                                   & 5219                              \\ \hline
+Google Auto Factory             & 1.0-beta9        & 1507                                                                   & 69                                   & 5832                              \\ \hline
+Google Auto Value               & 1.7.5            & 7466                                                                   & 745                                  & 16746                             \\ \hline
+Google Auto Service             & 1.0-rc6          & 750                                                                    & 0                                    & 475                               \\ \hline
+ScribeJava-Core                 & 8.1.0            & 5709                                                                   & 1358                                 & 5746                              \\ \hline
+Checkstyle                      & 8.41.1           & 38491                                                                  & 4529                                 & 100952                            \\ \hline
+Fastjson                        & 1.2.75           & 43405                                                                  & 70577                                & 116188                            \\ \hline
+Jfreechart                      & 1.5.3            & 91876                                                                  & 0                                    & 350741                            \\ \hline
+Commons-lang                    & 3.12.0           & 29836                                                                  & 31145                                & 134764                            \\ \hline
+Commons-codec                   & 1.15.0           & 9656                                                                   & 21719                                & 54804                            \\ \hline
+Commons-text                   & 1.9.0             & 9781                                                                   & 10403                                & 48490                            \\ \hline
+Commons-io                     & 2.8.0             & 13947                                                                  & 13245                                & 44631                            \\ \hline
+Gson                           & 2.8.6             & 8078                                                                   & 9198                                 & 28485                            \\ \hline
+Commons-cli                    & 1.4.0             & 2782                                                                   & 831                                  & 7193                             \\ \hline
+Commons-csv                    & 1.8.0             & 1855                                                                   & 2081                                 & 6906                             \\ \hline
+\end{tabular}
+\caption{\label{tab:mutants-generated}Amount of mutants generated by MuJava and PIT, LOC = Lines of code}
+\end{table}
 
 MuJava is not actively maintained and has not been updated in the last few years.
-It does not support JUnit 4 and all versions of TestNG\cite{mujava}.
-These are test engines and are crucial for executing tests.
-MuJava also does not support source projects with java version 1.7 or higher\cite{mujava}.
-Conforming to these requirements would reduce the set of projects we could use in or experiments.
-\newline
+It does not support source projects with java version 1.7 or higher, nor JUnit 4 or any version of TestNG~\cite{mujava}.
+These are test engines and are crucial for executing tests..
+All these constraints would reduce the set of projects we could use in our experiments.
+
 While Major supports JUnit 4,
-we did not succeed in generating mutants with this tool[CITE OWN PAPER HERE].
-It would be too time consuming to customize all source projects to work with Major.
-Furthermore there is not much documentation available for this tool.
-\newline
-PIT targets the industry, is actively maintained and is open source\cite{Kintis2016AnalysingStudy}.
+we did not succeed in generating mutants with this tool~\cite{thesis}.
+It would require customizing each source project, and
+there is not much documentation available for this tool.
+
+PIT is industry-minded, actively maintained and open source~\cite{Kintis2016AnalysingStudy}.
 It supports Maven, Gradle, has a command line interface and has a faster execution time than the other tools.
-For example PIT provides a plugin system in which you can inject your own code in various stages of mutation testing process\cite{pit}.
-PIT also generated significantly more mutants in every project we have tested[CITE OWN PAPER HERE].
-Based on the information presented in this chapter we decide that PIT is the best choice for generating and executing mutants for this research.
+For example PIT provides a plugin system in which you can inject your own code in various stages of mutation testing process~\cite{pit}
+We can use this feature to adjust the mutation process and include our experiments.
+PIT also generated significantly more mutants(see Table \ref{tab:mutants-generated}) in every project we have tested~\cite{thesis}.
+We therefore decide to use PIT for generating and executing mutants in this research.
 
 \section{Qualitative clustering approach}
+\label{ch:reserach_question1}
+Our goal is to cluster every generated mutant while maintaining effectiveness.
+We perform a qualitative analysis on mutants to identify characteristics that represent the mutants and we devise a methodology for clustering mutants based on these characteristics.
+
+\subsection{Identifying characteristics}
+\label{ch:identifying_characteristics}
+Zhang et al. identify several mutant characteristics~\cite{Zhang2019PredictiveTesting}. 
+The goal of their research is detecting as accurately as possible whether a mutant survives before executing the mutant.
+The mutant characteristics identified by Zhang et al., were used in other research with promising results~\cite{Oonk2021}.
+Our goal is to group mutants based on characteristics, this is different than that of Zha, however we can still use these characteristics for clustering.
+We select characteristics relevant to our research from the research of Zhang et al., and Oonk~\cite{Oonk2021}.
+We extend the list of characteristics with data we can extract from the selected mutation testing tool.
+
+PIT generates and executes mutants on byte code level~\cite{pitestBytecode}.
+By mutating in Java byte code we can identify characteristics specific to Java byte code.
+PIT also gathers mutant metadata.
+We can extract this mutant metadata and identify more characteristics.
+With the combination of information extracted from PIT and from existing research we identify the following characteristics~\cite{thesis}:
+\begin{enumerate}
+    \item Mutant similarity.
+    \item Amount of tests the mutant is challenged by~\cite{Zhang2019PredictiveTesting, Oonk2021, pit}.
+    \item Mutator identifier~\cite{pit}.
+    \item Mutant opcode~\cite{Oonk2021, pit}.
+    \item Mutant return type~\cite{Zhang2019PredictiveTesting, Oonk2021, pit}.
+    \item Mutant contains local variables~\cite{pit}.
+    \item Mutant is in try catch block~\cite{pit}.
+    \item Mutant is in finally block~\cite{pit}.
+    \item Name of class that contains the mutant~\cite{Oonk2021, pit}.
+    \item Name of function that contains the mutant~\cite{Oonk2021, pit}.
+    \item Line number of the start of the block that contains the mutant~\cite{pit}.
+    \item Line number on which the mutation occur~\cite{pit}.
+\end{enumerate}
+
+\subsection{Hierarchical clustering}
+Research shows that hierarchical clustering performs better when clustering with at least ten features and that varying the parameters of hierarchical clustering improves the performance compared to that of the default settings of the algorithm~\cite{Rodriguez2019}.
+
+We identified more than ten characteristics and can use them as features. We adjust the parameters of the algorithm based on the characteristics in our data set.
+Based on these observations and prior research, we select hierarchical clustering\todo{do you have experiments showing best performance in terms of partitioning?}.
+% as it gives us the best performance in terms of partitioning.
+
+Next we explain the configuration we use for clustering the mutants. 
+Hierarchical clustering is subdivided into agglomerative and divisive. 
+The agglomerative hierarchical technique follows a bottom up approach whereas divisive follows top-down approaches.
+Hierarchical clustering uses different metrics which calculates the Euclidean distance between two clusters and the linkage criteria~\cite{Rajalingam2011}. 
+The linkage criteria specifies the dissimilarity in the sets as a function of the pair-wise distances of observations in those sets.
+
+Research shows that the complete linkage outperforms the single linkage method~\cite{Vijaya2019ComparativeClustering}.
+The ward linkage and complete linkage methods perform the same when clusters are well separated, however, if the clusters overlap the ward linkage outperforms the complete linkage~\cite{Vijaya2019ComparativeClustering}.
+
+Each mutant is represented by its characteristics, the mutants start in their own cluster of size one.
+Starting out with each mutant as a separate cluster we can use the agglomerative form of hierarchical clustering.
+Since we cannot assume that our clusters are well separated we chose to use the ward linkage method for our clustering algorithm.
+In summary, we select the agglomerative hierarchical clustering algorithm with ward linkage to cluster mutants represented by the characteristics we gathered\todo{perhaps a figure would help explain it better?}.
+
+\subsection{Categorical data}
+There are different categorical variable encoding techniques available~\cite{Potdar2017AClassifiers}.
+The selected categorical characteristics have  no specific order and no particular ranking compared to each other, e.g., a return type void is not better or worse than a return type string.
+% There is also no specific order to the characteristics.
+The same applies to the location characteristics, there is no location that should have a bigger weight than the other locations.
+
+The individual characteristics do contain a finite set of values.
+For example multiple mutants may contain the same class name.
+The hierarchical clustering algorithm needs all characteristics in a numerical form~\cite{Vijaya2019ComparativeClustering}.
+The characteristics mutator identifier, class name, method name and return type are non numerical.
+To deal with this problem we apply categorical variable encoding to these specific features.
+Taking into account the properties of our categorical characteristics the nominal variable encoding fits our requirements.
+Nominal encoding comprises a finite set of discrete values with no relationship between values~\cite{Potdar2017AClassifiers}.
+Therefore we implement this type of encoding in the experiment\todo{if time allows, revisit for clarity between categorical variable encoding and nominal variable encoding: which do we use? or both?}.
+
+\subsection{Number of clusters}
+Agglomerative hierarchical clustering can continue to cluster until there is one cluster left. 
+Naturally this cluster will contain all the mutants.
+We can cut off the clustering algorithm at any point.
+We decide on the number of clusters based on the amount of mutants generated in a full set.
+We select the number of clusters from a performance increasing perspective.
+Mutation clustering increases the performance by reducing the amount of mutants executed~\cite{Pizzoleto2019}.
+By reducing the amount of mutants, for example by half, we increase the performance by 50\%~\cite{Just2014AreTesting}.
+The amount of mutants inside a cluster and the number of clusters is be decided by the clustering algorithm.
+As a starting point we perform three experiments where the cluster size is 25\%, 50\% and 75\% of the total amount of mutants respectively.
+
 \section{Quantitative clustering approach}
-\section{validation}
+\label{ch:reserach_question2}
+In the following sections we devise a black box approach for clustering mutants.
+While our goal remains the same as in Section~\ref{ch:reserach_question1}, the method to achieve it is different.
+We select a machine learning model by first choosing between supervised or unsupervised learning.
+Then we look at the machine learning models that are discussed in the survey conducted by K.L. Du~\cite{Du2010Clustering:Approach}\todo{this is a 10 year old survey, why not something more recent?}.
+We reuse of the characteristics identified in 
+Section~\ref{ch:identifying_characteristics}.
 
+\subsection{Supervised or unsupervised learning}
+Supervised machine learning models require a labeled data set.
+We could label our data set with the results of our first approach.
+Each mutant would then be labeled with the identity of the cluster assigned by the hierarchical clustering model.
+If we would train a supervised model with the data labeled as described, it may result in a model that can do the same as hierarchical clustering.
+However, we did not set out to prove whether we can train a model to cluster mutants the same way as hierarchical clustering.
+The goal is to train a model to find the relations between the characteristics without having to define these relations manually.
+
+Unsupervised learning models do not need a labeled data set and can recognize relations or discover hidden patterns in the characteristics we identified.
+This fits better the goal of our research and
+therefore we choose unsupervised learning.
+
+\subsection{Self organizing map, C-means and fuzzy c-means}
+\label{ch:topThreeModels}
+We reviewed the models discussed by K.L. Du~\cite{Du2010Clustering:Approach, thesis} and select three models which best fit our use case:
+the self-organizing map, C-means and fuzzy c-means. 
+To choose between these models we look at their performance. Mingoti et al. compared the performance of these three models~\cite{Mingoti2006ComparingAlgorithms}.
+% They performed a Monte Carlo simulation where the cluster sizes and amount of random numbers in the data set varied each simulation.
+Other variables such as cluster boundaries and overlap were also controlled variables in the experiment.
+They observe that the C-means and fuzzy c-means had good performance for non overlapping situations.
+The best results for average recovery and internal dispersion rates were found for fuzzy c-means which was stable in all situations achieving recovery averages over 90\%\cite{Mingoti2006ComparingAlgorithms}
+% The C-means method was affected by the presence of a large amount of outliers.
+The best results for average recovery and internal dispersion rates were found for fuzzy c-means which was stable in all situations achieving recovery averages over 90\%.
+The C-means method was affected by the presence of a large amount of outliers, and the self-organizing map did not perform well in many cases being affected by the amount of variables and clusters even for the non overlapping cases.
+We therefore select the fuzzy c-means model.
+
+\subsection{Training strategy}
+\label{ch:training_strategy}
+Our data set consists of mutants represented by their characteristics. 
+The characteristics of the mutants are generated per project and have a project specific context.
+For example, a mutant with a mutator in a certain project can have a different effect on the source code and can give other results than a mutator in a different project.
+The relations between the mutants are scoped to their own project. 
+In other words if a specific mutant has a relation to another mutant in the same project, this relation may not exist between the same mutant and a mutant from a different project.
+To preserve these relations and context we decided to train a separate model per project.
+
+To validate each model we need to split our data set into a training set and a validation set.
+Research has shown that if the data set is big enough a ratio of 80/20 is sufficient~\cite{Guyon1997ARatio}.
+We divide each source projects into training and validation groups to reflect the 80/20 ratio as best as possible.
+Depending on the amount of mutants per project we might not achieve a perfect 80/20 ratio.
+The division of the mutants is done randomly, using a controlled random seed (see Section~\ref{ch:exp1_validation}).
+% that is included in the results for reproducibility (file \textit{hierarchical-clustering/main.py}, line 182~\cite{rbasarat-repo}).
+
+\subsection{Algorithm parameters}
+\label{ch:algorithm_parameters}
+The fuzzy c-means algorithm has to initialize the centroids of the clusters. The implementation library uses a random number that can be controlled.
+% To make the initialization of the centroid reproducible we use the same generated seeds as mentioned above.
+
+Bezdek et al. also researched cluster validity for the fuzzy c-means model~\cite{Bezdek1995OnModel}.
+They researched what effect which parameters could have on the validity of the clusters and 
+% . They specifically analyzed the role of weighting exponent \textit{m}(fuzziness parameter).
+concluded that the best choice for \textit{m} is in the interval [1.5, 2.5], whose mean and midpoint is \textit{m}=2.
+Wu proposed a new guideline for selecting parameter \textit{m}~\cite{Wu2012}.
+His point of view was that a large \textit{m} value will make the fuzzy c-means more robust to noise and outliers.
+He suggested implementing fuzzy c-means with \textit{m}=[1.5, 4].
+When the data set contains noise and outliers, the fuzzifier \textit{m}=4 is recommended in a theoretical upper bound case.
+Since our data set does not contain noise(all mutants consists of every characteristic) we choose to select the midpoint as recommended by Bezdek et al.~\cite{Bezdek1995OnModel}.
+The value of \textit{m} in this experiment is 2.
+
+Finding the correct amount of clusters has been an ongoing problem for clustering algorithms~\cite{Du2010Clustering:Approach}.
+Fortunately there are techniques available to estimate the number of clusters.
+The subtractive clustering model is specifically designed to approximate the number of clusters for fuzzy system models~\cite{Chiu1994FuzzyEstimation}.
+It approximates the centroids and amount of the clusters based on the data points in the data set~\cite{Chiu1994FuzzyEstimation}.
+We select and implement the subtractive clustering model to estimate the number of clusters.
+We use this number as the value for the number of clusters parameter in the fuzzy c-means algorithm.
+
+\section{Experimental setup and design}
+We design an experiment to cluster mutants and measure the mutation score with the clustered set. 
+We execute the following steps:
+\begin{enumerate}
+    \item Gather all mutant characteristics from PIT.
+    \item Execute full set of mutants to gather baseline measurement.
+    \item Cluster the mutants according the black and white box methodologies.
+    \item Execute one random mutant per cluster and gather result.
+    \item Calculate weighted mutation score.
+\end{enumerate}
+Both research questions have the same goal, but with different approaches.
+We devise one hypothesis for both research questions: Mutants can be clustered with the set of selected characteristics while maintaining effectiveness, thus reducing the amount executed by executing one mutant from each cluster that is randomly selected.
+Our goal is to achieve a weighted mutation score that is as close as possible to the mutation score of a full set of executed mutants.
+We repeat the experiment for both the white and black box approaches.
+
+\subsection{Baseline measurement}
+The baseline for our experiment is the mutation score of the set of all mutants generated by PIT.
+% To extract this score we need to execute a full set of mutants within PIT.
+For each source we have added the PIT plugin and required configuration in the \textit{pom.xml}, this code instructs PIT to generate and execute mutants with all mutators available.
+We use the command line interface and a maven command(see Listing \ref{lst:mvn-full}) to execute PIT.
+\begin{lstlisting}[label=lst:mvn-full,caption=Command to execute full set of mutants with PIT.]
+mvn -U org.pitest:pitest-maven:mutationCoverage
+\end{lstlisting}
+
+\subsection{Weighted mutation score}
+\label{ch:weighted_score}
+Our hypothesis entails that each executed mutant should represent that whole cluster.
+With a mutant executed from each cluster we can calculate a weighted mutation score as the product of the result of a mutant (1 for killed and 0 for survived) and the amount of mutants in the cluster it represents.
+The weighted mutation score is then comparable to the score of a full set as the total number of mutants will be the same.
+
+For example, take a full set with a score of 75/100 killed mutants. This gives us a mutation score of 75\%. 
+Assume mutants are clustered in four clusters consisting of 12, 30, 38 and 20 mutants, respectively.
+We randomly select one mutant of each cluster(four in total) and execute them. The mutants representing cluster one and four survive, while two and three are killed.
+If we calculate the weighted score we get 68/100 which is 68\%.
+We can then compare the weighted score to the score of a full set because the amount of mutants in each score is the same: 75/100(75\%) and 68/100(68\%).
+
+\subsection{Random mutant selection}
+\label{ch:mutant_selection}
+To validate the random selection of mutants we make use of statistical hypothesis testing~\cite{Emmert-Streib2019UnderstandingInference}.
+Our null hypothesis states that there is no relation between the characteristics identified and the results of a mutant with the chosen methodology.
+To test this hypothesis we select an alpha value of 0.05.
+% To translate: if more than five percent of the values in a sample deviates more than the significance level(see Chapter \ref{ch:exp1_validation}) we cannot reject the null hypothesis.
+To make our results reproducible we select the random mutant based on a generated seed.
+The seed generated for each run will be included in the results and can be found in the Github repository in the file \textit{hierarchical-clustering/main.py} at line 182\cite{rbasarat-repo}.
+We repeat our experiment 30 times with 30 different seeds.
+Achieving consistent results while applying random selection contributes to the validity of the experiment.
+
+\subsection{Evaluation}
+\todo[inline]{by the way, you mention nothing about collecting the runtimes?}
+\label{ch:exp1_validation}
+The accurate way to measure test effectiveness with mutation testing is by executing all mutants that a mutation testing tool can possibly generate.
+% The goal of this research is to reduce the amount of mutants executed while maintaining effectiveness.
+To reduce the amounts executed we cluster the mutants. 
+We measure the reduction factor by counting the number of clusters we generate and compare it to the number of total mutants generated by the selected tool.
+
+To validate how effective our method of clustering is we compare the weighted mutation score (see Section~\ref{ch:weighted_score}) of the clustered set to the mutation score of the full set.
+The closer the weighted score is to that of a full set the more effective our set of characteristics and clustering algorithm proves to be.
+% In other words our goal is to achieve a mutation score that is as close as possible as to that of a full set.
+We select a statistical significance level of $\leq 0.05$.
+This is the conventional threshold for declaring statistical significance~\cite{Kirk1996PracticalCome}.
+
+Depending on the performance of our clustering algorithm, in terms of clustering accuracy, we may lose accuracy in test effectiveness. 
+This can happen if a cluster contains mutants of both results; killed and sruvived. 
+We can measure the accuracy inside a cluster by calculating a percentage of all mutants that have survived against the ones that have been killed in a cluster.
+If the majority of the mutants in a cluster is killed then we consider that cluster to represent a killed cluster and the other way around for survived mutants.
+We consider the mutants that do not have the same result as the majority of results of the cluster as inaccuracy.
+If the weighted mutation score of our clustered set deviates by more than 5\% from the score of a full set we reject our hypothesis \todo{why 5?}.
+
+\section{Results}
+\label{ch:results}
+% In this chapter we present the results of our experiments.
+We divide the results in two sections.
+The first section presents the results of clustering with hierarchical clustering.
+The second section presents the results of clustering with the fuzzy c-means model.
+We select the Levenshtein distance for our similarity characteristic\cite{thesis}.
+While extracting the set of characteristics (Section~\ref{ch:identifying_characteristics}, we observed that calculating the Levenshtein distance is computationally expensive.
+% Calculating the Levenshtein distance for the four biggest projects in the data set cost days\cite{thesis}.
+% Even for the smaller projects the extraction of the characteristics cost hours\cite{thesis}.
+Including the Levenshtein distance defeats the purpose of increasing performance as every bit of performance we gain would be replaced by the cost of calculating the Levenshtein distance~\cite{thesis}.
+All results displayed are in percentages and ms=mutation score, ca=cluster accuracy.
+
+\subsection{Hierarchical clustering}
+We present the results of our clustering algorithm when executed without the Levenshtein distance.
+The results are grouped based on the reduced percentage of mutants executed.
+We also display the accuracy of the clusters per reduction amount.
+The data we present are the average of the results of the 30 repetitions of the experiment.
+The result of each single repetition can be found in our Github repository in the folder\textit{experiment{\_}results}~\cite{rbasarat-repo}.
+
+\begin{figure}[ht]
+\includegraphics[width=0.5 \textwidth]{images/boxplot_summary/boxplot_hc_no_distance_0.25.png}
+\caption{\label{box:clustering_no_distance_25}Box-plots containing weighted mutation score of clustering mutants without Levensthein distance and cluster size n*0.25}
+\end{figure}
+
+\begin{table*}[ht]
+\centering
+\begin{tabular}{|l|c|c|c|c|c|}
+\hline
+\multicolumn{1}{|c|}{Project} & \begin{tabular}[c]{@{}c@{}}ms\\ full set\end{tabular} & \begin{tabular}[c]{@{}c@{}}ms \\ clustered set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg.
+ca\end{tabular} & \begin{tabular}[c]{@{}c@{}}Min.
+ca\end{tabular} & \begin{tabular}[c]{@{}c@{}}ca\end{tabular} \\ \hline
+Google Auto Service           & 0.00                                                                                  & 0.00                                                                                      & 0.00                                                                               & 0.00                                                                                & 0.00                                                                               \\ \hline
+ScribeJava-Core               & 33.52                                                                                 & 33.72                                                                                     & 92.24                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+Google Auto Factory           & 48.41                                                                                 & 48.32                                                                                     & 94.21                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+Google Auto Common            & 59.59                                                                                 & 60.03                                                                                     & 91.19                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+Google Auto Value             & 61.42                                                                                 & 62.58                                                                                     & 93.90                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+Google Gson                   & 66.83                                                                                 & 66.89                                                                                     & 91.75                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+commons-io                    & 69.70                                                                                 & 66.08                                                                                     & 90.20                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+commons-cli                   & 79.07                                                                                 & 78.97                                                                                     & 91.11                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+commons-text                  & 79.29                                                                                 & 79.30                                                                                     & 90.95                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+commons-codec                 & 84.56                                                                                 & 84.59                                                                                     & 94.56                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+commons-csv                   & 85.16                                                                                 & 85.24                                                                                     & 92.76                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+\end{tabular}
+\caption{\label{tab:clustering_no_distance_25}Results of clustering mutants without Levensthein distance and cluster size n*0.25}
+\end{table*}
+
+Figure \ref{box:clustering_no_distance_25} displays the box-plots of the weighted mutation score obtained from each individual sample.
+We observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
+Table~\ref{tab:clustering_no_distance_25} displays the results of clustering mutants without Levensthein distance and the number of clusters equal to the total amount of mutants * 0.25. 
+We observe that the maximum and minimum differences in mutation score between a full set and a clustered set  3.61\%, and 0.09\%, respectively.
+The average of the differences between full set score and clustered score is 0.53\%.
+
+\begin{figure}[ht]
+\includegraphics[width=0.5 \textwidth]{images/boxplot_summary/boxplot_hc_no_distance_0.5.png}
+\caption{\label{box:clustering_no_distance_50}Box-plots containing weighted mutation score of clustering mutants without Levensthein distance and cluster size n*0.50}
+\end{figure}
+
+\begin{table*}[ht]
+\centering
+\begin{tabular}{|l|c|c|c|c|c|}
+\hline
+\multicolumn{1}{|c|}{Project} & \begin{tabular}[c]{@{}c@{}}ms\\ full set\end{tabular} & \begin{tabular}[c]{@{}c@{}}ms \\ clustered set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg.
+ca\end{tabular} & \begin{tabular}[c]{@{}c@{}}Min.
+ca\end{tabular} & \begin{tabular}[c]{@{}c@{}}ca\end{tabular} \\ \hline
+Google Auto Service           & 0.00                                                                                  & 0.00                                                                                      & 0.00                                                                               & 0.00                                                                                & 0.00                                                                               \\ \hline
+ScribeJava-Core               & 33.52                                                                                 & 33.55                                                                                     & 97.16                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+Google Auto Factory           & 48.41                                                                                 & 48.44                                                                                     & 97.71                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+Google Auto Common            & 59.59                                                                                 & 59.86                                                                                     & 96.35                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+Google Auto Value             & 61.42                                                                                 & 62.62                                                                                     & 97.31                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+Google Gson                   & 66.83                                                                                 & 66.81                                                                                     & 95.29                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+commons-io                    & 69.70                                                                                 & 66.05                                                                                     & 95.08                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+commons-cli                   & 79.07                                                                                 & 78.97                                                                                     & 95.55                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+commons-text                  & 79.29                                                                                 & 79.30                                                                                     & 95.34                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+commons-codec                 & 84.56                                                                                 & 84.59                                                                                     & 96.95                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+commons-csv                   & 85.16                                                                                 & 85.11                                                                                     & 95.87                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+\end{tabular}
+\caption{\label{tab:clustering_no_distance_50}Results of clustering mutants without Levensthein distance and cluster size n*0.50}
+\end{table*}
+
+Figure~\ref{box:clustering_no_distance_50} displays the box-plots of the weighted mutation score obtained from each individual sample. 
+We observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
+Table~\ref{tab:clustering_no_distance_50} displays the results of clustering mutants without Levensthein distance and the number of clusters equal to the total amount of mutants * 0.50.
+The maximum and minimum differences between the score of a full set and clustered set are 3.65\%, and 0.03\%, respectively.
+We observe that the average cluster accuracy is higher than the average cluster accuracy of the 25\% set. 
+The average of the differences between full set score and clustered score is 0.49\% which is smaller than the average difference than that of the 25\% set.
+
+\begin{figure}[ht]
+\includegraphics[width=0.5 \textwidth]{images/boxplot_summary/boxplot_hc_no_distance_0.75.png}
+\caption{\label{box:clustering_no_distance_75}Box-plots containing weighted mutation score of clustering mutants without Levensthein distance and cluster size n*0.75}
+\end{figure}
+
+\begin{table*}[ht]
+\centering
+\begin{tabular}{|l|c|c|c|c|c|}
+\hline
+\multicolumn{1}{|c|}{Project} & \begin{tabular}[c]{@{}c@{}}ms\\ full set\end{tabular} & \begin{tabular}[c]{@{}c@{}}ms \\ clustered set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg.
+ca\end{tabular} & \begin{tabular}[c]{@{}c@{}}Min.
+ca\end{tabular} & \begin{tabular}[c]{@{}c@{}}ca\end{tabular} \\ \hline
+Google Auto Service           & 0.00                                                                                  & 0.00                                                                                      & 0.00                                                                               & 0.00                                                                                & 0.00                                                                               \\ \hline
+ScribeJava-Core               & 33.52                                                                                 & 33.48                                                                                     & 98.63                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+Google Auto Factory           & 48.41                                                                                 & 48.34                                                                                     & 98.41                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+Google Auto Common            & 59.59                                                                                 & 59.71                                                                                     & 97.08                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+Google Auto Value             & 61.42                                                                                 & 62.66                                                                                     & 97.99                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+Google Gson                   & 66.83                                                                                 & 66.84                                                                                     & 97.37                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+commons-io                    & 69.70                                                                                 & 66.07                                                                                     & 97.34                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+commons-cli                   & 79.07                                                                                 & 79.12                                                                                     & 96.83                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+commons-text                  & 79.29                                                                                 & 79.33                                                                                     & 97.60                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+commons-codec                 & 84.56                                                                                 & 84.61                                                                                     & 98.43                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+commons-csv                   & 85.16                                                                                 & 85.07                                                                                     & 97.49                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+\end{tabular}
+\caption{\label{tab:clustering_no_distance_75}Results of clustering mutant without Levensthein distance and cluster size n*0.75}
+\end{table*}
+
+Figure~\ref{box:clustering_no_distance_75} displays the box-plots of the weighted mutation score obtained from each individual sample. 
+We observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
+Table~\ref{tab:clustering_no_distance_75} displays the results of clustering mutants without Levensthein distance and the number of clusters equal to the total amount of mutants * 0.50.
+The maximum and minimum differences between the score of a full set and clustered set are 3.63\%, and 0.04\%, respectively.
+We observe that the average accuracy is higher than the average accuracy of the other sets.
+The average of the differences between full set score and clustered score is  0.49\% which is higher than the average difference than that of the 25\% and the same of that of the 50\% set.
+
+\subsection{Fuzzy c-means}
+\label{ch:results_rq2}
+We present the results of our second experiment.
+We display the weighted mutation score to create box-plots.
+Similarly to Section~\ref{ch:results}, we removed the Levenshtein distance characteristic.
+
+The data presented is the average of the results of the 30 repetitions of the experiment.
+The result of each single repetition can be found in our Github repository in the folder \textit{experiment{\_}results}\cite{rbasarat-repo}.
+The fuzzy c-means model requires the amount of clusters as a parameter.
+Instead of displaying the reduction amount, as we did in the results of experiment one, we display the number of clusters that is calculated by the subtractive clustering model(See section \ref{ch:algorithm_parameters}.
+
+\begin{figure}[ht]
+\includegraphics[width=0.5 \textwidth]{images/fcm_summary/fcm_no_distance.png}
+\caption{\label{box:clustering_fcm_no_distance_25}Box-plots containing weighted mutation score of clustering mutants without Levenshtein distance characteristic.}
+\end{figure}
+
+\begin{table*}[ht]
+\centering
+\begin{tabular}{|l|l|l|l|l|l|l|}
+\hline
+Project               & \begin{tabular}[c]{@{}l@{}}No.
+clusters\end{tabular} & \begin{tabular}[c]{@{}l@{}}Ms full set\end{tabular} & \begin{tabular}[c]{@{}l@{}}Ms cluster- \\ed set\end{tabular} & \begin{tabular}[c]{@{}l@{}}Avg ca\end{tabular} & \begin{tabular}[c]{@{}l@{}}Min.
+ca\end{tabular} & \begin{tabular}[c]{@{}l@{}}Max.
+ca\end{tabular} \\ \hline
+Google Auto   Service & 17                   & 0.00                            & 0.00                                                 & 0.00                             & 0.00                                                                                                                                     & 0.00                                                                                                                                     \\ \hline
+ScribeJava-Core       & 32                   & 33.52                           & 23.64                                                & 76.80                            & 58.77                                                                                                                                    & 96.60                                                                                                                                    \\ \hline
+Google Auto   Factory & 22                   & 48.41                           & 57.40                                                & 59.20                            & 50.67                                                                                                                                    & 73.96                                                                                                                                    \\ \hline
+Google Auto   Common  & 31                   & 59.59                           & 37.64                                                & 62.95                            & 50.59                                                                                                                                    & 83.53                                                                                                                                    \\ \hline
+Google Auto   Value   & 41                   & 61.42                           & 59.61                                                & 61.48                            & 50.60                                                                                                                                    & 83.53                                                                                                                                    \\ \hline
+Google Gson           & 40                   & 66.83                           & 57.61                                                & 59.11                            & 50.84                                                                                                                                    & 76.68                                                                                                                                    \\ \hline
+Commons-io            & 40                   & 69.70                           & 57.95                                                & 57.25                            & 50.62                                                                                                                                    & 72.70                                                                                                                                    \\ \hline
+Commons-cli           & 26                   & 79.07                           & 81.53                                                & 81.77                            & 67.70                                                                                                                                    & 95.75                                                                                                                                    \\ \hline
+Commons-text          & 55                   & 79.29                           & 77.92                                                & 78.22                            & 61.80                                                                                                                                    & 93.92                                                                                                                                    \\ \hline
+Commons-codec         & 17                   & 84.56                           & 79.46                                                & 77.73                            & 74.39                                                                                                                                    & 81.66                                                                                                                                    \\ \hline
+Commons-csv           & 32                   & 85.16                           & 82.84                                                & 83.47                            & 65.58                                                                                                                                    & 97.80                                                                                                                                    \\ \hline
+Commons-lang          & 28                   & 78.31                           & 76.83                                                & 76.99                            & 73.03                                                                                                                                    & 80.56                                                                                                                                    \\ \hline
+Jfreechart            & 28                   & 27.02                           & 26.41                                                & 74.11                            & 71.83                                                                                                                                    & 76.36                                                                                                                                    \\ \hline
+Jodatime              & 25                   & 70.30                           & 65.17                                                & 64.75                            & 60.82                                                                                                                                    & 69.81                                                                                                                                    \\ \hline
+Zxing                 & 28                   & 70.64                           & 80.03                                                & 83.11                            & 80.10                                                                                                                                    & 86.03                                                                                                                                    \\ \hline
+\end{tabular}
+\caption{\label{tab:ml_no_distance}Results of clustering mutants with the fuzzy c-means model without Levenshtein distance}
+\end{table*}
+
+Figure~\ref{box:clustering_fcm_no_distance_25} displays the box-plots of the weighted mutation score obtained from each individual sample. 
+We observe that for every box-plot the \textit{p-value} is more than 0.05 or 5\%.
+Table~\ref{tab:ml_no_distance} displays the results of clustering mutants without Levenshtein distance.
+The maximum and minimum differences between the score of a full set and clustered set are 21.95\%, and 0.61\%, respectively.
+The average of the differences between full set score and clustered score is 4.44\% which is bigger than the average difference than that of the results with all characteristics.
+
+\section{Discussion}
+The maximum deviation from the original mutation score of the experiment with hierarchical clustering, is below the statistical threshold.
+We observe that the distances between the whiskers of the box-plots do not stretch more than the \textit{alpha-value}=0.05.
+This means that there is not one sample that deviated more than 5\% from the original mutation score.
+Thus we can reject the null hypothesis and accept our alternative hypothesis.
+
+As hypothesised we observe that it is possible to cluster mutants with the set of characteristics we identified while maintaining effectiveness and reducing the amount executed.
+The hypothesis holds for all the projects and variants of the experiment: 66\% of the displayed results have a deviation of less than 1\%.
+This includes both variants of the experiment.
+The results show that the set of characteristics can be used to cluster mutants efficiently.
+A higher cluster accuracy means a score closer to that of the original.
+If all clusters had an accuracy of 100\% we will achieve the same mutation score as that of the full set.
+However we do not observe this behaviour in our results.
+This is due to our random sampling method.
+If we would select a mutant in the majority of each cluster the accuracy per cluster would correlate to the deviation in mutation score.
+Unfortunately we do not know the majority of a cluster as we do not know if a mutant survives or is killed before execution.
+If we want to find out we need to execute every mutant that is generated which equals running a full suite.
+This defeats our purpose of increasing the performance by executing less mutants.
+
+The maximum deviation from the original mutation score of the experiment with the fuzzy c-means model, is above the statistical threshold.
+We observe that the distance between whiskers of the box-plots is more than the \textit{alpha-value} of 0.05.
+Thus we cannot reject the null hypothesis.
+
+The biggest deviation measured is 22.39\%.
+This is significantly higher than the biggest deviation measured in the first experiment.
+56\% of the results of both variants had a deviation in mutation score bigger than 5\%.
+Our hypothesis remains inconclusive for clustering the mutants with the fuzzy c-means model.
+
+We observe that the number of clusters does not have an obvious relation to the total amount of mutants. 
+This is expected as the subtractive clustering does not look at the total amount of data points(mutants) but to the contents(characteristics) of the data points.
+The number of clusters compared to the that of the hierarchical clustering is significantly lower.
+We did not research the exact clusters centroids, due to a limitation of resources.
+The subtractive model calculated a low number of clusters.
+This may indicate a close relationships between certain mutants, more research is needed.
+
+We observe that the minimum lowest minimum cluster accuracy measured over all variants and projects is 50.54\% with and average of 62\%. 
+This is higher than the minimum of the hierarchical clustering.
+This means that in all project there was not one cluster that was evenly divided.
+While the minimum accuracy was higher, the maximum accuracy was lower compared to the maximum accuracy of the hierarchical clustering results.
+The highest accuracy measured over all variants and projects is 97.80\% with an average of 83.79\%
+This means that in all project there was not one cluster that consisted of only survived or killed mutants.
+This is caused by the relatively low number of clusters.
+One cluster contains more mutants which results in a bigger weight per cluster.
 \section{Related work}
-\label{sec:rw}
+\label{ch:related_work}
+Clustering mutants has been done before.
+The research that exists tries to cluster mutants by defining the centroid of a cluster and cluster mutants on that definition.
+Our research extends mutation clustering by defining a centroid that is dynamic.
+In other words the centroids of our clusters are defined by the mutants itself.
+
+\subsection{Clustering overlapped mutants}
+\label{ch:overlapping_mutants}
+Ma et al., clustered mutants by overlap\cite{Ma2016}.
+They defined the term overlap as as mutants that are functionally equivalent.
+This is close to an equivalent mutant but not the same.
+They explain that an equivalent mutants is functionally identical to the original source code while an overlapped mutant is functionally identical to at least one other mutant.
+If a mutant does not have an overlapped mutant that mutant would be a cluster on its own.
+The achieved a reduction, in mutants executed, of 10\%.
+They write in their research that it is limited to the mutator operators on expression level. 
+If a mutator did not generate two or more mutants they could not detect overlapping mutants.
+Their future work includes adding more mutators and widening the scope level to statements or blocks.
+\newline
+The research does not include efficiency.
+While they show that they can cluster mutants based on overlap they do not show how efficient the mutation testing process is compared to executing a full set of mutants.
+A threat to their validity, which is not addressed, is the fact that two overlapping mutants may not be challenged by the same tests and thus can result in different results.
+Our research solves this problem by taking into account the location of the mutants.
+
+\subsection{Clustering by scope}
+Yu et al., extended the research about overlapping mutants(see section \ref{ch:overlapping_mutants})~\cite{Yu2019PossibilityScope}.
+They extended the scope by adding more mutators.
+The also extended the research to include statements and blocks.
+Their results shows an increase in clustered mutants.
+The research of Yu et al., suffers from the same limitations and threats  mentioned in Chapter \ref{ch:overlapping_mutants}.
+
+\subsection{Clustering Hamming distance}
+Ji et al., did a qualitative domain analysis on mutants\cite{Ji2009}.
+As a result they identified the Hamming distance to cluster with. 
+The Hamming distance is, like the Levenshtein distance, a similarity measure.
+They use the k-means algorithm to cluster mutants represented by Hamming distance.
+They write that the reduced test set in their experiment is still as strong as the original test set\cite{Ji2009}.
+\newline
+Ji et al., were successful in clustering mutants with Hamming distance and the k-means algorithm.
+They do acknowledge that their research still has problems with rationally determining the domains of their clusters.
+Our research does not have this problem, as the centre of the domains are decided by the characteristics of the mutants.
+
+\subsection{Spectral clustering}
+Wei et al., makes use of an intelligent technique, namely spectral clustering, to improve the efficacy of mutant reduction\cite{Wei2021SpectralTesting}.
+They defined multiple definitions for the mutants; similarity distance, distance between the mutants and a killing matrix of the mutants.
+With these definitions they reduced the mutants based on their proposed reduction method.
+This method is based on spectral clustering (SCMT), the determination method of the number of clusters, spectral clustering of mutants, and selection of representative mutants.
+The reduced mutants were then clustered with a classical clustering algorithm.
+Their results are promising and show high cluster accuracy.
+They write that there is still work left to do in optimizing the matrices and clustering algorithms.
+Our goal overlaps with that of Wei, the difference is that we used a different methodology to achieve that goal.
+Another difference is that our research does not require to know the result of a mutant(killed or survived).
+
+\subsection{Clustering similarity}
+Hussain et al., used the k-means and agglomerative clustering algorithm to cluster mutants according to a similarity measure. 
+They used the Hamming distance as similarity measure.
+They calculated the distance and used this as data to feed into the clustering algorithms.
+The number of clusters and the initial position of the cluster center in the k-means algorithm are difficult to determine, and the process of the agglomerative clustering algorithm makes it difficult to correct the existing cluster formation\cite{Hussain2008}.
+They did take into account efficiency.
+Hussain et al, had the same goals as our research. 
+The difference is that our research used different and more elaborated methods to cluster mutants.
+Hussain et al., calculated the mutation score by counting one cluster as one mutant.
+Our research calculated a weighted mutation score to reflect the clusters more accurately.
+
+\subsection{Generalizing mutants}
+Wilinski et al., tried to generalize the mutants by defining metrics.
+Each metric is calculated for a specific mutator.
+The three metrics are usefulness, frequency and dependency.
+Combining the results of usefulness and frequency metrics, they observed that reducing the number of generated mutants gives noticeable cost reduction without a loss of the mutation score accuracy.
+Wilinski et al., their research is narrowly scoped to the specific mutators they decided to do research on.
+Our research does not contain this limitation.
+The mutator is used as a characteristic in our research instead of limiting our research it helps define a mutant.
 
 \section{Conclusion}
-\label{sec:conclusion}
+\label{ch:conclusion}
+The high temporal cost requirement is often a barrier for adopting mutation testing~\cite{Pizzoleto2019}.
+An important technique to reduce the number of mutants, and thus the number of test case executions is mutation clustering~\cite{Ma2016,Yu2019PossibilityScope}, which has been researched with promising results~\cite{Ji2009,Wilinski2015,Ma2016}.
+However, the reduction in temporal cost is still insufficient when considering CI/CD pipelines, and the cost outweighs the benefits.
+
+To reduce the cost of mutation testing, we aim to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants.
+
+While preparing our research, we evaluated three mutation testing tools and found that two out of three mutation testing tools were not actively maintained anymore.
+The mutation testing tool PIT scored best on the metrics measured in our replication study, this is also the tool we decided to use for our research.
+
+We performed a qualitative analysis on mutant characteristics.
+As a result we created a list of identified characteristics based on existing research. We then extended this list by identifying characteristic with logic and data available.
+We used the identified characteristics to represent the mutants.
+
+We clustered the mutants with a white box approach and a black box approach.
+For our white box approach we selected the hierarchical clustering algorithm.
+The hierarchical clustering showed that we can cluster mutants while maintaining effectiveness.
+We achieved reductions of up to 75\% of the total amount of mutants without significantly reducing the accuracy.
+
+For the black box approach we selected a machine learning model, specifically the fuzzy c-means model.
+The fuzzy c-means model showed a significant reduction in effectiveness.
+We trained the fuzzy c-means model with parameters that were optimal according to existing research.
+We used a different machine learning model, specialized in calculating the number of clusters for fuzzy clustering, to calculate the number of clusters.
+While the minimum clustering accuracy was higher than that of the hierarchical clustering, the maximum cluster accuracy was lower than that of the hierarchical clustering.
+The cluster accuracy had a larger impact on the clusters of the machine learning model because there was a relatively small number of clusters.
+
+Our research shows that it is possible to determine the means of domains as the centre of a cluster within a programming language without narrowing down the scope to specific pieces of code.
+The centroids our research identified are decided by the characteristics of the mutants.
+It can be applied to all Java projects that can be run with PIT, which is a larger scope than in existing mutation clustering research.
 

### 2022-01-25 mutation-workshop/paper.tex
fixed tables

@@
 rasjaad@nivero.io}
-
+\and
+\IEEEauthorblockN{Jelle Mouissie}
+\IEEEauthorblockA{\textit{Informatics Institute} \\ %% TODO: CHANGE THIS
+\textit{University of Amsterdam}\\
+Amsterdam, the Netherlands \\
+jelle.mouissie@xs4all.nl}
+\and
+\IEEEauthorblockN{Thomas Biesaart}
+\IEEEauthorblockA{\textit{} \\ %% TODO: CHANGE THIS
+Amsterdam, the Netherlands \\
+thomas.biesaart@protonmail.com}
 \and
@@
 a.m.oprescu@uva.nl}
-\and
-\IEEEauthorblockN{Thomas Biesaart}
-\IEEEauthorblockA{\textit{} \\ %% TODO: CHANGE THIS
-Amsterdam, the Netherlands \\
-thomas.biesaart@protonmail.com}
 }
@@
 Including the Levenshtein distance defeats the purpose of increasing performance as every bit of performance we gain would be replaced by the cost of calculating the Levenshtein distance~\cite{thesis}.
-All results displayed are in percentages and ms=mutation score, ca=cluster accuracy.
 
-\subsection{Hierarchical clustering}
+\subsection{Hierarchical clustering and more}
 We present the results of our clustering algorithm when executed without the Levenshtein distance.
@@
 The result of each single repetition can be found in our Github repository in the folder\textit{experiment{\_}results}~\cite{rbasarat-repo}.
+We compare the cluster accuracy of hierarchical clustering with different clustering algorithms, the results can be found in Table \ref{tab:clustering_extra}. 
+Due to a limitation of resources the results only include the 75\% reduction variant of the experiment.
+We can observe that hierarchical clustering has the smallest deviation in cluster accuracy compared to the other clustering algorithms.
 
@@
 \centering
-\begin{tabular}{|l|c|c|c|c|c|}
+\begin{tabular}{|l|c|c|c|c|c|c|c|}
 \hline
-\multicolumn{1}{|c|}{Project} & \begin{tabular}[c]{@{}c@{}}ms\\ full set\end{tabular} & \begin{tabular}[c]{@{}c@{}}ms \\ clustered set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg.
-ca\end{tabular} & \begin{tabular}[c]{@{}c@{}}Min.
-ca\end{tabular} & \begin{tabular}[c]{@{}c@{}}ca\end{tabular} \\ \hline
-Google Auto Service           & 0.00                                                                                  & 0.00                                                                                      & 0.00                                                                               & 0.00                                                                                & 0.00                                                                               \\ \hline
-ScribeJava-Core               & 33.52                                                                                 & 33.72                                                                                     & 92.24                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-Google Auto Factory           & 48.41                                                                                 & 48.32                                                                                     & 94.21                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-Google Auto Common            & 59.59                                                                                 & 60.03                                                                                     & 91.19                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-Google Auto Value             & 61.42                                                                                 & 62.58                                                                                     & 93.90                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-Google Gson                   & 66.83                                                                                 & 66.89                                                                                     & 91.75                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-commons-io                    & 69.70                                                                                 & 66.08                                                                                     & 90.20                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-commons-cli                   & 79.07                                                                                 & 78.97                                                                                     & 91.11                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-commons-text                  & 79.29                                                                                 & 79.30                                                                                     & 90.95                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-commons-codec                 & 84.56                                                                                 & 84.59                                                                                     & 94.56                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-commons-csv                   & 85.16                                                                                 & 85.24                                                                                     & 92.76                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
+\multicolumn{1}{|c|}{Project} & \begin{tabular}[c]{@{}c@{}}ms.\\ full set\end{tabular} & \begin{tabular}[c]{@{}c@{}}ms.\\ clustered set\end{tabular} & Min.
+ca.
+& max.
+ca.
+& \begin{tabular}[c]{@{}c@{}}Avg.
+ca.\\ size*0.25\end{tabular} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Avg.
+ca.\\ size*0.50\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Avg.
+ca.\\ size*0.75\end{tabular}} \\ \hline
+Google Auto Service           & 0.00                                                   & 0.00                                                        & 0.00     & 0.00     & 0.00                                                           & 0.00                                                                              & 0.00                                                                              \\ \hline
+ScribeJava-Core               & 33.52                                                  & 33.72                                                       & 50.00    & 100.00   & 92.24                                                          & 97.16                                                                             & 98.63                                                                             \\ \hline
+Google Auto Factory           & 48.41                                                  & 48.32                                                       & 50.00    & 100.00   & 94.21                                                          & 97.71                                                                             & 98.41                                                                             \\ \hline
+Google Auto Common            & 59.59                                                  & 60.03                                                       & 50.00    & 100.00   & 91.19                                                          & 96.35                                                                             & 97.08                                                                             \\ \hline
+Google Auto Value             & 61.42                                                  & 62.58                                                       & 50.00    & 100.00   & 93.90                                                          & 97.31                                                                             & 97.99                                                                             \\ \hline
+Google Gson                   & 66.83                                                  & 66.89                                                       & 50.00    & 100.00   & 91.75                                                          & 95.29                                                                             & 97.37                                                                             \\ \hline
+commons-io                    & 69.70                                                  & 66.08                                                       & 50.00    & 100.00   & 90.20                                                          & 95.08                                                                             & 97.34                                                                             \\ \hline
+commons-cli                   & 79.07                                                  & 78.97                                                       & 50.00    & 100.00   & 91.11                                                          & 95.55                                                                             & 96.83                                                                             \\ \hline
+commons-text                  & 79.29                                                  & 79.30                                                       & 50.00    & 100.00   & 90.95                                                          & 95.34                                                                             & 97.60                                                                             \\ \hline
+commons-codec                 & 84.56                                                  & 84.59                                                       & 50.00    & 100.00   & 94.56                                                          & 96.95                                                                             & 98.43                                                                             \\ \hline
+commons-csv                   & 85.16                                                  & 85.24                                                       & 50.00    & 100.00   & 92.76                                                          & 95.87                                                                             & 97.49                                                                             \\ \hline
 \end{tabular}
-\caption{\label{tab:clustering_no_distance_25}Results of clustering mutants without Levensthein distance and cluster size n*0.25}
+\caption{\label{tab:clustering_no_distance}Results of clustering mutants with hierarchical clustering where and ms=mutation score in \%, ca=cluster accuracy in \% and size the total amount of mutants generated.}
 \end{table*}
 
-Figure \ref{box:clustering_no_distance_25} displays the box-plots of the weighted mutation score obtained from each individual sample.
+\begin{table*}[ht]
+\centering
+\begin{tabular}{|l|c|c|c|c|c|c|c|}
+\hline
+\multicolumn{1}{|c|}{Project} & \begin{tabular}[c]{@{}c@{}}Ms.
+\\ full set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg.
+ca.\\ random\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg.
+ca.\\ K-means\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg.
+ca.\\ Birch\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg.
+ca.\\ DBScan\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg.
+ca.\\ Optics\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg.
+ca.\\ MeanShift\end{tabular} \\ \hline
+Google Auto Service           & 0.00                                                  & 0.00                                                    & 0.00                                                     & 0.00                                                   & 0.00                                                    & 0.00                                                    & 0.00                                                       \\ \hline
+ScribeJava-Core               & 33.52                                                 & 76.65                                                   & 93.21                                                    & 93.26                                                  & 90.91                                                   & 92.09                                                   & 94.51                                                      \\ \hline
+Google Auto Factory           & 48.41                                                 & 72.56                                                   & 94.36                                                    & 94.20                                                  & 91.41                                                   & 92.22                                                   & 95.00                                                      \\ \hline
+Google Auto Common            & 59.76                                                 & 73.89                                                   & 91.09                                                    & 91.12                                                  & 87.36                                                   & 88.09                                                   & 93.73                                                      \\ \hline
+Google Auto Value             & 62.68                                                 & 75.27                                                   & 94.23                                                    & 94.33                                                  & 90.59                                                   & 91.77                                                   & 94.90                                                      \\ \hline
+Google Gson                   & 66.84                                                 & 76.56                                                   & 91.77                                                    & 91.62                                                  & 87.93                                                   & 88.30                                                   & 93.03                                                      \\ \hline
+commons-io                    & 65.47                                                 & 76.08                                                   & 90.73                                                    & 90.43                                                  & 86.26                                                   & 87.39                                                   & 92.73                                                      \\ \hline
+commons-cli                   & 79.10                                                 & 83.32                                                   & 91.36                                                    & 91.36                                                  & 86.85                                                   & 87.88                                                   & 92.77                                                      \\ \hline
+commons-text                  & 79.33                                                 & 83.48                                                   & 90.28                                                    & 89.95                                                  & 86.30                                                   & 87.11                                                   & 91.96                                                      \\ \hline
+commonc-codec                 & 84.61                                                 & 87.08                                                   & 94.46                                                    & 94.44                                                  & 92.08                                                   & 91.69                                                   & 95.41                                                      \\ \hline
+commons-csv                   & 85.07                                                 & 87.28                                                   & 92.26                                                    & 92.08                                                  & 87.95                                                   & 89.65                                                   & 93.10                                                      \\ \hline
+\end{tabular}
+\caption{\label{tab:clustering_extra}Results of clustering mutants with different clustering algorithms where and ms=mutation score in \%, ca=cluster accuracy in \%}
+\end{table*}
+
+Figure \ref{box:clustering_no_distance_25}, \ref{box:clustering_no_distance_50}, \ref{box:clustering_no_distance_75} displays the box-plots of the weighted mutation score obtained from each individual sample.
 We observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
-Table~\ref{tab:clustering_no_distance_25} displays the results of clustering mutants without Levensthein distance and the number of clusters equal to the total amount of mutants * 0.25. 
-We observe that the maximum and minimum differences in mutation score between a full set and a clustered set  3.61\%, and 0.09\%, respectively.
-The average of the differences between full set score and clustered score is 0.53\%.
+Table~\ref{tab:clustering_no_distance} displays the results of clustering mutants 
+We observe that the maximum differences in mutation score between a full set and a clustered set are 3.61\%, 3.65\% and 3.63\% with a reduction of 75\%, 50\% and 25\% respectively.
+We observe that the minimum differences in mutation score between a full set and a clustered set are 0.49\%, 0.03\% and 10.04\%  with a reduction of 75\%, 50\% and 25\% respectively.
+The average of the differences between full set score and clustered score are 0.58\% 0.49\%, 0.49\% with a reduction of 75\%, 50\% and 25\% respectively.
 
@@
 
-\begin{table*}[ht]
-\centering
-\begin{tabular}{|l|c|c|c|c|c|}
-\hline
-\multicolumn{1}{|c|}{Project} & \begin{tabular}[c]{@{}c@{}}ms\\ full set\end{tabular} & \begin{tabular}[c]{@{}c@{}}ms \\ clustered set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg.
-ca\end{tabular} & \begin{tabular}[c]{@{}c@{}}Min.
-ca\end{tabular} & \begin{tabular}[c]{@{}c@{}}ca\end{tabular} \\ \hline
-Google Auto Service           & 0.00                                                                                  & 0.00                                                                                      & 0.00                                                                               & 0.00                                                                                & 0.00                                                                               \\ \hline
-ScribeJava-Core               & 33.52                                                                                 & 33.55                                                                                     & 97.16                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-Google Auto Factory           & 48.41                                                                                 & 48.44                                                                                     & 97.71                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-Google Auto Common            & 59.59                                                                                 & 59.86                                                                                     & 96.35                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-Google Auto Value             & 61.42                                                                                 & 62.62                                                                                     & 97.31                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-Google Gson                   & 66.83                                                                                 & 66.81                                                                                     & 95.29                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-commons-io                    & 69.70                                                                                 & 66.05                                                                                     & 95.08                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-commons-cli                   & 79.07                                                                                 & 78.97                                                                                     & 95.55                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-commons-text                  & 79.29                                                                                 & 79.30                                                                                     & 95.34                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-commons-codec                 & 84.56                                                                                 & 84.59                                                                                     & 96.95                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-commons-csv                   & 85.16                                                                                 & 85.11                                                                                     & 95.87                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-\end{tabular}
-\caption{\label{tab:clustering_no_distance_50}Results of clustering mutants without Levensthein distance and cluster size n*0.50}
-\end{table*}
-
-Figure~\ref{box:clustering_no_distance_50} displays the box-plots of the weighted mutation score obtained from each individual sample. 
-We observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
-Table~\ref{tab:clustering_no_distance_50} displays the results of clustering mutants without Levensthein distance and the number of clusters equal to the total amount of mutants * 0.50.
-The maximum and minimum differences between the score of a full set and clustered set are 3.65\%, and 0.03\%, respectively.
-We observe that the average cluster accuracy is higher than the average cluster accuracy of the 25\% set. 
-The average of the differences between full set score and clustered score is 0.49\% which is smaller than the average difference than that of the 25\% set.
-
 \begin{figure}[ht]
@@
 \end{figure}
-
-\begin{table*}[ht]
-\centering
-\begin{tabular}{|l|c|c|c|c|c|}
-\hline
-\multicolumn{1}{|c|}{Project} & \begin{tabular}[c]{@{}c@{}}ms\\ full set\end{tabular} & \begin{tabular}[c]{@{}c@{}}ms \\ clustered set\end{tabular} & \begin{tabular}[c]{@{}c@{}}Avg.
-ca\end{tabular} & \begin{tabular}[c]{@{}c@{}}Min.
-ca\end{tabular} & \begin{tabular}[c]{@{}c@{}}ca\end{tabular} \\ \hline
-Google Auto Service           & 0.00                                                                                  & 0.00                                                                                      & 0.00                                                                               & 0.00                                                                                & 0.00                                                                               \\ \hline
-ScribeJava-Core               & 33.52                                                                                 & 33.48                                                                                     & 98.63                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-Google Auto Factory           & 48.41                                                                                 & 48.34                                                                                     & 98.41                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-Google Auto Common            & 59.59                                                                                 & 59.71                                                                                     & 97.08                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-Google Auto Value             & 61.42                                                                                 & 62.66                                                                                     & 97.99                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-Google Gson                   & 66.83                                                                                 & 66.84                                                                                     & 97.37                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-commons-io                    & 69.70                                                                                 & 66.07                                                                                     & 97.34                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-commons-cli                   & 79.07                                                                                 & 79.12                                                                                     & 96.83                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-commons-text                  & 79.29                                                                                 & 79.33                                                                                     & 97.60                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-commons-codec                 & 84.56                                                                                 & 84.61                                                                                     & 98.43                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-commons-csv                   & 85.16                                                                                 & 85.07                                                                                     & 97.49                                                                              & 50.00                                                                               & 100.00                                                                             \\ \hline
-\end{tabular}
-\caption{\label{tab:clustering_no_distance_75}Results of clustering mutant without Levensthein distance and cluster size n*0.75}
-\end{table*}
-
-Figure~\ref{box:clustering_no_distance_75} displays the box-plots of the weighted mutation score obtained from each individual sample. 
-We observe that for every box-plot the \textit{p-value} is below 0.05 or 5\%.
-Table~\ref{tab:clustering_no_distance_75} displays the results of clustering mutants without Levensthein distance and the number of clusters equal to the total amount of mutants * 0.50.
-The maximum and minimum differences between the score of a full set and clustered set are 3.63\%, and 0.04\%, respectively.
-We observe that the average accuracy is higher than the average accuracy of the other sets.
-The average of the differences between full set score and clustered score is  0.49\% which is higher than the average difference than that of the 25\% and the same of that of the 50\% set.
 
